\documentclass[11pt,a4paper]{article}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{geometry}
\usepackage{hyperref}
\usepackage{booktabs}
\usepackage{enumitem}
\geometry{margin=2.5cm}

\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{remark}[theorem]{Remark}

\title{The Cascade Series --- Part I\\[0.5em]
\large The Cosmological Constant from the Observer's Frame}
\author{RTAC}
\date{March 2026}

\begin{document}
\maketitle

\begin{abstract}
Part 0~\cite{part0} derives the cascade invariant
$I_0 = \Omega_{19}\times\Omega_{217} = 1.2051\times 10^{-120}$ from pure
mathematics. This paper adds one input --- the hypothesis that the
cascade's geometry is indistinguishable from our universe --- and
applies two cascade-native sphere-area corrections to convert the
pure-number invariant into the dimensionless cosmological constant
measured in reduced-Planck units.

The observer at $d = 4$ lives on the $S^3$ shell of a host 5D black
hole, experiencing ``time'' as an asymptotic fall toward the hole's
pseudo-horizon: the cascade's own descent direction (Part~II
Section~7.1). Time is \emph{locked} as this infall, not a free fourth
dimension. Consequently only three dimensions --- the angular
dimensions of the spatial $S^3$ --- are free to receive cascade
content, and the relevant 3-volume at the observer's frame is
$\Omega_3 = 2\pi^2$.

Two corrections connect the cascade invariant to the physical ratio:
(i)~the host-frame rescaling $(\Omega_5/\Omega_7)^2 = 9/\pi^2$ from the
cascade's natural reference at $d_0 = 7$ to the observer's host at
$d_V = 5$ (both cascade-internal sphere ratios); (ii)~the 3D projection
factor $\Omega_2/\Omega_3 = 2/\pi$ at the observer's frame, which is
the latitudinal integral $\int_0^\pi\sin^2\theta\,d\theta = \pi/2$
expressed as a cascade sphere-area ratio and arises because the Morse
foliation of $S^3$ by 2-spheres is the projection from per-$S^2$
content density (the cascade-native unit at each latitude of the
locked-time $S^3$ slice) to integrated $S^3$ 3-volume. Combined:
\[
\frac{\rho_\Lambda}{M^4_{\mathrm{Pl,red}}}
= \frac{\Omega_2}{\Omega_3}\cdot\left(\frac{\Omega_5}{\Omega_7}\right)^{\!2}\cdot\Omega_{19}\Omega_{217}
= \frac{2}{\pi}\cdot\frac{9\;\Omega_{19}\;\Omega_{217}}{\pi^2}
= \frac{18\;\Omega_{19}\;\Omega_{217}}{\pi^3}
= 6.996\times 10^{-121}.
\]
Observed (Planck 2018): $(7.150\pm 0.13)\times 10^{-121}$. Leading
deviation: $-2.2\%$, placing $\Lambda$ in the descent-dependent
population. After the Part~0 Supplement first-order Gram correction
$\exp(\sum(1-C^2_{d,d+1}))=\exp(0.02108)$, the residual is $-0.07\%$,
well inside the $1.9\%$ Planck $1\sigma$ uncertainty.

Every factor is a cascade-native sphere-area ratio: six sphere areas
($\Omega_2,\Omega_3,\Omega_5,\Omega_7,\Omega_{19},\Omega_{217}$), zero
imported objects. The QFT comparison in Section~\ref{sec:qft-comparison}
traces the ``$10^{120}$ cosmological constant problem'' to three
structural disagreements: (i) the number of free dimensions at the
observer's frame (4 for QFT, 3 for the cascade, because time is
locked), (ii) continuous vs discrete mode spectrum, and
(iii) hand-imposed UV cutoff vs Gamma-function terminus.
\end{abstract}

\section{The Hypothesis}

\begin{definition}[Hypothesis]\label{def:hyp}
The infinite-dimensional unit ball, descended to four dimensions, is
indistinguishable from our universe.
\end{definition}

This is the single input beyond Part 0. It is tested by every prediction
in the series. It is not an identification, not a model, not an
approximation. It is a factual claim: no measurement can distinguish the
cascade's geometry from the observed universe.

\section{The Observer's Frame}

The observer at $d = 4$ lives on the $S^3$ shell of a host 5D black
hole (Cover Sheet, ``The Thought Experiment''). The 3 angular
dimensions of $S^3$ are the observer's genuinely free spatial
dimensions; time is the infall direction toward the hole's
pseudo-horizon, asymptotic and never completing --- the cascade's own
descent parameter under the identification of Part~II~\cite{paper2},
Section~7.1.

\begin{remark}[Why the observer's host is $d_V = 5$]\label{rem:sp12-status}
The identification ``observer's host is $d_V = 5$'' is not an ad hoc
modelling choice imported from the Cover Sheet's thought experiment
alone: it is forced by combining one empirical input with three
cascade-internal theorems. The forcing chain is:
\begin{enumerate}
\item \emph{Empirical input.} The observer's spacetime is
4-dimensional. This single empirical fact is the only non-cascade
input in what follows (CLAUDE.md's acknowledged ``observer at
$d=4$'').
\item \emph{Foliation structure (Part~III, \S4).} The cascade's
slicing of a unit ball $B^d$ foliates it into $(d-1)$-ball
cross-sections $B^{d-1}$ parameterised by the slicing axis; under
the Part~II identification of the slicing direction with time, the
observer's spatial cross-section at each instant is the boundary
$S^{d-2} = \partial B^{d-1}$ of the cross-section.
\item \emph{Dimensional matching.} A 4D observer's spatial
cross-section is 3-dimensional, so $S^{d-2}$ must be an $S^3$,
forcing $d - 2 = 3$, hence $d = 5$. The observer's host ball is
therefore $B^5$, independent of any further cascade input.
\item \emph{Cascade confirmation.} That $d = 5$ independently
coincides with $d_V$, the volume maximum of Part~0's tower
(Paper~0, Theorem~7.1 ``Tower completeness''), is a \emph{structural
confirmation}, not an assumption: the 4D observer's host, forced by
steps (1)--(3), happens to sit at the cascade's geometrically
richest layer. Boundary dominance ($\Omega_{d-1}/V_d = d$) then
assigns $4/5$ of the $d=5$ content to the $S^3$ shell on which the
observer resides, as stated below.
\end{enumerate}
Under this chain, the ``observer's host'' identification licenses
Correction~1 below without the Cover Sheet's thought experiment
being taken as a premise of Part~I: the thought experiment
illustrates steps (1)--(3), and the cascade's tower structure supplies
step (4). The step that remains properly empirical is only the
dimensionality of the observer; every geometric consequence drawn
from it is cascade-internal.
\end{remark}

By Part~0's boundary dominance ($\Omega_{d-1}/V_d = d$), the $S^3$
shell carries $4/5$ of the $d=5$ content. The dimension $d_V = 5$ is
the volume maximum of Part~0's tower: the geometrically richest
interior in the cascade. Part~0's threshold pair ($\Omega_{19}$,
$\Omega_{217}$) is generated by the natural zero at $d_0 = 7$. The
invariant $I_0 = \Omega_{19}\times\Omega_{217}$ is intrinsically
referenced to this area maximum. The observer is not at $d_0 = 7$.
The observer is at $d_V = 5$, with only 3 free spatial dimensions.

Two cascade-native sphere-area corrections are required to connect the
pure-number cascade invariant to the physical cosmological-constant
ratio measured in reduced-Planck units. The first rescales between the
cascade's natural reference at $d_0 = 7$ and the observer's host at
$d_V = 5$; the second projects the cascade's per-$S^2$ content density
onto the observer's 3D spatial slice under locked time.

\subsection{Correction 1: the host frame at $d_V = 5$}

The ratio of sphere areas at the two maxima:
\[
\frac{\Omega_5}{\Omega_7} = \frac{\pi^3}{\pi^4/3} = \frac{3}{\pi}.
\]
Applied once per threshold sphere area (both factors of the bilinear
invariant pass through the host rescaling; Part~0, Lemma~8.8), this
contributes $(\Omega_5/\Omega_7)^2 = 9/\pi^2$.

\subsection{Correction 2: projection onto the observer's 3D spatial slice}

Time is not a free fourth dimension at the observer's frame. In the
cover sheet's thought experiment, the observer sits on the $S^3$ shell of a
host 5D black hole and experiences ``time'' as the asymptotic fall
toward the hole's pseudo-horizon. Part~II~\cite{paper2}, Section~7.1
makes this identification explicit at the cascade-mathematical level:
time is the cascade's own slicing direction, the parameter along which
the descent proceeds. The infall is asymptotic --- the observer never
crosses --- so the total ``time'' available at any finite moment is
bounded in the cascade's frame. Time is \emph{locked} into the cascade's
descent parameter, not a free direction the observer can choose.

\medskip\noindent\textbf{The structural consequence.}
Cascade content descending from the higher-dimensional layers has only
\emph{three} free dimensions to project onto at the observer's frame,
not four. Content arriving from the compactified interior layers
$d = 5, \ldots, 217$ passes through the locked-time direction (the
infall coordinate) and lands on the observer's 3-sphere $S^3$, which is
their entire receiving volume at any given moment of the infall.

The natural 3-volume at the observer's frame is therefore
$\Omega_3 = 2\pi^2$, the area of the unit $S^3$ --- a cascade-native
quantity (Part~0, Corollary~3.2: sphere area is the unique independent
cascade quantity), with no imported cube-volume normalisation.

\medskip\noindent\textbf{The projection factor from the Morse foliation.}
The 3-sphere admits a canonical Morse foliation by 2-spheres at constant
latitude: writing $S^3 = \{(\cos\theta,\, \sin\theta\,\vec n) :
\theta\in[0,\pi],\; \vec n\in S^2\}$, each latitude $\theta$ carries an
equatorial 2-sphere of area $\sin^2\theta \cdot \Omega_2$. Integrating
latitude:
\[
\Omega_3 \;=\; \int_0^\pi \sin^2\theta\cdot\Omega_2 \, d\theta \;=\; \frac{\pi}{2}\cdot\Omega_2,
\]
which is the elementary identity
$\int_0^\pi \sin^2\theta\,d\theta = \pi/2$ expressed in sphere-area
units. Inverted:
\[
\frac{\Omega_2}{\Omega_3} \;=\; \frac{4\pi}{2\pi^2} \;=\; \frac{2}{\pi}.
\]

This is the projection factor. Cascade content arriving at the
observer's frame is natively per-$S^2$ (at each latitude of the $S^3$
foliation); converting to integrated $S^3$ 3-volume introduces the
$\pi/2$ latitudinal integral, and its inverse $2/\pi$ is the factor by
which per-$S^2$ content densities are compressed when spread over the
full 3-sphere receiving volume. The conversion is purely in cascade-native
sphere-area terms --- no cube, no QFT mode-counting measure. Both
$\Omega_2$ and $\Omega_3$ are unit-sphere areas at dimensions
immediately adjacent to the observer; their ratio is a Gamma-function
identity.

\begin{remark}[Why 3 dimensions, not 4]\label{rem:why-3d}
The cosmological constant is an energy density, and an energy density
is an extensive quantity per receiving volume. The receiving volume at
the observer's frame has dimension equal to the number of \emph{free}
dimensions the observer actually has --- not the number of spacetime
dimensions in the abstract. In the cascade picture, three dimensions
are free (the angular dimensions of $S^3$) and one is locked (time, as
the cascade's own descent / the observer's asymptotic infall). The
projection is therefore over three dimensions, and the natural 3-volume
$\Omega_3 = 2\pi^2$ is the integrated Morse foliation of the per-$S^2$
content density.

The appearance that time is free --- that the observer can ``choose'' a
4-volume $d^4 x$ reference --- is a feature of local flat-space
approximations on the observer's patch of $S^3$, not of the cascade's
global structure. Over any cosmological scale where the cascade's
asymptotic infall becomes visible (e.g., in $\rho_\Lambda$, which
involves the full cascade tower $\Omega_{19}\Omega_{217}$), the
locked-time character of the cascade's descent forbids projection onto
a free fourth dimension and enforces the 3D projection factor
$\Omega_2/\Omega_3 = 2/\pi$.

Section~\ref{sec:qft-comparison} develops the QFT contrast: QFT
implicitly assumes time is free (4D Minkowski with unbounded time
evolution) and therefore counts modes over 4 free dimensions; the
cascade has time locked and therefore projects over only 3. This is
the structural origin of the $\pi/2$ mismatch at the observer's frame.
\end{remark}

\subsection{Unified descent: the three corrections are one operation}
\label{subsec:unified-descent}

Sections~3.1 and~3.2 introduced the host frame correction
$(\Omega_5/\Omega_7)^2 = 9/\pi^2$ and the 3D projection correction
$\Omega_2/\Omega_3 = 2/\pi$ as separate cascade-native sphere-area
identities. This subsection shows they are not separate: both are
instances of Part~0's slicing recurrence evaluated at different
cascade layers, with the Part~0 bilinear invariant
$\Omega_{19}\Omega_{217}$ as the remaining threshold content. The
three corrections of Theorem~\ref{thm:cc} collapse to one operation
applied at the cascade's distinguished dimensions; no second operation
enters.

\medskip\noindent\textbf{The slicing recurrence as the sole operation.}
Part~0~Theorem~3.1 establishes the cascade's unique first-order
multiplicative map between sphere areas:
\begin{equation}\label{eq:recurrence-forward}
\Omega_d \;=\; \Omega_{d-1}\cdot\sqrt{\pi}\cdot R(d-1),
\qquad R(d) := \frac{\Gamma((d+1)/2)}{\Gamma((d+2)/2)}.
\end{equation}
Inverting, the backward step is
\begin{equation}\label{eq:recurrence-backward}
\frac{\Omega_{d-1}}{\Omega_d} \;=\; \frac{1}{\sqrt{\pi}\,R(d-1)}.
\end{equation}
The claim of this subsection is that both frame corrections are
instances of~\eqref{eq:recurrence-backward}, evaluated at different
cascade layers; no second operation enters.

\begin{lemma}[3D projection as one backward slicing step]\label{lem:projection-step}
The 3D projection correction equals one backward step of the slicing
recurrence, evaluated at $d=3$:
\[
\frac{\Omega_2}{\Omega_3}
\;=\; \frac{1}{\sqrt{\pi}\,R(2)}
\;=\; \frac{2}{\pi}.
\]
\end{lemma}

\begin{proof}
Direct evaluation using Gamma function values:
$R(2) = \Gamma(3/2)/\Gamma(2) = (\sqrt{\pi}/2)/1 = \sqrt{\pi}/2$.
Therefore $1/(\sqrt{\pi}\,R(2)) = 1/(\sqrt{\pi}\cdot\sqrt{\pi}/2) = 2/\pi$.
By~\eqref{eq:recurrence-backward} at $d = 3$ this equals
$\Omega_2/\Omega_3$. Equivalently, $\Omega_2 = 4\pi$ and
$\Omega_3 = 2\pi^2$ give the same value directly.
\end{proof}

\begin{lemma}[Host frame as two backward slicing steps]\label{lem:host-step}
The host frame correction per threshold sphere area equals the composition
of two backward slicing steps, evaluated at $d = 7$ and $d = 6$:
\[
\frac{\Omega_5}{\Omega_7}
\;=\; \frac{1}{\sqrt{\pi}\,R(6)}\cdot\frac{1}{\sqrt{\pi}\,R(5)}
\;=\; \frac{1}{\pi\,R(5)\,R(6)}
\;=\; \frac{3}{\pi}.
\]
\end{lemma}

\begin{proof}
By~\eqref{eq:recurrence-backward}: $\Omega_6/\Omega_7 = 1/(\sqrt{\pi} R(6))$
and $\Omega_5/\Omega_6 = 1/(\sqrt{\pi} R(5))$. Composing:
$\Omega_5/\Omega_7 = 1/(\pi R(5) R(6))$. Gamma function evaluation:
$R(5) = \Gamma(3)/\Gamma(7/2) = 2/((15\sqrt{\pi})/8) = 16/(15\sqrt{\pi})$
and $R(6) = \Gamma(7/2)/\Gamma(4) = (15\sqrt{\pi}/8)/6 = 5\sqrt{\pi}/16$.
Their product: $R(5)\,R(6) = (16/(15\sqrt{\pi}))\cdot(5\sqrt{\pi}/16) = 5/15 = 1/3$.
Therefore $\Omega_5/\Omega_7 = 1/(\pi/3) = 3/\pi$. Squared for the
bilinear invariant: $(\Omega_5/\Omega_7)^2 = 9/\pi^2$, as used in
Section~3.1.
\end{proof}

The numerical values in Lemmas~\ref{lem:projection-step}
and~\ref{lem:host-step} are not opaque Gamma-function outputs --- they
are the index structure of the cascade's observer layers expressed
through a telescoping identity.

\begin{lemma}[Consecutive slicing steps telescope]\label{lem:telescoping}
For any $d$ with $(d+1)/2$ a positive integer, two consecutive
slicing coefficients satisfy
\[
R(d)\,R(d+1)
\;=\; \frac{\Gamma((d+1)/2)}{\Gamma((d+3)/2)}
\;=\; \frac{2}{d+1}.
\]
\end{lemma}

\begin{proof}
$R(d)\,R(d+1) = \Gamma((d+1)/2)\Gamma((d+2)/2) / [\Gamma((d+2)/2)\,\Gamma((d+3)/2)] = \Gamma((d+1)/2)/\Gamma((d+3)/2)$.
The intermediate Gamma factor $\Gamma((d+2)/2)$ cancels. Applying
$\Gamma(x+1) = x\,\Gamma(x)$ with $x = (d+1)/2$ gives
$\Gamma((d+3)/2) = ((d+1)/2)\,\Gamma((d+1)/2)$, hence the ratio is
$1/((d+1)/2) = 2/(d+1)$.
\end{proof}

\begin{remark}[The ``3'' in $3/\pi$ is the observer-host index]
\label{rem:host-index}
Lemma~\ref{lem:telescoping} at $d = 5$ gives $R(5)\,R(6) = 2/6 = 1/3$,
so the two-step cascade descent from the area maximum $d_0 = 7$ to the
observer's host $d_V = 5$ evaluates to
\[
\frac{\Omega_5}{\Omega_7}
\;=\; \frac{1}{\pi\,R(5)\,R(6)}
\;=\; \frac{d_V + 1}{2\pi}
\;=\; \frac{6}{2\pi}
\;=\; \frac{3}{\pi}.
\]
The factor~3 in the host frame correction is $(d_V + 1)/2 = 3$: the
observer-host dimension plus one, halved. Each slicing step
contributes one $\sqrt{\pi}$ and the telescoping removes the
intermediate Gamma function exactly. Similarly the single-step
3D projection factor uses
$R(2) = \Gamma(3/2)/\Gamma(2) = \sqrt{\pi}/2$, the simplest non-trivial
half-integer Gamma ratio --- the observer's receiving dimension
$d = 3$ encoded in one step. No Gamma function value in
Theorem~\ref{thm:cc} is structurally opaque; each is a telescoping
ratio at a cascade-distinguished dimension.
\end{remark}

\begin{theorem}[Unified descent]\label{thm:unified}
Every factor in Theorem~\ref{thm:cc} is either a threshold sphere area
at a Part~0 distinguished dimension or a backward slicing step of
\eqref{eq:recurrence-backward} at a cascade layer. Explicitly:
\[
\frac{\rho_\Lambda}{M^4_{\mathrm{Pl,red}}}
\;=\; \underbrace{\frac{1}{\sqrt{\pi}\,R(2)}}_{\text{step at }d=3}
\;\cdot\;\underbrace{\left(\frac{1}{\sqrt{\pi}\,R(5)\,\sqrt{\pi}\,R(6)}\right)^{\!2}}_{\text{two steps at }d=6,7,\text{ squared}}
\;\cdot\;\underbrace{\Omega_{19}}_{d_1}
\;\cdot\;\underbrace{\Omega_{217}}_{d_2}.
\]
\end{theorem}

\begin{proof}
Substitute Lemmas~\ref{lem:projection-step} and~\ref{lem:host-step}
into Theorem~\ref{thm:cc}. Every remaining factor is a sphere area
$\Omega_d$ at one of Part~0's four distinguished dimensions
(Part~0~Theorem~7.1): $d_V = 5$ (host; implicitly through the host
frame), $d_0 = 7$ (cascade reference; implicitly through the host
frame), $d_1 = 19$, $d_2 = 217$. The factor at $d = 3$ is the
observer's spatial dimension, forced by locked time
(Section~3.2). No additional operations enter.
\end{proof}

\begin{remark}[No foreign machinery]
Theorem~\ref{thm:unified} is a cascade-internal identity, proved from
Part~0's slicing recurrence alone. No Kaluza--Klein reduction, no
semiclassical integration over compactified dimensions, no external
normalisation convention. The recurrence is a theorem about the Gamma
function (Part~0~Theorem~3.1); the distinguished dimensions are theorems
about the Gamma function (Part~0~Theorem~7.1); the Part~I formula is
the product of recurrence backward-steps at a specific set of these
dimensions, times the threshold sphere areas themselves. The three
``separate'' corrections originally introduced in Sections~3.1 and~3.2
are now seen as one operation applied at three distinct points along
the cascade tower.
\end{remark}

\begin{remark}[Which layers, structurally]
The set of layers at which the backward step is evaluated,
$\{3, 6, 7\}$ (with the $6$--$7$ pair appearing in squared form for
the bilinear transport), is determined structurally:
\begin{itemize}[nosep]
\item $d = 3$: the observer's spatial receiving dimension under
locked time (Section~3.2; equivalently, the sphere $\partial B^4$ at
the observer's own dimension $d = 4$).
\item $d = 6, 7$: the two slicing steps of the backward descent from
$d_0 = 7$ (cascade reference) to $d_V = 5$ (observer's host). Both
are interior layers of Part~0's Growth regime
(Part~0~Section~7.1), crossed during the observer's host-frame
rescaling.
\end{itemize}
The squaring in the host-frame factor is the bilinear passage of the
two threshold sphere areas $\Omega_{19}, \Omega_{217}$ through the
same host rescaling, as established in Part~0~Lemma~8.8.
\end{remark}

\begin{remark}[Numerical verification]
The unified descent decomposition is verified numerically in
\texttt{tools/cascade\_unified\_descent.py}: all backward steps agree
with their closed-form values to machine precision, and the
assembled product reproduces
$\rho_\Lambda / M^4_{\mathrm{Pl,red}} = 6.996\times 10^{-121}$
identically to Theorem~\ref{thm:cc}. The Gram-corrected value
$7.145\times 10^{-121}$ (Part~0 Supplement) remains at $-0.07\%$
from Planck 2018.
\end{remark}

\section{The Cosmological Constant}

\begin{theorem}\label{thm:cc}
\[
\frac{\rho_\Lambda}{M^4_{\mathrm{Pl,red}}}
\;=\; \underbrace{\frac{\Omega_2}{\Omega_3}}_{\substack{\text{3D projection}\\ \text{(locked time)}}}
\;\cdot\;\underbrace{\left(\frac{\Omega_5}{\Omega_7}\right)^{\!2}}_{\substack{\text{host frame}\\ d_V\to d_0}}
\;\cdot\;\Omega_{19}\;\Omega_{217}
\;=\; \frac{2}{\pi}\cdot\frac{9}{\pi^2}\cdot\Omega_{19}\Omega_{217}
\;=\; \frac{18\;\Omega_{19}\;\Omega_{217}}{\pi^3}
\;=\; 6.996\times 10^{-121}.
\]
\end{theorem}

\begin{proof}
The invariant $I_0 = \Omega_{19}\times\Omega_{217}$ is the product of
two threshold sphere areas (Part 0, Theorem~8.5). Each is referenced
to $d_0 = 7$. The observer at $d_V = 5$ re-references each through one
factor of $\Omega_5/\Omega_7 = 3/\pi$, contributing
$(3/\pi)^2 = 9/\pi^2$ (Correction~1, host frame). Cascade content from
the compactified higher-dimensional layers then projects onto the
observer's 3D spatial slice $S^3$, with time locked as the cascade's
own descent / the observer's asymptotic infall toward the 5D-BH
pseudo-horizon (Correction~2). The projection factor from the
per-$S^2$ content density at each latitude of the $S^3$ Morse
foliation to the integrated $S^3$ 3-volume is
$\Omega_2/\Omega_3 = 2/\pi$. Combining both corrections:
\[
\frac{\rho_\Lambda}{M^4_{\mathrm{Pl,red}}}
= \frac{\Omega_2}{\Omega_3}\cdot\left(\frac{\Omega_5}{\Omega_7}\right)^{\!2}\cdot\Omega_{19}\Omega_{217}
= \frac{2}{\pi}\cdot\frac{9}{\pi^2}\times 1.20516\times 10^{-120}
= \frac{18}{\pi^3}\times 1.20516\times 10^{-120}
= 6.996\times 10^{-121}.\qedhere
\]
\end{proof}

Observed: $(7.150\pm 0.13)\times 10^{-121}$ (Planck 2018 $H_0=67.36$,
$\Omega_\Lambda=0.6847$; uncertainty $\pm 1.9\%$).

Leading deviation: $-2.2\%$. This places $\Lambda$ in the
\emph{descent-dependent} population of observables (negative sign at
leading order), alongside $\alpha_s$, $m_\tau/m_\mu$, $v$, $\ell_A$,
and $\Omega_m^{\mathrm{Bott}}$, all of which the Part~0 Supplement
first-order Gram correction closes from $\sim\!2\%$ to sub-$0.5\%$.
Applying the same correction to $\Lambda$:
\[
\frac{\rho_\Lambda}{M^4_{\mathrm{Pl,red}}}\bigg|_{\mathrm{Gram}}
= \frac{18}{\pi^3}\,\Omega_{19}\,\Omega_{217}\cdot\exp\!\bigl(\textstyle\sum_{d=5}^{216}(1-C^2_{d,d+1})\bigr)
= 6.996\times 10^{-121}\cdot e^{0.02108}
= 7.145\times 10^{-121}.
\]
Residual: $-0.07\%$, well inside the $\pm 1.9\%$ Planck $1\sigma$
uncertainty. The cosmological constant is not a Tier~1 structural
exception; it is a Tier~2 descent-dependent observable that closes to
sub-experimental precision through the same first-order correction
that closes the other descent-dependent members of the series.

\noindent\textbf{Why the sum runs over $d=5..216$.} The range of
adjacent-pair indices is not imported by convention; it is forced by
the austerity principle of the Prelude
(Principle~2.2~\cite{paper0pre}). Austerity permits inclusion of only
those layers that are in the cascade's descent to the observer, and no
others:
\begin{itemize}[nosep]
\item The cascade descends from the terminus $d_2 = 217$ down to the
observer's host $d_V = 5$. Every layer in this range is a compactified
cascade layer whose sphere-area content has participated in the
descent; their adjacent-pair Gram deficits are physical.
\item Layers below the observer's host ($d = 1, 2, 3$) are not cascade
descent layers. The cascade does not pass through them --- they are
below the observer's position, not reached by the descent from
infinity. Including them would introduce unjustified inputs
(Principle~2.2 clause~(i)).
\item The observer's own layer $d=4$ is the \emph{endpoint} of the
descent, not a descent step. The pair $(4,5)$ crosses the
observer-boundary; it is not a step within the cascade.
\item Above the terminus ($d > 217$), sphere areas collapse below
$10^{-120}$ (Part~0's ``oblivion'' regime, Section on the regime
partition); their adjacent-pair deficits contribute negligibly.
\end{itemize}
The austerity-compatible range of adjacent-pair indices is therefore
exactly $\{(5,6), (6,7), \ldots, (216, 217)\}$, summed as
$d = 5, 6, \ldots, 216$ --- 212 pairs, the range used in the Gram
correction above. No other range is austerity-compatible: including
$d < 5$ violates clause~(i) (layers below the descent); excluding
$d \in [5, 216]$ drops descent layers without justification. The
$0.02108$ sum value and the $-0.07\%$ residual are therefore
consequences of a principled range, not artefacts of a tuned one.

Part 0 matches $\log_{10}|\Omega_{19}\Omega_{217}|$ to within $\sim 0.3$
without any frame correction. Both frame corrections ($9/\pi^2$ from
the host-frame rescaling and $2/\pi$ from the 3D projection under locked
time) are Gamma-function identities at specified cascade sphere
dimensions, not fits. The host-frame correction is internal to the
compactified cascade tower; the 3D projection correction arises where
the cascade meets the observer's frame and accounts for the fact that
only three of the observer's apparent four spacetime dimensions are
genuinely free to receive cascade content.

\section{The Formula Reads the Tower}

Part 0 establishes exactly four distinguished dimensions (Theorem~7.1).
The cosmological constant uses all four in the pure-cascade numerator
and adds one further sphere-area ratio --- the 3D projection factor
from the locked-time $S^3$ Morse foliation --- to convert from the
cascade's per-$S^2$ content density to the observer's integrated
3-sphere 3-volume:
\[
\frac{\rho_\Lambda}{M^4_{\mathrm{Pl,red}}}
\;=\;
\underbrace{\frac{\Omega_2}{\Omega_3}}_{\substack{\text{3D projection}\\\text{(locked time)}}}
\;\cdot\;
\underbrace{\frac{\Omega_{5}^{\,2}\;\Omega_{19}\;\Omega_{217}}{\Omega_7^{\,2}}}_{\text{Part 0 bilinear invariant}}
\;=\;
\frac{18\;\Omega_{19}\;\Omega_{217}}{\pi^3}.
\]

Every factor is a ratio or product of cascade-native sphere areas. No
cubes, no imported QFT measures, no external normalisation conventions.

\begin{center}
\begin{tabular}{clll}
\toprule
Sphere & $\dim$ & Cascade role & Formula role \\
\midrule
$\Omega_2$ & 2 & Observer equatorial 2-sphere & 3D projection numerator \\
$\Omega_3$ & 3 & Observer spatial slice $S^3$ & 3D projection denominator \\
$\Omega_5$ & 5 & Volume maximum region ($d_V=5$) & Host frame \\
$\Omega_7$ & 7 & Area maximum region ($d_0=7$) & Cascade reference \\
$\Omega_{19}$ & 19 & Phase boundary ($d_1=19$) & Near threshold \\
$\Omega_{217}$ & 217 & Terminus ($d_2=217$) & Far threshold \\
\bottomrule
\end{tabular}
\end{center}

Frame from the two cascade maxima. Scale from the two thresholds.
Projection from the observer's locked-time $S^3$ slice, expressed as
the Morse-foliation ratio $\Omega_2/\Omega_3$. Four distinguished
cascade dimensions plus two sphere-areas at the observer's spatial
receiving volume --- six cascade sphere areas in total, all
Gamma-function values.

\section{The Discrete Vacuum}

Between the two thresholds, the cascade has $217 - 19 = 198$ levels. Each
carries a specific sphere area, monotonically decreasing from
$\Omega_{18} = 0.59$ to $\Omega_{216}\approx 10^{-120}$. The vacuum is
198 discrete layers of geometry. The smallest fluctuation is
$\Omega_{217}\approx 10^{-120}$: the geometric floor, forced by the Gamma
function at $p(d) = \sqrt{\pi}$. The detailed contrast with the
continuous divergent-integral picture of quantum field theory is
developed in the next section.

\section{Comparison with Quantum Field Theory}
\label{sec:qft-comparison}

The cascade and quantum field theory are two different ways of counting
vacuum content. They give different answers by roughly $10^{120}$ --- the
``cosmological constant problem'' in its standard framing. This section
argues that the difference decomposes into three distinct structural
disagreements, each of which is resolved by a specific feature of the
cascade's architecture: (i) the number of free dimensions at the
observer's frame (4 for QFT, 3 for the cascade, because time is locked
as the observer's asymptotic infall), (ii) the spectrum of modes
(continuous divergent integral for QFT, finite discrete tower for the
cascade), and (iii) the UV regulator (hand-imposed Planck cutoff for
QFT, Gamma-function terminus at $d_2 = 217$ for the cascade).

\subsection{How many dimensions does the observer have?}

QFT's zero-point vacuum energy density sums the ground-state energies
of every mode of every free field on flat Minkowski space
$\mathbb{R}^{3,1}$:
\begin{equation}\label{eq:qft-vacuum}
\rho_\Lambda^{\mathrm{QFT}} \;=\; \sum_{\text{fields}}\int \frac{d^3k}{(2\pi)^3}\,\frac{\hbar\omega_k}{2}
\;=\; \frac{1}{4\pi^2}\int_0^{k_{\max}} k^2\sqrt{k^2+m^2}\,dk\,+\,\cdots
\end{equation}
The integration is over 3-momentum space at fixed time, with the
implicit assumption that time is a free fourth dimension: the
Hamiltonian $H$ generates unbounded unitary time evolution, and the
4D energy-per-volume $\rho_\Lambda$ is extracted by dividing out an
infinite time translation. QFT counts modes over \emph{four} free
spacetime dimensions.

The cascade's analogue is a finite sum:
\begin{equation}\label{eq:cascade-vacuum}
\rho_\Lambda^{\mathrm{cascade}} \;\propto\; \sum_{d=d_1}^{d_2}\Omega_{d-1}
\;=\; \sum_{d=19}^{217}\Omega_{d-1},
\end{equation}
where $\Omega_{d-1}$ is the unit $(d{-}1)$-sphere area at cascade layer
$d$ (Part~0, Corollary~3.2). And the receiving volume at the observer's
frame is \emph{not} 4-dimensional: time is locked as the cascade's
descent parameter, equivalently the observer's asymptotic infall
toward the host 5D-BH pseudo-horizon (Cover Sheet thought experiment,
Part~II~\cite{paper2} Section~7.1). Only the three angular dimensions
of the observer's spatial slice $S^3$ are free. Cascade content from
the higher-dimensional layers projects onto those three dimensions ---
not four.

The numerical cost of this dimensional-accounting difference is
exactly the $2/\pi$ factor of Theorem~3.1. In the $S^3$ Morse foliation
by 2-spheres at constant latitude, the elementary integral
$\int_0^\pi\sin^2\theta\,d\theta = \pi/2$ gives
\[
\Omega_3 \;=\; \frac{\pi}{2}\,\Omega_2 \qquad\Longleftrightarrow\qquad \frac{\Omega_2}{\Omega_3} \;=\; \frac{2}{\pi}.
\]
Per-$S^2$ content density (the natural cascade-frame object at each
latitude) and integrated-$S^3$ 3-volume (the observer's full receiving
volume) differ by exactly this $2/\pi$ Morse-integral factor. The
factor is cascade-native and sphere-valued throughout; no cube, no
imported QFT measure.

The observation is sharp: QFT's apparent 4-dimensional integration
region and the cascade's 3-dimensional $S^3$ receiving volume are not
two normalisation conventions. They reflect a genuine structural
disagreement about whether time is free or locked. If time is free
(QFT), vacuum energy accounts over 4D spacetime and no $\pi/2$
projection appears. If time is locked as the cascade's own descent
(cascade), vacuum content accounts only over the spatial $S^3$ and the
$\pi/2$ factor is forced by the Morse foliation.

\begin{remark}[Time as observer artefact]
The 4-dimensional spacetime picture is the observer's local
approximation: on any patch of $S^3$ small compared to the cosmological
curvature scale and any time interval short compared to the infall
timescale, the observer sees flat Minkowski space with a free time
direction. QFT is built on this local approximation and is internally
consistent within it. The cascade argues that the local approximation
is exactly that --- local. At the cosmological scales where
$\rho_\Lambda$ lives (involving the full cascade tower
$\Omega_{19}\Omega_{217}$), the locked-time character of the cascade's
global descent becomes visible, and the naive 4D accounting overcounts
by a factor that resolves exactly to $\pi/2$.
\end{remark}

\subsection{Two regulators for the same divergence}

The integral \eqref{eq:qft-vacuum} diverges at $k\to\infty$ as
$k_{\max}^4$. Standard QFT imposes a hand-chosen cutoff, typically
$k_{\max}\sim M_{\mathrm{Pl}}$, giving
$\rho_\Lambda^{\mathrm{QFT}}\sim M_{\mathrm{Pl}}^4$ and the standard
$120$-orders-of-magnitude fine-tuning problem. The cutoff is external
to the formalism: nothing inside QFT selects $k_{\max}=M_{\mathrm{Pl}}$
as opposed to any other scale; the choice is justified by dimensional
analysis and the assumption that quantum gravity will regulate the UV.

The cascade has no divergence to regulate. The sum
\eqref{eq:cascade-vacuum} contains exactly $198$ terms, because Part~0
proves (Theorem~6.5, orbit termination; Section~5, the four regimes of
$p(d)$) that the cascade has exactly two thresholds and terminates
structurally at $d_2 = 217$. Beyond $d_2$ the decay rate $p(d)$
satisfies $p(d)>\sqrt{\pi}$ and sphere areas collapse below
$10^{-120}$: the ``oblivion regime'' where the cascade has no
geometric content to contribute. The terminus is a theorem about the
Gamma function, not an imposed UV cutoff.

Concretely: QFT's $k_{\max}$ is a free parameter disguised as a
physical input; the cascade's $d_2 = 217$ is the unique integer at
which the recurrence's second canonical constant $c_2 = \sqrt{\pi}$ is
reached by the decay rate (Part~0, Theorem~6.7). The ``$10^{120}$
fine-tuning'' that arises from QFT's divergent integral does not arise
in the cascade because the cascade's sum is convergent by construction.
What appears to be a miraculous near-cancellation in QFT is, in the
cascade, the natural value of a finite sum dominated by its smallest
term ($\Omega_{217}\approx 10^{-120}$).

\subsection{Architectural comparison}

\begin{center}
\renewcommand{\arraystretch}{1.35}
\begin{tabular}{lll}
\toprule
 & Quantum Field Theory & Cascade \\
\midrule
Spacetime topology at the observer & Flat $\mathbb{R}^{3,1}$ (locally assumed)
 & Compact $S^3\times$ infall (globally) \\
Free dimensions at the observer & 4 (time free, unbounded $H$)
 & 3 (time locked as infall) \\
Mode counting & Continuous $\displaystyle\int\!\frac{d^3k}{(2\pi)^3}$
 & Discrete $\displaystyle\sum_{d=19}^{217}\!\Omega_{d-1}$ \\
Receiving volume & 4D spacetime, $d^4x$
 & 3D spatial $S^3$, area $\Omega_3=2\pi^2$ \\
Projection factor at the observer & none (time is free)
 & $\displaystyle\frac{\Omega_2}{\Omega_3}=\frac{2}{\pi}$ (Morse foliation of $S^3$) \\
UV regulator & Imposed cutoff $k_{\max}\sim M_{\mathrm{Pl}}$
 & Structural terminus $d_2 = 217$ (Gamma theorem) \\
Number of modes summed & Continuous (divergent) & 198 discrete layers \\
Leading $\rho_\Lambda$ estimate & $\sim M_{\mathrm{Pl}}^4$
 & $\sim \Omega_{217}\,M_{\mathrm{Pl,red}}^4 \sim 10^{-120}\,M_{\mathrm{Pl,red}}^4$ \\
Fine-tuning required & $\sim 10^{120}$ & None \\
\bottomrule
\end{tabular}
\end{center}

\subsection{Reframing the cosmological constant problem}

The standard statement of the cosmological constant problem assumes QFT
is giving a genuine estimate of $\rho_\Lambda$ that happens to be
$10^{120}$ too large, so that some unknown physics must cancel it down
to the observed value. The cascade reframes this as three separable
structural disagreements, none of which requires a cancellation:

\begin{enumerate}
\item \emph{The number of free dimensions is different.} QFT counts
modes over 4 free dimensions (flat $\mathbb{R}^{3,1}$, with time
unbounded). The cascade counts modes over 3 free dimensions (the
angular $S^3$), because time is locked as the observer's asymptotic
infall. The Morse foliation of $S^3$ by 2-spheres gives the
projection factor $\Omega_2/\Omega_3 = 2/\pi$, i.e., $O(1)$, not
$O(10^{120})$. This $\pi/2$ is a geometric cost of the 3D projection;
it has nothing to do with fine-tuning.
\item \emph{The spectrum of modes is different.} QFT integrates a
continuous mode density $d^3k/(2\pi)^3$ and diverges as
$k_{\max}^4$. The cascade sums a finite discrete tower of 198 layers
and converges by construction. This is the $\sim 10^{120}$ factor.
It is not a miracle cancellation in the cascade; it is the structural
fact that the cascade's tower is finite, while QFT's mode integral is
infinite before the cutoff is imposed.
\item \emph{The UV regulator is different.} QFT imposes $k_{\max}
\sim M_{\mathrm{Pl}}$ by hand. The cascade terminates at $d_2 = 217$
by theorem --- specifically, by Part~0 Theorem~6.7, which identifies
$d_2$ as the unique integer at which the recurrence's second
canonical constant $c_2 = \sqrt{\pi}$ is reached by the decay rate.
The terminus has no parameter-counting cost. QFT's cutoff does.
\end{enumerate}

This does not ``solve'' the cosmological constant problem in the sense
of showing how QFT's naive answer becomes the cascade's natural answer
through some field-theoretic mechanism. The cascade is not a refinement
of QFT's calculation. It is an alternative architecture for counting
vacuum content, grounded in three architectural commitments that QFT
does not make: time is locked as the observer's infall; the spectrum
of cascade layers is discrete and finite; the terminus is forced by
the Gamma function rather than imposed by hand. The $\pi/2$ projection
factor of Theorem~3.1 is the consequence of the first commitment
(locked time) and is the only such factor of order unity that the
cascade needs at the observer's frame; the remaining $\sim 10^{120}$
difference between QFT's naive estimate and the observed value is
accounted for by the second and third commitments together.

\begin{remark}[The $\pi/2$ is Part~0's own lowest slicing integral]\label{rem:open-pi-over-two}
The Morse-foliation identification $\Omega_3 = (\pi/2)\Omega_2$ is not
a plausible conjecture awaiting independent verification. It is the
same integral as Part~0's slicing recurrence at its lowest nontrivial
step. Part~0's recurrence is
$V_{d+1}/V_d = \int_{-1}^{1}(1-x^2)^{d/2}\,dx$, and at $d = 1$ this
gives
\[
\frac{V_2}{V_1} \;=\; \int_{-1}^{1}(1-x^2)^{1/2}\,dx \;=\; \frac{\pi}{2}.
\]
Under the substitution $x = \cos\theta$, the integrand
$(1-x^2)^{1/2}$ becomes $\sin\theta$ and the measure $dx$ becomes
$\sin\theta\,d\theta$, so the integral is
$\int_0^\pi \sin^2\theta\,d\theta = \pi/2$. This is the Morse
foliation of $S^3$ by 2-spheres: at each latitude $\theta$ on $S^3$,
the equatorial 2-sphere has area $\sin^2\theta\cdot\Omega_2$, and
integrating latitude recovers $\Omega_3 = (\pi/2)\Omega_2$. The two
identities --- the Part~0 slicing step and the $S^3$ Morse foliation
--- are literally the same integral in different coordinates. The
$\pi/2$ at the observer's frame is Part~0's own lowest recurrence
step, applied inversely at the observer's spatial $S^3$ slice.

\medskip\noindent\textbf{The cascade does not do Kaluza--Klein
reduction.} Part~II=III makes the structural commitment explicit:
gravity is not quantised via semiclassical integration over a
background because the metric is never an operator on a fixed
background --- it is a state property on the cascade geometry. The
same commitment governs vacuum energy density. There is no
integration of continuous zero-point modes over compactified
dimensions, regulated by a hand-imposed cutoff, to be performed and
then matched to the cascade's prediction. The cascade's slicing
descent \emph{is} the replacement for such semiclassical procedures.
The $\pi/2$ projection factor is therefore not the output of a
Kaluza--Klein calculation awaiting completion; it is a direct sphere
identity in Part~0's recurrence, applied at the observer's locked-time
$S^3$ slice under the identification of time with the cascade's own
descent direction (Part~II~\cite{paper2}, Section~7.1).

The three corrections of Theorem~\ref{thm:cc} (host frame $9/\pi^2$,
3D projection $2/\pi$, Part~0 bilinear invariant
$\Omega_{19}\Omega_{217}$) are unified in Section~\ref{subsec:unified-descent}
(Theorem~\ref{thm:unified}): each is an instance of Part~0's slicing
recurrence evaluated at a specific cascade layer, and the host-frame
and projection factors telescope to simple ratios of the observer's
cascade indices (Lemma~\ref{lem:telescoping}, Remark~\ref{rem:host-index}).
No second operation enters; the derivation stays within Part~0's
architecture. It is \emph{not} a Kaluza--Klein problem and \emph{not} a
semiclassical reduction --- the cascade replaces both.
\end{remark}

\section{What This Paper Does and Does Not Do}

\textbf{Does:}
\begin{itemize}
\item Introduces the hypothesis (the single physical input of the series).
\item Derives
$\rho_\Lambda/M^4_{\mathrm{Pl,red}} = 18\Omega_{19}\Omega_{217}/\pi^3$
from Part~0's bilinear invariant together with two cascade-native
sphere-area corrections: the host frame $(9/\pi^2)$ from $d_0=7$ to
$d_V=5$, and the 3D projection factor $\Omega_2/\Omega_3 = 2/\pi$ at
the observer's $S^3$ slice under locked time. Leading deviation from
Planck~2018: $-2.2\%$; after the Part~0 Supplement first-order Gram
correction, residual $-0.07\%$, inside the $\pm 1.9\%$ Planck
$1\sigma$ uncertainty.
\item Places $\Lambda$ in the cascade's descent-dependent observable
population, closed by the same structural correction as $\alpha_s$,
$m_\tau/m_\mu$, $v$, $\ell_A$, and $\Omega_m^{\mathrm{Bott}}$.
\item Identifies the vacuum as 198 discrete layers with a natural terminus.
\item Identifies time as the cascade's locked descent parameter (the
observer's asymptotic infall toward the 5D-BH pseudo-horizon), with
the consequence that cascade content projects over only 3 free
dimensions at the observer's frame.
\item Identifies the $\pi/2$ projection factor as the cascade-native
sphere-area ratio $\Omega_2/\Omega_3$, which coincides with the
latitudinal integral $\int_0^\pi\sin^2\theta\,d\theta = \pi/2$ of the
Morse foliation of $S^3$ by 2-spheres.
\item Unifies Theorem~\ref{thm:cc}'s three corrections (host frame,
3D projection, Part~0 bilinear invariant) as a single cascade-internal
operation: Part~0's slicing recurrence evaluated at cascade-distinguished
layers (Section~\ref{subsec:unified-descent}, Theorem~\ref{thm:unified}).
A Gamma-function telescoping identity
(Lemma~\ref{lem:telescoping}) reveals that the ``3'' in $3/\pi$ is
$(d_V + 1)/2$ --- the observer-host index, not an opaque constant.
\end{itemize}

\textbf{Does not:}
\begin{itemize}
\item Derive $d = 4$ from the cascade (used here as empirical fact).
\item Derive quantum mechanics, general relativity, the Standard Model, or
cosmological parameters beyond $\Lambda$.
\item Derive the Gram first-order correction from first principles; it
is imported from the Part~0 Supplement.
\end{itemize}

\textbf{Inputs:} One axiom (orthogonality, via Part 0). One hypothesis.
One empirical fact ($d = 4$). Zero free parameters.

\begin{thebibliography}{3}
\bibitem{paper0pre} RTAC, \textit{The Cascade Series --- Prelude:
Why Nothing Has Structure}, 2026.
\bibitem{part0} RTAC, \textit{The Cascade Series --- Part 0: Scale
Variance from Orthogonality}, 2026.
\bibitem{paper2} RTAC, \textit{The Cascade Series --- Part II:
Quantum Mechanics from the Cascade}, 2026.
\end{thebibliography}

\end{document}