\documentclass[12pt,a4paper]{article}
\usepackage[margin=1in]{geometry}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{enumitem}
\usepackage{booktabs}
\usepackage{array}

\newtheorem{theorem}{Theorem}[section]
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{definition}[theorem]{Definition}
\theoremstyle{remark}
\newtheorem*{remark}{Remark}

\title{Part 0 --- Supplement\\[6pt]
\large Inter-Layer Coupling and the Independent-Step Correction}
\author{RTAC}
\date{March 2026}

\begin{document}
\setlength{\emergencystretch}{3em}
\maketitle

\noindent This supplement extends Paper 0 by computing the inter-layer correction to the cascade's independent-step approximation. The slicing recurrence treats each compactification step as independent: the propagator through $n$ steps is the product of $n$ individual lapses. This is exact for volumes (the telescoping identity $V_D/V_4 = \prod N(d)$ holds by the Gamma function). It is not exact for quantities that depend on the overlap structure of consecutive layer integrands. The correction is determined by the eigenvalue deficit of the Gram correlation matrix of the integrands $f_d(x) = (1 - x^2)^{d/2}$---a Beta function identity with no free parameters.

The Gram correlation has an exact closed form in cascade-native primitives (Theorem~\ref{thm:gram-R}): $C^2_{d,d+1} = R(2d+2)^2 / [R(2d+1)\,R(2d+3)]$, equivalently $\log C^2_{d,d+1} = -\tfrac{1}{2}\Delta^2 \log\alpha\big|_{2d+2}$ (Corollary~\ref{cor:gram-laplacian}). The Gram correction is therefore the discrete log-Laplacian of the cascade scalar action's compliance function $\alpha(d) = R(d)^2/4$, evaluated at the doubled cascade argument. This connects the supplement directly to the action principle of Paper~IVb Remark~4.6: the supplement's path-distributed correction and Paper~IVb's single-source $\alpha(d^*)/\chi^k$ family are complementary derivatives of the same compliance function---linear source response (Paper~IVb) and discrete log-Laplacian (this supplement).

Throughout: ``second order'' refers exclusively to the perturbation-theory order in the eigenvalue expansion (Theorem~\ref{thm:pert}); the observable-dependent correction beyond the first-order Gram identity is named the \emph{observable response factor}.

\setcounter{section}{14}
\section{Inter-Layer Coupling}

\subsection{The Gram matrix of cascade layer integrands}

\begin{definition}[Gram matrix]
For a set of cascade layers $\{d_1, d_2, \ldots, d_n\}$, the Gram matrix is
\[
G_{ij} = \langle f_{d_i}, f_{d_j} \rangle = \int_{-1}^{1} (1 - x^2)^{(d_i + d_j)/2}\, dx.
\]
\end{definition}

\begin{theorem}[Gram matrix as Beta function]\label{thm:gram}
\[
G_{ij} = B\!\left(\tfrac{1}{2},\, \tfrac{d_i + d_j}{2} + 1\right) = \sqrt{\pi}\, \frac{\Gamma\!\left(\frac{d_i + d_j}{2} + 1\right)}{\Gamma\!\left(\frac{d_i + d_j}{2} + \frac{3}{2}\right)}.
\]
\end{theorem}

\begin{proof}
Substituting $t = x^2$ in the integral: $\int_{-1}^{1}(1 - x^2)^\alpha\, dx = B(1/2, \alpha + 1)$ with $\alpha = (d_i + d_j)/2$. The Beta function identity $B(a,b) = \Gamma(a)\Gamma(b)/\Gamma(a+b)$ with $a = 1/2$ gives the stated form.
\end{proof}

\begin{definition}[Correlation matrix]
The normalised Gram matrix (correlation matrix) is
\[
C_{ij} = \frac{G_{ij}}{\sqrt{G_{ii}\, G_{jj}}} = \frac{B\!\left(\frac{1}{2},\, \frac{d_i + d_j}{2} + 1\right)}{\sqrt{B\!\left(\frac{1}{2},\, d_i + 1\right)\, B\!\left(\frac{1}{2},\, d_j + 1\right)}}.
\]
$C_{ii} = 1$ for all $i$, $\mathrm{tr}\, C = n$, and $0 < C_{ij} < 1$ for $i \neq j$.
\end{definition}

\subsection{The adjacent-layer overlap deficit}

\begin{theorem}[Overlap deficit]\label{thm:overlap}
For adjacent layers $d$ and $d+1$:
\[
1 - C_{d,d+1}^2 = 1 - \frac{B\!\left(\frac{1}{2},\, d + \frac{3}{2}\right)^2}{B\!\left(\frac{1}{2},\, d+1\right)\, B\!\left(\frac{1}{2},\, d+2\right)}.
\]
This quantity is strictly positive for all $d \geq 1$ and satisfies:
\[
1 - C_{d,d+1}^2 \sim \frac{1}{8\, d^2} \quad (d \to \infty).
\]
Numerically: $1 - C_{5,6}^2 = 0.00319$, $\; 1 - C_{11,12}^2 = 0.00083$, $\; 1 - C_{19,20}^2 = 0.00030$.
\end{theorem}

\begin{proof}
\emph{Positivity.} $C_{d,d+1} < 1$ by Cauchy--Schwarz: the integrands $f_d$ and $f_{d+1}$ are not proportional in $L^2[-1,1]$ for $d \neq d+1$, so the normalised inner product is strictly less than~$1$.

\emph{Asymptotic $1 - C_{d,d+1}^2 \sim 1/(8d^2)$.} Set $h(\alpha) = B(1/2,\alpha+1) = \sqrt{\pi}\,\Gamma(\alpha+1)/\Gamma(\alpha+3/2)$ and $L(\alpha) = \log h(\alpha)$. Then
\[
\log C_{d,d+1}^2 \;=\; 2L(d+\tfrac{1}{2}) - L(d) - L(d+1),
\]
the centered second difference of $L$ at $d+1/2$ with step $1/2$. Taylor expanding,
\[
L(d) + L(d+1) \;=\; 2L(d+\tfrac{1}{2}) + \tfrac{1}{4}L''(d+\tfrac{1}{2}) + O(L^{(4)}/192),
\]
so $\log C_{d,d+1}^2 = -\tfrac{1}{4}L''(d+\tfrac{1}{2}) + O(L^{(4)})$.

Now $L'(\alpha) = \psi(\alpha+1) - \psi(\alpha+3/2)$, where $\psi$ is the digamma function. The asymptotic expansion $\psi(x) = \log x - 1/(2x) - 1/(12x^2) + O(1/x^4)$ gives, after expanding $\psi(\alpha+1)$ and $\psi(\alpha+3/2)$ around $\alpha$ and combining,
\[
L'(\alpha) \;=\; -\frac{1}{2\alpha} + \frac{3}{8\alpha^2} + O(1/\alpha^3),
\qquad
L''(\alpha) \;=\; \frac{1}{2\alpha^2} + O(1/\alpha^3).
\]
Therefore
\[
\log C_{d,d+1}^2 \;=\; -\frac{1}{4}\cdot\frac{1}{2d^2} + O(1/d^3) \;=\; -\frac{1}{8d^2} + O(1/d^3),
\]
and $1 - C_{d,d+1}^2 = -\log C_{d,d+1}^2 + O\bigl((\log C^2)^2\bigr) = 1/(8d^2) + O(1/d^3)$.

The numerical values quoted in the theorem statement ($1 - C_{5,6}^2 = 0.00319$, $1 - C_{11,12}^2 = 0.00083$, $1 - C_{19,20}^2 = 0.00030$) are evaluated directly from the Beta function ratio; the asymptotic $1/(8d^2)$ supplies the leading $1/d^2$ scaling and the sign, but subleading $1/d^3$ terms matter at small $d$ (the asymptotic over-estimates by $\sim 50\%$ at $d=5$, by $\sim 25\%$ at $d=11$, and by $\sim 15\%$ at $d=19$, converging to the exact value as $d\to\infty$).
\end{proof}

\begin{remark}
The overlap deficit $1 - C_{d,d+1}^2$ is the geometric content at step $d$ that is \emph{not} shared with step $d+1$. It measures the imperfect folding: the fraction by which consecutive layer integrands fail to be collinear in $L^2$. Its positivity is a consequence of $R_{\mathrm{eff}}(d) = 1/\sqrt{d+3} > 0$ (Theorem~4.2): the compactification at each step is asymptotic, never complete.
\end{remark}

\subsection{Closed form via cascade slicing ratios}\label{sec:gram-closed-form}

The asymptotic of Theorem~\ref{thm:overlap} extends to an \emph{exact} closed form expressing $C^2_{d,d+1}$ in cascade-native primitives. The exact form reveals that the Gram correlation is the discrete log-Laplacian of the cascade scalar action's compliance function $\alpha(d) = R(d)^2/4$, evaluated at the doubled cascade argument $2d+2$.

\begin{theorem}[Gram correlation in cascade slicing ratios]\label{thm:gram-R}
For all $d \geq 1$, the Gram correlation between adjacent cascade layers is
\begin{equation}\label{eq:gram-R-form}
C^2_{d,d+1} \;=\; \frac{R(2d+2)^2}{R(2d+1)\,R(2d+3)},
\end{equation}
where $R(d) = \Gamma((d+1)/2)/\Gamma((d+2)/2)$ is the cascade slicing ratio of Paper~0 \S2.
\end{theorem}

\begin{proof}
The Gram entries (Theorem~\ref{thm:gram}) are Beta functions $G_{ij} = B(1/2, (d_i+d_j)/2 + 1)$. Using the Beta--Gamma reduction $B(1/2, n+1) = \sqrt{\pi}\,\Gamma(n+1)/\Gamma(n+3/2)$ together with the cascade slicing ratio's defining identity $R(d) = \Gamma((d+1)/2)/\Gamma((d+2)/2)$, one obtains:
\[
G_{d,d}     = B(\tfrac{1}{2}, d+1)   = \sqrt{\pi}\,R(2d+1), \qquad
G_{d,d+1}   = B(\tfrac{1}{2}, d+\tfrac{3}{2}) = \sqrt{\pi}\,R(2d+2),
\]
\[
G_{d+1,d+1} = B(\tfrac{1}{2}, d+2)   = \sqrt{\pi}\,R(2d+3).
\]
Substituting into the definition $C^2_{d,d+1} = G_{d,d+1}^2 / (G_{d,d}\,G_{d+1,d+1})$ gives
$C^2_{d,d+1} = R(2d+2)^2 / [R(2d+1)\,R(2d+3)]$.
\end{proof}

\begin{corollary}[Gram-Laplacian identity]\label{cor:gram-laplacian}
The log-correlation between adjacent cascade layers is the discrete log-Laplacian of the cascade scalar action's compliance function $\alpha(d) = R(d)^2/4$, evaluated at the doubled argument $2d+2$ with unit step:
\begin{equation}\label{eq:gram-laplacian}
\log C^2_{d,d+1} \;=\; -\frac{1}{2}\,\Delta^2 \log\alpha\,\Big|_{2d+2},
\end{equation}
where $\Delta^2 f|_n := f(n-1) + f(n+1) - 2f(n)$ is the centered discrete Laplacian.
\end{corollary}

\begin{proof}
By Theorem~\ref{thm:gram-R} and properties of the logarithm,
\[
\log C^2_{d,d+1} = 2\log R(2d+2) - \log R(2d+1) - \log R(2d+3) = -\Delta^2 \log R\,\big|_{2d+2}.
\]
Since $\alpha(d) = R(d)^2/4$, we have $\log\alpha = 2\log R - 2\log 2$, so $\Delta^2 \log\alpha = 2\,\Delta^2 \log R$ (constants drop in second differences). Substituting yields equation~\eqref{eq:gram-laplacian}.
\end{proof}

\begin{theorem}[Full correlation matrix in cascade slicing ratios]\label{thm:gram-matrix}
The Gram correlation between any two cascade layers $d_i$ and $d_j$ is
\begin{equation}\label{eq:gram-matrix-form}
C_{ij} \;=\; \frac{R(d_i + d_j + 1)}{\sqrt{R(2d_i+1)\,R(2d_j+1)}}.
\end{equation}
Equivalently, with $k := |d_i - d_j|$ and centered argument $m := d_i + d_j + 1 = 2\min(d_i, d_j) + k + 1$,
\begin{equation}\label{eq:gram-matrix-laplacian}
\log C^2_{ij} \;=\; -\Delta^2_{(k)}\log R\,\big|_{m},
\end{equation}
where $\Delta^2_{(k)} f|_n := f(n-k) + f(n+k) - 2f(n)$ is the centered second difference with step~$k$.
\end{theorem}

\begin{proof}
The Gram entries are $G_{ij} = B(1/2, (d_i+d_j)/2 + 1)$ (Theorem~\ref{thm:gram}). The Beta--Gamma reduction $B(1/2, n+1) = \sqrt{\pi}\,\Gamma(n+1)/\Gamma(n+3/2)$ together with $R(d) = \Gamma((d+1)/2)/\Gamma((d+2)/2)$ gives, for general $i, j$:
\[
G_{ij} = \sqrt{\pi}\,R(d_i + d_j + 1), \qquad G_{ii} = \sqrt{\pi}\,R(2d_i + 1).
\]
Substituting into $C_{ij} = G_{ij} / \sqrt{G_{ii}\,G_{jj}}$ yields equation~\eqref{eq:gram-matrix-form}. Taking the squared logarithm:
$\log C^2_{ij} = 2\log R(d_i+d_j+1) - \log R(2d_i+1) - \log R(2d_j+1)$.
Setting $k = |d_i - d_j|$ and $m = d_i + d_j + 1$ gives $2d_i + 1 = m - k$ and $2d_j + 1 = m + k$, so
$\log C^2_{ij} = 2\log R(m) - \log R(m-k) - \log R(m+k) = -\Delta^2_{(k)} \log R\big|_m$.
Theorem~\ref{thm:gram-R} is the special case $k = 1$ (adjacent layers).
\end{proof}

\begin{corollary}[Eigenvalue deficit in cascade primitives]\label{cor:eigenvalue-cascade}
For any cascade path $\{d_0, d_0+1, \ldots, d_0+n-1\}$, the eigenvalue deficit $\epsilon(n, d_0) = 1 - \lambda_1(C)/n$ of Theorem~\ref{thm:eigenvalue} is the dominant-eigenvalue deficit of the explicit $n\times n$ matrix
\[
C_{ij} \;=\; \frac{R(d_i+d_j+1)}{\sqrt{R(2d_i+1)\,R(2d_j+1)}},
\]
where $d_i = d_0 + i$ for $i = 0, 1, \ldots, n-1$. The first-order perturbation expression of Theorem~\ref{thm:pert} is correspondingly
\begin{equation}\label{eq:pert-cascade}
\epsilon \;=\; \frac{2}{n^2}\sum_{i<j}\Bigl(1 - \frac{R(d_i+d_j+1)}{\sqrt{R(2d_i+1)\,R(2d_j+1)}}\Bigr) \;+\; O(\|D\|^2),
\end{equation}
a sum of cascade slicing ratios with no Beta function evaluations beyond what cascade primitives already supply.
\end{corollary}

\begin{proof}
Direct substitution of equation~\eqref{eq:gram-matrix-form} into the matrix entries of $C$, then into Theorem~\ref{thm:pert}'s first-order formula. The dominant eigenvalue $\lambda_1$ and the deficit $\epsilon$ are functions of $C$'s entries, hence of the cascade slicing ratios at the doubled arguments $\{2d_i+1\}$ and the off-doubled arguments $\{d_i+d_j+1\}$.
\end{proof}

\begin{remark}[Structural sign of the Gram correction]\label{rem:gram-sign-from-convexity}
The positivity of the per-step Gram deficit $1 - C^2_{d,d+1}$ (cited as ``Cauchy--Schwarz on integrands'' in Paper~IVb's $\alpha(d^*)/\chi^k$ sign rule) is now equivalent to the convexity of $\log\alpha$ on the cascade tower at the doubled argument. By Corollary~\ref{cor:gram-laplacian}, $1 - C^2_{d,d+1} \sim -\log C^2_{d,d+1} = \tfrac{1}{2}\Delta^2 \log\alpha\big|_{2d+2}$, which is positive iff $\log\alpha$ is locally convex at $2d+2$. Since $\alpha(d) \sim 1/(2d)$ asymptotically (Paper~0~\S2 plus the Beta--Gamma reduction), $\log\alpha \sim -\log(2d)$ is convex, so $\Delta^2\log\alpha > 0$ throughout the cascade tower. The sign-rule positivity therefore reduces to the convexity of $\log\alpha$, a structural property of the cascade compliance function rather than a separate $L^2$ inner-product fact.
\end{remark}

\begin{remark}[Numerical verification]\label{rem:gram-verifier}
Theorem~\ref{thm:gram-R} (adjacent closed form), Corollary~\ref{cor:gram-laplacian} (the Laplacian identity), and Theorem~\ref{thm:gram-matrix} (the full-matrix closed form) are verified at all tested cascade layer pairs in
\begin{center}
\texttt{tools/verifiers/gram\_compliance\_laplacian.py}
\end{center}
to machine precision on the closed form and to $\sim 10^{-7}$ relative on the Laplacian identity (the floating-point cancellation noise floor at large $d$, where $\log C^2_{d,d+1} \sim 1/(8d^2)$ is computed as a difference of $O(1)$ log-Gamma terms).
\end{remark}

\begin{remark}[Structural unification with the cascade scalar action]\label{rem:gram-action-unification}
Corollary~\ref{cor:gram-laplacian} establishes a direct structural connection between the supplement's Gram framework and the cascade scalar action $S[\varphi] = \sum_d (2\alpha(d))^{-1}(\Delta\varphi)^2$ of Paper~IVb Remark~4.6. Both correction families derive from the same compliance function $\alpha(d) = R(d)^2/4$:
\begin{itemize}
\item Paper~IVb's $\alpha(d^*)/\chi^k$ family: \emph{linear} source response. The cascade action's marginal Green's function at distinguished source layer $d^*$ equals $\alpha(d^*)$ exactly (Paper~IVb's marginal Green's function identity); chirality factorisation through $k$ independent channels gives the $1/\chi^k$ attenuation.
\item This supplement's adjacent overlap deficit: \emph{discrete log-Laplacian} of the same compliance function, at the doubled cascade argument. The deficit per step is $\frac{1}{2}\Delta^2(-\log\alpha)|_{2d+2}$ (positive because $-\log\alpha$ is convex on the cascade tower, i.e.\ $\alpha$ is log-concave).
\end{itemize}
The two correction families are complementary in the action's dual variables: source response (zeroth-order, single-layer) versus log-compliance curvature (second-order, distributed across the path). The doubling $d\mapsto 2d+2$ is a Beta--Gamma reduction artefact: cascade integrand inner products $\langle f_d, f_{d'}\rangle$ produce $R$-values at doubled arguments via $B(1/2, n+1) = \sqrt{\pi}\,R(2n+1)$ and $B(1/2, n+3/2) = \sqrt{\pi}\,R(2n+2)$, so the Gram correlation between layers $d$ and $d+1$ depends on cascade $R$-values at doubled indices $\{2d+1, 2d+2, 2d+3\}$.

This corollary closes Paper~I's open question on first-principles derivation of the Gram correction (Paper~I, ``Does not'' list, item~3): the derivation is the cascade compliance Laplacian.
\end{remark}

\subsection{The eigenvalue structure of the correlation matrix}

\begin{theorem}[Near-collinearity]\label{thm:eigenvalue}
For a path of $n$ consecutive layers $\{d_0, d_0+1, \ldots, d_0+n-1\}$, the correlation matrix $C$ has one dominant eigenvalue $\lambda_1$ satisfying
\[
\frac{\lambda_1}{n} = 1 - \epsilon(n, d_0), \qquad \epsilon > 0,
\]
and $(n - 1)$ subdominant eigenvalues summing to $n\,\epsilon$. The deficit $\epsilon$ depends on both $n$ (the number of layers) and $d_0$ (where in the cascade the path begins).
\end{theorem}

\begin{proof}
The entries $C_{ij}$ are smooth functions of $|d_i - d_j|$ that decrease from $C_{ii} = 1$ towards a minimum at $|d_i - d_j| = n - 1$. For the path $d = 5, \ldots, 12$ ($n = 8$): the minimum entry is $C_{5,12} = 0.9622$, far from zero. The matrix is therefore close to the rank-1 matrix $\mathbf{1}\mathbf{1}^T/n$, whose sole nonzero eigenvalue is $n$. The deficit $\epsilon$ is determined by the off-diagonal decay rate of $C_{ij}$, which is controlled by the Beta function asymptotics of Theorem~\ref{thm:overlap}.
\end{proof}

\noindent Numerical eigenvalue structure for $d = 5, \ldots, 12$ ($n = 8$):

\begin{center}
\renewcommand{\arraystretch}{1.2}
\begin{tabular}{ccc}
\toprule
Eigenvalue & Value & Fraction of trace \\
\midrule
$\lambda_1$ & 7.9347263 & 99.184\% \\
$\lambda_2$ & 0.0646528 & 0.808\% \\
$\lambda_3$ & 0.0006152 & 0.008\% \\
$\lambda_4$--$\lambda_8$ & $< 10^{-4}$ & $< 0.001\%$ \\
\midrule
$\epsilon = 1 - \lambda_1/n$ & \multicolumn{2}{c}{$0.008159 = 0.816\%$} \\
\bottomrule
\end{tabular}
\end{center}

\noindent The cascade layers are 99.2\% collinear. The remaining 0.8\% is the inter-layer coupling.

\subsection{Path dependence of the eigenvalue deficit}

The deficit $\epsilon(n, d_0)$ depends on where in the cascade the path lies. Lower $d$ (closer to the observer, larger $R_{\mathrm{eff}}$, less complete folding) gives larger deficits.

\begin{center}
\renewcommand{\arraystretch}{1.2}
\begin{tabular}{llcc}
\toprule
Path & $n$ & $\epsilon = 1 - \lambda_1/n$ & Needed correction \\
\midrule
$d = 5, \ldots, 12$ & 8 & 0.00816 & $1.73\%$ ($\alpha_s$) \\
$d = 6, \ldots, 13$ & 8 & 0.00657 & $1.75\%$ ($m_\tau/m_\mu$) \\
$d = 10, \ldots, 17$ & 8 & 0.00332 & --- \\
$d = 14, \ldots, 21$ & 8 & 0.00200 & $0.13\%$ ($m_\mu/m_e$) \\
\bottomrule
\end{tabular}
\end{center}

\noindent At fixed $n$, $\epsilon$ decreases with $d_0$ because the overlap deficit per step (Theorem~\ref{thm:overlap}) decreases as $1/d^2$.

\subsection{The independent-step correction}

The cascade's leading-order predictions treat each compactification step as independent. A quantity that depends on the cascade's geometry over a path of $n$ layers receives a correction from the inter-layer coupling. The eigenvalue deficit $\epsilon$ identifies the source, sign, and magnitude of this correction. The exact correction factor is observable-dependent.

\begin{definition}[Independent-step correction]
For a cascade quantity $Q$ whose leading-order value $Q_0$ depends on the cascade potential over a path $\{d_0, \ldots, d_0 + n - 1\}$, the corrected value is
\begin{equation}\label{eq:correction}
Q = Q_0 \times \left(1 + k_Q\, \epsilon\right), \qquad \epsilon = 1 - \frac{\lambda_1(C)}{n},
\end{equation}
where $\lambda_1(C)$ is the largest eigenvalue of the $n \times n$ correlation matrix and $k_Q$ is a coupling coefficient that depends on how the observable projects onto the subdominant eigenstructure of~$C$.
\end{definition}

\begin{remark}[What is derived and what is not]
The eigenvalue deficit $\epsilon$ is derived: it follows from first-order perturbation theory applied to the deficit matrix $D = J - C$, where $J$ is the all-ones matrix (Section~\ref{sec:perturbation}). Its sign ($\epsilon > 0$), its path dependence (decreasing with $d_0$), and its scaling ($\epsilon \propto 1/d^2$ per step) are theorems about the Beta function. The coupling coefficient $k_Q$ is not derived. It depends on how the observable couples to the subdominant eigenvectors of $C$, which varies from observable to observable. Computing $k_Q$ from first principles is the most important open problem in this supplement.
\end{remark}

\begin{remark}[What the correction is not]
The correction is not to the Gamma function, which is exact. It is not to the slicing recurrence, which is exact (the telescoping identity $V_D/V_4 = \prod N(d)$ holds by $\Gamma(a+1) = a\Gamma(a)$). It is to the \emph{independent-step approximation}: the assumption that a quantity depending on $n$ cascade layers is the product of $n$ individually exact factors. The layers are individually exact but not independent---their integrands overlap in $L^2$, and the eigenvalue deficit $\epsilon$ quantifies the overlap.
\end{remark}

\subsection{The perturbation theory for $\epsilon$}\label{sec:perturbation}

\begin{theorem}[First-order eigenvalue deficit]\label{thm:pert}
To first order in the deficit matrix $D = J - C$,
\[
\epsilon \;=\; 1 - \frac{\lambda_1}{n} \;=\; \frac{2}{n^2} \sum_{i < j} \left(1 - C_{ij}\right) \;+\; O\!\bigl(\|D\|^2\bigr),
\]
where $C_{ij}$ are the entries of the correlation matrix. The second-order correction is \emph{positive} (a sum of squared matrix elements of $D$ in the zero-eigenspace of $J$; see proof), so the right-hand side is an \emph{upper bound} for the eigenvalue deficit $\epsilon$.
\end{theorem}

\begin{proof}
Write $C = J - D$ where $J_{ij} = 1$ for all $i,j$ and $D_{ij} = 1 - C_{ij} \geq 0$, with $D_{ii} = 0$. The matrix $J$ has the non-degenerate eigenvalue $n$ with eigenvector $u = (1, \ldots, 1)/\sqrt{n}$, and eigenvalue $0$ with multiplicity $n - 1$ (eigenspace $u^\perp$). By standard Rayleigh--Schr\"odinger perturbation theory for the non-degenerate eigenvalue $n$:
\[
\lambda_1(C) \;=\; n - \langle u | D | u \rangle \;+\; \frac{1}{n}\sum_{v \in u^\perp}|\langle v | D | u \rangle|^2 \;+\; O\!\bigl(\|D\|^3\bigr).
\]
The first-order term evaluates to
\[
\langle u | D | u \rangle \;=\; \frac{1}{n}\sum_{i,j}D_{ij} \;=\; \frac{2}{n}\sum_{i<j}(1 - C_{ij}),
\]
giving the stated first-order formula after dividing by $n$. The second-order term is a sum of squared matrix elements, hence non-negative; it vanishes only when $D$ has no component in the $u^\perp$ sector (the rank-1 case, when $C$ is a constant-off-diagonal matrix). In all non-trivial cases it is strictly positive, so $\lambda_1(C) > n - \langle u|D|u\rangle$ at second order, equivalently $\epsilon < (2/n^2)\sum_{i<j}(1 - C_{ij})$: the first-order formula over-estimates the deficit, and the right-hand side is an upper bound.
\end{proof}

\noindent Verification: for $d = 5, \ldots, 12$ ($n = 8$), the first-order formula gives $\epsilon^{(1)} = 0.008173$; the exact eigenvalue gives $\epsilon = 0.008159$. Agreement: $0.17\%$; the exact value is below the first-order formula as predicted by the upper-bound direction of the theorem. The formula is verified to comparable precision for all paths tested.

\subsection{Numerical evidence for the correction}\label{sec:verification}

Three quantities with unambiguous cascade paths test the correction. The coupling coefficient $k_Q$ is fitted independently for each.

\begin{center}
\renewcommand{\arraystretch}{1.3}
\begin{tabular}{llcccccc}
\toprule
Observable & Path & $n$ & $\epsilon$ & $k_Q$ & Leading & Corrected & Observed \\
\midrule
$\alpha_s$ & $d\!=\!5\text{--}12$ & 8 & 0.00816 & 2.11 & 0.1159 & 0.1179 & 0.1179 \\
$\ell_A$ & $d\!=\!5\text{--}12$ & 8 & 0.00816 & 1.66 & 297.6 & 301.6 & 301.6 \\
$m_\tau/m_\mu$ & $d\!=\!6\text{--}13$ & 8 & 0.00657 & 2.67 & 16.53 & 16.82 & 16.82 \\
\bottomrule
\end{tabular}
\end{center}

\noindent All three residuals vanish when $k_Q$ is fitted individually. The coupling coefficients cluster between 1.7 and 2.7. Their variation reflects the different ways each observable couples to the subdominant eigenstructure.

\begin{remark}[The approximate universality of $k_Q \approx 2$]
Setting $k_Q = 2$ for all three quantities gives residuals of $-0.09\%$ ($\alpha_s$), $+0.27\%$ ($\ell_A$), and $-0.43\%$ ($m_\tau/m_\mu$)---sub-half-percent in every case, and a significant improvement over the uncorrected deviations of $-1.70\%$, $-1.34\%$, and $-1.72\%$. The value $k_Q = 2$ is the best integer approximation (RMS residual $0.30\%$, compared to $0.87\%$ for $k = 1$ and $0.76\%$ for $k = 3$). The optimal least-squares value across all three is $k = 2.08$.
\end{remark}

\subsection{Sign structure of the correction}

The correction $k_Q \epsilon$ is always positive ($\epsilon > 0$, $k_Q > 0$), so it increases the predicted value. This produces the observed sign structure across all cascade predictions:

\begin{itemize}[nosep]
\item \emph{Descent-dependent quantities} (those traversing multiple cascade layers) have negative leading-order deviations, because the independent-step approximation underestimates the propagated geometric content. The correction $+k_Q\epsilon$ closes the gap.
\item \emph{Geometric quantities} (single-level ratios, topological constants) have positive leading-order deviations, because they do not traverse multiple layers and receive no inter-layer correction. Their deviations reflect the subleading structure of the Gamma function at a single level, which is a separate (and smaller) effect.
\end{itemize}

The clean sign separation---negative for descent, positive for geometric---is not an empirical observation. It is a consequence of the Gram eigenvalue deficit being positive: $\epsilon > 0$ because $C_{ij} < 1$ for $i \neq j$ (Cauchy--Schwarz), so the deficit matrix $D$ has positive entries and $\langle u | D | u \rangle > 0$.

\subsection{The correction as compactification residual}

Theorem~4.2 establishes that each compactification step retains a residual $R_{\mathrm{eff}}(d) = 1/\sqrt{d+3}$. The overlap deficit (Theorem~\ref{thm:overlap}) scales as $1-C_{d,d+1}^2 \sim 1/(8d^2) \sim R_{\mathrm{eff}}^4/8$. The eigenvalue deficit $\epsilon$ is the cumulative effect of these per-step residuals over the path.

The independent-step approximation corresponds to the limit $R_{\mathrm{eff}} \to 0$ (complete compactification). The correction $k_Q\epsilon$ is the leading-order cost of $R_{\mathrm{eff}} > 0$---the imperfect folding of the cascade's own geometry.

\subsection{The first-order correction: a Beta function identity}\label{sec:first-order}

\begin{theorem}[Adjacent overlap deficit correction]\label{thm:first-order}
For a descent-dependent observable traversing path $\{d_0,\ldots,d_0+n-1\}$:
\begin{equation}\label{eq:first-order}
\frac{\delta Q}{Q_0} = \sum_{k=0}^{n-2}\bigl(1 - C_{d_k,d_{k+1}}^2\bigr),
\end{equation}
where each $C_{d,d+1}$ is the adjacent-layer correlation (Theorem~\ref{thm:overlap}).
This is a sum of Beta function ratios. No free parameter enters.
\end{theorem}

\begin{proof}
The observable propagates through $n$ layers. The independent-step approximation treats each layer as perfectly collinear with the next ($C_{d,d+1}=1$). The first-order departure from collinearity at each step is the squared overlap deficit $1-C_{d,d+1}^2$, which measures the fraction of the integrand at step $d$ that is orthogonal to step $d+1$ (Theorem~\ref{thm:overlap}). For a multiplicative propagator, the leading correction is the sum of per-step deficits.
\end{proof}

\begin{corollary}[Exponential resummation forced by multiplicative composition]\label{cor:exp-resum}
For a multiplicative propagator $Q_0 = \prod_k (\text{step}_k)$
traversing the path of Theorem~\ref{thm:first-order}, the unique
closed-form leading-order Gram-corrected observable respecting
multiplicative composition under path splitting is
\begin{equation}\label{eq:exp-form}
Q \;=\; Q_0\cdot\exp\!\Bigl(\sum_{k=0}^{n-2}\bigl(1-C_{d_k,d_{k+1}}^2\bigr)\Bigr) \;+\; O(\delta^2),
\end{equation}
where $\delta = \max_k (1 - C_{d_k,d_{k+1}}^2)$. The exponential
matches Theorem~\ref{thm:first-order} at first order in $\delta$ per
step.
\end{corollary}

\begin{proof}
\emph{Setup.} Because the cascade recurrence is first-order
multiplicative (Corollary~3.2 of Paper~0), the Gram-corrected step at
layer $d_k$ receives a multiplicative factor
$1 + (1 - C_{d_k,d_{k+1}}^2) + O((1-C^2)^2)$, accumulating as a product:
\[
Q \;=\; Q_0\prod_{k=0}^{n-2}\bigl(1 + \delta_k + O(\delta_k^2)\bigr),
\qquad \delta_k := 1 - C_{d_k,d_{k+1}}^2.
\]
This product is the exact corrected observable (modulo $O(\delta^2)$
per-step residuals from Theorem~\ref{thm:first-order}'s first-order
truncation). Define the path-functional
$\Phi(P) := \sum_{k\in P} \delta_k$ and the correction factor
$K(P) := Q/Q_0$ over a path $P$.

\emph{Composition law.} Multiplicative composition under path
splitting is a structural property of the cascade recurrence: for a
path $P = P_1 \cup P_2$ formed by concatenating two consecutive
sub-paths, the corrected observable composes as
$Q(P) = Q_0(P_1)\cdot Q_0(P_2) \cdot K(P_1) \cdot K(P_2) = Q_0(P)\cdot K(P_1)\cdot K(P_2)$.
Therefore $K(P) = K(P_1)\cdot K(P_2)$, and since $\Phi(P) = \Phi(P_1) + \Phi(P_2)$
(the path-functional is additive under concatenation), $K$ as a
function of $\Phi$ must satisfy
\begin{equation}\label{eq:cauchy}
K(\Phi_1 + \Phi_2) \;=\; K(\Phi_1) \cdot K(\Phi_2).
\end{equation}

\emph{Cauchy's exponential functional equation forces the form.}
Equation~\eqref{eq:cauchy} together with continuity (a closed-form
correction is a smooth function of its path-functional argument) and
the boundary condition $K(0) = 1$ (no correction for an empty path)
has, by the standard solution of Cauchy's exponential functional
equation, the unique family of solutions $K(\Phi) = e^{c\Phi}$ for
some constant $c$. Matching the first-order behaviour to
Theorem~\ref{thm:first-order} ($K(\Phi) = 1 + \Phi + O(\Phi^2)$ for
small $\Phi$, hence $K'(0) = 1$) fixes $c = 1$. Therefore
\[
K(\Phi) \;=\; e^\Phi \;=\; \exp\!\Bigl(\sum_k \delta_k\Bigr),
\]
and the corrected observable is~\eqref{eq:exp-form}.

\emph{Why the linear form fails.} The linear closed-form
$K_{\mathrm{lin}}(\Phi) = 1 + \Phi$ matches at first order but
violates~\eqref{eq:cauchy}: $K_{\mathrm{lin}}(\Phi_1) K_{\mathrm{lin}}(\Phi_2)
= (1 + \Phi_1)(1 + \Phi_2) = 1 + \Phi_1 + \Phi_2 + \Phi_1\Phi_2 \neq
1 + (\Phi_1 + \Phi_2) = K_{\mathrm{lin}}(\Phi_1 + \Phi_2)$. The cross-term
$\Phi_1\Phi_2$ breaks additivity. The exponential is the unique
continuous closed form respecting composition; the linear form is the
unique closed form failing it.

\emph{Residual versus the exact product.} The exact corrected
observable is the product $Q_0\prod(1 + \delta_k + O(\delta_k^2))$, not
the exponential. The exponential and the product agree at first order
in $\delta$ per step but differ at second order: the exact product
expands as $1 + \sum\delta_k + \sum_{j<k}\delta_j\delta_k + O(\delta^3)$,
while the exponential expands as $1 + \sum\delta_k + \tfrac{1}{2}(\sum\delta_k)^2 + O(\delta^3)
= 1 + \sum\delta_k + \tfrac{1}{2}\sum\delta_k^2 + \sum_{j<k}\delta_j\delta_k + O(\delta^3)$.
The exponential overestimates the exact product by
$\tfrac{1}{2}\sum\delta_k^2$ at second order. The linear closed-form
$1 + \sum\delta_k$, by contrast, underestimates the product by
$\sum_{j<k}\delta_j\delta_k$ at second order.

For cascade paths starting at $d_0\geq 5$ the per-step deficits scale
as $\delta_k = 1 - C^2_{d_k,d_{k+1}} \sim 1/(8d^2)$
(Theorem~\ref{thm:overlap}), so
$\sum_{k}\delta_k^2 \lesssim \tfrac{1}{64}\sum_{d\geq 5} d^{-4} \approx
6\times 10^{-5}$, and the exponential overshoots the exact product by
at most $\tfrac{1}{2}(6\times 10^{-5}) \approx 3\times 10^{-5}$ in the
multiplicative correction (i.e.\ $\sim 0.003\%$ of $Q$). The linear
form underestimates the product by approximately
$\tfrac{1}{2}(\sum\delta)^2 \approx 2\times 10^{-4}$ ($\sim 0.022\%$
of $Q$) for the longest path ($d=5$ to $d=217$). The exponential is
therefore not only the unique composition-respecting closed form but
also the closer numerical approximation to the exact product than the
linear form, by an order of magnitude.

Both gaps are below the reported observational precisions of all
currently-tested cascade observables ($\sim 0.1\%$ or coarser). The
exponential is therefore the cascade's natural closed-form correction
--- forced by composition, exact at first order in $\delta$, and within
observational precision of the exact product for all paths considered.
It is the form used downstream (Paper~I, $\rho_\Lambda$ closure;
Paper~IVb, the $v$ and $m_\tau$ absolute-mass shifts).
\end{proof}

\begin{center}
\renewcommand{\arraystretch}{1.3}
\begin{tabular}{llcccccc}
\toprule
Observable & Path & $n$ & $\sum(1-C^2)$ & Leading & Corrected & Observed & Dev \\
\midrule
$\alpha_s$ & $d\!=\!5$--$12$ & 8 & 0.01186 & 0.1159 & 0.1173 & 0.1179 & $-0.5\%$ \\
$m_\tau/m_\mu$ & $d\!=\!6$--$13$ & 8 & 0.00938 & 16.53 & 16.69 & 16.82 & $-0.8\%$ \\
$m_\mu/m_e$ & $d\!=\!14$--$21$ & 8 & 0.00272 & 206.50 & 207.06 & 206.77 & $+0.1\%$ \\
$m_\tau$~(MeV) & $d\!=\!5$--$12$ & 8 & 0.01186 & 1755 & 1776 & 1777 & $-0.1\%$ \\
$v$~(GeV) & $d\!=\!5$--$12$ & 8 & 0.01186 & 240.8 & 243.7 & 246.2 & $-1.0\%$ \\
$\Omega_m^{\rm Bott}$ & $d\!=\!5$--$217$ & 213 & 0.02108 & 0.3115 & 0.3181 & 0.315 & $+1.0\%$ \\
\bottomrule
\end{tabular}
\end{center}

\noindent The first-order correction reduces descent deviations from $\sim\!2\%$ to sub-$1\%$ for all tested quantities. The correction for $m_\mu/m_e$ is small ($0.27\%$) because its path lies at higher $d$ where the per-step deficit scales as $1/(8d^2)$.

\begin{remark}[Scope of Theorem~\ref{thm:first-order}; status of $\Omega_m^{\rm Bott}$]\label{rem:sp9-status}
Theorem~\ref{thm:first-order} is proved for a \emph{multiplicative
propagator} traversing a contiguous path---i.e., for an observable of
the form $Q = \prod_{k=0}^{n-1}(\text{step factor at }d_k)$. Five of the
six table entries ($\alpha_s$, $m_\tau/m_\mu$, $m_\mu/m_e$, $m_\tau$,
$v$) have this structure directly. The sixth entry,
$\Omega_m^{\rm Bott}$, is structurally different: by
Theorem~5.10 of Paper~V (``Bott partition matter fraction''),
\[
\Omega_m^{\rm Bott}
\;=\;
\frac
{\displaystyle\sum_{\substack{d=5\\ d\bmod 8\in\{5,6\}}}^{217}\Omega_{d-1}}
{\displaystyle\sum_{d=5}^{217}\Omega_{d-1}},
\]
a \emph{ratio of sums} over cascade layers rather than a single
multiplicative propagator. Its inclusion in the table under the
full-path correction $\sum_{d=5}^{217}(1-C^2_{d,d+1}) = 0.02108$
applies the theorem outside its proven scope:
\begin{itemize}
\item The rigorous per-term treatment would correct each
$\Omega_{d-1}$ in the sums by its own path-length-dependent factor,
$\Omega_{d-1}\mapsto\Omega_{d-1}\cdot
\exp\bigl(\sum_{k=4}^{d-1}(1-C^2_{k,k+1})\bigr)$, then re-form the
ratio. The sums are dominated by the early terms near $d\sim 5$--$7$
(where $\Omega_d$ peaks), which have short paths and small corrections
($\sim\exp(0.001)\approx 1.001$). The resulting correction to the
ratio is therefore much smaller than the full-path factor
$\exp(0.02108)$.
\item The observed numerics support this: the leading value
$\Omega_m^{\rm Bott,\;lead}=0.31150$ sits $-1.1\%$ below the
observed $0.315$; the full-path ``corrected'' value $0.3181$
overshoots by $+1.0\%$. A rigorous per-term correction would fall
between these and is consistent with the two-population systematic
to leading order. The table entry reports the uniform full-path
correction for comparability with the other rows; it should be read
as an upper bound on the Gram shift for this observable, not as its
derived value.
\end{itemize}
A formal extension of Theorem~\ref{thm:first-order} to ratio-of-sums
observables---deriving the effective correction for
$Q=\sum_i a_i/\sum_j b_j$ where each $a_i,b_j$ is a multiplicative
propagator---is Paper~V Open Question~1 (and remains open: the
identification step from non-trivial-phase layers to matter is the
other half of that question). The $\Omega_m$ closure does not
\emph{require} the Gram correction: both the leading lapse identity
$\Omega_m=1/\pi=0.31831$ ($+1.1\%$, geometric population) and the
leading Bott value $0.31150$ ($-1.1\%$, descent-dependent population)
bracket observation to the cascade's stated precision.
\end{remark}

\begin{corollary}[Derived coupling coefficient]\label{cor:kQ-derived}
The first-order coupling coefficient is
\begin{equation}\label{eq:kQ-first}
k_Q^{(1)} = \frac{\displaystyle\sum_{\rm adj}(1-C^2_{d,d+1})}{\epsilon},
\end{equation}
where the numerator is the adjacent overlap deficit sum (Theorem~\ref{thm:first-order})
and the denominator is the eigenvalue deficit (Theorem~\ref{thm:pert}). Both are sums
of Beta function ratios. No free parameter enters.
\end{corollary}
\begin{proof}
$k_Q^{(1)}$ is defined by equating the first-order correction
$\delta Q/Q_0 = \sum(1-C^2)$ with the parametrisation $k_Q\,\epsilon$. The ratio is
$\sum(1-C^2)/\epsilon$, with both quantities derived.
\end{proof}

\begin{center}
\renewcommand{\arraystretch}{1.2}
\begin{tabular}{llcc}
\toprule
Path & $n$ & $k_Q^{(1)}$ & Residual after $k_Q^{(1)}$ \\
\midrule
$d = 5, \ldots, 12$ & 8 & 1.454 & sub-$0.6\%$ \\
$d = 6, \ldots, 13$ & 8 & 1.429 & sub-$0.8\%$ \\
$d = 14, \ldots, 21$ & 8 & 1.362 & sub-$0.2\%$ \\
\bottomrule
\end{tabular}
\end{center}

\begin{remark}[Status of the observable response factor]
The fitted $k_Q$ values (2.11, 1.66, 2.67) exceed $k_Q^{(1)}$ for some observables and
fall below it for others. The variation is observable-dependent: it reflects how each
quantity's functional form responds to the universal geometric correction
$\delta\Phi = \sum(1-C^2)$. For direct exponentials ($\alpha_s = \alpha_{\rm GUT}\times
e^{\Phi}$), the response is close to~1. For integrated quantities ($\ell_A$, involving the
comoving distance integral), the response is smoothed below~1. For quantities on higher-$d$
paths ($m_\mu/m_e$), the first-order over-corrects and the effective
$k_Q < k_Q^{(1)}$. The observable response factor $R_Q = k_Q/k_Q^{(1)}$ is a property of the observable's formula,
not of the cascade's geometry, and is distinct from the perturbation-theory second-order term in Theorem~\ref{thm:pert}.
\end{remark}

\subsection{What this section proves}

\begin{enumerate}[nosep]
\item The Gram matrix of cascade layer integrands is a matrix of Beta function values (Theorem~\ref{thm:gram}).
\item Adjacent layers have a strictly positive overlap deficit $1 - C_{d,d+1}^2 > 0$, scaling as $1/(8d^2)$ (Theorem~\ref{thm:overlap}). The correlation has the exact closed form $C^2_{d,d+1} = R(2d+2)^2 / [R(2d+1)\,R(2d+3)]$ in cascade slicing ratios (Theorem~\ref{thm:gram-R}), equivalently $\log C^2_{d,d+1} = -\tfrac{1}{2}\Delta^2 \log\alpha\big|_{2d+2}$ (Corollary~\ref{cor:gram-laplacian}): the Gram correction is the discrete log-Laplacian of the cascade scalar action's compliance function $\alpha(d) = R(d)^2/4$ at doubled argument, unifying the supplement with Paper~IVb's action principle (Remark~\ref{rem:gram-action-unification}). The closed form generalises to all layer pairs (Theorem~\ref{thm:gram-matrix}): $C_{ij} = R(d_i+d_j+1)/\sqrt{R(2d_i+1)\,R(2d_j+1)}$, which makes the entire correlation matrix---and hence the eigenvalue deficit $\epsilon$ (Corollary~\ref{cor:eigenvalue-cascade})---manifestly cascade-internal.
\item The correlation matrix has one dominant eigenvalue $\lambda_1 \approx n(1 - \epsilon)$ with $\epsilon > 0$ (Theorem~\ref{thm:eigenvalue}).
\item The eigenvalue deficit is derived from first-order perturbation theory as a sum of Beta function deficits (Theorem~\ref{thm:pert}).
\item The sign of the correction (positive) explains the clean sign separation between descent and geometric predictions.
\item The first-order coupling coefficient $k_Q^{(1)} = \sum(1-C^2)/\epsilon$ is derived (Corollary~\ref{cor:kQ-derived}): a ratio of Beta function sums with no free parameter. It reduces all descent deviations to sub-$1\%$.
\item The residual variation between observables (the observable response factor $R_Q = k_Q/k_Q^{(1)}$) is a property of each observable's functional form, not of the cascade's geometry. It is observable-dependent and open.
\item Every quantity in this section---including $k_Q^{(1)}$---is a Gamma function identity. No physics enters.
\end{enumerate}

\subsection{Open questions}\label{sec:open}

\begin{enumerate}[nosep]
\item \emph{Derive the observable response factor.} The first-order correction $\delta Q/Q_0 = \sum(1-C^2)$ (Theorem~\ref{thm:first-order}) and its coupling coefficient $k_Q^{(1)} = \sum(1-C^2)/\epsilon$ (Corollary~\ref{cor:kQ-derived}) are derived. They reduce descent deviations from $\sim\!2\%$ to sub-$1\%$ for all tested quantities. The residual variation is observable-dependent: for direct exponentials ($\alpha_s$) the first-order under-corrects by $\sim\!0.5\%$; for integrated quantities ($\ell_A$) it over-corrects slightly; for high-$d$ paths ($m_\mu/m_e$) it over-corrects by $\sim\!0.1\%$. The observable response factor $R_Q = k_Q/k_Q^{(1)}$ depends on how each observable's formula converts the geometric correction $\delta\Phi$ into a prediction shift $\delta Q/Q$, and is distinct from the perturbation-theory second-order eigenvalue term of Theorem~\ref{thm:pert}. Deriving $R_Q$ for each observable is a calculus problem (differentiate the observable's formula with respect to the cascade potential), not a cascade geometry problem.
\item \emph{Analytic formula for $\epsilon(n, d_0)$ --- partially closed.} The deficit is the dominant-eigenvalue deficit of the explicit cascade-native matrix $C_{ij} = R(d_i+d_j+1)/\sqrt{R(2d_i+1)\,R(2d_j+1)}$ (Theorem~\ref{thm:gram-matrix} and Corollary~\ref{cor:eigenvalue-cascade}), so the matrix's entries are now in cascade slicing ratios rather than opaque Beta function values. The first-order perturbation expression of Theorem~\ref{thm:pert} is correspondingly a sum over $i<j$ pairs of $1 - R(d_i+d_j+1)/\sqrt{R(2d_i+1)\,R(2d_j+1)}$ (equation~\eqref{eq:pert-cascade}). What remains: a closed-form expression for $\lambda_1$ itself in terms of $n$ and $d_0$ (eliminating the numerical eigenvalue computation entirely). The matrix is now manifestly cascade-internal, so any closed form would derive from properties of $R$ at consecutive doubled arguments rather than from Beta function asymptotics.
\item \emph{Higher-order corrections.} The subdominant eigenvalues $\lambda_3, \lambda_4, \ldots$ contribute at third order and beyond. For the $d = 5, \ldots, 12$ path, $\lambda_2/n = 0.81\%$ and $\lambda_3/n = 0.008\%$: the third-order correction is negligible. For longer paths, it may not be.
\item \emph{Application to mixed quantities.} The correction applies cleanly to quantities with a single unambiguous cascade path ($\alpha_s$, $\ell_A$, mass ratios). Quantities depending on multiple interacting paths (absolute masses, the electroweak scale) require decomposing the observable into its cascade-path components.
\end{enumerate}

\begin{remark}[Status of the correction]
The first-order correction (Theorem~\ref{thm:first-order}) is derived: it is a sum of Beta function ratios with no free parameters, reducing descent deviations from sub-$2\%$ to sub-$1\%$. Theorem~\ref{thm:gram-R} and Corollary~\ref{cor:gram-laplacian} sharpen this derivation: the per-step Gram correction is the discrete log-Laplacian of the cascade scalar action's compliance function $\alpha$, evaluated at the doubled cascade argument. This identifies the Gram framework as a corollary of Paper~IVb's action principle---both correction families (single-source $\alpha(d^*)/\chi^k$ and path-distributed Gram) derive from the same compliance $\alpha(d)=R(d)^2/4$. The observable response factor (the subleading correction beyond the first-order Gram identity) remains the open problem at the level of how observables couple to the geometric correction; the geometric correction itself is now derived from cascade-native first principles.
\end{remark}

\begin{remark}[Superseded for $\alpha_s$, $m_\tau/m_\mu$, and $m_\tau$]\label{rem:superseded-by-family}
The Gram-matrix framework above is a general structural account of how inter-layer coupling corrects the independent-step approximation. For three specific Standard Model precision observables---$\alpha_s(M_Z)$, $m_\tau/m_\mu$, and the $\tau$ absolute mass---a sharper structural identity has subsequently been identified in Part~IVb. Both deviations in these observables are closed \emph{at experimental precision} by cascade potential shifts of the form $\delta\Phi = \alpha(d^*)/\chi$, where $d^*$ is a distinguished dimension from this paper's four-dimension tower (Theorem~7.1) and $\chi=\chi(S^{2n})=2$ is the Euler characteristic of the $\mathrm{SU}(2)$ sphere:
\begin{itemize}
\item $\delta\Phi_{\mathrm{U}(1)}=\alpha(14)/\chi$ closes $\alpha_s$ and $m_\tau/m_\mu$ together (Part~IVb, Theorem~4.3);
\item $\delta\Phi_{\mathrm{phase}}=\alpha(19)/\chi$ closes the $\tau$ absolute mass at the phase-transition threshold $d_1=19$ (Part~IVb, Theorem~4.4).
\end{itemize}
These identities supersede the Gram-matrix corrections for the three observables named: the universal $k_Q\approx 2$ fit used for $\alpha_s$ and $m_\tau/m_\mu$ in Section~\ref{sec:verification} above is replaced by a parameter-free $\alpha(d^*)/\chi$ closed form that works at experimental precision rather than sub-$1\%$. The $m_\mu/m_e$ match through the Gram-matrix first-order correction remains the current best account for that observable. Open Question~1 (the observable response factor $R_Q$) is accordingly narrowed: the three highlighted observables no longer require it, but the remaining descent-dependent quantities ($v$, $m_W$, $\Omega_m$, $\ell_A$, other absolute masses) may be closed either by further members of the $\alpha(d^*)/\chi$ family or by a subleading Gram-matrix response, to be determined.
\end{remark}

\end{document}
