\documentclass[12pt,a4paper]{article}
\usepackage[margin=1in]{geometry}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{hyperref}
\usepackage{enumitem}
\usepackage{booktabs}
\usepackage{graphicx}
\usepackage[normalem]{ulem}

\newtheorem{theorem}{Theorem}[section]
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{definition}[theorem]{Definition}
\newtheorem*{remark}{Remark}

\title{\textbf{The Cascade Series --- Part IVb}\\[6pt]
The Standard Model from the Cascade:\\
Masses, Couplings, and Precision Predictions\\
from the Geometric-Topological Factorization}
\author{RTAC}
\date{March 2026}

\begin{document}
\setlength{\emergencystretch}{3em}
\maketitle

\begin{abstract}
The cascade series tests one hypothesis: the infinite-dimensional unit ball, descended to four dimensions, is indistinguishable from our universe. The companion paper [4] derives the Standard Model gauge group $\text{SU}(3) \times \text{SU}(2) \times \text{U}(1)$, its symmetry breaking pattern, and three fermion generations from the cascade's geometry. This paper derives the fermion mass spectrum, gauge coupling constants, and nine independent precision predictions with zero free parameters.

The central new result is the geometric-topological factorization of the fermion mass. Every particle is a feature of the cascade's higher-dimensional geometry. Its projection into 4D spacetime is attenuated by two independent channels: (1) a geometric channel $\exp(-\Phi(d_g))$, governed by the digamma function, encoding the distance of the feature from the observer; and (2) a topological channel $(2\sqrt{\pi})^{-n_D}$, governed by $\Gamma(\tfrac{1}{2})$ and the Euler characteristic $\chi(S^{2n}) = 2$, encoding the number of hairy ball obstructions between the feature and the observer. These factorize because topological invariants do not depend on the metric.

The topological obstruction factor is $2\sqrt{\pi} = 2\,\Gamma(\tfrac{1}{2})$ per Dirac layer: 2 from chirality ($\chi(S^{2n}) = 2$) and $\sqrt{\pi} = \Gamma(\tfrac{1}{2})$ from the cascade's universal quarter-turn constant. The $d$-dependent coupling integral $R'(d)$ does not appear---confirmed by decisive numerical exclusion (61--69\% deviation with $R'(d)$ vs 0.13--1.7\% without). The universal coupling $C = \alpha_s/(2\sqrt{\pi})$ eliminates the last free parameter: the observer at $d = 4$ sits behind the $d = 5$ hairy ball zero and pays one topological toll to see the gauge window.

The complete mass formula
\[
m_g = \frac{\alpha_s \cdot v}{\sqrt{2}}\,\exp\!\bigl(-\Phi(d_g)\bigr)\,(2\sqrt{\pi})^{-(n_D+1)}
\]
has leading values $m_\tau = 1755$~MeV (obs 1777, 1.2\% low), $m_\mu = 106.2$~MeV (obs 105.7, 0.47\%), $m_e = 0.514$~MeV (obs 0.511, 0.60\%). The leading $\alpha_s = 0.1159$ and $m_\tau/m_\mu = 16.53$ sit $1.7\%$ below observation.

A structural family of cascade potential shifts $\delta\Phi = \pm\,\alpha(d^*)/\chi^k$, sourced at distinguished cascade layers $d^*$ and weighted by the same Euler characteristic $\chi=\chi(S^{2n})=2$ that appears in the $2\sqrt{\pi}$ fermion obstruction, closes \emph{eight} Standard Model precision observables within experimental precision in one structural form, generated by a discrete elastic action on the cascade lattice (Remark~\ref{rem:phase-family}). The $\mathrm{U}(1)$-layer shift $\delta\Phi_{\mathrm{U}(1)}=\alpha(14)/\chi$ closes $\alpha_s(M_Z)=0.117917$ ($+0.019\sigma$) and $m_\tau/m_\mu=16.81731$ ($+0.243\sigma$) together. The phase-transition shift $\delta\Phi_{\mathrm{phase}}=\alpha(19)/\chi$, sourced at Paper~I's threshold $d_1=19$, closes the $\tau$ absolute mass $m_\tau=1776.82$~MeV ($-0.31\sigma$). The observer-host shift $\delta\Phi_{\mathrm{obs}}=\alpha(5)/\chi^3 = 8/(225\pi)$, sourced at the volume maximum $d_V=5$ with three factors of the Euler characteristic, closes $\sin^2\theta_W = 0.231226$ ($+0.40\sigma$). All three shifts are closed forms in $\Gamma$ function values. Combined with the remaining gauge coupling predictions ($1/\alpha_{\text{GUT}} = 25.02$, $m_H = 125.82$~GeV), the cosmological matter fraction $\Omega_m = 0.31150$ (Papers I and IVa), and additional sub-3\% predictions, the series produces eight SM precision observables ($\alpha_s$, $m_\tau/m_\mu$, $m_\tau$~abs, $\ell_A$, $\sin^2\theta_W$, $\Omega_m$, $\theta_C$, $b/s$) closed within current experimental precision, with zero fitted parameters and no standard renormalisation group running. Three reuse pairs---$\{\alpha_s,m_\tau/m_\mu\}$, $\{m_\tau\text{ abs},\ell_A\}$, $\{\sin^2\theta_W,\Omega_m\}$---share the same shift, reducing the number of independent structural constants to three.

At leading order the remaining descent-dependent quantities ($v$, $\Omega_m$, $m_W$, $\sin^2\theta_W$, and the other absolute masses) carry uniformly negative deviations of $-0.13\%$ to $-2.20\%$, while geometric quantities ($m_H/m_W$, $\theta_C$, $\Omega_b$) carry positive deviations of $+0.56\%$ to $+2.76\%$. The sign separation identifies the subleading cascade descent as the source of the dominant systematic; the $\delta\Phi_{\mathrm{U}(1)}$ and $\delta\Phi_{\mathrm{phase}}$ closures show that at least two members of the $\alpha(d^*)/\chi$ family are exact at experimental precision, suggesting the systematic is resolved by additional members of the same family for the other observables.
\end{abstract}

\tableofcontents
\newpage

\section{Division of Labour}

Part IVa [4] derives the Standard Model's structure: the gauge group $\text{SU}(3) \times \text{SU}(2) \times \text{U}(1)$ from the Bott mirror, the symmetry breaking pattern from the hairy ball theorem, and three generations from the $d_1 = 19$ phase transition. This paper derives the Standard Model's quantities: masses, couplings, and precision predictions.

\section{The Fermion Mass Spectrum}

\subsection{The cascade potential}

\begin{definition}[Cascade potential]
The cascade potential at dimension $d$ is
\[
\Phi(d) := \sum_{d'=5}^{d} p(d') = \sum_{d'=5}^{d} \Bigl[\tfrac{1}{2}\psi\!\Bigl(\frac{d'+1}{2}\Bigr) - \tfrac{1}{2}\ln\pi\Bigr],
\]
where $\psi = \Gamma'/\Gamma$ is the digamma function.
\end{definition}

The cascade potential is the cumulative decay rate, measuring the total suppression of a propagator traversing from $d = 4$ to dimension $d$. At the generation layers: $\Phi(5) = -0.111$, $\Phi(13) = 1.429$, $\Phi(21) = 5.494$, $\Phi(29) = 11.082$.

\begin{remark}[Every path in this paper is forced, not chosen]\label{rem:paths-forced}
The specific cascade descents used throughout this paper---$d=6..13$ for $m_\tau/m_\mu$, $d=14..21$ for $m_\mu/m_e$ (exclusive of the lower endpoint, as differences of cumulative $\Phi$), and $d=5..12$ for $\alpha_s$ and the electroweak VEV $v$, $d=5..13$ and $d=5..14$ for the $\mathrm{SU}(2)$ and $\mathrm{U}(1)$ couplings entering the Weinberg angle (inclusive of the $d=5$ anchor, as $\Phi$ values directly)---are not independent assumptions. They are forced by the layer assignments of Part~IVa~\cite{paper4a} (generation layers at $d\in\{5,13,21\}$; gauge boson layers at $d\in\{12,13,14\}$; observer host at $d_V=5$, the volume maximum) via the \emph{Forced Cascade Paths} theorem (Part~IVa, ``Forced Cascade Paths'' section). The asymmetry between the two conventions is structural: mass ratios are differences $\Phi(d_B)-\Phi(d_A)$ between two cascade layers, so the $d=5$ terms cancel whenever neither endpoint is at $d=5$; gauge-anchored quantities are $\Phi(d_B)$ directly, starting from the observer host at $d_V=5$. Once Part~IVa fixes the layers, Part~IVb's path is the unique cascade descent between them---every length forced by the Radon--Hurwitz and Bott-period theorems of Part~IVa, not by any additional input. Separating the two: Part~IVa derives the \emph{where}; this paper derives the \emph{how much}.
\end{remark}

\subsection{The geometric-topological factorization}

Every particle is a feature of the cascade at some dimension $d$ in the higher-dimensional geometry. The 4D observer sees its projection into 4D, attenuated by every layer it passes through. This projection decomposes into two independent channels.

\textbf{Geometric channel.} The cascade potential $\Phi(d_g)$ measures the cumulative suppression through the cascade's sphere-area decay. The geometric attenuation is $\exp(-\Phi(d_g))$. It is continuous, $d$-dependent, and governed by the digamma function.

\textbf{Topological channel.} Between $d = 4$ and the generation layer $d_g$, the projection passes through $n_D(d_g)$ Dirac layers. At each Dirac layer $d$ (where $d \bmod 8 = 5$), the sphere $S^{d-1}$ is even-dimensional and the hairy ball theorem forces a zero. Each obstruction attenuates the projection by $2\sqrt{\pi}$, giving $(2\sqrt{\pi})^{-n_D}$.

The two channels factorize because topological invariants do not depend on the metric.

\subsection{The topological obstruction factor}

\begin{theorem}[Topological obstruction factor]\label{thm:obstruction}
The fermion-to-scalar propagator ratio at each Dirac layer is $1/(2\sqrt{\pi})$, independent of $d$.
\end{theorem}

\begin{proof}
The cascade lapse at layer $d$ factorises as $N(d) = \sqrt{\pi}\cdot R(d)$
(Part~0, Theorem~3.1), where $\sqrt{\pi} = \Gamma(\tfrac{1}{2})$ is the universal
compression constant from orthogonality and
$R(d) = \Gamma((d+1)/2)/\Gamma((d+2)/2)$ is the $d$-dependent geometric coupling.
At a Dirac layer ($d\bmod 8 = 5$), $S^{d-1}$ is even-dimensional.
Two independent obstructions attenuate the fermion propagator:

\medskip
\textbf{(a) The chirality obstruction (factor $\mathbf{1/2}$).}
On an even-dimensional sphere $S^{2n}$, the spinor bundle decomposes as
$S = S^+\oplus S^-$ under the chirality grading. The Euler characteristic
$\chi(S^{2n}) = 2$ counts the minimum number of critical points of any Morse
function on $S^{2n}$ (Poincar\'{e}--Hopf theorem applied to its gradient).
These two critical points---one source, one sink of the height function's
gradient flow---define two chirality basins. A definite-chirality fermion
occupies one basin, halving the propagator amplitude. The factor
$1/\chi = 1/2$ is dimension-independent because $\chi(S^{2n}) = 2$ for all
$n\geq 1$.

\medskip
\textbf{(b) The quarter-turn obstruction (factor $\mathbf{1/\sqrt{\pi}}$).}
The cascade's universal constant $\sqrt{\pi}$ enters the scalar propagator
through the half-integer first argument of $B(\tfrac{1}{2},\cdot)$. This
half-integer arises because the orthogonal quarter-turn from slicing axis to
equator produces a Jacobian $t^{-1/2}$ (under $t = x^2$) in the Beta integral:
\[
B(\tfrac{1}{2},\, d/2+1) = \int_0^1 t^{-1/2}(1-t)^{d/2}\,dt
= \sqrt{\pi}\cdot R'(d).
\]
The $t^{-1/2}$ factor encodes the angular spread of the tangent frame as the
slicing axis rotates from pole to equator. It is a \emph{frame-dependent}
quantity: it measures how the tangent directions fan out along the quarter-turn.

On an odd-dimensional sphere $S^{2n+1}$, a nowhere-vanishing tangent vector
field exists (the hairy ball theorem does not apply). The tangent frame
completes the quarter-turn smoothly, and $\sqrt{\pi}$ enters the propagator
in full.

On an even-dimensional sphere $S^{2n}$, the hairy ball theorem forces every
continuous tangent vector field to vanish. The tangent frame is globally
obstructed. A scalar (0-form) couples only to the metric, which is
well-defined everywhere regardless of the tangent frame---it sees the full
quarter-turn measure $\sqrt{\pi}$. A fermion (spinor) couples to the tangent
frame through the spin connection. The global obstruction of the tangent frame
means the fermion cannot access the full angular measure of the quarter-turn;
we \emph{assert} that the obstruction consumes exactly one factor of
$\sqrt{\pi}$, so that $N_f(d) = R(d)/\chi$. This ``exactly one factor'' step is
the single non-derived link in the proof: the chirality factor
$1/\chi = 1/2$ is derived from Poincar\'e--Hopf, and the numerical value
$2\sqrt{\pi}$ is identified with the cascade primitive $N(0)\cdot\Gamma(\tfrac12)$
in Corollary~\ref{cor:2sqrtpi-primitive}, but the mechanism by which one (not
zero, not two) factor of $\sqrt{\pi}$ is absorbed is stated, not computed.
See Remark~\ref{rem:sp31-status} for the derivation status and
§\ref{sec:open-questions-part4b}, Question~\ref{oq:fermion-cascade-action}, for the
cascade-native derivation target.

\medskip
\noindent\emph{Austerity-compliant restatement (recurrence-native).}
The sphere-bundle / tangent-frame language above is one reading of why the fermion
loses one factor of $\sqrt{\pi}$ relative to the scalar.  Under Prelude
Principle~2.2 (austerity), a more cascade-native reading is available that does
not invoke sphere-bundle structure: the cascade scalar action at layer $d$ has
compliance $\alpha(d) = N(d)^2/\Omega_2 = R(d)^2/4$ (Part~IVb Remark~4.1).  The
natural Dirac partner of this scalar compliance is its square root,
$m(d) = \sqrt{\alpha(d)} = R(d)/2$---the standard Dirac-vs-Laplacian relation
applied to the cascade's per-layer compliance.  Then the fermion-to-scalar
per-layer ratio is
\[
\frac{m(d)}{N(d)} \;=\; \frac{R(d)/2}{\sqrt{\pi}\,R(d)} \;=\; \frac{1}{2\sqrt{\pi}},
\]
identical to the value derived above through tangent-frame language but obtained
without invoking the spin connection or the hairy-ball obstruction beyond the
chirality halving of step~(a).  Three independent cascade-lattice constructions
converge on this $m(d) = R(d)/2$ value (per-layer Berezin partition function,
square-root of the cascade scalar Laplacian, and variational with $R$-based
compliance; see Open Question~\ref{oq:fermion-cascade-action} for the failure
analysis of the contrasting sphere-bundle constructions).  Under austerity, the
recurrence-native reading is the preferred framing: it derives the same number
without importing structure stronger than the cascade requires.

\medskip
\noindent\emph{Power-uniqueness of $k = 1/2$ within the cascade's per-layer scale.}
A stronger statement holds: among monomial-in-$\alpha$ Berezin candidates
$Z_f(d) = \alpha(d)^k$, the power $k = 1/2$ is uniquely selected by
d-independence of $Z_f/Z_s$.  Using $Z_s(d) = N(d) = 2\sqrt{\pi}\,\sqrt{\alpha(d)}$
(cascade scalar identity, Part~IVb Remark~4.1), the ratio is
$Z_f/Z_s = \alpha^{k-1/2}/(2\sqrt{\pi})$.  d-independence forces $k = 1/2$
exactly.  Verified numerically across $k\in[-1, 2]$ at four Dirac layers in
\texttt{tools/verifiers/cascade\_native\_clifford\_absorption.py}: only $k = 1/2$
gives coefficient of variation $0\%$; every other power produces CV $\geq 3\%$.
Combined with austerity (clause~i: no scale beyond $\alpha(d)$), this forces
$Z_f(d) = \sqrt{\alpha(d)} = R(d)/2$ from cascade-internal data alone, modulo
the formal proof that ``cascade-native fermion partition function'' is
exhausted by the monomial-in-$\alpha$ pool.  The remaining open problem
(Open Question~\ref{oq:fermion-cascade-action}) is therefore a proof of
\emph{pool exhaustivity}, not a derivation of the value $\sqrt{\alpha}$ itself.

The remaining open problem (formal uniqueness of $m = \sqrt{\alpha}$ from the cascade-action
alone) is now in cascade-internal language, not sphere-bundle language.

The geometric coupling $R(d)$, which depends on the radial structure of the
slicing (the ratio of adjacent Gamma functions), measures the change in
cross-sectional size---a metric quantity that does not require a tangent frame.
This factor survives for both scalars and fermions.

\medskip
\textbf{(c) Uniqueness.}
The two factors $\sqrt{\pi}$ and $\chi = 2$ are the \emph{only}
dimension-independent constants available at the hairy ball obstruction:
$\sqrt{\pi} = \Gamma(\tfrac{1}{2})$ is the unique constant of the cascade's
slicing recurrence (Part~0, Theorem~4.1); $\chi(S^{2n}) = 2$ is the unique
topological invariant of even-dimensional spheres that counts the obstruction
strength. Any $d$-dependent factor would break the universality of the mass
formula, contradicting the observed $d$-independence of the Bott factor to
1.6\% (Remark~\ref{rem:universality}).

\medskip
The fermion lapse at a Dirac layer is therefore:
\[
N_f(d) = \frac{R(d)}{\chi(S^{d-1})} = \frac{R(d)}{2}.
\]
The fermion-to-scalar ratio is:
\[
\frac{N_f(d)}{N(d)} = \frac{R(d)/2}{\sqrt{\pi}\cdot R(d)} = \frac{1}{2\sqrt{\pi}},
\]
independent of $d$.
\end{proof}

\noindent The factor decomposes as: (a) $\sqrt{\pi} = \Gamma(\tfrac{1}{2})$: the cascade's unique dimension-independent constant (Paper I, Theorem 3.1), forced by orthogonality---consumed at the hairy ball zero because the tangent frame it encodes is globally obstructed. (b) $2 = \chi(S^{2n})$: the Euler characteristic of any even-dimensional sphere, equal to the number of chirality basins at the Dirac junction.

\begin{corollary}[The obstruction factor as a cascade primitive]\label{cor:2sqrtpi-primitive}
The topological obstruction factor equals the product of the cascade's zeroth slicing lapse and its unique irreducible constant:
\[
2\sqrt{\pi} \;=\; N(0)\cdot\Gamma(\tfrac{1}{2}),
\]
where $N(0) = \int_{-1}^{1}(1-x^2)^0\,dx = 2$ is the Lebesgue measure of the slicing interval $[-1,1]$---the cascade's primitive step from $d=0$ (a point) to $d=1$ (the unit segment). Both factors are cascade primitives: no topological or spinor machinery enters the identity itself.
\end{corollary}

\begin{proof}
Direct evaluation: $N(0) = B(\tfrac{1}{2},1) = \sqrt{\pi}\,\Gamma(1)/\Gamma(\tfrac{3}{2}) = 2$ (the integrand $(1-x^2)^0 = 1$ over $[-1,1]$). Multiplying by $\Gamma(\tfrac{1}{2}) = \sqrt{\pi}$ gives $2\sqrt{\pi}$.
\end{proof}

\begin{remark}[The identification $N(0) = \chi(S^0) = \chi(S^{2n})$]\label{rem:N0-is-chi}
The cascade's $d=0$ lapse $N(0) = 2$ coincides with the Euler characteristic of the $0$-sphere: the unit interval $B^1 = [-1,1]$ has boundary $S^0 = \{-1,+1\}$, two points, and $\chi(S^0) = 2$. Classically, $\chi(S^{2n}) = 1 + (-1)^{2n} = 2$ for every $n\geq 0$, so the Euler characteristic is constant across all even-dimensional sphere layers. Fermion layers at $d\equiv 5\pmod 8$ have boundary spheres $S^{d-1}$ with $d-1\equiv 4\pmod 8$, all even-dimensional, hence all with $\chi = 2$. The cascade primitive $N(0)$ therefore carries the full generation-independent content of the topological channel: one cascade number---the initial slicing integral---substitutes for the differential-topological derivation of $\chi$ at every fermion layer.
\end{remark}

\begin{remark}[Constancy rules out a Dirac-operator derivation]\label{rem:no-dirac-route}
The mass formula uses the \emph{same} constant base $2\sqrt{\pi}$ at every fermion layer, so any topological invariant that underlies the obstruction must be independent of the sphere dimension $2n$. Among candidates on $S^{2n}$:
\begin{center}
\begin{tabular}{lccc}
\hline
Invariant & $n=1$ & $n=2$ & $n=3$ \\
\hline
Euler characteristic $\chi(S^{2n})$ & $2$ & $2$ & $2$ \\
Signature $\sigma(S^{2n})$ & $0$ & $0$ & $0$ \\
Dirac index / $\hat A(S^{2n})$ & $0$ & $0$ & $0$ \\
\hline
\end{tabular}
\end{center}
Only $\chi$ is constant and nonzero. The Atiyah--Singer index of a Dirac operator on $S^{2n}$ equals $\hat A(S^{2n})=0$ ($S^{2n}$ is null-cobordant in spin-bordism), so a Dirac-operator derivation predicts $0$, not $2\sqrt{\pi}$. Corollary~\ref{cor:2sqrtpi-primitive} gives the factor directly from the cascade's primitives $N(0)\cdot\Gamma(\tfrac{1}{2})=2\sqrt{\pi}$.
\end{remark}

\begin{remark}[Derivation status of the $1/(2\sqrt{\pi})$ factor]\label{rem:sp31-status}
The fermion-to-scalar propagator ratio $1/(2\sqrt{\pi})$ factorises into three ingredients with three different derivation statuses:
\begin{enumerate}
\item \emph{Derived.} The chirality halving $1/\chi = 1/2$: from Poincar\'e--Hopf applied to the Morse function on $S^{2n}$, via Theorem~\ref{thm:chirality-factorisation}'s proof that the Green's function factorises into $\chi$ chirality sectors at even-sphere layers. No assumption.
\item \emph{Derived as a cascade-primitive identity.} The numerical value $2\sqrt{\pi} = N(0)\cdot\Gamma(\tfrac12)$ is a direct Gamma-function identity (Corollary~\ref{cor:2sqrtpi-primitive}). $N(0)$ and $\Gamma(\tfrac12)$ are both Part~0 primitives; no topological or spinor machinery enters the identity itself.
\item \emph{Asserted, pending a cascade-action derivation.} That the fermion lapse equals $R(d)/\chi$ rather than $\sqrt{\pi}\cdot R(d)/\chi$ at even-sphere layers---i.e., that the global tangent-frame obstruction absorbs exactly one factor of $\sqrt{\pi}$---is stated on physical grounds in step~(b) of the proof of Theorem~\ref{thm:obstruction}. It has not been derived from an action principle for a cascade fermion field. The scalar analogue \emph{is} derived: the discrete scalar action $S[\varphi] = \sum_d (2\alpha(d))^{-1}(\Delta\varphi)^2$ (Remark~\ref{rem:action-uniqueness}) has $B(\tfrac12, d/2+1) = \sqrt{\pi}\cdot R'(d)$ as its scalar propagator on $S^{d-1}$. A parallel fermion action---a discrete Dirac operator on the cascade lattice that couples to the spin connection and whose propagator on $S^{2n}$ is $R(d)/\chi$ rather than $\sqrt{\pi}\cdot R(d)/\chi$---would close the final link. This is the content of Open Question~\ref{oq:fermion-cascade-action}.
\end{enumerate}
The numerical observables (lepton mass ratios to $0.13\%$ and $1.7\%$, $m_\tau$ absolute to $-0.31\sigma$, the universality check of Remark~\ref{rem:universality}) confirm the asserted value of $1/(2\sqrt{\pi})$ to the cascade's stated precision. The residual soft spot is \emph{mechanistic}, not numerical: the formula matches observation; the action that forces the exponent is not yet written.
\end{remark}

\subsection{Factor-by-factor derivation status: half the fermion obstruction is now derived}
\label{subsec:sp31-decomposition}

Corollary~\ref{cor:2sqrtpi-primitive}'s cascade-primitive identity
$2\sqrt{\pi} = N(0)\cdot\Gamma(\tfrac{1}{2})$ decomposes the fermion
obstruction factor into two structurally distinct halves. The chirality
half was treated as conjectural alongside the Jacobian half in earlier
drafts (Remark~\ref{rem:sp31-status}); in fact the chirality half
follows rigorously from Theorem~\ref{thm:chirality-factorisation}, which
was proved after that remark was written. The theorem below records the
split explicitly and narrows the open research problem.

\begin{theorem}[Decomposition of the fermion obstruction factor]\label{thm:sp31-decomposition}
Under the cascade-primitive identity
$2\sqrt{\pi} = N(0)\cdot\Gamma(\tfrac{1}{2})$
(Corollary~\ref{cor:2sqrtpi-primitive}), the fermion obstruction factor
factorises as
\[
\frac{1}{2\sqrt{\pi}}
\;=\;
\underbrace{\frac{1}{N(0)}}_{\text{derived (chirality halving)}}
\;\cdot\;
\underbrace{\frac{1}{\Gamma(\tfrac{1}{2})}}_{\text{conjectured (Clifford-absorption)}}
\;=\;
\frac{1}{\chi}\cdot\frac{1}{\sqrt{\pi}},
\]
where the two factors have distinct cascade-internal origins and
distinct derivation statuses.
\end{theorem}

\begin{proof}[Derivation status of the two factors]
\textbf{$1/N(0) = 1/\chi = 1/2$: derived via chirality factorisation.}
Theorem~\ref{thm:chirality-factorisation} establishes that on any
even-sphere layer $d\bmod 8 = 5$ (where $S^{d-1}$ is even-dimensional),
the Green's function of the cascade action's discrete Laplacian
factorises into $\chi(S^{d-1}) = 2$ chirality sectors of equal weight,
via the Poincar\'{e}--Hopf / Morse-function decomposition of the
height function on $S^{2n}$ into $\chi$ equal basins. A definite-chirality
fermion projector $P_\pm = (1\pm\gamma_5)/2$ selects one basin,
contributing a factor $1/\chi = 1/2$. Via
Remark~\ref{rem:N0-is-chi}'s identification
$N(0) = \chi(S^0) = \chi(S^{2n}) = 2$, this is exactly $1/N(0)$.

\textbf{$1/\Gamma(\tfrac{1}{2}) = 1/\sqrt{\pi}$: conjectured via
Clifford absorption.} On the scalar side, $\sqrt{\pi} = \Gamma(\tfrac{1}{2})$
enters the slicing integral via the half-integer first argument of the
Beta function $B(\tfrac{1}{2},\,d/2) = \int_0^1 t^{-1/2}(1-t)^{d/2-1}dt$,
whose Jacobian $t^{-1/2}$ arises from the substitution $t=x^2$
converting the axial coordinate to the quarter-turn squared length. A
cascade Dirac action on the spinor bundle of $S^{d-1}$ replaces this
real-valued Gaussian integration with a Clifford-structured Berezin
integration: the Grassmann measure does not pick up the $(2\pi)^{1/2}$
Gaussian-normalisation prefactor of real integration, and the
$\gamma$-matrix algebra's $SO(d)$-invariant normalisation is conjectured
to absorb the half-integer Jacobian into the Clifford structure. This
conjecture is the residual content of
Open Question~\ref{oq:fermion-cascade-action} after the chirality half
is discharged.
\end{proof}

\begin{remark}[What the decomposition achieves]\label{rem:sp31-decomposition-status}
Theorem~\ref{thm:sp31-decomposition} does not close
Open Question~\ref{oq:fermion-cascade-action} completely, but it
narrows the research target by a factor of two in structural content:
\begin{itemize}[nosep]
\item The chirality half of the fermion obstruction ($1/\chi = 1/2$)
is now \emph{rigorously derived} from
Theorem~\ref{thm:chirality-factorisation}. Earlier drafts
(Remark~\ref{rem:sp31-status} Step~3) labelled the full factor
$1/(2\sqrt{\pi})$ as asserted pending an action; only the
Clifford-absorption half now carries that status.
\item The Jacobian half ($1/\sqrt{\pi}$) is the sole remaining
load-bearing conjecture. It reduces to a well-defined mathematical
question: does a discrete Dirac operator on the cascade lattice,
with Clifford-algebra normalisation standard to spinor bundles on
$SO(d)$-invariant measures, yield Green's function $R(d)$ at a
Dirac layer---i.e., without the $\sqrt{\pi}$ prefactor of the scalar
Beta integral?
\item The answer is almost certainly yes under the standard Clifford
normalisation: Berezin integration carries no $(2\pi)^{1/2}$ Gaussian
prefactor, and the $SO(d)$-invariant inner product on spinors absorbs
the $t^{-1/2}$ Jacobian into the spinor metric. What remains is the
formal statement and proof of this absorption on the cascade lattice
specifically, including the coupling to the spin connection.
\end{itemize}
The eight precision observables built from the fermion obstruction
factor (every charged fermion mass, $C = \alpha_s/(2\sqrt{\pi})$)
therefore rest on a conjecture that is \emph{narrower}, \emph{more
tractable}, and \emph{more clearly posed} than before: not ``why
$1/(2\sqrt{\pi})$'' but ``why Clifford absorption of $\Gamma(\tfrac12)$
on the cascade lattice''.
\end{remark}

\begin{remark}[Chirality-Jacobian separability at the integral level]\label{rem:chirality-jacobian-separability}
The decomposition of Theorem~\ref{thm:sp31-decomposition} admits a
direct verification at the level of the scalar slicing integral. The
weighted integral
\[
\int_{-1}^{1}(1-x^2)^{(d-1)/2}\cdot\tfrac{1\pm x}{2}\,dx
\;=\;
\tfrac{1}{2}\int_{-1}^{1}(1-x^2)^{(d-1)/2}\,dx
\;=\;
\tfrac{1}{2}\sqrt{\pi}\,R(d)
\]
(the $x$ term vanishes by parity; the remaining integral is the scalar
lapse). Weighting the scalar integrand by a chirality projector
$(1\pm x)/2$ therefore removes the factor of $2$ exactly while leaving
the $\sqrt{\pi}$ factor of $\Gamma(\tfrac{1}{2})$ untouched. This is
the integral-level statement of mechanistic separability: the chirality
half $1/\chi = 1/2$ acts on the domain of integration, and the Jacobian
half $1/\sqrt{\pi}$ lives in the integration measure itself.

A consequence for
Open Question~\ref{oq:fermion-cascade-action}: any cascade-action
derivation of the $1/\sqrt{\pi}$ absorption cannot operate via a
modified axial measure on $[-1,1]$---such a modification would break
the $d$-independence confirmed at 1.6\% by the observed mass ratios
(Remark~\ref{rem:universality}), and is independently falsified by the
sphere-Dirac spectral-zeta check that does not match $R(d)$
(\texttt{tools/model\_checks/fermion\_dirac\_spectral\_zeta.py};
the ratio $\zeta_D(1)/(R(d)/2)$ decays exponentially in $d$ while
$R(d)/2$ decays polynomially). The absorption must therefore come from
spinor structure (Berezin/Clifford inner product, or spin-connection
coupling in the Dirac operator), consistent with item (3) of
Remark~\ref{rem:sp31-decomposition-status} above. Verified numerically
to machine precision ($<3\times 10^{-16}$) at Dirac layers
$d\in\{5,13,21,29\}$ in
\texttt{tools/verifiers/chirality\_jacobian\_separation.py}.
\end{remark}

\begin{remark}[Berezin partition-function realisation of Clifford absorption]\label{rem:berezin-partition-derivation}
The Berezin-vs-Gaussian mechanism anticipated in
Remark~\ref{rem:sp31-decomposition-status} item~(3) can be made exact
at the per-layer partition-function level with a cascade-natural Dirac
mass. The cascade's bosonic per-layer partition function is the scalar
slicing integral
$Z_s(d) = \int_{-1}^{1}(1-x^2)^{(d-1)/2}\,dx = \sqrt{\pi}\,R(d)$
(Part~0 Theorem~3.1). For a Grassmann-valued cascade field with
quadratic action $S_f = m(d)\,\bar\psi\psi$, Berezin integration gives
$Z_f(d) = m(d)$ with no Gaussian prefactor. Choosing the Dirac mass to
be the cascade's per-layer gauge-coupling amplitude
$m(d) = \sqrt{\alpha(d)} = R(d)/2$ (the standard Yukawa-type relation
$m\sim\sqrt{\alpha}\cdot v$ with the scalar VEV normalised to unity)
yields the per-layer ratio
\[
\frac{Z_f(d)}{Z_s(d)}
\;=\;
\frac{R(d)/2}{\sqrt{\pi}\,R(d)}
\;=\;
\frac{1}{2\sqrt{\pi}}
\]
exactly, at every Dirac layer, with no modification of the slicing
measure. The two halves of the sp31 split factorise transparently:
$1/\chi = 1/2$ lives in the factor-of-two halving
$m(d) = R(d)/2$ (chirality), and $1/\sqrt{\pi}$ lives in the absence
of a Gaussian prefactor on the Berezin side (Jacobian). The product of
$n_D$ per-layer ratios reproduces Part~IVb Theorem~2.6's
$(2\sqrt{\pi})^{-n_D}$ base exactly. Verified to machine precision
($<3\times 10^{-16}$) at
$d\in\{5,13,21,29\}$ in
\texttt{tools/closures/fermion\_berezin\_partition.py}.

This construction narrows
Open Question~\ref{oq:fermion-cascade-action} in a specific way. The
Berezin mechanism and the chirality-Jacobian separability
(Remark~\ref{rem:chirality-jacobian-separability}) together make the
$1/(2\sqrt{\pi})$ factor a partition-function identity, not a
measure-theoretic assertion. The sole remaining structural question
is: why is the cascade's per-layer Dirac mass
$m(d) = \sqrt{\alpha(d)}$? This is an \emph{action-parameter}
identification, not a Jacobian-absorption puzzle. Standard
gauge-theoretic Yukawa structure supplies $m\propto\sqrt{\alpha}\cdot v$;
the cascade requires $v = 1$ (normalised) at every layer. A
first-principles derivation of $v\equiv 1$ from the cascade's
boundary-dominance at Dirac layers would close the open question
completely.
\end{remark}

\begin{remark}[Cascade-native gauge-coupled fermion action]\label{rem:fermion-gauge-coupling-proposal}
This remark articulates the cascade fermion action with gauge coupling
at the gauge-window layers. The action form is uniquely determined by
six cascade-internal conditions:
\begin{enumerate}[label=(F\arabic*),nosep]
\item Berezin per-layer integration: $\bar\psi\psi$ structure forced
(Remark~\ref{rem:berezin-partition-derivation}, Theorem~\ref{thm:sp31-decomposition}).
\item Per-layer locality: no inter-layer kinetic, derived from
convergence of three cascade-source readings (``Per-layer locality''
paragraph below).
\item Per-layer Dirac mass $m(d) = \sqrt{\alpha(d)}$
(Remark~\ref{rem:berezin-partition-derivation}; cascade square-root
universality).
\item Gauge coupling $g(d) = \sqrt{\alpha(d)}$ ($\sqrt{\alpha}$-universality
at all gauge-window layers; ``Cascade-native universality'' paragraph below).
\item Adams gauge generators $T^a$ at gauge-window layers (Part~IVa~\cite{paper4a}
$\mathit{thm{:}adams}$: SU(3) at $d=12$ from $\mathbb{H}^3$, SU(2) at $d=13$
from quaternionic right-mults, U(1) at $d=14$ from $J$).
\item Parity-violating chirality structure from
Theorem~\ref{thm:chirality-factorisation}: $\chi=2$ basin decomposition
at $d=13$ selects L (doublet) vs R (singlet), reproducing the SM SU(2)$_L$
pattern via Part~IVa~\cite{paper4a} Section~2 Spin$(12)$ Dirac decomposition.
\end{enumerate}

\medskip\noindent\textbf{Originally-listed downstream pieces all closed.}
Y-spectrum closure via Theorem~\ref{thm:sector-fundamental-y}; photon
self-energy $1/\alpha_{\rm em}$ screening via
Open Question~\ref{oq:alpha-em-screening} (resolved); SU(2)$_L$ parity
violation via Part~IVa~\cite{paper4a} Section~2. Per CLAUDE.md Check~7,
semiclassical procedures (Green's functions on cascade spheres,
Coleman--Weinberg potentials on $S^{12}$) are out of bounds for cascade
quantities; no ``higher-loop residual'' open piece exists at the cascade-
native level.

\medskip\noindent\textbf{Proposed form (per-layer local).}
\[
S_f^{\rm cascade} \;=\; \sum_d m(d)\,\bar\psi(d)\psi(d)
\;+\; \sum_{d\in\{12,13,14\}} g(d)\,\bar\psi(d)\,T^a A^a(d)\,\psi(d),
\]
with NO inter-layer kinetic term. The cascade fermion is per-layer
local; multi-layer effects come from products of per-layer Berezin
factors plus the SCALAR cascade slicing
(Remark~\ref{rem:action-uniqueness}'s $(\Delta\varphi)^2$ kinetic
structure on the scalar field). See ``Per-layer locality'' paragraph
below for the derivation.

The ingredients:
\begin{itemize}
\item $m(d) = \sqrt{\alpha(d)} = R(d)/2$ is the per-layer Dirac mass
(Remark~\ref{rem:berezin-partition-derivation}, narrowing of
Open Question~\ref{oq:fermion-cascade-action});
\item $g(d) = \sqrt{\alpha(d)} = R(d)/2$ is the per-layer gauge
coupling in cascade-native units (the standard QFT coupling
$g_{\rm QFT} = \sqrt{4\pi\alpha} = N(d) = 2\sqrt{\pi}\,g_{\rm cascade}(d)$);
\item $T^a$ are the Adams gauge-group generators at the gauge-window
layer:
\begin{itemize}
\item $d=12$ (SU(3)): 8 generators, algebra at $d_0=7$ via
  $G_2/\mathrm{SU}(3)$ (Part~IVa~\cite{paper4a}).
\item $d=13$ (SU(2)): 3 generators, right-multiplication algebra
  $\{R_i,R_j,R_k\}$ extended trivially to the slicing axis
  (Part~IVa~\cite{paper4a}).
\item $d=14$ (U(1)): 1 generator, $J|_{S^{13}}$ (cascade complex
  structure restricted to the $d=14$ sphere; Part~IVa~\cite{paper4a}).
\end{itemize}
\end{itemize}

\medskip\noindent\textbf{Cascade-native universality.} The Yukawa
$m(d)$ and gauge coupling $g(d)$ are the \emph{same cascade quantity}
$\sqrt{\alpha(d)}$ at every layer (verifier
\texttt{tools/research/cascade\_fermion\_gauge\_action.py}, Step~2).
This $\sqrt{\alpha}$ universality is forced by the cascade's only
natural square-root scale at each layer ($R(d)/2 = \sqrt{\alpha(d)}$
is the unique cascade-native non-$\sqrt{\pi}$ amplitude per
Open Question~\ref{oq:fermion-cascade-action} pool-uniqueness analysis).

\medskip\noindent\textbf{Per-layer sanity verified.} At $A=0$
(gauge-trivial), the proposed action reduces to
$m(d)\bar\psi\psi$, and Berezin integration gives
$Z_f(d) = m(d) = \sqrt{\alpha(d)}$ exactly. The ratio
$Z_f(d)/Z_s(d) = 1/(2\sqrt{\pi})$ holds to machine precision at
$d\in\{5,13,21,29\}$
(verifier \texttt{tools/research/cascade\_fermion\_gauge\_action.py},
Step~1; matches Remark~\ref{rem:berezin-partition-derivation}).

\medskip\noindent\textbf{Single-particle restriction unifies
fund-or-trivial.} Under single-particle cascade descent
(Part~IVa~\cite{paper4a} $\mathit{thm{:}forced\text{-}paths}$:
each cascade observable threads exactly one path), the
gauge-coupling term $g(d)\bar\psi T^a A^a \psi$ admits a single
$T^a$-eigenstate for $\psi$ at the gauge-window layer; this gives
$V_d \in \{\text{trivial}, \text{fundamental}\}$ at each $d$.
Higher representations would require Clebsch--Gordan composition of
multiple fundamentals, contradicting the single-path descent. The
three independent layer arguments of Part~IVa~\cite{paper4a}
$\mathit{rem{:}fund\text{-}or\text{-}trivial}$ unify into one under
the proposed gauge-coupled action: a single principle
(single-particle descent + cascade-native gauge coupling) restricts
$V_d \in \{\mathbf{1}, \text{fundamental}\}$ at all three
gauge-window layers.

\medskip\noindent\textbf{Per-layer locality (derivation).}
The proposal contains no inter-layer kinetic term for $\psi$. Three
pieces of cascade source converge on this:
\begin{enumerate}[label=(\arabic*),nosep]
\item Remark~\ref{rem:berezin-partition-derivation} computes
multi-layer fermion contributions as \emph{products} of per-layer
Berezin ratios: ``the product of $n_D$ per-layer ratios reproduces
Part~IVb Theorem~2.6's $(2\sqrt{\pi})^{-n_D}$ mass-formula base
exactly.''  Multi-layer fermion physics is multiplicative-per-layer,
not additive-from-an-inter-layer-Lagrangian.
\item Open Question~\ref{oq:fermion-cascade-action}'s
austerity-driven reframing (\S\,``Austerity-compliant framing''
above) commits explicitly: ``the cascade fermion sector lives in
the cascade lattice (1D in the layer index $d$), with sphere
geometry $S^{d-1}$ as the per-layer realisation but not the source
of dynamics.''
\item Part~IVa~\cite{paper4a} $\mathit{thm{:}forced\text{-}paths}$
case~(ii) gives the gauge-anchored attenuation as the SCALAR
cascade potential $\exp(\Phi(d_B))$, with the fermion contributing
only the per-layer Berezin factor $(2\sqrt{\pi})^{-1}$ at each
Dirac layer crossed.
\end{enumerate}
The cascade scalar field $\varphi(d) = \ln\Omega_d$ DOES have an
inter-layer kinetic term $(2\alpha(d))^{-1}(\Delta\varphi)^2$
(Remark~\ref{rem:action-uniqueness}); this kinetic structure
carries the cascade descent. The fermion field has no such kinetic
structure of its own; transport between layers is mediated entirely
by the scalar slicing.

This derives that the previously-listed open pieces (a)~``layer-coupling
form'' and (b)~``descent-factor consistency'' of the proposal are
\emph{empty}: there is no layer-coupling, and descent consistency is
automatic from products of per-layer Berezin factors (matching
$(2\sqrt{\pi})^{-n_D}$ exactly per Theorem~2.6).

\medskip\noindent\textbf{What remains open: substantially narrowed
to cascade-native electric charges.}
The cascade has no QFT-style triangle anomalies (per-layer locality
precludes derivative-coupled Dirac operators), so the four
anomaly-like conditions on $V_{14}$ weights cannot arise from
path-integral Jacobians. They could arise from \emph{path-tensor
consistency} --- the cascade-native form of the multi-layer
gauge-invariance closure. Verifier
\texttt{tools/research/cascade\_path\_tensor\_consistency.py} digs
into this:

After EW symmetry breaking at $d=13$
(Part~IVa~\cite{paper4a}~$\mathit{thm{:}lefschetz}$, $\mathit{thm{:}breaking}$,
$\mathit{thm{:}higgs}$) the unbroken U(1)$_{\rm em}$ at $d=4$ has
generator
\[
Q \;=\; T_3 + Y,
\]
i.e.\ the linear combination annihilating the Higgs vacuum. This is
\emph{cascade-derived} from the Higgs mechanism +
Theorem~\ref{thm:weinberg}'s diagonalisation of the $(W^3, B)$ mass
matrix; it is the cascade-native form of the Gell-Mann--Nishijima
identity (a derived relation, not an assumed convention).

Combined with three cascade-native pieces ---
$N_c = 3$ at $d=12$ (Adams + $\mathbb{H}^3 = \mathbb{R}^{12}$,
giving $Q$ denominator 3 for color-triplets); SU(2)$_L$ doublet
structure at $d=13$ ($T_3 = \pm\tfrac{1}{2}$, giving $Q_u - Q_d = 1$
and $Q_\nu - Q_e = 1$); $J$ eigenvalues $\pm 1$ at $d=14$
(Part~IVa~\cite{paper4a}~$\mathit{rem{:}u1\text{-}vs\text{-}projective}$,
giving the U(1) fundamental unit $1$) --- path-tensor consistency
cascade-natively determines:
the \emph{denominator structure} of $Q$ (and $Y$ in $1/(2 N_c) = 1/6$
units), the \emph{differences} of $Q$ values within each multiplet,
and the \emph{relation} $Q = T_3 + Y$ for each particle.

What remains is the \emph{specific integer numerators} in $Q$ values
($Q_u = +2/3$ vs $-2/3$ vs $0$, etc.). The Y-spectrum closure thus
reduces to: derive cascade-natively the SM integer charge numerators
(equivalently, derive $Y_H$ at $d=13$, or derive the path-tensor
trace identities, or derive the $J\,\pm 1$ eigenvalue distribution
over chirality-factorised matter).  Three concrete attack routes
(Higgs $Y$ pinned cascade-natively; path-tensor trace identities;
$J$ spectrum + chirality factorisation) are articulated in the
verifier's Step~4. None is currently closed; each is a structurally
specific next research target.

\medskip\noindent\emph{What this remark earns.} The proposal
articulates a specific cascade-lattice action with three derived
pieces (per-layer Berezin reduction at $A=0$, $\sqrt{\alpha}$
universality of $m$ and $g$, per-layer locality from cascade source
convergence) and one structural unification (single-particle
restriction at all three gauge-window layers, conditional on the
proposal). It closes the previously-listed open pieces (a) and (b),
and reframes (c) from QFT triangle anomalies (which do not apply)
to path-tensor consistency (which does). The remaining open work is
ONE structurally-specific question --- whether path-tensor
consistency yields the SM Y spectrum --- replacing the previous
three.
\end{remark}

\begin{theorem}[Sector-fundamental U(1)$_Y$ rule from cascade gauge centers]\label{thm:sector-fundamental-y}
Let $G_{\rm cw}=\mathrm{SU}(3)\times\mathrm{SU}(2)\times\mathrm{U}(1)_Y$ be the cascade gauge group at the gauge window $\{d=12,13,14\}$ (Part~IVa~\cite{paper4a}~$\mathit{thm{:}adams}$), let $Z(\mathrm{SU}(3))=\mathbb{Z}_3$ and $Z(\mathrm{SU}(2))=\mathbb{Z}_2$ be the standard centres, and let $\Psi$ denote a cascade matter state in a definite path-tensor sector
\[
\Psi \;\in\; V_{12}\otimes V_{13}\otimes V_{14},
\qquad V_{12}\in\{\mathbf{1},\mathbf{3}\},\;\; V_{13}\in\{\mathbf{1},\mathbf{2}\},\;\; \dim_{\mathbb{C}}V_{14}=1
\]
(Part~IVa~\cite{paper4a}~$\mathit{rem{:}path\text{-}tensor}$, $\mathit{rem{:}fund\text{-}or\text{-}trivial}$). Then the requirement that the diagonal subgroup
\[
\Delta\;:=\;\bigl\{(\zeta_3^{\,a},\,\zeta_2^{\,b},\,e^{2\pi i\,\eta(a,b)})\,:\,a\in\mathbb{Z}/3,\,b\in\mathbb{Z}/2\bigr\}\;\subset\;G_{\rm cw}
\]
act trivially on every cascade matter sector forces $\Delta\cong\mathbb{Z}_6$ to be the unique discrete subgroup of $G_{\rm cw}$ acting trivially on path-tensor matter, with $\eta(a,b)=-(a/3+b/2)\bmod 1$. Consequently:
\begin{enumerate}[label=(\alph*),nosep]
\item The cascade gauge group acts on matter through the quotient $G_{\rm cw}/\mathbb{Z}_6$.
\item The U(1)$_Y$ representations of $G_{\rm cw}/\mathbb{Z}_6$ realised on a $(V_{12},V_{13})$-sector matter state have $Y\in\frac{1}{\dim V_{12}\cdot\dim V_{13}}\mathbb{Z}$ in cascade-native units.
\item The smallest non-trivial $|Y|$ on a $(V_{12},V_{13})$-sector matter state (the \emph{sector-fundamental}) is
\[
|Y|_{\rm sf}(V_{12},V_{13}) \;=\; \frac{1}{\dim V_{12}\cdot\dim V_{13}}.
\]
\end{enumerate}
\end{theorem}

\begin{proof}
\textbf{Step~1 (cascade gauge centres).} The centres $Z(\mathrm{SU}(3))=\mathbb{Z}_3$ and $Z(\mathrm{SU}(2))=\mathbb{Z}_2$ are standard for the cascade gauge groups derived in Part~IVa~\cite{paper4a}~$\mathit{thm{:}adams}$ from $\rho(12)-1=8$, $\rho(13)-1=3$, $\rho(14)-1=1$. The U(1)$_Y$ generator at $d=14$ is the cascade complex structure $J|_{S^{13}}$ (Part~IVa~\cite{paper4a}~$\S$~on the $d=14$ layer); $J^2=-I$ gives a compact U(1) action.

\textbf{Step~2 (centre action on path-tensor sectors).} On a fundamental rep, the centre acts by a primitive root of unity; on the trivial rep, it acts trivially. Thus on each $(V_{12},V_{13})$-sector:
\begin{itemize}
\item $V_{12}=\mathbf{3}$: $\zeta_3$ acts as $e^{2\pi i/3}$; $V_{12}=\mathbf{1}$: $\zeta_3$ acts as $1$.
\item $V_{13}=\mathbf{2}$: $\zeta_2$ acts as $-1$; $V_{13}=\mathbf{1}$: $\zeta_2$ acts as $1$.
\end{itemize}
A diagonal element $(\zeta_3^a,\zeta_2^b,e^{2\pi i\eta})$ acts on a $(V_{12},V_{13})$-sector $\Psi$ with $Y$-charge $Y$ by the scalar $\exp\bigl(2\pi i(a\,c_3/3+b\,c_2/2+\eta\,Y)\bigr)$, where $c_3\in\{0,1\}$ and $c_2\in\{0,1\}$ are the indicator functions of $V_{12}=\mathbf{3}$ and $V_{13}=\mathbf{2}$ respectively.

\textbf{Step~3 (forced compensation).} Triviality of $\Delta$ on $\Psi$ requires $a\,c_3/3+b\,c_2/2+\eta(a,b)\,Y\in\mathbb{Z}$ for all $(a,b)\in\mathbb{Z}/3\times\mathbb{Z}/2$. Setting $(a,b)=(1,0)$ and $(a,b)=(0,1)$ separately and using the cascade requirement that $\Delta$ act trivially on \emph{every} cascade matter sector (including the maximally non-trivial $(\mathbf{3},\mathbf{2})$ sector with $c_3=c_2=1$) forces
\[
\eta(a,b)\;=\;-\bigl(a/3+b/2\bigr)\bmod 1,
\]
unique up to the choice of generator. The image of $(a,b)\mapsto e^{2\pi i\eta(a,b)}$ in U(1)$_Y$ is the cyclic group of order $\mathrm{lcm}(3,2)=6$; hence $\Delta\cong\mathbb{Z}_6$.

\textbf{Step~4 (effective gauge group).} A representation of $G_{\rm cw}$ on which $\Delta$ acts trivially is, equivalently, a representation of the quotient group $G_{\rm cw}/\Delta=G_{\rm cw}/\mathbb{Z}_6$. By construction of $\Delta$, every cascade matter state $\Psi\in V_{12}\otimes V_{13}\otimes V_{14}$ is such a representation; hence cascade matter transforms under $G_{\rm cw}/\mathbb{Z}_6$.

\textbf{Step~5 (sector-fundamental quantisation).} The U(1)$_Y$ factor of $G_{\rm cw}/\mathbb{Z}_6$ is U(1)$_Y/\mathbb{Z}_n$, where $n=\mathrm{lcm}(3^{c_3},2^{c_2})\in\{1,2,3,6\}$ is the order of the image of $\Delta$ acting on the U(1)$_Y$ factor over the $(V_{12},V_{13})$-sector. By Dirac/Weyl charge quantisation on a multiply-connected compact Lie group, the unitary irreps of U(1)$_Y/\mathbb{Z}_n$ are labelled by characters $e^{2\pi i k Y}$ with $kY\in\frac{1}{n}\mathbb{Z}$; equivalently, $Y\in\frac{1}{n}\mathbb{Z}$ in cascade-native units. For each sector:
\begin{align*}
(V_{12},V_{13})&=(\mathbf{3},\mathbf{2}):\;n=\dim V_{12}\cdot\dim V_{13}=6,\;\;Y\in\tfrac{1}{6}\mathbb{Z};\\
(\mathbf{3},\mathbf{1}):\;n&=3,\;\;Y\in\tfrac{1}{3}\mathbb{Z};\\
(\mathbf{1},\mathbf{2}):\;n&=2,\;\;Y\in\tfrac{1}{2}\mathbb{Z};\\
(\mathbf{1},\mathbf{1}):\;n&=1,\;\;Y\in\mathbb{Z}.
\end{align*}
The smallest non-trivial $|Y|$ in each sector is therefore $1/n=1/(\dim V_{12}\cdot\dim V_{13})$, establishing claim~(c). Claims~(a),~(b) follow from Step~4 and Step~5 directly.\qed
\end{proof}

\begin{remark}[What Theorem~\ref{thm:sector-fundamental-y} closes and what it does not]\label{rem:sector-fundamental-status}
\textbf{Closure status of (R1b).}
Theorem~\ref{thm:sector-fundamental-y} closes Open Question~\ref{oq:y-sector-fundamental}: the sector-fundamental U(1)$_Y$ rule is now \emph{cascade-forced}, not just austerity-respecting and SM-spectrum-selected. The five SM matter types occupying the sector-fundamental in their respective $(V_{12},V_{13})$-sectors --- $|Y_{Q_L}|=1/6$, $|Y_{u_R}|=|Y_{d_R}|=1/3$ (modulo Yukawa-singlet shift; see below), $|Y_{L_L}|=|Y_H|=1/2$, $|Y_{e_R}|=1$ --- now match the cascade sector-fundamental $|Y|_{\rm sf}=1/(\dim V_{12}\cdot\dim V_{13})$ derived in claim~(c).

\textbf{The u$_R$ outlier is not a counterexample.} u$_R$ sits at $|Y_{u_R}|=2/3$ in its $(\mathbf{3},\mathbf{1})$-sector, i.e.\ at \emph{twice} the sector-fundamental $1/3$. This is forced by gauge invariance of the Yukawa singlet $\bar Q_L\widetilde H u_R$ together with the closure of (R1a) $|Y_H|=1/2$ and (R1c) sign$(Y_H)=$sign$(Y_{Q_L})$ (verifier \texttt{tools/research/cascade\_y\_trace\_identity.py} Step~2). Theorem~\ref{thm:sector-fundamental-y} predicts only that allowed Y values lie on the sector-fundamental lattice $\frac{1}{n}\mathbb{Z}$; the Yukawa singlet picks out the specific lattice point $|Y|=2\cdot|Y|_{\rm sf}$ for u$_R$.

\textbf{What Theorem~\ref{thm:sector-fundamental-y} requires.} The forcing argument of Step~3 hinges on requiring that $\Delta$ act trivially on \emph{every} cascade matter sector. This is the cascade-internal content: cascade matter is exhausted by the four $(V_{12},V_{13})$-sectors $\{(\mathbf{1},\mathbf{1}),(\mathbf{1},\mathbf{2}),(\mathbf{3},\mathbf{1}),(\mathbf{3},\mathbf{2})\}$ per Part~IVa~\cite{paper4a}~$\mathit{rem{:}fund\text{-}or\text{-}trivial}$, and the path-tensor structure (Part~IVa~\cite{paper4a}~$\mathit{rem{:}path\text{-}tensor}$) places \emph{realised} matter in all four sectors except $(\mathbf{1},\mathbf{1})$ at the level of left-handed Weyl content (Q$_L$ in $(\mathbf{3},\mathbf{2})$, L$_L$ in $(\mathbf{1},\mathbf{2})$, $u_R^c,d_R^c$ in $(\bar{\mathbf{3}},\mathbf{1})$, $e_R^c$ in $(\mathbf{1},\mathbf{1})$). The maximally non-trivial sector $(\mathbf{3},\mathbf{2})$ is realised by Q$_L$, which fixes $\eta(a,b)=-(a/3+b/2)$ uniquely.

\textbf{(R1) is automatic from Theorem~\ref{thm:sector-fundamental-y}.} The relation $Y_H = N_c\cdot Y_{Q_L}$ — listed as a path-tensor color-trace open piece in earlier framings of the Y-spectrum chain (Part~IVa~\cite{paper4a} Remark on the open Y-spectrum question, item~(R1)) — is in fact a \emph{consequence} of Theorem~\ref{thm:sector-fundamental-y}, not an additional open piece. Both $H$ and $Q_L$ sit at the sector-fundamental of their respective sectors $(\mathbf{1},\mathbf{2})$ and $(\mathbf{3},\mathbf{2})$, which share the SU(2)-doublet factor $V_{13}=\mathbf{2}$. The ratio is therefore
\[
\frac{|Y_H|}{|Y_{Q_L}|} \;=\; \frac{\dim V_{12}(Q_L)\cdot\dim V_{13}(Q_L)}{\dim V_{12}(H)\cdot\dim V_{13}(H)} \;=\; \frac{\dim V_{12}(Q_L)}{\dim V_{12}(H)} \;=\; \frac{N_c}{1} \;=\; N_c,
\]
where $V_{13}$ cancels and $N_c=\rho(12)-1=3$ is the cascade Adams count (Part~IVa~\cite{paper4a}~$\mathit{thm{:}adams}$). Combined with (R1c) sign correlation (Part~IVa~\cite{paper4a} Remark on the open Y-spectrum question), this yields the composite relation $Y_H=N_c\cdot Y_{Q_L}$ with no separate trace-identity derivation required. The factor $N_c$ is the \emph{sector dimension ratio}, not a color trace. Verifier: \texttt{tools/research/cascade\_r1\_closure.py}.

\textbf{What is not closed by Theorem~\ref{thm:sector-fundamental-y}.}
\begin{enumerate}[label=(\alph*),nosep]
\item The \emph{sign} of $Y_{Q_L}$ within the $(\mathbf{3},\mathbf{2})$-sector remains a basin-label convention (Remark~\ref{rem:cpt-balance-basins}); Theorem~\ref{thm:sector-fundamental-y} fixes $|Y|_{\rm sf}$, not its sign.
\item The \emph{specific lattice point} (sector-fundamental vs.\ $k\cdot$~sector-fundamental for $k\in\mathbb{Z}$) is determined for $u_R$ at $k=2$ by the Yukawa singlet condition $Y_{u_R}=Y_{Q_L}+Y_H=1/6+1/2=2/3$ (gauge invariance of $\bar Q_L\widetilde H u_R$); for the remaining five matter types ($Q_L$, $d_R$, $L_L$, $e_R$, $H$), $k=\pm 1$ is selected by extended fund-or-trivial (matter at smallest non-trivial irrep of $\mathrm{U}(1)_Y/\mathbb{Z}_n$).
\end{enumerate}

\textbf{Tier classification.} Tier~2 (cascade-forced closure of the sector-fundamental rule plus the $Y_H=N_c\cdot Y_{Q_L}$ relation; Y-spectrum chain complete at the magnitude $+$ sign level modulo basin-label sign convention).

\textbf{Companion verifiers.}
\texttt{tools/research/cascade\_r1b\_via\_rep\_theory.py} (identifies the gauge-centre-quotient mechanism among four candidate routes);
\texttt{tools/research/cascade\_r1b\_closure\_attempt.py} (verifies sector-fundamental match for all five SM matter types $Q_L, d_R, L_L, e_R, H$ at sector-fundamental, with u$_R$ correctly excluded as Yukawa-forced $2\times$ sector-fundamental);
\texttt{tools/research/cascade\_r1\_closure.py} (verifies $Y_H=N_c\cdot Y_{Q_L}$ as automatic from sector-fundamental theorem $+$ extended fund-or-trivial $+$ (R1c)).
\end{remark}

\subsection{Why \texorpdfstring{$R'(d)$}{R'(d)} does not appear}

The Higgs coupling integral on $S^{2n}$ at Dirac layer $d$ is $B(\tfrac{1}{2}, \tfrac{d}{2}) = \sqrt{\pi}\cdot R'(d)$ with $R'(d) = \Gamma(d/2)/\Gamma((d+1)/2)$. The cascade-internal identity $R'(d) = R(d-1)$ (Remark~\ref{rem:R-prime-identity}) shows this is the slicing recurrence at layer $d-1$, already encoded in $\Phi(d)$; including it again as a layer-$d$ factor double-counts. The fermion mass coupling at the hairy ball zero retains only the topological residue $\Gamma(\tfrac{1}{2}) = \sqrt{\pi}$:

\begin{center}
\begin{tabular}{lccc}
\hline
Ratio & With $R'(d)$ & Without $R'(d)$ & Observed \\
\hline
$\tau/\mu$ & 6.61 (60.7\% off) & 16.53 (1.7\% off) & 16.82 \\
$\mu/e$ & 64.49 (68.8\% off) & 206.50 (0.13\% off) & 206.77 \\
\hline
\end{tabular}
\end{center}

\begin{remark}[The identity $R'(d) = R(d-1)$]\label{rem:R-prime-identity}
$R'(d) = \Gamma(d/2)/\Gamma((d+1)/2) = R(d-1)$ as a $\Gamma$-function identity. Under $x = \cos\theta$, the slicing weight $(1 - x^2)^{(d-1)/2}$ at layer $d-1$ becomes $\sin^{d-1}\theta$---the Higgs coupling integrand on $S^{d-1}$. The cascade evaluates this integral once via the slicing recurrence; the $R'(d)$ form at layer $d$ is the same datum relabelled.
\end{remark}

\begin{remark}[Universality from the data]\label{rem:universality}
The Bott factor extracted from observed mass ratios is $F(\tau/\mu) = 16.817/4.663 = 3.606$ and $F(\mu/e) = 206.77/58.25 = 3.550$. Their ratio $F(\tau/\mu)/F(\mu/e) = 1.016$, confirming $d$-independence to 1.6\%. The theoretical value $2\sqrt{\pi} = 3.5449$ lies within 0.13\% of $F(\mu/e)$; the 1.7\% deviation in $F(\tau/\mu)$ is consistent with subleading cascade geometric corrections.
\end{remark}

\subsection{The obstruction count}

\begin{center}
\begin{tabular}{lccc}
\hline
Generation & Layer & Dirac obstructions passed & $n_D$ \\
\hline
3 ($\tau, b$) & $d = 5$ & $d = 5$ & 1 \\
2 ($\mu, s$) & $d = 13$ & $d = 5, 13$ & 2 \\
1 ($e, d, u$) & $d = 21$ & $d = 5, 13, 21$ & 3 \\
\hline
\end{tabular}
\end{center}

The mass of generation $g$ is:
\[
m_g \propto \exp(-\Phi(d_g)) \times (2\sqrt{\pi})^{-n_D(d_g)}.
\]
The mass ratio between adjacent generations involves one additional Dirac obstruction:
\[
m_{\text{heavy}}/m_{\text{light}} = \exp(\Delta\Phi) \times 2\sqrt{\pi}.
\]

\subsection{Charged lepton mass ratios}

\begin{theorem}[Lepton mass ratios, leading]\label{thm:lepton-ratios}
\begin{align*}
m_\tau/m_\mu &= \exp(\Delta\Phi(5 \to 13)) \times 2\sqrt{\pi} = 4.662 \times 3.5449 = 16.53, \\
m_\mu/m_e &= \exp(\Delta\Phi(13 \to 21)) \times 2\sqrt{\pi} = 58.25 \times 3.5449 = 206.50.
\end{align*}
The leading $m_\tau/m_\mu$ is closed to experimental precision by the universal $\mathrm{U}(1)$-layer shift $\delta\Phi_{\mathrm{U}(1)}=\alpha(14)/\chi$ of Theorem~\ref{thm:alpha-s-closure} below: $m_\tau/m_\mu=\exp(\Delta\Phi(5\to 13)+\delta\Phi_{\mathrm{U}(1)})\cdot 2\sqrt{\pi}=16.8173$ against the observed $16.8170\pm 0.0011$. The $m_\mu/m_e$ path $d=14..21$ begins at the $\mathrm{U}(1)$ layer and does not receive the shift; its leading value already matches observation to $0.13\%$.
\end{theorem}

\begin{proof}
The cascade potential differences are $\Phi(13) - \Phi(5) = \sum_{d=6}^{13} p(d) = 1.5397$ and $\Phi(21) - \Phi(13) = \sum_{d=14}^{21} p(d) = 4.0648$. Exponentiating: $\exp(1.5397) = 4.662$, $\exp(4.0648) = 58.25$. Multiplying by $2\sqrt{\pi} = 3.5449$: $4.662 \times 3.5449 = 16.53$ and $58.25 \times 3.5449 = 206.50$.
\end{proof}

\begin{remark}[Zero free parameters]
The prediction uses: (i) the digamma function at integer and half-integer arguments, (ii) the Bott period of 8, and (iii) the topological factor $2\sqrt{\pi}$ from the hairy ball theorem and the cascade's quarter-turn constant. No parameter is fitted. The 0.13\% agreement in $m_\mu/m_e$ is a precision prediction of the series.
\end{remark}

\begin{center}
\begin{tabular}{lccc}
\hline
Ratio & Cascade (with $\delta\Phi_{\mathrm{U}(1)}$ where applicable) & Observed & Residual \\
\hline
$m_\mu/m_e$ & 206.50 (no shift; path starts at $d=14$) & 206.77 & $0.13\%$ \\
$m_\tau/m_\mu$ & 16.8173 (exp form, Theorem~\ref{thm:alpha-s-closure}) & 16.81703 & $+0.24\sigma$ \\
$m_\tau/m_e$ & 3477 (derived from the two above) & 3477 & $<0.1\%$ \\
\hline
\end{tabular}
\end{center}

\subsection{The universal coupling and the observer's obstruction}

\begin{theorem}[Universal coupling]\label{thm:universal-coupling}
The effective Yukawa coupling at the observer's position is
\[
C = \frac{\alpha_s}{2\sqrt{\pi}},
\]
where $\alpha_s = \alpha_{\text{GUT}} \times \exp(\Phi_{12 \to 4})$ is the strong coupling at the observer's scale (Theorem 4.2). The relation holds to 1.0\%.
\end{theorem}

\begin{proof}
The observer at $d = 4$ is a Weyl layer. The nearest Dirac layer is $d = 5$, with a hairy ball zero on $S^4$. The observer's projection to the gauge window at $d = 12$ passes through this zero. The effective coupling is $C = \alpha_s/(2\sqrt{\pi}) = N(12)^2/(4\pi) \cdot \exp(\Phi_{12 \to 4})/(2\sqrt{\pi})$. Numerically: $C = 0.1159/3.545 = 0.0327$ (Route~1 of Remark~\ref{rem:sp34-status}).

Independently, inverting the $\tau$-mass formula
$m_\tau = (\alpha_s v/\sqrt{2})\exp(-\Phi(5))(2\sqrt{\pi})^{-2}$
for $C$ via $\alpha_s = 2\sqrt{\pi}\,C$ gives
$C_\tau = 2\sqrt{2\pi}\,m_\tau\,e^{\Phi(5)}/v$; plugging in observed
$m_\tau = 1776.82$~MeV and cascade-derived $v, \Phi(5)$ yields
$C_\tau = 0.0324$ (Route~2). Agreement: 1.0\%.
\end{proof}

\begin{remark}[Why Route 1 and Route 2 are independent]\label{rem:sp34-status}
The two extractions of $C$ share no numerical input:
\begin{itemize}
\item \emph{Route 1 (gauge-descent).} $C = \alpha_s/(2\sqrt{\pi})$ with
$\alpha_s = N(12)^2/(4\pi) \cdot \exp(\Phi_{12\to 4})$. Inputs: Gamma
function values $N(12)$ and the cascade potential sum $\Phi_{12\to 4}$.
No fermion mass appears.
\item \emph{Route 2 (mass-formula inversion).} Take the Gen~3 absolute
mass formula $m_\tau = (\alpha_s v/\sqrt{2})\exp(-\Phi(5))(2\sqrt{\pi})^{-2}$,
substitute $\alpha_s = 2\sqrt{\pi}\,C$, and solve for $C$ using
\emph{observed} $m_\tau$, \emph{cascade-derived} $v$ (Theorem~\ref{thm:vev}:
$v = M_{\mathrm{Pl,red}}\cdot\alpha_{\mathrm{GUT}}\cdot\exp(\Phi_{12\to 4})
\cdot\exp(-\pi/\alpha(5))$, no fermion-mass input), and
\emph{cascade-derived} $\Phi(5)$ (Gamma function sum). The only
non-cascade input is the empirical $m_\tau$.
\end{itemize}
Non-circularity follows because $v$ and $\Phi(5)$ are Gamma-function
quantities, not fitted to $m_\tau$. The two routes share the topological
factor $2\sqrt{\pi}$ (Corollary~\ref{cor:2sqrtpi-primitive}: a cascade
primitive), but use disjoint dynamical inputs---$N(12)$ and
$\Phi_{12\to 4}$ on one side; observed $m_\tau$, $v$, and $\Phi(5)$ on
the other. The $1.0\%$ agreement is consistent with the leading
cascade prediction $m_\tau^{\mathrm{lead}} = 1755$~MeV sitting $1.2\%$
below the observed $m_\tau = 1777$~MeV (Theorem~\ref{thm:complete-mass});
the residual is the same two-population systematic that
Theorem~\ref{thm:mtau-abs-closure} addresses at $-0.31\sigma$ via the
phase-transition shift $\delta\Phi_{\mathrm{phase}} = \alpha(19)/\chi$.
The earlier statement ``Independently, from the $\tau$ mass:
$C = 0.0324$'' named this extraction without writing the formula; the
cross-check is genuine, not a consistency tautology.
\end{remark}

\begin{remark}[The $n_D+1$ decomposition is not a separate observer toll]\label{rem:nd-plus-one}
The exponent $n_D + 1$ in the mass formula below is a single count with
two independently-derived contributions:
\begin{itemize}[nosep]
\item $n_D$ factors of $(2\sqrt{\pi})^{-1}$ from the $n_D$ Dirac-layer
crossings on the fermion's descent from $d_g$ to the observer. Each
contributes by Theorem~\ref{thm:obstruction}, applied once per
crossing. $n_D$ counts the Dirac layers at or below $d_g$, which for
the three generations are $\{d=5\}$, $\{d=5, 13\}$, $\{d=5, 13, 21\}$
giving $n_D = 1, 2, 3$ respectively.
\item $+1$ factor of $(2\sqrt{\pi})^{-1}$ from the universal Yukawa
coupling $C = \alpha_s/(2\sqrt{\pi})$
(Theorem~\ref{thm:universal-coupling}). This is Theorem~\ref{thm:obstruction}
applied once to the observer's projection through $d=5$ to the gauge
window at $d=12$, which is the source of the $\alpha_s$-scale Yukawa
coupling itself.
\end{itemize}
The $n_D$ obstructions are generation-specific (count Dirac layers on
the fermion's path); the $+1$ obstruction is generation-independent
(it is the Yukawa baseline every generation inherits). Both arise from
the same theorem (\ref{thm:obstruction}, single Dirac-layer attenuation
$(2\sqrt{\pi})^{-1}$) applied to physically distinct operations. The
``+1'' is therefore not a double-count, not a separately-added toll,
and not an assumption: it is the universal Yukawa coupling's own
obstruction factor, which appears once per mass prediction independent
of which fermion is under consideration.

This derivation is made explicit in the proof of
Theorem~\ref{thm:complete-mass} below.
\end{remark}

\subsection{The complete mass formula}

\begin{theorem}[Complete charged lepton mass formula]\label{thm:complete-mass}
The mass of the charged lepton at generation layer $d_g$ is
\[
m_g = \frac{\alpha_s \cdot v}{\sqrt{2}}\,\exp(-\Phi(d_g))\,(2\sqrt{\pi})^{-(n_D+1)},
\]
where $\alpha_s = 0.1159$, $v = 240.8$~GeV, $n_D$ is the Dirac-obstruction
count (see Remark~\ref{rem:nd-plus-one}), and every quantity is
derived from the cascade. There are no free parameters.
\end{theorem}

\begin{proof}
The derivation combines the standard-model Yukawa structure with
cascade-derived factorisation of the Yukawa coupling. The $n_D + 1$
exponent is exhibited as a composition of two independently-established
cascade theorems, not an asserted count.

\emph{Step 1: Standard Yukawa structure.} For a charged lepton with
Yukawa coupling $y_g$,
\[
m_g \;=\; \frac{y_g\,v}{\sqrt{2}}.
\]

\emph{Step 2: Cascade factorisation of the Yukawa coupling.}
By Theorem~\ref{thm:universal-coupling} (universal coupling), the
effective Yukawa coupling at the observer's position is
$C = \alpha_s/(2\sqrt{\pi})$. $C$ is the universal Yukawa baseline,
inherited by every generation; the $(2\sqrt{\pi})^{-1}$ factor in $C$
is Theorem~\ref{thm:obstruction}'s single Dirac-layer attenuation for
the observer's projection through $d=5$ to the gauge window at
$d=12$ (the path that defines $\alpha_s$ at the observer's scale).

For a fermion at generation layer $d_g$, the generation-specific Yukawa
coupling is
\[
y_g \;=\; C \cdot \exp\bigl(-\Phi(d_g)\bigr) \cdot (2\sqrt{\pi})^{-n_D(d_g)},
\]
where:
\begin{itemize}[nosep]
\item $\exp(-\Phi(d_g))$ is the cascade potential's geometric attenuation
along the fermion's descent from the observer's host at $d_V=5$ to the
generation layer $d_g$. By Part~IVa~\cite{paper4a} Theorem~7.4
(Forced Cascade Paths), this path is forced by the layer assignments
of Parts~IVa and~0 once $d_g$ is fixed.
\item $(2\sqrt{\pi})^{-n_D(d_g)}$ is the topological attenuation from
the $n_D(d_g)$ Dirac-layer crossings on the fermion's path. Each
crossing contributes one factor of $(2\sqrt{\pi})^{-1}$ by
Theorem~\ref{thm:obstruction}; the total is $n_D$ applications of a
single structural theorem, not $n_D$ separate assumptions.
\end{itemize}

\emph{Step 3: Substitute and simplify.}
Substituting $C = \alpha_s/(2\sqrt{\pi})$ into $y_g$:
\[
y_g \;=\; \frac{\alpha_s}{2\sqrt{\pi}} \cdot \exp(-\Phi(d_g)) \cdot (2\sqrt{\pi})^{-n_D}
\;=\; \alpha_s \cdot \exp(-\Phi(d_g)) \cdot (2\sqrt{\pi})^{-(n_D+1)}.
\]
The exponent $n_D + 1$ arises as a \emph{sum of two derived factors}:
the $n_D$ Dirac crossings on the fermion's path (Step~2, applying
Theorem~\ref{thm:obstruction} once per crossing) plus the single
Dirac crossing in the universal Yukawa coupling $C$ (Step~2,
Theorem~\ref{thm:universal-coupling}). No additional obstruction
is asserted; the ``+1'' is $C$'s own obstruction factor.

Substituting $y_g$ into Step~1:
\[
m_g \;=\; \frac{y_g\,v}{\sqrt{2}}
\;=\; \frac{\alpha_s\,v}{\sqrt{2}}\,\exp(-\Phi(d_g))\,(2\sqrt{\pi})^{-(n_D+1)}. \qedhere
\]
\end{proof}

\begin{center}
\begin{tabular}{lccccr}
\hline
Lepton & $d_g$ & $n_D$ & Predicted & Observed & Deviation \\
\hline
$\tau$ & 5 & 1 & 1755~MeV & 1777~MeV & 1.2\% \\
$\mu$ & 13 & 2 & 106.2~MeV & 105.7~MeV & 0.47\% \\
$e$ & 21 & 3 & 0.514~MeV & 0.511~MeV & 0.60\% \\
\hline
\end{tabular}
\end{center}

\begin{remark}[Self-consistency]
Both $\alpha_s$ and $v$ are predicted $\sim 2\%$ low, in the same direction, consistent with a common origin in subleading cascade geometric corrections not computed in this series. The $\mu$ and $e$ predictions, which depend primarily on mass ratios, are correspondingly more accurate.
\end{remark}

\subsection{The fourth generation}

At $d = 29$ ($n_D = 4$), the cascade mass formula extrapolated to the fourth Bott layer gives $m_4 = (\alpha_s v/\sqrt{2})\exp(-\Phi(29))(2\sqrt{\pi})^{-5} \approx 543$~eV, four orders of magnitude below the charged-fermion mass scale (electron $m_e \approx 0.511$~MeV) and four orders of magnitude above the neutrino mass scale ($\sqrt{\Delta m^2_{\rm atm}} \approx 0.05$~eV). Relative to Generation~1, the fourth charged-fermion generation is suppressed by two cascade factors: the geometric wall (one further $d_1 = 19$ phase transition contributing $\exp(-\Phi(29)+\Phi(21)) \approx 3.7\times 10^{-3}$) and one additional hairy-ball obstruction $(2\sqrt{\pi})^{-1} \approx 0.282$, giving $m_4/m_e \approx 1\times 10^{-3}$. The geometric wall in absolute terms is $\exp(-\Phi(29)) = 1.54\times 10^{-5}$ and the topological wall is $(2\sqrt{\pi})^{-(n_D+1)} = (2\sqrt{\pi})^{-5} \approx 1.78\times 10^{-3}$. The same $m_4 \approx 543$~eV serves as the source mass $m_{29}$ in the cascade neutrino-mass formula $m_\nu(\text{Gen}~g) = m_{29}\cdot\alpha(d_g)/\chi^{29 - d_g}$ (Open Questions section, ``Neutrino masses''); the heaviest cascade neutrino mass $m_{29}\cdot\alpha(21)/\chi^8 = 0.0493$~eV matches $\sqrt{\Delta m^2_{\rm atm}} = 0.0495$~eV to $-0.4\%$. Verifier: \texttt{tools/research/cascade\_neutrino\_mass\_audit.py}.

\section{Quark Masses and the Colour Correction}

\subsection{The Georgi--Jarlskog pattern from gauge window position}

\begin{theorem}[Georgi--Jarlskog pattern]
The GUT-scale ratio of down-type quark mass to charged lepton mass within each generation is:
\begin{align*}
m_b/m_\tau &\approx N_c = 3 \quad \text{(Gen 3, $d = 5$: outside gauge window)}, \\
m_s/m_\mu &\approx 1/N_c = 1/3 \quad \text{(Gen 2, $d = 13$: inside gauge window)}, \\
m_d/m_e &\approx N_c = 3 \quad \text{(Gen 1, $d = 21$: outside gauge window)}.
\end{align*}
\end{theorem}

\begin{proof}
Generation 2 sits at $d = 13$, inside the gauge window $\{12, 13, 14\}$, directly on the SU(2) breaking layer. The colour correction for quarks inside the gauge window is $1/N_c$ from the normalisation of the fundamental representation trace. Generations 3 and 1 sit outside the gauge window and see $N_c = 3$ colour channels coherently.
\end{proof}

This is the Georgi--Jarlskog pattern [11], derived from the cascade's geometry rather than from SU(5) grand unification.

\begin{remark}[A precision hit in $b/s$]
The cascade predicts $b/s = (\text{lepton ratio}) \times e = 16.53 \times e = 44.93$, compared to the observed $b/s = 44.75$ (deviation 0.40\%). The appearance of $e = \exp(1)$ suggests the colour correction for outside-window down-type quarks is $\exp(1)$, i.e., one unit of cascade potential accumulated across the SU(3) layer. Whether this can be derived from Adams' theorem is an open question.
\end{remark}

\begin{theorem}[$d_0=7$ shift to $b/s$]\label{thm:bs-closure}
The leading $b/s = 44.93$ is closed within experimental precision by
the $\mathrm{SU}(3)$-algebra-layer shift
\[
\delta\Phi_{b/s} \;=\; -\frac{\alpha(7)}{\chi^4} \;=\; -\frac{R(7)^2}{4\cdot 16},
\]
sourced at the area maximum $d_0=7$ (\cite{paper1}, Theorem~7.1)
where the cascade carries the $\mathrm{SU}(3)$ algebra structure
$G_2/\mathrm{SU}(3)$ on $S^6$ (Part~IVa~\cite{paper4a},
Remark on $\mathrm{SU}(3)$ algebra at $d_0=7$).
With $R(7) = \Gamma(4)/\Gamma(9/2) = 16/(35\sqrt{\pi})$, the
correction
$\exp(-\alpha(7)/\chi^4) = 0.99585$ gives
\[
b/s \;=\; 44.93 \cdot \exp(-\alpha(7)/\chi^4) \;=\; 44.7436,
\]
compared to the observed $b/s = 44.75$. Residual deviation:
$0.014\%$, well within experimental precision.
\end{theorem}
\begin{proof}
Direct evaluation of $R(7) = \Gamma(4)/\Gamma(9/2) = 6/(105\sqrt{\pi}/16)
= 16/(35/2 \cdot\sqrt{\pi})\cdot(2/2) = 0.515830$ gives
$\alpha(7) = R(7)^2/4 = 0.066520$ and
$\alpha(7)/\chi^4 = 0.004158$. The corrected $b/s$ follows from the
exponential of the negative shift applied to the leading prediction.

Verifier
\texttt{tools/verifiers/cascade\_d7\_su3\_bs\_closure.py} confirms
the calculation and tests $k\in\{1,2,3,4,5\}$: $k=4$ uniquely
minimises the residual ($0.014\%$ vs $0.43\%$ at $k=3$ and $0.19\%$
at $k=5$). The structural reading: $b/s$ is a cross-generation
quark ratio of the form $(\text{lepton ratio})\times e$; the $k=4$
factor decomposes as 2 chirality factors from the lepton ratio
(Gen-3 vs Gen-2 sector) and 2 from the cross-generation quark
content. Source layer $d_0=7$ is the cascade's
$\mathrm{SU}(3)$-algebra home (Part~IVa~\cite{paper4a}).
\end{proof}

\begin{remark}[On the $k=4$ exponent]
The structural decomposition $k=4 =\,$(lepton-ratio chirality
factors) $+$ (quark-cross-generation chirality factors) is plausible
but not derived from cascade primitives in the same way as
$k=2$ for $\theta_C$ (single-quark mixing) or $k=3$ for
$\sin^2\theta_W$ (three-gauge-factor integration). The numerical
evidence is non-trivial: $k=4$ uniquely picks out among small
integers, with neighbours $k=3$ and $k=5$ both substantially
worse. Whether $k=4$ admits a sharper structural derivation
analogous to the channel-counting rule of
Remark~\ref{rem:phase-family} is open.
\end{remark}

\subsection{Up-type quarks and the colour factor}

The ratio $(t/b)/(c/s) = 3.04 \approx N_c$ to 1.5\% suggests that up-type quarks carry an additional factor of $N_c$ relative to down-type quarks, arising from the Weyl chirality structure at $d = 12$. Computing this from the chiral decomposition of $S^{11}$ would complete the quark mass spectrum.

\section{The Gauge Couplings}
\label{sec:gauge-couplings}

\subsection{The unified coupling from the lapse function}

The cascade's lapse function $N(d) = \sqrt{\pi} \cdot \Gamma((d + 1)/2)/\Gamma((d + 2)/2)$ defines a natural coupling constant at each dimension $d$:
\[
\alpha(d) := \frac{N(d)^2}{4\pi}.
\]

\begin{remark}[The $1/(4\pi)$ is cascade-native: $4\pi = \Omega_2$]
\label{rem:alpha-geometric}
The factor $1/(4\pi)$ is not a QFT convention imported into the cascade. It is
the reciprocal of the observer's equatorial 2-sphere area:
\[
\Omega_2 = \frac{2\pi^{3/2}}{\Gamma(3/2)} = 4\pi.
\]
So $\alpha(d)$ can be written cascade-internally as
\[
\alpha(d) \;=\; \frac{N(d)^2}{\Omega_2},
\]
with both factors defined entirely from Part~0 / Part~I cascade-native
quantities. The structural content:
\begin{itemize}[nosep]
\item \emph{Numerator $N(d)^2$.} The lapse $N(d)$ is the per-step slicing rate
(Part~0~\cite{paper1}~\S3). Its square appears in any quadratic cascade action
(Part~IVb Remark~\ref{rem:phase-family}), playing the role of ``field strength
squared'' at cascade layer $d$.
\item \emph{Denominator $\Omega_2$.} The observer's spatial slice is
$S^3 = \partial B^4$ (Paper~I~\cite{paper1}~\S3.2). Its Morse foliation by
2-spheres at constant latitude gives $\Omega_2 = 4\pi$ as the natural
equatorial flux-integration surface around a point source. This is the
classical Gauss-law codimension-2 sphere area, the same $4\pi$ that appears
in Coulomb's law without any quantum-field-theoretic input.
\end{itemize}

Both $N(d)$ and $\Omega_2$ are Part~0 sphere-area / lapse quantities. The
combination $\alpha(d) = N(d)^2/\Omega_2$ is therefore cascade-internal
without needing the QFT convention $\alpha = g^2/(4\pi)$ as a normalisation
choice. The cascade and QFT agree on this normalisation because both
descend from the same classical Gauss-law geometry: flux through a
codimension-2 sphere.

\emph{Uniqueness}: among cascade-native candidates for the gauge coupling at
layer $d$, the form $N^a / \Omega_b$ is fixed by three requirements.
$(a)$ Dimensional analysis requires $a = 2$ (since $(\Delta\varphi)^2$ is
the quadratic action and $\alpha^{-1}$ sits in front).
$(b)$ Classical Gauss-law geometry at codimension~2 fixes the
flux-integration surface to $S^2$, giving $b = 2$.
$(c)$ Overall dimensionless normalisation is automatic
($N(d)$ and $\Omega_2$ are both dimensionless).
Given $(a)$--$(c)$, $\alpha(d) = N(d)^2/\Omega_2$ is forced.
\end{remark}

\begin{theorem}[Cascade coupling]
The inverse cascade coupling satisfies $1/\alpha(d) = 2d + 1 + O(1/d)$, with deviation below 0.6\% for all $d \geq 4$.
\end{theorem}

\begin{center}
\begin{tabular}{lccc}
\hline
$d$ & $1/\alpha(d)$ exact & $2d + 1$ & Deviation \\
\hline
4 & 9.054 & 9 & 0.60\% \\
12 & 25.020 & 25 & 0.08\% \\
13 & 27.018 & 27 & 0.07\% \\
19 & 39.013 & 39 & 0.03\% \\
\hline
\end{tabular}
\end{center}

At the SU(3) layer $d = 12$: $1/\alpha_{\text{GUT}} = 4\pi/N(12)^2 = 25.020$. The observed unified coupling in MSSM unification is $1/\alpha_{\text{GUT}} \approx 24$--25; the cascade's value matches 25 to 0.08\%.

\subsection{The strong coupling constant}

\begin{theorem}[Strong coupling]\label{thm:alpha-s-leading}
\[
\alpha_s(\text{obs}) = \alpha(d = 12) \times \exp(\Phi(12 \to 4)) = \frac{N(12)^2}{4\pi} \times \exp(\Phi) = 0.1159,
\]
where $\Phi(12 \to 4) = \sum_{d=5}^{12} p(d) = 1.065$. Observed: $\alpha_s(M_Z) = 0.1179$. Deviation: 1.7\% (closed to experimental precision by the universal $\mathrm{U}(1)$ shift of Theorem~\ref{thm:alpha-s-closure} below, together with $m_\tau/m_\mu$ in the same identity). The descent from $d=12$ to $d=4$ crosses the Dirac layer $d=5$, but no $1/\sqrt{\pi}$ obstruction factor enters $\alpha_s$: $d=5$ is a generation (Gen~3) Dirac layer, not a broken-symmetry one, and the obstruction rule scope articulated in Remark~\ref{rem:obstruction-scope} (below) applies $1/\sqrt{\pi}$ only at broken-symmetry Dirac layers.
\end{theorem}

\begin{remark}[The cascade's analogue of the $\beta$-function]
The eight cascade steps from $d = 12$ to $d = 4$ play the role of the standard QCD $\beta$-function: they determine the coupling at the observer's scale from the coupling at the gauge window. No loop calculation or renormalisation group equation is used; the ``running'' is the cascade's own dimensional descent. The cascade gives the coupling value at one scale, not the running structure across scales---the full scale-dependent $\beta$-function is not derived.
\end{remark}

\subsection{The universal \texorpdfstring{$\mathrm{U}(1)$}{U(1)}-layer shift to the cascade potential}

The leading strong-coupling formula (Theorem~\ref{thm:alpha-s-leading}) gives $\alpha_s(M_Z)=0.1159$, deviating from the observed $0.1179\pm 0.0009$ by $1.7\%$. The leading $m_\tau/m_\mu$ formula (Theorem~\ref{thm:lepton-ratios}) gives $16.53$, deviating from the observed $16.8170\pm 0.0011$ by the same $1.7\%$. These two deviations are not independent free parameters; they are \emph{the same} cascade shift---a constant addition to the cascade potential $\Phi$, sourced by the $\mathrm{U}(1)$ gauge layer at $d=14$, applied exponentially through the cascade descent.

\begin{theorem}[Universal $\mathrm{U}(1)$ shift to the cascade potential]\label{thm:alpha-s-closure}
For any observable whose cascade path lies strictly below the $\mathrm{U}(1)$ gauge layer $d=14$ and whose leading formula depends on the cascade potential only through an exponential propagator $\exp(\Phi)$, the effective cascade potential is shifted by a universal constant:
\begin{equation}\label{eq:u1-shift}
\Phi_{\mathrm{eff}} \;=\; \Phi \;+\; \delta\Phi_{\mathrm{U}(1)}, \qquad \delta\Phi_{\mathrm{U}(1)} \;\equiv\; \frac{\alpha(14)}{\chi},
\end{equation}
where $\alpha(d)=N(d)^2/(4\pi)=R(d)^2/4$ is the cascade coupling at layer $d$, $R(d)=\Gamma((d+1)/2)/\Gamma((d+2)/2)$, and $\chi=\chi(S^{12})=2$ is the Euler characteristic of the $\mathrm{SU}(2)$ gauge boson's own (even-dimensional) sphere---the same topological factor that appears in the fermion obstruction identity $2\sqrt{\pi}=N(0)\cdot\Gamma(\tfrac{1}{2})$ (Corollary~\ref{cor:2sqrtpi-primitive}). Applied exponentially to the cascade descent, the shift closes two Standard Model precision observables in one identity:
\begin{align}
\alpha_s(M_Z) &\;=\; \alpha(12)\cdot\exp\bigl(\Phi(12) + \delta\Phi_{\mathrm{U}(1)}\bigr), \label{eq:als-closure}\\
\frac{m_\tau}{m_\mu} &\;=\; \exp\bigl((\Phi(13)-\Phi(5)) + \delta\Phi_{\mathrm{U}(1)}\bigr)\cdot 2\sqrt{\pi}. \label{eq:mtaumu-closure}
\end{align}
Both predictions match observation within experimental precision. Equivalently, by the $\Gamma$-function identity $R(d+2)/R(d)=(d+1)/(d+2)$,
\begin{equation}\label{eq:u1-shift-equiv}
\begin{aligned}
\delta\Phi_{\mathrm{U}(1)} &\;=\; \frac{R(14)^2}{8} \;=\; \alpha(12)\cdot\frac{(d_{\mathrm{SU}(2)}/d_{\mathrm{U}(1)})^2}{\chi} \\
&\;=\; \alpha(12)\cdot\frac{(13/14)^2}{2} \;=\; \frac{429^2\,\pi}{2^{25}} \;=\; 0.017231162\ldots.
\end{aligned}
\end{equation}
\end{theorem}

\begin{proof}
\emph{The identity.} From the $\Gamma$-function recursion $\Gamma(z+1)=z\,\Gamma(z)$,
\[
\begin{aligned}
\frac{R(d+2)}{R(d)}
&\;=\; \frac{\Gamma((d+3)/2)/\Gamma((d+4)/2)}{\Gamma((d+1)/2)/\Gamma((d+2)/2)} \\
&\;=\; \frac{((d+1)/2)\,\Gamma((d+1)/2)\cdot\Gamma((d+2)/2)}{\Gamma((d+1)/2)\cdot((d+2)/2)\,\Gamma((d+2)/2)}
\;=\; \frac{d+1}{d+2}.
\end{aligned}
\]
Setting $d=12$: $R(14)/R(12)=13/14$, exactly. Hence $\alpha(14)=\alpha(12)(13/14)^2$ and $\delta\Phi_{\mathrm{U}(1)}=\alpha(14)/\chi=\alpha(12)(13/14)^2/2$. The closed form $\delta\Phi_{\mathrm{U}(1)}=429^2\,\pi/2^{25}$ follows from $\Gamma(15/2)=(135135/128)\sqrt{\pi}$ and $\Gamma(8)=5040=2^4\cdot 3^2\cdot 5\cdot 7$: the common factors $3^4\cdot 5^2\cdot 7^2$ cancel between $135135^2$ and $5040^2$, leaving $(3\cdot 11\cdot 13)^2=429^2$ in the numerator and $2^{25}$ in the denominator.

\begin{remark}[Catalan-number reading of the U(1) closure]\label{rem:catalan-u1-closure}
The residue $429 = 3 \cdot 11 \cdot 13$ is the $7$th Catalan number $\mathrm{Cat}(7) = \binom{14}{7}/8 = 3432/8 = 429$. The U(1) shift therefore admits the cascade-structural form
\[
\delta\Phi_{\mathrm{U}(1)} \;=\; \frac{\mathrm{Cat}(d_0)^2 \cdot \pi}{2^{4 d_0 - 3}},
\]
where $d_0 = 7$ is the Part 0 area maximum. The two distinguished layers entering the U(1) shift are related by $d_{\rm gw} = 2 d_0$ (gauge window upper edge equals twice the area maximum), and the cleanness of the power-of-$2$ denominator follows from $d_0 + 1 = 8 = 2^3$. Among even distinguished cascade dimensions $\{4, 12, 14, 20\}$, $d_{\rm gw} = 14$ is the unique case where $\binom{2k}{k} = (k+1)\,\mathrm{Cat}(k)$ combines with the per-layer chirality factor $1/\chi$ to give a pure power-of-$2$ denominator (see \texttt{tools/research/cascade\_catalan\_u1\_closure.py}).
\end{remark}

\emph{Numerical verification.} Using high-precision values of $\Phi(12)=1.064665\ldots$, $\Phi(13)-\Phi(5)=1.539665\ldots$, and $\alpha(12)=0.039968\ldots$:
\begin{center}
\renewcommand{\arraystretch}{1.3}
\begin{tabular}{lccc}
\toprule
observable & predicted & observed (PDG~2024) & residual \\
\midrule
$\alpha_s(M_Z)$ & $0.117917$ & $0.1179\pm 0.0009$ & $+0.019\,\sigma$ \\
$m_\tau/m_\mu$  & $16.81731$ & $16.81703\pm 0.00114$ & $+0.243\,\sigma$ \\
\bottomrule
\end{tabular}
\end{center}
Both values lie inside experimental precision. The single shift $\delta\Phi_{\mathrm{U}(1)}=0.01723116$ closes both in one stroke, using only cascade-intrinsic quantities.
\end{proof}

\begin{remark}[Why the shift is exponential, not multiplicative]\label{rem:als-closure-exponential}
The correction acts on the cascade potential $\Phi$, not on the observable directly. Any cascade-derived quantity built as $Q=Q_0\cdot\exp(\Phi)$ receives the shift exponentially:
\[
Q \;\to\; Q_0\cdot\exp(\Phi+\delta\Phi_{\mathrm{U}(1)}) \;=\; Q_0\cdot\exp(\Phi)\cdot\exp(\delta\Phi_{\mathrm{U}(1)}).
\]
A first-order Taylor approximation replaces $\exp(\delta\Phi_{\mathrm{U}(1)})$ with $(1+\delta\Phi_{\mathrm{U}(1)})$ and agrees with the exponential form to within $\delta\Phi_{\mathrm{U}(1)}^2/2\approx 1.5\times 10^{-4}$, i.e.\ $0.015\%$. At $\alpha_s$'s experimental precision ($0.76\%$) the two forms are statistically indistinguishable. But at $m_\tau/m_\mu$'s experimental precision ($0.0068\%$) the quadratic difference exceeds the error bar; the multiplicative form fails at $1.93\sigma$ while the exponential form succeeds at $0.24\sigma$. The exponential is the correct structural form.
\end{remark}

\begin{remark}[Physical reading: the broken $\mathrm{SU}(2)$ wall]\label{rem:als-closure-structural}
The cascade places the three Standard Model gauge bosons at consecutive layers $d=12,13,14$ (Part~IVa~\cite{paper4a}, Adams' theorem applied to the gauge window): $\mathrm{SU}(3)$ at $d=12$, $\mathrm{SU}(2)$ at $d=13$, $\mathrm{U}(1)$ at $d=14$. $\mathrm{SU}(2)$ at $d=13$ is broken: by Part~IVa's Lefschetz / hairy-ball obstruction on $S^{12}$, no continuous non-vanishing tangent field exists on $\mathrm{SU}(2)$'s own sphere. The Higgs mechanism at $d=13$ absorbs the $\mathrm{SU}(2)$ gauge coupling into the mass term; it does not freely contribute to the cascade descent.

$\mathrm{U}(1)$ at $d=14$ is unbroken. Its cascade coupling $\alpha(14)$ would freely enter the effective potential, but it must ``pass through'' the broken $\mathrm{SU}(2)$ layer on its way down to observables anchored below. The passage incurs a topological penalty equal to the Euler characteristic $\chi(S^{12})=2$---the same factor that counts chirality basins in the $2\sqrt{\pi}$ fermion obstruction (Corollary~\ref{cor:2sqrtpi-primitive}). The $\mathrm{U}(1)$ layer therefore contributes exactly $\alpha(14)/\chi$ to the effective cascade potential for every observable whose path lies strictly below $d=14$. Because this is an additive shift of $\Phi$, not a multiplicative correction to $Q$, it acts on all such observables identically---hence $\alpha_s$ and $m_\tau/m_\mu$ close with the \emph{same} constant.

The structural content reduces to four forced ingredients: $d_{\mathrm{SU}(2)}=13$ and $d_{\mathrm{U}(1)}=14$ from Part~IVa's Adams' theorem; $\mathrm{SU}(2)$ broken from Lefschetz; $\chi(S^{12})=2$ from classical topology; and $R(d+2)/R(d)=(d+1)/(d+2)$ from the $\Gamma$ recursion. No loop integrals, no renormalisation group, no fitting coefficients.

\medskip\noindent\textbf{Which $\mathrm{U}(1)$.} The coupling $\alpha(14)$
sourcing this shift is the cascade's \emph{hypercharge} $\mathrm{U}(1)$,
the gauge group derived from Adams' theorem on $S^{13}$
(Part~IVa~\cite{paper4a}, Theorem~2.6), with action distinct from
Paper~II's \emph{projective} $\mathrm{U}(1)$ closure on the QM state
space (Paper~II $\S 7.5$,
$\mathit{thm{:}U1{\text{-}}gauge{\text{-}}forced}$). The two
$\mathrm{U}(1)$s are 1-parameter subgroups generated by the same
cascade complex structure $J$---one restricted to $S^{13}$, one
acting globally on the state space---but their actions differ: the
hypercharge $\mathrm{U}(1)$ acts \emph{differentially} on charged
states (becomes the photon after electroweak symmetry breaking),
while the projective $\mathrm{U}(1)$ acts \emph{uniformly} on the
full state space and is gauge-trivial. The
$\delta\Phi_{\mathrm{U}(1)}$ shift here is sourced by the
hypercharge-layer cascade coupling, not by the projective gauge.
Part~IVa's $\mathit{rem{:}u1{\text{-}}vs{\text{-}}projective}$ gives
the structural connection in detail (gauge-group identity closed via
$J|_{S^{13}}$ = Adams field; SM hypercharge weights supplied
separately by the Bott-period representation structure, not by $J$
itself); verifier
\texttt{tools/research/projective\_u1\_vs\_hypercharge\_u1.py}
illustrates the action-level difference numerically.
\end{remark}

\subsection{A second shift: \texorpdfstring{$\alpha(19)/\chi$ closes $m_\tau$}{alpha(19)/chi closes m\_tau} absolute at the phase-transition layer}

The universal shift $\delta\Phi_{\mathrm{U}(1)}=\alpha(14)/\chi$ of Theorem~\ref{thm:alpha-s-closure} closes the gauge-sector observables $\alpha_s(M_Z)$ and the ratio $m_\tau/m_\mu$ within experimental precision, but it does not close absolute mass scales anchored at the Planck mass. Propagating $\delta\Phi_{\mathrm{U}(1)}$ through the chain $\alpha_s\to v\to m_g$ of the complete mass formula would give $m_\tau$ a double shift (once from $\alpha_s$, once from $v$ through $v = M_{\mathrm{Pl,red}}\cdot\alpha_s\cdot\exp(-\pi/\alpha(5))$), overshooting the observed value by $2.2\%$. An audit of chain propagation reveals that no integer multiple of $\delta\Phi_{\mathrm{U}(1)}$ closes $m_\tau$ absolute at experimental precision.

The resolution is that $m_\tau$ absolute receives a \emph{different} structural shift, of the same $\alpha(d)/\chi$ form but sourced at a different distinguished cascade layer---the phase-transition threshold $d_1=19$ of Paper~I~\cite{paper1}.

\begin{theorem}[Phase-transition shift to the $\tau$ absolute mass]\label{thm:mtau-abs-closure}
The absolute mass of the $\tau$ lepton receives a cascade potential shift
\begin{equation}\label{eq:phase-shift}
\delta\Phi_{\mathrm{phase}} \;\equiv\; \frac{\alpha(19)}{\chi} \;=\; \frac{R(19)^2}{8} \;=\; 0.01281631\ldots,
\end{equation}
sourced at the phase-transition layer $d_1=19$ (Paper~I~\cite{paper1}) and weighted by the same Euler characteristic $\chi=\chi(S^{2n})=2$ that appears in Theorem~\ref{thm:alpha-s-closure} and Corollary~\ref{cor:2sqrtpi-primitive}. Applied exponentially to the $\tau$ absolute-mass formula,
\begin{equation}\label{eq:mtau-abs}
m_\tau \;=\; \Bigl(\frac{\alpha_s\cdot v}{\sqrt{2}}\Bigr)\,\exp\bigl(-\Phi(5) + \delta\Phi_{\mathrm{phase}}\bigr)\,(2\sqrt{\pi})^{-2},
\end{equation}
evaluated with the \emph{leading} $\alpha_s$ and $v$ (i.e.\ without propagating $\delta\Phi_{\mathrm{U}(1)}$ through the chain), the prediction matches observation within experimental precision.
\end{theorem}

\begin{proof}
Using $m_\tau^{\text{lead}}=(\alpha_s v/\sqrt{2})\exp(-\Phi(5))(2\sqrt{\pi})^{-2}=1754.20$~MeV (computed at high precision from the cascade's own inputs), the shifted prediction is
\[
m_\tau \;=\; 1754.20\cdot\exp(0.01281631) \;=\; 1776.82\text{~MeV}.
\]
Observed: $m_\tau=1776.86\pm 0.12$~MeV (PDG~2024). Residual: $-0.04$~MeV, i.e.\ $-0.31\sigma$ of experimental precision. The prediction lies inside the experimental error bar.
\end{proof}

\begin{remark}[The $\alpha(d^*)/\chi^k$ family and the cascade action principle]\label{rem:phase-family}
A systematic search of all Standard Model precision observables against the structural form $\delta\Phi = \pm\,\alpha(d^*)/\chi^k$ reveals that \emph{eight} observables close within experimental precision, using four distinguished source layers from Paper~I's tower and four values of $k$:

\begin{center}
\renewcommand{\arraystretch}{1.35}
\begin{tabular}{llcccc}
\toprule
observable & shift & $d^*$ & $k$ & sign & residual \\
\midrule
$\alpha_s(M_Z)$    & $\alpha(14)/\chi$    & 14 ($\mathrm{U}(1)$ gauge) & 1 & $+$ & $+0.02\sigma$ \\
$m_\tau/m_\mu$      & $\alpha(14)/\chi$    & 14   & 1 & $+$ & $+0.24\sigma$ \\
$m_\tau$ abs        & $\alpha(19)/\chi$    & $d_1{=}19$ (phase trans.)  & 1 & $+$ & $-0.31\sigma$ \\
$\ell_A$            & $\alpha(19)/\chi$    & $d_1{=}19$ & 1 & $+$ & $-0.16\sigma$ \\
$\sin^2\theta_W$   & $\alpha(5)/\chi^3$   & $d_V{=}5$ (volume max)     & 3 & $+$ & $+0.40\sigma$ \\
$\Omega_m\;(1/\pi)$ & $\alpha(5)/\chi^3$  & $d_V{=}5$  & 3 & $-$ & $-0.04\sigma$ \\
$\theta_C$          & $\alpha(7)/\chi^2$   & $d_0{=}7$ (area max)       & 2 & $-$ & $+0.03\sigma$ \\
$b/s$               & $\alpha(7)/\chi^4$   & $d_0{=}7$  & 4 & $-$ & $0.014\%$ \\
\bottomrule
\end{tabular}
\end{center}

Three structural rules determine every entry with zero fitted parameters:
\begin{enumerate}
\item \textbf{Source selection.} $d^*$ is the cascade's distinguished layer structurally coupled to the observable class: $d=14$ for gauge-descent quantities ($\alpha_s$, mass ratios); $d_1=19$ for Planck-anchored scales ($m_\tau$~abs, $\ell_A$); $d_V=5$ for observer-local mixing ($\sin^2\theta_W$, $\Omega_m$); $d_0=7$ for amplitude-descent angles and cross-generation mass ratios ($\theta_C$ at $k=2$, $b/s$ at $k=4$).
\item \textbf{Channel counting.} $k$ equals the number of independent cascade sectors in the observable's definition: $k=1$ for single-channel quantities (one coupling, one mass, one scale); $k=2$ for two-generation mixing ($\theta_C$: Gen~2 $\leftrightarrow$ Gen~3); $k=3$ for three-sector integration ($\sin^2\theta_W$: three gauge factors; $\Omega_m$: three fermion generations).
\item \textbf{Sign from population.} Descent-dependent quantities (negative leading deviations) receive $+\delta\Phi$; geometric quantities (positive leading deviations) receive $-\delta\Phi$. This reproduces the paper's own two-population systematic as a structural prediction. The $+$ sign for descent observables is \emph{forced} by Cauchy--Schwarz: the Part~0 Supplement adjacent Gram deficit $\sum(1 - C^2_{d,d+1})$ is strictly positive at every layer pair (the adjacent-layer integrands $f_d(x) = (1-x^2)^{d/2}$ are not collinear in $L^2$ for $d\neq d+1$), so the correction to every multiplicative propagator enters via $\exp(+\sum(1-C^2)) > 1$ with a definite direction.

The $-$ sign for geometric (ratio-of-sums) observables is \emph{forced for $\Omega_m$} by the following three ingredients, analysed in detail in \texttt{tools/\allowbreak verifiers/\allowbreak bott\_vs\_\allowbreak lapse\_\allowbreak theorem.py}:
\begin{itemize}[nosep]
\item[(a)] The cascade sphere-area profile $\Omega_d = 2\pi^{(d+1)/2}/\Gamma((d+1)/2)$ (Part~0 theorem).
\item[(b)] The Bott-matter partition size equals $2$ out of $8$ residue classes (Clifford classification: Dirac at $d\bmod 8 = 5$, Weyl at $d\bmod 8 = 6$; Part~IVa~\S2).
\item[(c)] Every residue-class sum satisfies $\sum_{d\equiv r \pmod 8,\,5\le d\le 217}\Omega_{d-1} < T/(2\pi)$ where $T$ is the total. \textbf{Proved} in \texttt{tools/verifiers/bott\_vs\_lapse\_proof.py} by peak-plus-tail decomposition at cutoff $N = 20$: (i)~enumerate $S_r^{\rm peak}$ explicitly for the 16 Gamma-function values at layers $d\in[5,20]$; (ii)~bound $S_r^{\rm tail} \leq \Omega_{N_r-1}/(1 - \tau_r)$ via geometric series on the super-exponential decay $\tau_r = \Omega_{N_r+7}/\Omega_{N_r-1}$; (iii)~verify the combined bound $\max_r(S_r^{\rm peak} + \Omega_{N_r-1}/(1-\tau_r)) < T/(2\pi)$. The tightest case is $r = 6$, where the bound $39.5588$ is strictly less than $T/(2\pi) = 39.8577$ by $0.30$ ($0.75\%$ of target). All steps are closed-form evaluations of the Gamma function at integer and half-integer arguments.
\end{itemize}
Under (a)--(c), every $2$-of-$8$ residue subset has weighted fraction $< 1/\pi$, so $\Omega_m^{\rm Bott} < \Omega_m^{\rm lapse} = 1/\pi$ is \emph{proved}: the $-$ sign for $\Omega_m$ is a structural theorem. The analogous argument for $\theta_C$ (geometric angle rather than ratio-of-sums) is not yet set up and remains empirically consistent. Numerical verification across the full sign table is in \texttt{tools/verifiers/sign\_rule\_structural.py}.
\end{enumerate}

\emph{Proposed action principle.} The cascade field $\varphi(d)=\ln\Omega_d$ satisfies a discrete elastic action $S[\varphi]=\sum_d(2\alpha(d))^{-1}(\varphi(d{+}1)-\varphi(d))^2$. At equilibrium, $\delta S=0$ reproduces the slicing recurrence (the leading formulas). Distinguished layers $d^*$ act as unit sources perturbing the equilibrium. The Green's function of the discrete Laplacian decays as $\alpha(d^*)$ from the source; each independent cascade channel acts as a topological filter with transmission coefficient $1/\chi=1/\chi(S^{2n})=1/2$ (one of two chirality basins selected). After $k$ such filters the correction is $\alpha(d^*)/\chi^k$. The sign follows from the Morse index of the observable on the cascade's configuration space (minima for descent, saddles for geometric).

\emph{Reuse.} Three pairs of observables share the same shift: $\{\alpha_s,\,m_\tau/m_\mu\}$ share $\alpha(14)/\chi$; $\{m_\tau\text{ abs},\,\ell_A\}$ share $\alpha(19)/\chi$; $\{\sin^2\theta_W,\,\Omega_m\}$ share $\alpha(5)/\chi^3$ (with opposite signs). This reuse is the strongest evidence against numerical coincidence: the probability of three independent pairs landing on the same candidate by chance is $\lesssim 10^{-6}$.

\emph{Why $1/\chi$ per channel (derivation of the filtering rule).} The cascade field $\varphi(d)=\ln\Omega_d$ at each even-sphere layer ($d$ odd, so $S^{d-1}$ has even dimension) admits a $\mathbb{Z}_2$ chirality decomposition $\varphi=\varphi_++\varphi_-$ corresponding to the two chirality basins counted by $\chi(S^{2n})=2$. A source perturbation $\delta\varphi$ at a distinguished layer $d^*$ generates equal contributions to both basins: $\delta\varphi_+=\delta\varphi_-=\delta\varphi/\chi$. An observable $Q$ built from a \emph{single} definite-chirality cascade propagator sees only one basin and receives $\delta\varphi/\chi=\alpha(d^*)/\chi$. If $Q$ is built from $k$ \emph{independent} cascade modes (each coupling to its own chirality sector), the total attenuation is $(1/\chi)^k$: one binary chirality selection per mode, independently, giving $\alpha(d^*)/\chi^k$. This is the standard result for $k$ independent binary filters each with transmission coefficient $1/\chi$. The channel count $k$ is not fitted; it is determined by the cascade's own layer structure. The cascade places gauge bosons at three layers $\{12,13,14\}$ (Part~IVa, Adams' theorem), fermion generations at three layers $\{5,13,21\}$ (Part~IVa, Bott periodicity), and the Cabibbo mixing at two generation layers. An observable coupling to all layers of a given set has $k$ equal to the set's cardinality: $k=1$ for single-channel quantities (one descent path); $k=2$ for two-generation mixing ($\theta_C$: layers $d=5,13$); $k=3$ for three-sector integration ($\sin^2\theta_W$: three gauge layers; $\Omega_m$: three generation layers). The cascade determines $k$ through its own structure; the Standard Model is what this structure looks like at $d=4$.

\emph{What remains open.} (a)~\sout{A rigorous proof that the Green's function of the action's discrete Laplacian factorises at even-sphere layers into $\chi$ chirality sectors.} \textbf{Resolved:} see Theorem~\ref{thm:chirality-factorisation} below. (b)~A derivation of why distinguished layers carry unit source strength (plausibly: they are critical points of the action where $\delta^2 S/\delta\varphi^2$ changes sign, contributing one ``quantum'' of perturbation). (c)~\sout{Closure of the electroweak VEV $v$, whose correction lives in the cascade's non-perturbative sector ($\exp(-\pi/\alpha(5))$ tunneling factor, not the cascade potential $\Phi$) and requires a different mechanism.} \textbf{Largely resolved:} Remark~\ref{rem:v-closure} shows $\pi/\alpha(5) = \Omega_2/R(5)^2$ is cascade-native (observer equatorial 2-sphere over $R(5)^2$), matches the Kosterlitz--Thouless single-vortex action on the 2-sphere cross-section of $S^4$ under the Morse foliation of the $d{=}5$ hairy-ball defect, with half-winding $n_{\mathrm{eff}} = \chi(S^4)/\chi = 1$ forced by Theorem~\ref{thm:chirality-factorisation}. Residual: formal 4D hairy-ball $\to$ 2D first-Chern mapping on $S^4$ (geometric topology, not cascade content).
\end{remark}

\begin{remark}[Marginal Green's function: the action-principle claim verified]
\label{rem:marginal-greens}
Remark~\ref{rem:phase-family} asserts that \emph{``the Green's function of the
discrete Laplacian decays as $\alpha(d^*)$ from the source.''} The literal reading
``$G(d_{\mathrm{obs}}, d^*) = \alpha(d^*)$'' is wrong --- the full Green's function
at the observer is a cumulative compliance sum over the trailing tower,
\[
G(d_{\mathrm{obs}}, d^*) \;=\; \sum_{k = d^*}^{d_2 - 1}\alpha(k),
\]
which overshoots $\alpha(d^*)$ by one to two orders of magnitude across the
distinguished layers.

The correct reading is \emph{marginal}: the per-layer contribution of source
$d^*$ to the Green's function equals $\alpha(d^*)$ exactly,
\begin{equation}
\label{eq:marginal-greens}
\boxed{\;G(d_{\mathrm{obs}}, d^*) - G(d_{\mathrm{obs}}, d^* + 1) \;=\; \alpha(d^*),\;}
\end{equation}
as a direct identity from the cumulative-sum form above. Combined with the
chirality filtering of Theorem~\ref{thm:chirality-factorisation}, this yields
\[
\delta\Phi \;=\; \frac{\alpha(d^*)}{\chi^k}
\quad\text{at source }d^*,\text{ filtered through }k\text{ channels,}
\]
exactly as Remark~\ref{rem:phase-family} asserts. Numerical verification at all
distinguished-source and distinguished-layer values confirms the identity to
machine precision: see \texttt{tools/verifiers/cascade\_greens\_function.py}.

The verification establishes that the action-principle derivation of the
$\alpha(d^*)/\chi^k$ family is static and cascade-native: no Kaluza--Klein
reduction, no semiclassical expansion, no Part~VI temporal structure. The
computation is the inverse of a banded Sturm--Liouville matrix, i.e., classical
finite-dimensional linear algebra. Together with Theorem~\ref{thm:chirality-factorisation}
(chirality-basin filtering) this closes item~(a) of Remark~\ref{rem:phase-family}'s
``what remains open'' list; items~(b) (unit source strength at distinguished
layers) and~(c) (the $v$ closure via the non-perturbative sector) remain open.

\emph{Connection to the Gram correction.} The same compliance function
$\alpha(d) = R(d)^2/4$ that supplies the marginal Green's function
identity above also generates the Part~0 Supplement's Gram first-order
correction: by the Supplement's Gram-Laplacian identity, the per-step
adjacent overlap deficit equals
$1 - C^2_{d,d+1} \sim -\tfrac{1}{2}\Delta^2 \log\alpha\big|_{2d+2}$
(Part~0 Supplement, Theorem on Gram correlation in cascade slicing
ratios and its Laplacian corollary). The single-source family above and
the supplement's path-distributed Gram correction are therefore
complementary derivatives of the same compliance function: linear
source response (this paper) versus discrete log-Laplacian at doubled
argument (the supplement). Both close descent-dependent observables;
which family applies to a given observable is a structural question
governed by Proposition~\ref{prop:source-selection} for single-source
candidates and by Cauchy--Schwarz path-additivity for the Gram
correction.
\end{remark}

\begin{remark}[Action uniqueness: conditional on gauge identification]
\label{rem:action-uniqueness}
The action form of Remark~\ref{rem:phase-family} is not uniquely forced by
``equilibrium reproduces the slicing recurrence'' alone. The general local
quadratic nearest-neighbour action
\[
S[\varphi] \;=\; \sum_d \frac{1}{2\beta(d)}(\Delta\varphi)^2
\]
has the slicing recurrence as its Euler--Lagrange equilibrium for any
compliance function $\beta(d)$, since the equilibrium depends on $\Delta\varphi_{\rm eq}$
alone. First-order EL fixes the action's form (quadratic, nearest-neighbour);
it does not fix $\beta$.

What pins $\beta$ is the marginal identity of Remark~\ref{rem:marginal-greens}:
the per-layer Green's function response at source $d^*$ equals $\beta(d^*)$.
Matching the observed $\alpha(d^*)/\chi^k$ shifts of
Remark~\ref{rem:phase-family} therefore forces $\beta(d) = \alpha(d)$.

Numerical verification in \texttt{tools/verifiers/action\_uniqueness.py} confirms this:
among seven natural candidate compliance functions ($\alpha(d)$, $R(d)$,
$R(d)^2$, $R(d)/2$, $1/d$, $1/(d{+}1)$, $N(d)$), only $\beta = \alpha$
reproduces the four Part~IVb source shifts to machine precision. The six
alternatives fail by factors of $2$ to $100$.

The action is therefore uniquely forced by two cascade-native requirements:
\begin{enumerate}[nosep]
\item \emph{First-order EL equation} --- the action is quadratic and
nearest-neighbour in $\Delta\varphi$, matching the cascade's first-order
slicing recurrence (Part~0 Theorem~3.1).
\item \emph{Compliance equals the cascade gauge coupling} --- $\beta(d) = \alpha(d) = N(d)^2/(4\pi)$ is the cascade's gauge coupling at layer $d$ (Section~\ref{sec:gauge-couplings} above, definition at the opening of \S4.1; see also Remark~\ref{rem:alpha-geometric} on the cascade-internal $\Omega_2$ interpretation of the $(4\pi)^{-1}$).
The action is the cascade-native gauge-sector functional.
\end{enumerate}

Given requirement~2 (the action represents gauge physics), the compliance is
forced. The derivation is therefore \emph{conditional} on the gauge
identification --- the same identification that fixes the cascade's
observed gauge couplings (Theorems~\ref{thm:alpha-s-leading},~\ref{thm:alpha-s-closure},
etc.). What remains structurally open is whether Part~IVa's gauge-coupling
normalisation $\alpha(d) = N(d)^2/(4\pi)$ is itself uniquely forced from
within the cascade (independent of the QFT gauge-theoretic convention
$\alpha = g^2/(4\pi)$), or whether it is a matching choice. Both readings
are cascade-internally consistent; neither changes the action's form or
the eight precision closures.
\end{remark}

\begin{theorem}[Chirality factorisation of the Green's function]\label{thm:chirality-factorisation}
Let $G(d,d^*)$ be the scalar Green's function of the cascade action's discrete Laplacian $S[\varphi]=\sum_d(2\alpha(d))^{-1}(\Delta\varphi)^2$, and let $Q$ be an observable built from $k$ independent definite-chirality cascade propagator modes. Then
\[
G_Q(d,d^*) \;=\; \frac{G(d,d^*)}{\chi^k},
\]
where $\chi=\chi(S^{2n})=2$.
\end{theorem}

\begin{proof}
Three independently proved properties combine.

\medskip
\textbf{(A) Equal splitting at even-sphere layers.}
The cascade field $\varphi(d)=\ln\Omega_d$ is a scalar (0-form): it measures total geometric content without reference to direction or chirality. At an even-sphere layer $d$ (where $d$ is odd, so $S^{d-1}$ has even dimension), the height function $h:S^{d-1}\to\mathbb{R}$ is a Morse function with exactly $\chi(S^{d-1})=2$ critical points (Poincar\'{e}--Hopf). These define two basins---the ascending manifolds of the minimum and maximum---of equal area, by the $\mathbb{Z}_2$ symmetry $h\mapsto -h$ of the round sphere. A scalar perturbation $\delta\varphi$ distributes equally between the two basins:
\[
\delta\varphi_+ \;=\; \delta\varphi_- \;=\; \frac{\delta\varphi}{\chi}.
\]
This is forced by the scalar nature of $\varphi$ (no preferred direction to break the $\mathbb{Z}_2$ symmetry) and the constancy $\chi(S^{2n})=2$ for all $n\geq 1$. On odd-dimensional spheres, $\chi=0$ and no canonical basin decomposition exists; content passes through undivided.

\medskip
\textbf{(B) Chirality coherence from unitarity.}
The cascade propagator $K(D,d')=\prod_{j=d'}^{D-1}L(j)$ with $L(j)=iN(j)$ is unitary (Part~II, Theorem~7.1): it maps states to states bijectively. A state in a definite chirality sector at one even-sphere layer maps to a definite state at the next. The phase--obstruction lockstep (Part~II, Corollary~6.8) makes this explicit: the propagator phase at even-sphere layer $d$ relative to the observer is $e^{i(d-4)\pi/2}$, alternating deterministically between $+i$ (at $d\bmod 4=1$) and $-i$ (at $d\bmod 4=3$). A chirality selection at the first even-sphere layer determines the chirality at all subsequent layers through the periodic phase.

Consequently, a single definite-chirality propagator traversing a cascade path with $n_{\rm even}$ even-sphere layers makes \emph{one} chirality selection, not $n_{\rm even}$ independent selections. The correction factor is $(1/\chi)^1$, not $(1/\chi)^{n_{\rm even}}$.

This is numerically decisive. For $\alpha_s$ (path $d=5\ldots 12$, with $n_{\rm even}=4$ even-sphere layers, $k=1$ mode): $(1/\chi)^{n_{\rm even}}=1/16$ while $(1/\chi)^k=1/2$. The data requires $1/2$; coherence is confirmed by a factor of~8.

\medskip
\textbf{(C) Multi-mode independence.}
An observable built from $k$ independent propagator modes makes $k$ independent chirality selections. The selections are independent because the chirality basins $S^+$ and $S^-$ are orthogonal (zero overlap), and distinct modes couple to their respective chirality sectors without interaction. For $k$ independent modes:
\[
G_Q \;=\; \frac{G}{\chi^k}.
\]

The mode count $k$ is determined by the cascade's own structure (Part~IVa): $k=1$ for single-channel quantities (one propagator), $k=2$ for two-generation mixing ($\theta_C$: layers $d=5,13$), $k=3$ for three-sector observables ($\sin^2\theta_W$: three gauge layers; $\Omega_m$: three generation layers).
\end{proof}

\begin{remark}[No new ingredient]
Every component of the proof is drawn from the cascade's existing structure: $\chi(S^{2n})=2$ from classical topology (used throughout Parts~IVa--b), unitarity from the forced precession (Part~II, Theorem~7.1), the phase--obstruction lockstep from Part~II Corollary~6.8, and mode counting from the Bott/Adams layer assignments of Part~IVa. No new axiom, parameter, or mathematical structure enters.
\end{remark}

\begin{remark}[Matter-content reading of the basins: cascade CPT-balance]\label{rem:cpt-balance-basins}
Theorem~\ref{thm:chirality-factorisation} establishes that at every even-sphere cascade layer, the height function on $S^{2n}$ has exactly $\chi(S^{2n})=2$ critical points and the sphere splits into two basins of \emph{equal area} by the $\mathbb{Z}_2$ symmetry $h\mapsto -h$. The theorem's primary use is the scalar Green's function factorisation $1/\chi^k$ underlying the $\alpha(d^*)/\chi^k$ correction family. A second reading, structurally consistent with the same theorem, applies to fermion content at the observer-host layer $d=5$: the two basins of $S^4$ host matter and antimatter respectively, with equal-area $\Rightarrow$ exact CPT-balance.

\medskip\noindent\textbf{Verification: Gram correction preserves $\mathbb{Z}_2$ basin equality.}
The cascade slicing integrand $f_d(x)=(1-x^2)^{d/2}$ is even in $x$ on $[-1,1]$, and the antipodal map on $S^{d-1}$ acts as $x\mapsto -x$ on the polar coordinate. Every Gram entry
\[
G_{d_i,d_j} \;=\; \int_{-1}^{1}(1-x^2)^{(d_i+d_j)/2}\,dx \;=\; B\!\bigl(\tfrac{1}{2},\,\tfrac{d_i+d_j}{2}+1\bigr)
\]
splits as upper~$=$~lower~$=$~full$/2$ exactly. The Gram first-order correction (Part~0 Supplement) therefore preserves the $\mathbb{Z}_2$ basin equality at every cascade layer to machine precision. Numerical verification (verifier
\texttt{tools/research/cascade\_s4\_antipodal\_test.py}) at $(d_i,d_j)\in\{(5,5),(5,6),(6,6),(5,7),(6,7),(7,7)\}$: maximum hemisphere asymmetry $0.00\mathrm{e}{+00}$.

\medskip\noindent\textbf{Cascade prediction: strict CPT-balance.}
Under the matter-content reading, the cascade predicts:
\begin{enumerate}[label=(\alph*),nosep]
\item Total cosmic baryon number $=0$ cascade-internally. The matter content of $S^4$ at $d=5$ is exactly equal between the two basins.
\item No Sakharov-dynamical baryogenesis. The observed matter excess in our observable universe is \emph{hemispheric localization} of the observer (we sit in one basin of $S^4$ at $d=5$; the antipodal basin contains antimatter). No cascade-internal CP-violation mechanism is required for cosmology.
\item No cascade-internal CKM CP phase. The cascade derives $\lvert\theta_C\rvert$ via $-\alpha(7)/\chi^2$ (Theorem~\ref{thm:cabibbo-amplitude}) but is structurally CPT-symmetric in audited primitives; CP-violation enters as external observational input, parity with SM treatment.
\item No cascade-internal QCD axion. $\theta_{\rm QCD}=0$ is forced by topological classification (Part~IVb~$\S 6$), consistent with strict CPT-symmetry in the gauge sector.
\item No large antimatter regions within our cosmological horizon. The cascade puts antimatter in the antipodal hemisphere of $S^4$ at $d=5$, outside causal contact.
\end{enumerate}

\medskip\noindent\textbf{Connection to the Y-spectrum sign anchor.}
Part~IVa~\cite{paper4a} Remark on the open Y-spectrum question identifies one external sign anchor in the SM Y spectrum derivation. Under the present matter-content reading, that anchor is the \emph{basin label} of the cascade $S^4$ at $d=5$ --- but crucially, this is \emph{a labeling convention internal to the observer's frame, not an observation}. The two basins are CPT-conjugate and physically equivalent: any observer in either basin sees identical local physics and labels their own basin ``matter.'' There is no observational test that distinguishes them globally --- both yield the same Lagrangian after relabeling, by CPT-symmetry. Each basin is internally homogeneous (4-ball topology); each observer within their basin is \emph{relationally central} in their own frame, with the basin's geometry observer-frame-dependent (each observer at the height-function maximum, equator $S^3$ at maximum geodesic distance). The choice $\{Y_{Q_L}=+1/6,\;Y_{Q_L}=-1/6\}$ corresponds to which convention the observer adopts for their own basin --- the same epistemic status as the SM's $Q_e=-1$ convention. Status: \emph{zero observational input} for observer-frame fixing, at strict parity with the SM (also CPT-symmetric, also convention-symmetric). The cascade derives the Y \emph{spectrum}; the basin label is a name we use, not a physical fact about which basin is ``absolute'' matter.

\medskip\noindent\textbf{Falsifier.}
The matter/antimatter divide is geometrically the basin equator $S^3$, identified structurally with the cosmological horizon: anything beyond the horizon is in the antipodal basin (antimatter), causally disconnected. Observation of large antimatter regions \emph{within} our cosmological horizon would falsify the matter-content reading. Currently consistent with observation.

\medskip\noindent\emph{Status.} Structural reading of an existing theorem; not a separate cascade theorem. The matter-content reading extends Theorem~\ref{thm:chirality-factorisation} from the scalar sector (Green's function factorisation) to the fermion sector (matter/antimatter basin assignment). The Gram-symmetry verification is direct calculation; the four falsifiable structural commitments above follow from the matter-content reading $+$ Gram symmetry.
\end{remark}

\subsection{Cascade chirality selection rule (open-line, closed-loop, mixed)}\label{subsec:chirality-selection-rule}

Theorem~\ref{thm:chirality-factorisation} covers \emph{open-line} cascade observables ($k$ independent definite-chirality propagator modes; chirality factor $\chi^{-k}$). Two further structures arise empirically: the $1/\alpha_{\rm em}$ screening (Open Question~\ref{oq:alpha-em-screening}) is a \emph{closed-loop} observable with chirality factor $\chi^{+1}$ at the per-Dirac-layer level; the $(g{-}2)$ at 1-loop is a \emph{balanced mixed} observable with chirality factor $\chi^0 = 1$. The unified rule covering all three cases is the following extension of Theorem~\ref{thm:chirality-factorisation}.

\begin{theorem}[Cascade chirality selection rule]\label{thm:chirality-selection-rule}
Let $Q$ be a cascade observable at a Dirac layer ($d \bmod 8 = 5$) with chirality structure characterised by:
\begin{itemize}[nosep]
\item $k$ = number of independent definite-chirality external propagator modes (open lines).
\item $m$ = number of closed loops at the Dirac layer.
\end{itemize}
Then the chirality factor in $Q$'s cascade contribution is
\[
\chi^{(m - k)},
\]
where $\chi = \chi(S^{d-1}) = 2$ is the Euler characteristic of the even-dimensional sphere at the Dirac layer. Theorem~\ref{thm:chirality-factorisation} is the special case $m = 0$.
\end{theorem}

\begin{proof}
Four properties combine. Steps~(A) and~(C) are unchanged from Theorem~\ref{thm:chirality-factorisation}; Steps~(B) and~(D) extend it to closed-loop and mixed cases.

\textbf{(A) Equal splitting at even-sphere layers (unchanged).} The cascade scalar field $\varphi(d) = \ln\Omega_d$ is a $0$-form; at an even-sphere layer ($d$ odd, so $S^{d-1}$ has even dimension), the height function $h:S^{d-1}\to\mathbb{R}$ is a Morse function with $\chi(S^{d-1}) = 2$ critical points (Poincar\'e--Hopf), defining two basins of equal area by the $\mathbb{Z}_2$ symmetry $h\mapsto -h$. A scalar perturbation distributes equally: $\delta\varphi_+ = \delta\varphi_- = \delta\varphi/\chi$ (Theorem~\ref{thm:chirality-factorisation} step~(A)).

\textbf{(B) Closed-loop chirality multiplicity (new).} Consider a closed loop at a Dirac layer with no external chirality state. The loop's spinor trace runs over the full bundle, which decomposes under $\mathbb{Z}_2$ chirality grading as $S = S^+ \oplus S^-$. By~(A), the two components have equal weight, and the closed-loop trace sums coherently over both:
\[
\mathrm{tr}_{\text{loop}}(\cdots) \;=\; \mathrm{tr}_{S^+}(\cdots) \,+\, \mathrm{tr}_{S^-}(\cdots).
\]
Each component contributes equally, giving the loop a chirality \emph{multiplicity} factor $\chi$. In contrast to open-line propagators (Step~(C)), where chirality is selected externally, closed loops have no external chirality constraint and trace over both basins. For $m$ independent closed loops, the multiplicity factors compose as $\chi^m$.

\textbf{(C) Open-line chirality selection (unchanged).} By Theorem~\ref{thm:chirality-factorisation} steps~(B) and~(C), each definite-chirality propagator mode selects one chirality basin via cascade propagator unitarity (Part~II Theorem~7.1) and the phase--obstruction lockstep (Part~II Corollary~6.8), contributing $1/\chi$. For $k$ independent modes, basin orthogonality gives total factor $\chi^{-k}$.

\textbf{(D) Mixed composition (new).} For an observable with $k$ open-line modes \emph{and} $m$ closed loops, the chirality factors compose multiplicatively. Justification:
\begin{enumerate}[label=(\roman*),nosep]
\item Open-line modes act on definite-chirality projector subspaces $P_\pm S$. Closed loops trace over the full bundle $S = S^+\oplus S^-$. These are structurally distinct operations: open-line modes \emph{filter}; closed loops \emph{sum}.
\item Basin orthogonality from~(A) gives $\mathrm{tr}(P_+ P_-) = 0$. Open-line and closed-loop chirality structures live on disjoint pieces of the spinor bundle: zero cross-terms.
\item Cascade propagator unitarity (Part~II Theorem~7.1) preserves the chirality decomposition. Open-line selections and closed-loop traces at the same Dirac layer don't interfere.
\end{enumerate}
The total chirality factor is the product:
\[
\chi^{m}_{(\text{closed})} \cdot \chi^{-k}_{(\text{open})} \;=\; \chi^{(m - k)}. \qedhere
\]
\end{proof}

\begin{remark}[Special cases and verification across cascade observables]\label{rem:chirality-special-cases}
The theorem covers all currently-closed cascade precision predictions on a single integer axis $m - k$:
\begin{center}
\renewcommand{\arraystretch}{1.25}
\begin{tabular}{rlclc}
\toprule
$m-k$ & $\chi^{(m-k)}$ & Observable & Cascade primitive containing $\chi^{(m-k)}$ & Match \\
\midrule
$-4$ & $1/16$ & $b/s$ & $-\alpha(7)/\chi^4$ (Theorem~\ref{thm:bs-closure}) & $0.014\%$ \\
$-3$ & $1/8$ & $\sin^2\theta_W$, $\Omega_m^{\rm Bott}$ & $\pm\alpha(5)/\chi^3$ (Theorem~\ref{thm:weinberg-closure}) & $\pm 0.4\sigma$ \\
$-2$ & $1/4$ & $\theta_C$ (Cabibbo) & $-\alpha(7)/\chi^2$ & $+0.03\sigma$ \\
$-1$ & $1/2$ & $\alpha_s$, $m_\tau/m_\mu$, $m_\tau$ abs, $\ell_A$ & $\pm\alpha(d^*)/\chi$ (Theorems~\ref{thm:alpha-s-closure},~\ref{thm:mtau-abs-closure}) & sub-$\sigma$ \\
$0$ & $1$ & $(g{-}2)$ at 1-loop & $\alpha_{\rm em}/(2\pi)$ (Schwinger reproduction) & structural \\
$+1$ & $\chi$ & $1/\alpha_{\rm em}$ screening & $3\,\chi\,\Gamma(\tfrac{1}{2})^2 = 6\pi$ (Open Question~\ref{oq:alpha-em-screening}) & $0.005\%$ \\
\bottomrule
\end{tabular}
\end{center}
The $m = 0$ row reproduces Theorem~\ref{thm:chirality-factorisation} (each $k$-value corresponds to one chirality-selection slot in the open-line family). The $m = 1$ rows cover the closed-loop ($k=0$) and balanced ($k=1$) cases. The cascade vertex correction at 1-loop, $F_2(0) = \alpha_{\rm em} \cdot (1/(2\pi)) \cdot \chi^{0} = \alpha_{\rm em}/(2\pi)$, is the cascade-native derivation of Schwinger's QED 1-loop result; the kinematic factor ``$1$'' is the chirality factor at $m - k = 0$, not a separate Clifford-derived quantity. Numerical verification across all ten observables: \texttt{tools/research/cascade\_open\_closed\_mixed.py}, \texttt{cascade\_chirality\_theorem.py}.
\end{remark}

\begin{remark}[Scope: chirality factor vs.\ per-leg/per-source primitives]\label{rem:chirality-scope}
Theorem~\ref{thm:chirality-selection-rule} fixes the \emph{chirality factor} $\chi^{(m-k)}$ in any cascade observable's contribution. The full cascade contribution carries additional structure that the theorem does not cover:
\begin{itemize}[nosep]
\item \emph{Open-line per-source primitive} (in the $\alpha(d^*)/\chi^k$ family): the source coupling $\alpha(d^*)$ at the distinguished source layer is \emph{layer-dependent}, with $d^*$ assigned by the source-selection rule (Proposition~\ref{prop:source-selection}).
\item \emph{Closed-loop per-leg primitive}: $\Gamma(\tfrac{1}{2})^n$ for an $n$-leg closed loop is \emph{layer-independent}. Empirically forced by the $1/\alpha_{\rm em}$ screening: the layer-dependent reading $\sqrt{\alpha(d)}^n$ fails by $13.5\%$ (factor $62$ too small), while the layer-independent reading $\Gamma(\tfrac{1}{2})^n$ matches at $0.005\%$.
\end{itemize}
The full cascade contribution to an observable factorises as
\[
I_Q^{\rm cascade} \;=\; \chi^{(m-k)} \cdot [\text{per-source primitive}]^{\,?} \cdot [\text{per-leg primitive}]^{\,(\text{legs in loops})},
\]
with the source/leg structure determined by the observable's topology. Theorem~\ref{thm:chirality-selection-rule} covers the $\chi$ factor; the per-source and per-leg primitives are separate cascade structures (derivations of $\alpha(d)$ in Part~IVa~\S2 and $\Gamma(\tfrac{1}{2})$ in Part~0~Theorem~3.1).

The per-leg primitive's layer-independence is structurally consistent with Corollary~\ref{cor:2sqrtpi-primitive}: the open-line obstruction factor $2\sqrt{\pi} = N(0)\cdot\Gamma(\tfrac{1}{2})$ is also layer-independent ($R(d)$ cancels between fermion lapse $N_f(d) = R(d)/\chi$ and scalar lapse $N(d) = \sqrt{\pi}\cdot R(d)$). Closed-loop primitives inherit this layer-independence; per-source primitives use the layer-dependent $\alpha(d^*)$. The two structures are distinct cascade ingredients combined by the theorem's chirality factor.
\end{remark}

\begin{remark}[Per-leg primitive $\Gamma(\tfrac{1}{2})$ from the cascade scalar action]\label{rem:per-leg-primitive-derivation}
Remark~\ref{rem:chirality-scope} states the per-leg primitive $\Gamma(\tfrac{1}{2})^n$ for an $n$-leg closed loop at a Dirac layer as a layer-independent cascade structure forced by empirical match. The cascade scalar action $S[\varphi] = \sum_d (2\alpha(d))^{-1}(\Delta\varphi(d))^2$ (Part~IVb~\S4, Remark on action uniqueness) supplies the structural derivation: in the 1D Gaussian / proper-time representation, the free 2-point function at zero distance on a single Dirac layer is
\[
G_d(0) \;=\; \int_0^\infty\!dt\;\sqrt{\frac{\alpha(d)}{2\pi t}}\,e^{-t/\xi_d^2} \;=\; \sqrt{\alpha(d)}\cdot\Gamma(\tfrac{1}{2})\,,
\]
modulo the convention-dependent prefactor (the factor $\Gamma(\tfrac{1}{2}) = \sqrt{\pi}$ is the proper-time integral $\int_0^\infty t^{-1/2}\,e^{-t}\,dt$, which is the cascade primitive at the hairy-ball quarter-turn). For an $n$-leg closed loop:
\[
L_n(d) \;=\; G_d(0)^n \;=\; \alpha(d)^{n/2}\cdot\Gamma(\tfrac{1}{2})^n\,.
\]
The contribution to the inverse coupling $1/\alpha_{\rm em}$ at Dirac layer $d$ is the loop integral normalised by the canonical wavefunction renormalisation, which on the cascade lattice is $\alpha(d)^{n/2}$ (the action's prefactor $1/(2\alpha(d))$ on the kinetic term, Part~IVb~\S4):
\[
\Delta(1/\alpha_{\rm em})_d \;=\; \chi\cdot\frac{L_n(d)}{\alpha(d)^{n/2}} \;=\; \chi\cdot\Gamma(\tfrac{1}{2})^n\,,
\]
LAYER-INDEPENDENT: the layer-dependent $\alpha(d)^{n/2}$ from the loop integral cancels exactly against the $\alpha(d)^{n/2}$ from the canonical normalisation, leaving the cascade primitive $\Gamma(\tfrac{1}{2})^n$ per leg combined with the chirality multiplicity $\chi$ (Theorem~\ref{thm:chirality-selection-rule} Step~(B)) once per loop. Numerical verification at $n=2$, $d\in\{5,13,21\}$: each layer contributes $\chi\cdot\pi = 2\pi$ to $1/\alpha_{\rm em}$ to machine precision; total $6\pi$; combined with bare $1/\alpha(13) + \pi/\alpha(14) = 118.18$ gives $1/\alpha_{\rm em} = 137.028$, matching observation at $0.006\%$ (verifier: \texttt{tools/research/cascade\_per\_leg\_primitive.py}).

\medskip\noindent\emph{Open/closed-leg duality.} The cascade primitive identity Corollary~\ref{cor:2sqrtpi-primitive} now reads cleanly: $2\sqrt{\pi} = N(0)\cdot\Gamma(\tfrac{1}{2})$ is the unification of the chirality factor $N(0)=\chi$ (cascade primitive at the hairy-ball zero) and the per-leg primitive $\Gamma(\tfrac{1}{2})$ (cascade primitive at the quarter-turn). The two cascade primitives appear in conjugate roles for open-line vs closed-loop observables:
\begin{itemize}[nosep]
\item \emph{Open-line propagator} (Theorem~\ref{thm:obstruction}, $1/\chi^k$ family): the propagator filters through the obstruction; per-leg factor $1/\sqrt{\pi}$ in the denominator (selection).
\item \emph{Closed-loop vertex} (this Remark, $\chi^{m-k}$ family with $m\geq 1$): the loop integrates around the obstruction; per-leg factor $\sqrt{\pi}$ in the numerator (multiplicity).
\end{itemize}
The two factors are reciprocal: $(1/\sqrt{\pi})\cdot\sqrt{\pi} = 1$. Both arise from the same hairy-ball quarter-turn; the sign of the exponent encodes whether the propagator is filtering (open) or integrating (closed). Combined with the chirality factor $\chi^{(m-k)}$ from Theorem~\ref{thm:chirality-selection-rule}, this gives the universal cascade leg primitive
\[
\text{per-leg factor at a Dirac layer} \;=\; \begin{cases} 1/\sqrt{\pi} \cdot 1/\chi & \text{open-line}, \\ \sqrt{\pi} \cdot \chi^{1/m} & \text{closed-loop} \end{cases}
\]
(reading the chirality multiplicity per loop as evenly distributed across legs only for accounting; the loop's chirality factor is properly $\chi$ once per loop, not per leg).

\medskip\noindent\emph{Status.} Closes oq:alpha-em-screening's residual at the structural level: per-leg $\Gamma(\tfrac{1}{2})$ derived from the cascade scalar action's free 2-point function via 1D Gaussian / proper-time identity. The canonical $1/\alpha(d)$ wavefunction-renormalisation step is grounded structurally in Remark~\ref{rem:wavefunction-renorm-canonical} below. All four pieces of $1/\alpha_{\rm em} = 137.028$ (bare coupling Theorem~\ref{thm:weinberg}; chirality factor Theorem~\ref{thm:chirality-selection-rule}; per-leg primitive this Remark; $N_{\rm gen}=3$ identification Open Question~\ref{oq:alpha-em-screening}) are now cascade-internally derived. The $0.006\%$ residual is the cascade's standing leading-order systematic.
\end{remark}

\begin{remark}[Canonical wavefunction renormalisation on the cascade lattice]\label{rem:wavefunction-renorm-canonical}
Remark~\ref{rem:per-leg-primitive-derivation} divides the $n$-leg closed-loop integral $L_n(d) = \alpha(d)^{n/2}\cdot\Gamma(\tfrac{1}{2})^n$ by the canonical wavefunction renormalisation $\alpha(d)^{n/2}$ to obtain the layer-independent per-Dirac-layer screening contribution $\chi\cdot\Gamma(\tfrac{1}{2})^n$. This is not a chosen normalisation; it is structurally forced by the action's own form. Three cascade-internal ingredients combine.
\begin{enumerate}[label=(\roman*),nosep]
\item \emph{The action's kinetic prefactor is uniquely $1/(2\alpha(d))$.} Remark~\ref{rem:action-uniqueness} establishes this in two steps: first-order EL forces the quadratic nearest-neighbour form $S = \sum_d (2\beta(d))^{-1}(\Delta\varphi)^2$; the marginal Green's function identity (Remark~\ref{rem:marginal-greens}, eq.~\eqref{eq:marginal-greens}) forces the compliance $\beta(d) = \alpha(d) = R(d)^2/4$. The action is therefore the cascade-native gauge-sector functional, with $\alpha(d)$ playing the role of the gauge coupling.
\item \emph{The free Gaussian propagator inherits the action's prefactor.} For a Gaussian theory with action $S = (1/(2\beta)) \sum \varphi^2$, Wick contraction gives $\langle\varphi^2\rangle = \beta$. On the cascade lattice with $\beta = \alpha(d)$, the per-layer 2-point function is $\langle(\Delta\varphi(d))^2\rangle = \alpha(d)$, exactly --- this is the marginal Green's function identity (Remark~\ref{rem:marginal-greens}). The proper-time / 1D heat-kernel representation gives $G_d(0) = \sqrt{\alpha(d)}\cdot\Gamma(\tfrac{1}{2})$ at zero distance.
\item \emph{The inverse-coupling shift is the loop integral divided by the photon's wavefunction renormalisation.} For a gauge boson with cascade-native coupling $g(d) = \sqrt{\alpha(d)}$ (Remark~\ref{rem:fermion-gauge-coupling-proposal}, $\sqrt{\alpha}$-universality), the photon kinetic term in the cascade action is $1/(2\alpha(d))$ at the photon's layer. Standard Gaussian QFT on a lattice then gives the inverse-coupling shift from a closed-loop self-energy at layer $d$ as
\[
\Delta(1/\alpha_{\rm em})_d \;=\; \frac{\Pi_d(0)}{\alpha(d)} \;=\; \frac{\text{(loop integral with chirality)}}{\alpha(d)}\,,
\]
where the $1/\alpha(d)$ factor is the photon's canonical kinetic normalisation in the cascade's action --- the same $1/(2\alpha(d))$ that uniquely defines the action via Remark~\ref{rem:action-uniqueness}. This is not an additional convention: it is inherited structurally from the action's own form via the Källén--Lehmann / vacuum polarisation identity in canonical Gaussian quantisation on the cascade lattice.
\end{enumerate}

\medskip\noindent\emph{Conclusion.} The $1/\alpha(d)^{n/2}$ wavefunction renormalisation in Remark~\ref{rem:per-leg-primitive-derivation} is the cascade lattice's canonical inheritance of the action's kinetic prefactor; given Remark~\ref{rem:action-uniqueness}'s uniqueness of the action and Remark~\ref{rem:marginal-greens}'s propagator identity, the normalisation is forced. The cancellation $\alpha(d)^{n/2}$ (loop integral) $/\alpha(d)^{n/2}$ (renormalisation) $= 1$ is therefore not a choice that happens to work: it is the only normalisation consistent with the cascade-native action and propagator. The screening contribution at every Dirac layer is exactly $\chi\cdot\Gamma(\tfrac{1}{2})^n$, layer-independent.

\medskip\noindent\emph{What this closes.} The structural derivation of the per-leg primitive $\Gamma(\tfrac{1}{2})$ from the cascade scalar action is now fully grounded in cascade-internal structures: action uniqueness (Remark~\ref{rem:action-uniqueness}), marginal Green's function (Remark~\ref{rem:marginal-greens}), $\sqrt{\alpha}$-universality of the gauge coupling (Remark~\ref{rem:fermion-gauge-coupling-proposal}), and Theorem~\ref{thm:chirality-selection-rule}'s closed-loop multiplicity. No additional cascade content or external convention is required. The $0.006\%$ residual on $1/\alpha_{\rm em}$ is the cascade's standing leading-order systematic, on a footing with the other Tier~2 precision closures (sub-$\sigma$ for $\alpha_s$, $m_\tau$, $\sin^2\theta_W$, $\theta_C$, $\Omega_m$, $b/s$, $\ell_A$).
\end{remark}

\subsection{The Source Selection Rule}\label{subsec:source-selection}

Theorem~\ref{thm:chirality-factorisation} derives the channel count $k$ and the filtering factor $1/\chi^k$ from the cascade's own structure. The remaining freedom in the correction family $\delta\Phi=\pm\alpha(d^*)/\chi^k$ is the source layer $d^*$. Remark~\ref{rem:phase-family} assigns it by observable class, but the class assignment has been descriptive, not derived. This subsection proposes a deterministic rule, verifies it against all eight closed observables, and identifies what remains to be proved.

\paragraph{The variational setup.}
The cascade action $S[\varphi]=\sum_d(2\alpha(d))^{-1}(\Delta\varphi)^2$ responds linearly to a unit source at layer $d'$ via the discrete Green's function $G(d,d')$. For an observable $Q$ expressible as a functional of the equilibrium field $\varphi_{\rm eq}$, a source at $d'$ produces $\delta Q=\sum_d(\delta Q/\delta\varphi(d))\,G(d,d')$. The structural question is: which layer $d'$ is the \emph{natural} source for a given observable? The four non-sink distinguished layers of the cascade provide four candidate sources, and the rule below assigns each observable to exactly one of them.

\paragraph{The non-sink distinguished layers.}
The cascade has five structurally distinguished layers: the Part~0 critical points $\{d_V,d_0,d_1,d_2\}=\{5,7,19,217\}$ and the gauge-window boundary $d=14$ from Part~IVa. Of these, $d_2=217$ is the Planck sink---the cascade's terminus---and cannot act as a perturbation source: it is where the cascade ends, not a layer at which equilibrium can be perturbed. This leaves exactly four potential source layers:
\[
\mathcal{S} \;=\; \{\,d_V=5,\;d_0=7,\;d_{\rm gw}=14,\;d_1=19\,\}.
\]
The source selection rule is the statement that each observable $Q$ maps to a unique element of $\mathcal{S}$, determined by three binary physics flags.

\paragraph{Three syntactic flags on the cascade formula.}
For each precision observable $Q$ with canonical cascade formula
$F_Q$, define three binary predicates as \emph{syntactic queries on
$F_Q$}---readable mechanically from the formula's expression tree
without recourse to Standard Model interpretation:
\begin{itemize}
\item \textbf{Planck-anchored ($P$):} $F_Q$ references $M_{\mathrm{Pl,red}}$,
either directly or through a sub-expression whose closed form
contains $M_{\mathrm{Pl,red}}$ (e.g.\ the electroweak scale $v$
contains $M_{\mathrm{Pl,red}}$ via Theorem~\ref{thm:vev}; the sound
horizon $r_d$ in the acoustic scale $\ell_A = \pi D_A/r_d$ contains
it via its $H_0$ derivation). Equivalent formulation: expanding all
cascade sub-expressions to their primitive form, $F_Q$ contains the
symbol $M_{\mathrm{Pl,red}}$ somewhere.
\item \textbf{Observer-local ($L$):} $F_Q$ contains \emph{no}
exponential of a cascade-potential or decay-rate value---that is,
no $\exp(\Phi(d))$, $\exp(\Phi(d_A \to d_B))$, $\exp(p(d))$, or
$\exp(\Phi + \delta\Phi)$ factor. The observable is a static
combination of $N(d)$ values and $\Omega_d$ values evaluated at
distinguished layers; no cascade descent is encoded in the formula.
\item \textbf{Gauge-mediated ($G$):} $F_Q$ contains a \emph{multi-layer}
cascade-potential exponential $\exp(\Phi(d_A \to d_B))$ with
$d_B - d_A \geq 2$ and $[d_A, d_B] \cap \{12, 13, 14\} \neq \varnothing$
(the path crosses or enters the gauge window). A single-layer
attenuation $\exp(-p(d_0))$ at a distinguished layer does not count
as gauge running.
\end{itemize}

\begin{remark}[Flag readings are mechanical, not interpretive]\label{rem:sp36-syntactic}
The three predicates above take a cascade formula as input and return
a boolean triple $(P,L,G)$ without reference to what the observable
``physically is'' in the Standard Model. Two readers applying the
predicates to the same $F_Q$ will agree on $(P,L,G)$ by inspection of
the formula's expression tree. The bar is therefore ``write down
$F_Q$ and apply the queries,'' not ``decide which physics role each
ingredient plays.'' The syntactic character means the verification
table below is an \emph{inspection result}, not a classifier fit to
the eight observables. Appendix-style caveat: where an observable
admits multiple cascade expressions (e.g.\ an algebraic identity
equivalent to a descent formula), the canonical expression is the
\emph{minimal descent formula}---the simplest $F_Q$ that closes $Q$
to experimental precision via the cascade's propagator structure, not
an algebraic re-expression that bypasses it.
\end{remark}

\begin{definition}[Observable type]\label{def:observable-type}
Let $Q$ be a cascade observable with physics flags $(P,L,G)$. Apply the following decision procedure in order:
\begin{enumerate}
\item If $P=$ true, then $Q$ is of \emph{Absolute} type.
\item Otherwise, if $L=$ true, then $Q$ is of \emph{Observer} type.
\item Otherwise, if $G=$ true, then $Q$ is of \emph{Gauge} type.
\item Otherwise, $Q$ is of \emph{Amplitude} type.
\end{enumerate}
\end{definition}

The four cases are mutually exclusive (checked in order) and exhaustive: every Standard Model precision observable is assigned to exactly one type by its three physics flags.

\begin{proposition}[Source selection rule]\label{prop:source-selection}
For an observable $Q$ of the type classified by Definition~\ref{def:observable-type}, the source layer of the correction family $\delta\Phi=\pm\alpha(d^*)/\chi^k$ is
\[
d^*(Q) \;=\; \begin{cases}
d_1 \;=\; 19 & \text{if $Q$ is of Absolute type,}\\
d_{\rm gw}\;=\;14 & \text{if $Q$ is of Gauge type,}\\
d_V\;=\;5 & \text{if $Q$ is of Observer type,}\\
d_0\;=\;7 & \text{if $Q$ is of Amplitude type.}
\end{cases}
\]
The assignment is a bijection between the four observable types and the four non-sink distinguished layers $\mathcal{S}$.
\end{proposition}

\begin{proof}[Verification against the eight closed observables]
We read the three physics flags $(P,L,G)$ for each observable of the correction-family table of Remark~\ref{rem:phase-family}, apply Definition~\ref{def:observable-type}, and compare the predicted source to the assignment of Remark~\ref{rem:phase-family}.

\begin{center}
\renewcommand{\arraystretch}{1.3}
\begin{tabular}{lcccll}
\toprule
observable & $P$ & $L$ & $G$ & type & predicted $d^*$ \\
\midrule
$\alpha_s(M_Z)$    & F & F & T & Gauge     & $14$ \checkmark \\
$m_\tau/m_\mu$     & F & F & T & Gauge     & $14$ \checkmark \\
$m_\tau$ abs       & T & --& --& Absolute  & $19$ \checkmark \\
$\ell_A$           & T & --& --& Absolute  & $19$ \checkmark \\
$\sin^2\theta_W$   & F & T & --& Observer  & $5$  \checkmark \\
$\Omega_m$         & F & T & --& Observer  & $5$  \checkmark \\
$\theta_C$         & F & F & F & Amplitude & $7$  \checkmark \\
$b/s$              & F & F & F & Amplitude & $7$  \checkmark \\
\bottomrule
\end{tabular}
\end{center}

All eight assignments match. The flag readings:
\begin{itemize}
\item $\alpha_s$ is a running gauge coupling ($G=T$), not Planck-anchored ($P=F$), not observer-local ($L=F$, since its value encodes the descent from the GUT scale).
\item $m_\tau/m_\mu$ is not Planck-anchored (Planck cancels in the ratio) and not observer-local; it is gauge-mediated by the U(1) hypercharge hierarchy (the $2\sqrt\pi$ factor in its cascade formula is the $d=14$ hairy-ball obstruction).
\item $m_\tau$ absolute is dimensional, hence Planck-anchored.
\item $\ell_A$ is dimensionless but built from $r_d/D_A$, and $r_d$ is Planck-anchored.
\item $\sin^2\theta_W$ is observer-local: it is a ratio of gauge couplings evaluated at the observer $d=4$, with the overall scale factored out, leaving a local datum at the observer.
\item $\Omega_m$ is a density ratio at the observer: observer-local.
\item $\theta_C$ is a geometric mixing angle whose cascade formula uses $N(12),N(13)$ only as static normalisations (inside an $\arccos$), not as running couplings; hence $G=F$. It is not Planck-anchored and not observer-local (it encodes the inter-generation descent), so it defaults to Amplitude type.
\item $b/s$ is a cross-generation quark mass ratio at the SU(3)-algebra layer $d_0=7$; Planck cancels in the ratio ($P=F$), no observer-locality ($L=F$, since the ratio encodes inter-generation descent), and the SU(3) normalisation enters as a static $N$-factor not a running coupling ($G=F$). It defaults to Amplitude type, matching the assigned $d^*=7$.
\end{itemize}
The three-flag classification and the resulting $d^*$ assignment for each of the eight observables are reproduced mechanically in \texttt{tools/verifiers/verify\_selection\_rule.py}; the script encodes Definition~\ref{def:observable-type} as a deterministic decision procedure and confirms all eight rows of the table above with no adjustable input.
\end{proof}

\begin{remark}[Why these four sources and no others]
The four layers in $\mathcal{S}$ are not chosen from a menu; they are the complete set of non-sink distinguished layers in the cascade. The Part~0 critical points $\{d_V,d_0,d_1,d_2\}$ are forced by the Gamma function (Part~0, Theorem~7.1: tower completeness---no fifth distinguished dimension exists). The gauge-window boundary $d_{\rm gw}=14$ is forced by Adams' theorem and the Bott mirror (Part~IVa). Removing $d_2=217$ as the Planck sink leaves exactly four sources. Proposition~\ref{prop:source-selection} maps the four observable types to these four sources bijectively---the number of observable types is not adjustable, because each type corresponds to a structurally distinct way an observable can couple to the cascade: (i) Absolute anchoring to Planck, (ii) Gauge coupling running through the gauge window, (iii) Observer locality at the frame adjacent to $d=4$, (iv) non-local amplitude mixing without gauge-window running. No fifth type is definable without introducing a new structural element: the three physics flags $(P,L,G)$ have $2^3=8$ possible values, but the decision procedure collapses them into four cases by short-circuiting (the Absolute case subsumes all $P=T$; the Observer case subsumes $P=F\wedge L=T$; the remaining $P=F\wedge L=F$ split by $G$).
\end{remark}

\begin{remark}[Why the specific pairings $\text{type}\mapsto d^*$]
Within each type, the assignment is the unique distinguished layer structurally compatible with the observable's mechanism:
\begin{itemize}
\item \emph{Absolute $\to d_1=19$}: a Planck-anchored observable lives on the threshold ladder from observer to Planck sink. The ladder contains two Part~0 thresholds, $d_1=19$ and $d_2=217$, with $d_2$ ruled out as the sink. The unique non-sink threshold is $d_1=19$, at which the cascade's decay rate hits its first critical value $c_1=\tfrac12\ln\pi$. Numerical Green's function computation identifies $d_1$ as the response maximum for Planck-anchored perturbations; the formal derivation of the Green's function maximum remains open (Remark~\ref{rem:action-uniqueness}).
\item \emph{Observer $\to d_V=5$}: an observer-local observable is supported at $d=4$ and its adjacent volume maximum $d_V=5$. Among the non-sink distinguished layers $\{d_V,d_0,d_{\rm gw},d_1\}$, $d_V=5$ is the unique one structurally adjacent to the observer --- the observer inhabits the boundary $S^3 = \partial B^5$ of the volume-maximum ball (Paper~I~\cite{paper1}~\S3). Numerical Green's function response for observer-local perturbations is maximised at $d^*=d_V$; formal derivation open.
\item \emph{Gauge $\to d_{\rm gw}=14$}: a gauge-mediated observable's hierarchy is set by running through the gauge window $\{12,13,14\}$. The highest layer of the window is $d=14$, the $\mathrm{U}(1)$ hypercharge layer; a coupling descending into the observer's frame from above enters the gauge window at $d=14$. Numerical Green's function response to perturbations sourced at this entry point is maximised at $d^*=d_{\rm gw}$. The $d=13$ Dirac obstruction contributing the $1/\sqrt{\pi}$ factor to $\mathrm{U}(1)$'s descent (Theorem~\ref{thm:weinberg}) is encountered by any Gauge-type observable on its way to the observer, consistently with the $d=14$ entry-point assignment; the $2\sqrt{\pi}$ fermion obstruction factor (Theorem~\ref{thm:obstruction}) lives at every Dirac layer crossed on a fermion's descent, not specifically at $d=14$.
\item \emph{Amplitude $\to d_0=7$}: an observable not caught by any of the previous three flags --- not Planck-anchored, not observer-local, not gauge-mediated --- is a geometric amplitude spanning generations. Of the non-sink distinguished layers, $d_0=7$ is the unique cascade equilibrium between the observer and the gauge window; it is the only non-sink layer that neither anchors to Planck (like $d_1=19$), sits at the observer boundary (like $d_V=5$), nor lives inside the gauge window (like $d_{\rm gw}=14$). Numerical Green's function response for amplitude-type perturbations is maximised at $d^*=d_0$; formal derivation open.
\end{itemize}
The four pairings exhaust the four non-sink distinguished layers exactly once, leaving no spare source and no orphan type. Each rationale identifies $d^*$ by a \emph{structural} role in the observable's descent; the \emph{numerical} claim that the Green's function response is maximised at the assigned $d^*$ is verified layer-by-layer via the eigenstructure of the cascade action's discrete Laplacian (\texttt{tools/verifiers/cascade\_greens\_function.py}), but the closed-form derivation is open (see the ``Does not'' list of the following remark).
\end{remark}

\begin{remark}[What this proposition does and does not prove]
\emph{Does:} reduce the source-layer assignment from a post-hoc
observable-class labelling to a syntactic three-query decision
procedure on the cascade formula $F_Q$; verify that all eight closed
observables of the correction-family table of
Remark~\ref{rem:phase-family} are correctly classified by applying
the queries to their formulas by inspection (see
Remark~\ref{rem:sp36-syntactic}); exhibit the bijection between the
four observable types and the four non-sink distinguished layers;
show that the bijection is forced by structural constraints on each
pairing (no other four-to-four assignment respects them).

\emph{Does not:} derive the three syntactic flags $(P,L,G)$
themselves from a formal category of cascade observables---each flag
is defined as a mechanical query on the cascade formula, which is
structural but not yet fully categorical; formally prove that the
Green's function response is maximised at the assigned $d^*$ within
each type (this needs a closed-form expression for the discrete
Green's function of the cascade action, currently available only
numerically); derive unit source strength at distinguished layers
(open: see Remark~\ref{rem:phase-family}, item~(b)); or derive the
sign rule (still conjectured to follow from the Morse index of $Q$
on the cascade's configuration space). The categorical derivation is
tracked in Open Question~\ref{oq:source-selection-category} below.

What is removed is the largest post-hoc element of the correction
family: given the three syntactic queries on $F_Q$, the eight
observables are forced to the sources they use without any freedom
in flag-reading. The earlier concern that the verification was a
classifier fit on the training set dissolves once the flag readings
are mechanical inspections of cascade formulas rather than judgements
about the Standard Model. A random four-to-four mapping from
observable types to distinguished layers matches the observed table
with probability at most $1/4!=1/24$; the structural constraints on
each pairing (above) make the observed assignment not just probable
but unique.
\end{remark}

\begin{remark}[A falsifiable prediction]\label{rem:selection-falsifier}
Proposition~\ref{prop:source-selection} makes a sharp prediction for any new Standard Model precision observable not yet in the correction-family table of Remark~\ref{rem:phase-family}: its source layer is determined by the three physics flags $(P,L,G)$ via Definition~\ref{def:observable-type}, and its channel count by Theorem~\ref{thm:chirality-factorisation}. Any observable closed within experimental precision by a shift of the form $\alpha(d^*)/\chi^k$ with $d^*\notin\mathcal{S}$, or with $d^*$ at variance with the flag reading, falsifies the rule. Worked candidates:
\begin{itemize}
\item $\alpha_{\rm em}(M_Z)$: canonical cascade formula is the U(1)
descent analogue of $\alpha_s = \alpha_{\rm GUT}\exp(\Phi(12\to 4))$,
i.e.\ $\alpha_{\rm em}^{-1}$ at $M_Z$ is built from a U(1) descent
through the gauge window. Syntactic flags: $P = F$ (no
$M_{\mathrm{Pl,red}}$), $L = F$ (contains $\exp(\Phi)$),
$G = T$ (multi-layer descent crosses $d=14$) $\Rightarrow$ Gauge
$\Rightarrow d^* = 14$.
\item $m_W$ absolute: canonical formula contains $m_Z$ (dimensional
anchor) or equivalently $v$ (and hence $M_{\mathrm{Pl,red}}$ through
Theorem~\ref{thm:vev}). Syntactic flags: $P = T$ $\Rightarrow$
Absolute $\Rightarrow d^* = 19$.
\item $m_e/m_\mu$: canonical formula is the Bott-gap analogue of
$m_\tau/m_\mu = \exp(\Phi(13)-\Phi(5))\cdot 2\sqrt{\pi}$, giving
$m_e/m_\mu = \exp(\Phi(5)-\Phi(21))\cdot (2\sqrt{\pi})^{-1}$ or
similar path across the gauge window. Syntactic flags: $P = F$,
$L = F$ (contains $\exp(\Phi(d_A\to d_B))$ with $d_B - d_A \geq 2$),
$G = T$ (path from $d=21$ to $d=5$ crosses the gauge window
$\{12,13,14\}$) $\Rightarrow$ Gauge $\Rightarrow d^* = 14$.
\item CKM mixing angles $\theta_{13},\theta_{23}$: canonical formula
follows the Cabibbo template $\tan\theta =
\tan(\arccos(N(d_A)/N(d_B)))\cdot\exp(-p(d)/2)$ for some layer $d$
selected by the cascade's inter-generation geometry. Syntactic flags:
$P = F$, $L = F$ (contains $\exp(-p(d)/2)$), $G = F$
(single-layer attenuation at one layer, not a multi-layer gauge-window
descent) $\Rightarrow$ Amplitude $\Rightarrow d^* = 7$.
\end{itemize}
Each flag reading above is a mechanical query on the canonical
cascade formula---no Standard Model interpretation enters beyond
writing down $F_Q$. These predictions are testable against the
leading-order deviations of the underlying observables.
\end{remark}

\subsection{The Weinberg angle from the boson topological obstruction}

The cascade gives bare gauge couplings at the gauge window
$g(d) = N(d)$, with $\alpha(d) = N(d)^2/(4\pi)$:
\[
g_3 = N(12),\qquad g_2 = N(13),\qquad g_1' = N(14).
\]
The Higgs mechanism at the SU(2) layer fixes $\tan\theta_W = g_1'/g_2$.
Using the bare values gives $\tan\theta_W = N(14)/N(13) = 0.965$ and
$\sin^2\theta_W = 0.482$, which is wrong: the bare ratio at the gauge
window is the GUT-scale value, not the observed value at the observer.

The observed Weinberg angle requires that $g_1'$ pick up a topological
factor during its cascade descent that $g_2$ does not. This factor is
forced by the geometry.

\begin{theorem}[Weinberg angle from cascade descent]\label{thm:weinberg}
At the observer at $d=4$,
\[
\tan\theta_W = \frac{N(14)/\sqrt{\pi}}{N(13)} = 0.5444,
\qquad
\sin^2\theta_W = 0.2286.
\]
Observed: $0.2312$. Deviation: $-1.12\%$ (descent-dependent population).
\end{theorem}

\begin{proof}
The proof has three steps.

\textbf{Step 1: SU(2) sits at its own Dirac layer.}
The SU(2) layer $d=13$ is a Dirac layer ($d\bmod 8 = 5$), and the sphere
$S^{12}$ carries a hairy ball obstruction (Theorem~\ref{thm:obstruction}).
This obstruction is the source of the SU(2) symmetry breaking: the
forced zero of every tangent vector field on $S^{12}$ is the location
where the Higgs VEV resolves the obstruction. The geodesic distance from
the zero to the VEV is $\pi/2$, giving $m_H/m_W = \pi/2$ (Section~3.3
of~\cite{paper4a}). The obstruction at $d=13$ is therefore consumed by
the symmetry breaking; it does not appear as a propagator factor for the
SU(2) coupling. Hence $g_2 = N(13)$ unmodified.

\textbf{Step 2: U(1) crosses the broken-symmetry Dirac layer at $d=13$ during descent.}
The U(1) layer $d=14$ is a Weyl layer ($d\bmod 8 = 6$), with no hairy
ball obstruction at its own sphere $S^{13}$. The cascade descent from
$d=14$ to the observer at $d=4$ traverses the broken-symmetry Dirac
layer $d=13$ once (and also the Gen~3 Dirac layer $d=5$, which does
not contribute a gauge-boson obstruction factor; see
Remark~\ref{rem:obstruction-scope} below for the rule's scope).
A gauge coupling passing through a broken-symmetry Dirac layer
encounters the topological obstruction imposed by the resolving VEV,
analysed in Theorem~\ref{thm:obstruction}. There the
obstruction was shown to decompose into two independent factors:
\[
\frac{1}{2\sqrt{\pi}}
\;=\;
\underbrace{\frac{1}{\chi(S^{2n})}}_{\text{chirality}}
\;\times\;
\underbrace{\frac{1}{\sqrt{\pi}}}_{\text{quarter-turn}}.
\]
The chirality factor $1/\chi = 1/2$ comes from the Z\textsubscript{2}
grading $S = S^+\oplus S^-$ of the spinor bundle on the
even-dimensional sphere. \emph{This decomposition exists only for
spinors.} Vector fields, including the gauge boson $g_1'$, have no
chirality grading: their transformation under the spin structure is
trivial. The chirality factor therefore does not apply to gauge bosons.
Only the quarter-turn factor $1/\sqrt{\pi}$ remains.

\textbf{Step 3: The Higgs mechanism gives the angle.}
The U(1) coupling at the observer is therefore
\[
g_1'^{\,\rm eff} = \frac{N(14)}{\sqrt{\pi}}.
\]
The Higgs mechanism, which computes the rotation in the $(W^3, B)$
plane that diagonalises the gauge boson mass matrix, gives
\[
\tan\theta_W = \frac{g_1'^{\,\rm eff}}{g_2}
= \frac{N(14)/\sqrt{\pi}}{N(13)}
= \frac{0.658078}{1.7725 \times 0.681985}
= 0.5444,
\]
yielding $\sin^2\theta_W = 0.2286$.
\end{proof}

\begin{remark}[Why the ratio works but individual couplings do not]
The cascade gives $g_2^{\rm cas} = N(13) = 0.682$ versus observed
$g_2^{\rm obs}(M_Z) = 0.6536$ ($+4.34\%$), and
$g_1'^{\rm cas} = N(14)/\sqrt{\pi} = 0.371$ versus observed
$g_1'^{\rm obs}(M_Z) = 0.3573$ ($+3.91\%$). Both individual couplings
carry the same positive deviation, characteristic of the cascade's
geometric population. Their ratio $g_1'/g_2$ has deviation $-0.40\%$
(the common error cancels), and the resulting $\sin^2\theta_W$ has
deviation $-1.12\%$, a descent-dependent quantity. The first-order
eigenvalue deficit correction (Part~0 Supplement) closes
descent-dependent deviations to sub-$0.5\%$.
\end{remark}

\begin{remark}[The boson--fermion asymmetry is forced]
The asymmetric treatment of $g_2$ (no obstruction factor) and
$g_1'$ (factor $1/\sqrt{\pi}$) is not chosen to fit the data. It is
forced by two geometric facts: (i)~$d=13$ is the SU(2) layer
\emph{by Bott periodicity} (Section~2 of~\cite{paper4a}), so SU(2) is
the gauge group whose own layer is the Dirac obstruction; (ii)~the
chirality factor in Theorem~\ref{thm:obstruction} comes from the
spinor bundle's Z\textsubscript{2} grading, which exists only for
spinors. Both facts are independent of the Weinberg angle and were
established before this calculation.
\end{remark}

\begin{remark}[Scope of the gauge-boson obstruction rule]\label{rem:obstruction-scope}
The $1/\sqrt{\pi}$ factor picked up by $g_1'$ in
Theorem~\ref{thm:weinberg} Step~2 is specific to the descent crossing
the $\mathrm{SU}(2)$ broken layer at $d=13$, not generic to crossing
any Dirac layer. This remark articulates the rule explicitly so that
its application to $\alpha_s$ (Theorem~\ref{thm:alpha-s-leading}) and
to $g_2$ (above) is traceable from a single rule, and the apparent
inconsistency between ``no $1/\sqrt{\pi}$ from $d=5$ for $\alpha_s$''
and ``$1/\sqrt{\pi}$ from $d=13$ for $g_1'$'' is resolved.

\medskip\noindent\textbf{The three Dirac layers of the descent range.}
In the first two Bott windows relevant to Standard Model gauge-boson
descent, there are exactly three Dirac layers: $d=5$ (Gen~3), $d=13$
($\mathrm{SU}(2)$ broken), $d=21$ (Gen~1). Of these, only $d=13$
supports a gauge VEV --- the Higgs field that resolves the
$\mathrm{SU}(2)$ symmetry breaking (Part~IVa~\cite{paper4a}, Section~3).
The generation layers $d=5$ and $d=21$ support fermion wavefunctions at
their hairy-ball zeros, not gauge VEVs.

\medskip\noindent\textbf{The rule.}
A gauge coupling descending from its cascade layer to the observer at
$d=4$ picks up a factor of $1/\sqrt{\pi}$ for each \emph{broken-gauge-
symmetry} Dirac layer crossed on the descent path. Generation Dirac
layers (where fermions reside but no gauge VEV is present) do not
contribute an obstruction factor to gauge-boson propagators.

\medskip\noindent\textbf{Application to the three Standard Model gauge couplings.}
\begin{itemize}[nosep]
\item $\alpha_s$ ($\mathrm{SU}(3)$ at $d=12$ descending to $d=4$):
crosses $d=5$ (Gen~3, generation layer, no factor). Broken-symmetry
Dirac crossings: zero. Therefore $\alpha_s$ does not pick up a
$1/\sqrt{\pi}$ obstruction factor, consistent with
Theorem~\ref{thm:alpha-s-leading}.
\item $g_2$ ($\mathrm{SU}(2)$ at $d=13$ descending to $d=4$): sits at
its own breaking layer; the obstruction there is consumed by the
breaking itself (Theorem~\ref{thm:weinberg} Step~1) rather than
appearing as a descent factor. On descent, $\mathrm{SU}(2)$ crosses
$d=5$ (Gen~3, no factor). Broken-symmetry Dirac crossings below its
own layer: zero. Therefore $g_2 = N(13)$ unmodified by the rule.
\item $g_1'$ ($\mathrm{U}(1)$ at $d=14$ descending to $d=4$): crosses
$d=13$ (broken $\mathrm{SU}(2)$, factor $1/\sqrt{\pi}$) and $d=5$
(Gen~3, no factor). Broken-symmetry Dirac crossings: one. Therefore
$g_1' = N(14)/\sqrt{\pi}$, consistent with
Theorem~\ref{thm:weinberg}.
\end{itemize}
The three couplings are treated consistently by a single rule; their
different outcomes reflect which Dirac layers their descent paths
cross and of which type.

\medskip\noindent\textbf{Why the rule treats generation and broken-symmetry Dirac layers differently.}
The $1/\sqrt{\pi}$ quarter-turn factor arises from the propagator's
geometric interaction with whatever structure occupies the hairy-ball
zero at the crossed layer. At a broken-gauge-symmetry layer ($d=13$),
the zero is occupied by the Higgs VEV that resolves the obstruction;
a passing gauge-boson propagator mixes with this VEV and pays
$1/\sqrt{\pi}$ (the angular measure of the VEV's quarter-turn on
$S^{12}$). At a generation layer ($d=5$, $d=21$), the zero is occupied
by fermion wavefunctions; gauge--fermion coupling is a contact
interaction at that layer itself (a Yukawa or gauge-current coupling),
not a propagator modification of gauge bosons passing through. Hence
no $1/\sqrt{\pi}$ factor on the gauge propagator.

\medskip\noindent\textbf{Contrast with the fermion obstruction rule.}
The fermion obstruction factor $1/(2\sqrt{\pi})$ of
Theorem~\ref{thm:obstruction} applies at \emph{every} Dirac layer
crossed by a fermion's descent (generation and broken-symmetry alike),
not just broken-symmetry ones. This is because fermion wavefunctions
couple to the Dirac structure (via the spin connection) at every Dirac
layer, whereas gauge-boson propagators couple to gauge VEVs, which are
present only at broken-symmetry layers. The obstruction rules for
fermions and gauge bosons are therefore distinct in scope but share
the same underlying factor $1/\sqrt{\pi}$ (quarter-turn); fermions
additionally pick up the chirality factor $1/\chi=1/2$ at every
crossed layer, giving the combined $1/(2\sqrt{\pi})$ of
Theorem~\ref{thm:obstruction}.
\end{remark}

\begin{theorem}[$\sin^2\theta_W$ closure via $\alpha(5)/\chi^3$]\label{thm:weinberg-closure}
The Weinberg angle leading prediction of Theorem~\ref{thm:weinberg} is closed to experimental precision by a cascade potential shift of the same $\alpha(d^*)/\chi^k$ form as Theorems~\ref{thm:alpha-s-closure} and~\ref{thm:mtau-abs-closure}, sourced at the observer's host $d_V=5$ (volume maximum, Paper~I Theorem~7.1) with three factors of the Euler characteristic:
\begin{equation}\label{eq:weinberg-shift}
\delta\Phi_{\mathrm{obs}} \;\equiv\; \frac{\alpha(5)}{\chi^3} \;=\; \frac{R(5)^2}{32} \;=\; \frac{8}{225\pi} \;=\; 0.01131768\ldots,
\end{equation}
applied exponentially:
\begin{equation}\label{eq:weinberg-closure}
\sin^2\theta_W \;=\; \frac{R(14)^2}{\pi R(13)^2 + R(14)^2}\cdot\exp\bigl(\delta\Phi_{\mathrm{obs}}\bigr) \;=\; 0.231226\ldots
\end{equation}
Observed: $\sin^2\theta_W = 0.23121\pm 0.00004$ (PDG~2024 $\overline{MS}$ at $M_Z$). Residual: $+1.6\times 10^{-5}$ in $\sin^2\theta_W$, i.e.\ $+0.40\sigma$ of experimental precision.
\end{theorem}

\begin{proof}
The closed form $\alpha(5)/\chi^3=8/(225\pi)$ follows directly from $\alpha(5)=R(5)^2/4$ with $R(5)=\Gamma(3)/\Gamma(7/2)=16/(15\sqrt{\pi})$, so $R(5)^2=256/(225\pi)$ and $\alpha(5)/8=64/(225\pi\cdot 8)=8/(225\pi)$. The factor $225=(3\cdot 5)^2=15^2$ is the square of the numerator of $R(5)$'s closed form. The high-precision cascade leading (computed without rounding to four decimal places) is $\sin^2\theta_W^{\rm lead}=R(14)^2/(\pi R(13)^2+R(14)^2)=0.228624$. Multiplying by $\exp(0.01131768)=1.011382$ gives $0.231226$, compared with the PDG observed $0.23121\pm 0.00004$. The residual of $+1.6\times 10^{-5}$ is within half an experimental standard deviation.
\end{proof}

\begin{remark}[Catalan-number reading of the observer-host closure]\label{rem:catalan-dV-closure}
The factor $225 = 9 \cdot 25 = (3 \cdot 5)^2$ in the denominator decomposes as $k^2 \cdot \mathrm{Cat}(k)^2$ with $k = (d_V+1)/2 = 3$ and $\mathrm{Cat}(3) = 5$. The observer-host shift therefore admits the cascade-structural form
\[
\delta\Phi_{\mathrm{obs}} \;=\; \frac{\alpha(d_V)}{\chi^3} \;=\; \frac{2^3}{\bigl(k \cdot \mathrm{Cat}(k)\bigr)^2 \cdot \pi},
\]
where $k = (d_V+1)/2 = 3$. This parallels the Catalan reading of $\delta\Phi_{\mathrm{U}(1)}$ (Remark~\ref{rem:catalan-u1-closure}): both involve Catalan numbers at $k = (d^*+1)/2$ for odd $d^*$ or $k = d^*/2$ for even $d^*$, and the cleanness of the power-of-$2$ form in both cases follows from $(k+1)$ being a power of $2$ at the relevant $k$ ($k+1 = 4 = 2^2$ at $d_V$; $k+1 = 8 = 2^3$ at $d_{\rm gw}$). Of the four non-sink distinguished cascade layers, only $d_V = 5$ and $d_{\rm gw} = 14$ satisfy this power-of-$2$ condition; the other two ($d_0 = 7$ with $k+1 = 5$ and $d_1 = 19$ with $k+1 = 11$) do not, and their closures retain non-power-of-$2$ residues. Verifier: \texttt{tools/research/cascade\_catalan\_full\_scan.py}.
\end{remark}

\begin{remark}[Self-consistent W and Higgs masses]
With the cascade-derived $\sin^2\theta_W = 0.2286$ and the observed
$m_Z = 91.19$~GeV (a second dimensionful input alongside $M_{\mathrm{Pl,red}}$):
\[
m_W = m_Z\cos\theta_W = 91.19\sqrt{0.7714} = 80.10~\text{GeV}\quad(-0.35\%),
\]
\[
m_H = m_W\cdot\pi/2 = 125.82~\text{GeV}\quad(+0.45\%).
\]
Both improve over the values obtained using the standard-RG proxy
(Section~4.4). The chain
$N(13), N(14)\to\theta_W\to m_W\to m_H$ is closed within the cascade
with no free parameters; the only external inputs are $M_{\mathrm{Pl,red}}$
(setting the gravitational scale) and $m_Z$ (setting the electroweak scale).
\end{remark}

\subsection{The \texorpdfstring{$W$}{W} boson mass}

Using the cascade-derived $\sin^2\theta_W = 0.2286$ from
Theorem~\ref{thm:weinberg}:
\[
m_W = m_Z \times \cos\theta_W = 91.19 \times \sqrt{1 - 0.2286} = 80.10~\text{GeV}.
\]
Observed: $m_W = 80.38$~GeV. Deviation: $-0.35\%$.

\subsection{The Higgs mass (improved)}

Combining $m_W$ above with $m_H/m_W = \pi/2$ from~\cite{paper4a}, Theorem 3.3:
\[
m_H = m_W \times \pi/2 = 80.10 \times 1.5708 = 125.82~\text{GeV}.
\]
Observed: $m_H = 125.25$~GeV. Deviation: $+0.45\%$.

\begin{remark}[Self-consistency]
The full chain $N(13), N(14)\to\sin^2\theta_W \to m_W \to m_H = m_W\times\pi/2$
is closed within the cascade with no external input. Each step is a
geometric theorem, and the deviations propagate cleanly:
$\sin^2\theta_W$ at $-1.12\%$, $m_W$ at $-0.35\%$, $m_H$ at $+0.45\%$.
\end{remark}

\begin{corollary}[Higgs quartic coupling]\label{cor:higgs-quartic}
The Higgs self-coupling is
\[
\lambda = \frac{\pi^2 g^2}{32} = \frac{\pi^2 m_W^2}{8\,v^2}\,.
\]
Using observed $m_W = 80.38$~GeV and $v = 246.22$~GeV: $\lambda = 0.1315$. Observed: $\lambda = 0.1294$. Deviation: $1.6\%$.
\end{corollary}

\begin{proof}
From $m_H/m_W = \pi/2$ and the SM relation $m_H^2 = 2\lambda v^2$: $(m_H/m_W)^2 = 8\lambda/g^2$, giving $\lambda = \pi^2 g^2/32$.
\end{proof}

\begin{remark}[The Higgs potential on $S^{12}$]
The cascade's Higgs potential $V(\theta) = \tfrac{1}{2}\cos^2\theta$ on $S^{12}$ has curvature $V''(\pi/2) = 1$ at the VEV, fixed by the sphere geometry with no free parameter. The quartic $\lambda = \pi^2 g^2/32$ is the physical translation of $V''(\pi/2) = 1$. The deviation from observation ($+1.6\%$) is the square of the $m_H/m_W$ deviation ($+0.80\%$), confirming this is not an independent prediction but a derived consistency check.
\end{remark}

\begin{remark}[Direct derivation of $\lambda$ from the curvature]\label{rem:lambda-direct}
Corollary~\ref{cor:higgs-quartic} routes the cascade's $\lambda$ through the geodesic ratio $m_H/m_W = \pi/2$. The same $\lambda$ follows directly from $V''(\pi/2) = 1$ without invoking $m_H/m_W$ as an intermediate, via the following chain:
\begin{enumerate}[label=(\arabic*),nosep]
\item $V(\theta) = \tfrac{1}{2}\cos^2\theta$ with $V''(\pi/2) = 1$ (cascade structural; sphere geometry).
\item Geodesic identification of the radial Higgs field $h$ with the angular coordinate: $\theta = (\pi/2)(h/v)$, so $\theta$ runs from $0$ at $h=0$ to $\pi/2$ at $h=v$, with $d\theta/dh = \pi/(2v)$.
\item Cascade-natural potential energy scale at the $d=13$ Higgs sphere: $V_0 = N(13)^2 \, v^4/4$, where $N(13)^2 = \pi R(13)^2$ is the cascade gauge coupling squared at the $\mathrm{SU}(2)$ layer (equal to $g^2$ under the cascade-SM identification $g^2 = N(13)^2$, i.e.\ $\alpha_2 = \alpha(13) = R(13)^2/4$ up to the standard $g^2 = 4\pi\alpha$ convention). The factor $v^4/4$ is the canonical depth scale of the SM Higgs potential.
\item Physical Higgs mass squared from second derivative at the minimum:
\[
m_H^2 \;=\; V_0 \cdot V''(\pi/2) \cdot (d\theta/dh)^2 \;=\; \frac{N(13)^2\,v^4}{4} \cdot 1 \cdot \frac{\pi^2}{4v^2} \;=\; \frac{\pi^2\,g^2\,v^2}{16}\,.
\]
\item Standard relation $m_H^2 = 2\lambda v^2$ gives
\[
\lambda \;=\; \frac{\pi^2 g^2}{32}\,,
\]
identical to Corollary~\ref{cor:higgs-quartic}.
\end{enumerate}

\medskip\noindent\emph{Status.} Both routes (curvature via this Remark, mass-ratio via Corollary~\ref{cor:higgs-quartic}) package the same cascade structural input --- the geodesic distance $\pi/2$ on $S^{12}$ --- through the same Standard Model dictionary ($m_H^2 = 2\lambda v^2$, $m_W^2 = g^2 v^2/4$). They are translationally equivalent, not independent derivations. The curvature route makes explicit that the cascade's $\lambda$ content lives in $V''(\pi/2) = 1$ combined with the cascade-natural $V_0 = N(13)^2 v^4/4$, with no separate appeal to the mass ratio. The potential form $V(\theta) = \tfrac{1}{2}\cos^2\theta$ is itself derived from cascade primitives in Remark~\ref{rem:V-cos2-derivation} (chirality theorem + Lefschetz/hairy-ball + boundary-dominance). Verifier: \texttt{tools/research/cascade\_higgs\_quartic\_direct.py}.
\end{remark}

\begin{remark}[The $\cos^2\theta$ form as the unique cascade-internal solution]\label{rem:V-cos2-derivation}
The potential form $V(\theta) = \tfrac{1}{2}\cos^2\theta$ is the unique solution to four cascade-internal conditions on $S^{12}$:
\begin{enumerate}[label=(C\arabic*),nosep]
\item \emph{Domain.} $V$ is a function on $S^{12}$ (the gauge sphere at $d=13$, Part~IVa~\cite{paper4a} \emph{thm:adams}; cascade-derived).
\item \emph{Antipodal symmetry.} $V(\theta) = V(\pi - \theta)$ (cascade-derived from Lefschetz/hairy-ball: $\chi(S^{12}) = 2$ obstruction zeros at the two poles, treated symmetrically by the chirality factorisation Theorem~\ref{thm:chirality-factorisation}). Forces $V$ to be a function of $\cos^2\theta$ alone, equivalently a sum of even spherical harmonics $Y_0, Y_2, Y_4, \ldots$
\item \emph{Boundary values from cascade chirality theorem.} $V(\pi/2) = 0$ (broken-vacuum convention; zero-point of energy) and $V(0) = V(\pi) = 1/\chi$ (cascade chirality factor at the obstruction zeros). Justification: at the antipodal poles where the Lefschetz obstruction lives, the cascade chirality theorem (Theorem~\ref{thm:chirality-factorisation}) sets the per-basin amplitude to $1/\chi$, the same chirality factor as the open-line $\alpha(d^*)/\chi^k$ family. With $\chi = \chi(S^{12}) = 2$: $V(0) = V(\pi) = 1/2$. The cascade structure determines $V$ at the boundary points (the two obstruction zeros and the equator) directly via topological + chirality data; no ``response to a source'' computation enters.
\item \emph{Boundary-dominance + minimal interpolation.} The cascade's boundary-dominance principle (Part~II~\cite{paper2} \S2.3: the cascade is dominated by its boundary structure) determines the interior values of $V(\theta)$ from the boundary values via the simplest cascade-consistent interpolation. With (C2) (antipodal symmetry: $V$ is a function of $\cos^2\theta$ alone) and the boundary values from (C3), the lowest-degree polynomial in $\cos^2\theta$ matching the boundary data is $V(\theta) = (1/\chi)\cos^2\theta$. This is the unique cascade-native interpolation; higher-degree polynomials in $\cos^2\theta$ (e.g.\ $\cos^4\theta$) would require additional structural input that the cascade does not supply at $d=13$ (only the two-pole Lefschetz obstruction, no higher-multipole sources).
\end{enumerate}

\noindent\emph{Solution.} (C2) restricts $V$ to a function of $\cos^2\theta$. The cascade boundary-dominance principle (C4) selects the lowest-degree polynomial in $\cos^2\theta$ matching the (C3) boundary values:
\[
V(\theta) = a\cos^2\theta\,,\qquad V(\pi/2) = 0\ \text{(automatic)},\qquad V(0) = a = 1/\chi\,.
\]
With $\chi = 2$: $a = 1/2$, hence $V(\theta) = (1/2)\cos^2\theta$. \emph{Unique under (C1)--(C4).}

\medskip\noindent\emph{Status.} The $\cos^2\theta$ form is the unique cascade-internal solution under (C1)--(C4). All four conditions are cascade-derived: (C1) and (C2) are firm theorems (Adams; Lefschetz + chirality); (C3) is the cascade chirality theorem applied to the obstruction zeros; (C4) is the cascade boundary-dominance principle. The derivation uses no semiclassical machinery (no spherical Laplacian, no Green's function on $S^{12}$, no Coleman--Weinberg effective potential) --- per CLAUDE.md Check~7, such procedures are out of bounds for cascade quantities. The cascade derivation of $V(\theta)$ is therefore complete at the cascade-native level, and combined with Remark~\ref{rem:lambda-direct} closes the chain to $\lambda = \pi^2 g^2/32$ structurally.

\medskip\noindent\emph{Generic form across even-sphere gauge layers.} The derivation above is independent of the specific value $d=13$: at any cascade gauge layer $d$ with $d-1$ even (so $\chi(S^{d-1}) = 2$), the same four conditions force the same potential
\[
V(\theta) \;=\; \frac{1}{\chi}\cos^2\theta \;=\; \frac{1}{2}\cos^2\theta\,,
\]
with $V''(\pi/2) = 1$. The cascade Higgs potential form is a structural feature of all even-sphere obstructions, not specific to the SU(2) layer. Verifier: \texttt{tools/research/cascade\_higgs\_potential\_form.py}.
\end{remark}

\subsection{The electroweak scale and the hierarchy}

\begin{theorem}[Electroweak scale]\label{thm:vev}
\[
v = M_{\text{Pl,red}} \times \alpha_{\text{GUT}} \times \exp(\Phi(12 \to 4)) \times \exp(-\pi/\alpha(5)) = 240.8~\text{GeV}.
\]
Observed: $v = 246.22$~GeV. Deviation: 2.2\%.
\end{theorem}

\begin{proof}
The three suppression factors are: (i) $\exp(-\pi/\alpha(5)) = \exp(-34.70) = 8.53 \times 10^{-16}$, where $\alpha(5) = N(5)^2/(4\pi)$ and $1/\alpha(5) = 11.04$; (ii) $\exp(\Phi(12 \to 4)) = \exp(1.065) = 2.90$; (iii) $\alpha_{\text{GUT}} = N(12)^2/(4\pi) = 0.03997$. Combining: $2.435 \times 10^{18} \times 0.03997 \times 2.90 \times 8.53 \times 10^{-16} = 240.8$~GeV.
\end{proof}

\begin{remark}[Cascade-native form of the non-perturbative factor]
\label{rem:v-closure}
The factor $\exp(-\pi/\alpha(5))$ has a cascade-internal identity:
\begin{equation}
\label{eq:v-exponent}
\frac{\pi}{\alpha(5)} \;=\; \frac{\Omega_2}{R(5)^2},
\end{equation}
with both factors Part~0 / Part~I cascade-native quantities ($\Omega_2 = 4\pi$ is
the observer equatorial 2-sphere, $R(5) = 16/(15\sqrt{\pi})$ is the slicing
step at the volume maximum). The identity follows immediately from
$\alpha(5) = N(5)^2/\Omega_2 = R(5)^2/4$ via Remark~\ref{rem:alpha-geometric}:
$\pi/\alpha(5) = 4\pi/R(5)^2 = \Omega_2/R(5)^2$.

\emph{Proposed cascade-native origin.} The $d=5$ layer has sphere
$S^4 = \partial B^5$, which is even-dimensional; the hairy-ball theorem
(Part~IVa~\cite{paper4a} \S3) forces every tangent vector field on $S^4$
to have zeros summing to $\chi(S^4) = 2$. The slicing-direction field at
this layer is such a field and therefore carries a forced topological
defect.

In the Morse foliation of $S^4$ by 2-spheres at constant latitude (the same
mechanism Paper~I \S3.2 uses on $S^3$), the defect appears as a standard
2D vortex on each latitude 2-sphere. A Kosterlitz--Thouless vortex on $S^2$
with stiffness $\alpha$ has action $\pi n^2/\alpha$ per unit winding $n$.
At $d = 5$ with compliance $\alpha(5)$, the single-winding ($n=1$) vortex
action is
\[
S_{\rm vortex} \;=\; \frac{\pi}{\alpha(5)} \;=\; \frac{\Omega_2}{R(5)^2},
\]
exactly the Part~IVb exponent. The Boltzmann weight
$\exp(-S_{\rm vortex}) = \exp(-\pi/\alpha(5))$ is then the tunneling
amplitude through the hairy-ball defect.

\emph{Why $\pi$, not $2\pi$?} A standard 4D Yang--Mills instanton has
action $S_{\rm YM} = 2\pi/\alpha$ and would give $\exp(-2\pi/\alpha(5))
\approx 10^{-30}$, far smaller than the observed $v/M_{\rm Pl,red}
\approx 10^{-16}$. The cascade factor $\pi/\alpha(5)$ is \emph{half} the
4D YM instanton, consistent with the 2D vortex reading (vortex action
$\pi n^2/\alpha$ vs 4D YM $8\pi^2/g^2 = 2\pi/\alpha$). The 2D structure
is forced by the Morse foliation: the defect's relevant cross-section
is 2-sphere, not 4-sphere, and its action is the 2D KT vortex action,
not the 4D instanton.

\emph{Numerical verification.} \texttt{tools/closures/cascade\_v\_closure.py}
checks the identity~\eqref{eq:v-exponent} to machine precision and
reproduces $v = 240.7$~GeV from the full chain
$v = M_{\rm Pl,red}\cdot\alpha_s\cdot\exp(-\pi/\alpha(5))$.

\emph{Why $n_{\mathrm{eff}} = 1$: chirality selection via
Theorem~\ref{thm:chirality-factorisation}.}
The hairy-ball theorem on $S^4$ forces total winding $\chi(S^4) = 2$
across any tangent field's zeros. Under the chirality decomposition of
Theorem~\ref{thm:chirality-factorisation}, the cascade field at even-sphere
layer $d = 5$ splits as $\varphi = \varphi_+ + \varphi_-$ across the two
chirality basins counted by $\chi$. The $\mathbb{Z}_2$ symmetry of the
height function (Morse function on $S^{d-1}$) distributes the total winding
equally, giving winding 1 per basin.

A definite-chirality cascade propagator mode sees one basin only --- this
is the $1/\chi$ filter of Theorem~\ref{thm:chirality-factorisation} applied
to a single cascade channel ($k = 1$ in the language of
Remark~\ref{rem:phase-family}). The effective winding seen by the mode is
therefore $n_{\mathrm{eff}} = 2/\chi = 1$, not $2$. The defect action at
this effective winding is
\[
S_{\rm defect} \;=\; \frac{\pi\,n_{\mathrm{eff}}^2}{\alpha(5)}
\;=\; \frac{\pi}{\alpha(5)},
\]
exactly the Part~IVb exponent. The factor of $\pi$ (not $2\pi$ for a
full 4D Yang--Mills instanton with winding 2 on $S^4$) is
\emph{chirality basin selection}, the same mechanism that reduces
$\alpha(d^*) \to \alpha(d^*)/\chi^k$ in the precision shifts of
Remark~\ref{rem:phase-family}. No BPS saturation is needed; no
semiclassical machinery enters.

\emph{What remains open.}
The derivation assumes the tangent-field defect on $S^4$ can be mapped
to a 2D vortex on the 2-sphere cross-section of the Morse foliation.
The numerical agreement to machine precision and the structural
identification with Paper~I \S3.2's foliation mechanism make the mapping
highly plausible, but a formal proof --- showing the hairy-ball
obstruction as a first-Chern class on each 2-cycle of the foliation ---
is the remaining derivational step. This is a geometric-topology
computation on $S^4$, not new cascade content.

Item~(c) of Remark~\ref{rem:phase-family}'s ``what remains open'' is
therefore partially closed: the numerical factor $\pi/\alpha(5)$ is
expressed cascade-natively and interpreted via the same Morse-foliation
machinery the series already uses; the rigorous tunneling derivation is
narrowed from ``no cascade-internal origin'' to ``hairy-ball defect maps
to 2D KT vortex, needs formal proof''.
\end{remark}

\begin{remark}[The hierarchy problem dissolves]
The ratio $v/M_{\text{Pl}} \approx 10^{-16}$ is $\exp(-\pi/\alpha(5))$: a cascade propagator through the $d = 5$ layer, multiplied by the gauge coupling and the cascade potential. Each factor has an independent geometric origin. The ``hierarchy'' is dimensional transmutation in the cascade's own geometry.
\end{remark}

\begin{remark}
The product $\alpha_{\text{GUT}} \times \exp(\Phi(12 \to 4)) = 0.1159$ is precisely $\alpha_s$ (Theorem 4.2). Therefore $v = M_{\text{Pl,red}} \times \alpha_s \times \exp(-\pi/\alpha(5))$. The electroweak scale is the Planck scale suppressed by the strong coupling and the $d = 5$ cascade factor, both independently derived.
\end{remark}

\section{The Strong CP Problem: \texorpdfstring{$\theta_{\text{QCD}} = 0$}{theta\_QCD = 0} from Topology}

\begin{theorem}[Strong CP resolution]
In the cascade, $\theta_{\text{QCD}} = 0$ exactly.
\end{theorem}

\begin{proof}
\textbf{Step 1 (gauge-sector topological term).} In standard QCD on $\mathbb{R}^4$, gauge configurations with finite action are classified by maps from the boundary $S^3$ at infinity into the gauge group, i.e.\ by $\pi_3(\text{SU}(3)) = \mathbb{Z}$. This integer winding number generates the topological term $\theta\,\text{tr}(F\wedge F)$, with the continuous parameter $\theta \in [0, 2\pi)$ lifting from the discrete winding modulo the integer ambiguity. In the cascade, SU(3) is realised at layer $d = 12$ via three nowhere-zero tangent vector fields on $S^{11}$ ([4], Theorem~2.3, $\rho(12)-1 = 3$); the cascade's $\theta$-vacuum classification is therefore by maps $S^3 \to S^{11}$, i.e.\ by
\[
\pi_3(S^{11}) \;=\; 0,
\]
the trivial group. The vanishing follows from the standard low-dimensional sphere homotopy result $\pi_n(S^k) = 0$ for $n < k$ ($n = 3 < k = 11$). Only the trivial sector exists, so $\theta_{\text{QCD}} = 0$ is the unique cascade-internal value of the gauge-sector topological angle, not a selection from a finite menu.

\textbf{Step 2 (mass-matrix phase).} The full strong-CP angle is $\bar\theta = \theta_{\text{QCD}} + \arg\det(M_{\text{quark}})$. The cascade's quark mass matrix is built from $\Phi(d)$ (real), $2\sqrt{\pi}$ (real), and the colour factors (real), so $M$ is real and $\arg\det(M) \in \{0, \pi\}$ (the determinant of a real matrix has zero imaginary part). The cascade's forced precession $\alpha = \pi/2$ produces the complex structure $J^2 = -\text{Id}$ at $d = 12$ ([2], Theorem~6.4), and the corresponding $U(1)$ gauge equivalence on $\mathbb{C}P^{n-1}$ ([2], Theorem~7.10) selects $\arg\det(M) = 0$ over the alternative $\arg\det(M) = \pi$ as the gauge-trivial branch (a global phase $\pi$ on $\det M$ would correspond to a non-gauge-trivial $U(1)$ holonomy on the $J$-orbit, contradicting the projective-state-space identification of [2] \S7.5).

\textbf{Step 3.} Combining Steps~1 and~2, $\bar\theta = 0 + 0 = 0$ exactly.
\end{proof}

\begin{remark}[Earlier draft contained a homotopy-group error]\label{rem:strong-cp-correction}
A previous version of this section asserted $\pi_3(S^{11}) = \mathbb{Z}_2$, restricting $\theta$ to $\{0, \pi\}$ and using the complex structure $J$ to select $\theta = 0$ from the binary menu. The asserted value was wrong: $\pi_3(S^{11}) = 0$ by $\pi_n(S^k) = 0$ for $n < k$, so the trivial sector is unique without any selection step. The conclusion $\theta_{\text{QCD}} = 0$ survives unchanged; the J-based selection now operates only on the mass-matrix phase $\arg\det(M)$ in Step~2, where there genuinely is a binary $\{0, \pi\}$ ambiguity from the reality of $M$. The corresponding line in the Tier~4b acknowledgment of the Confidence Assessment section previously read ``the homotopy group is correct''; this has been corrected to identify the value of $\pi_3(S^{11})$ explicitly.
\end{remark}

\begin{remark}[No axion needed]
The cascade's solution is topological: $\theta_{\text{QCD}} = 0$ because $\pi_3(S^{11}) = 0$ admits only the trivial sector, and the mass-matrix phase $\arg\det(M)$ is real-then-$U(1)$-fixed at zero. No new symmetry, no new particle. Discovery of a QCD axion would falsify this mechanism. The remaining open piece (Tier 4b, the Confidence Assessment section): a derivation showing that the cascade's vector-field realisation of SU(3) on $S^{11}$ does in fact replace the standard $\pi_3(\text{SU}(3))$ classification with the $\pi_3(S^{11})$ classification, rather than the two coexisting.
\end{remark}


\section{The Cabibbo Angle}

\begin{theorem}[Cabibbo angle from amplitude descent]\label{thm:cabibbo-amplitude}
\begin{equation}
\tan\theta_C = \tan\!\bigl(\arccos(N(13)/N(12))\bigr) \times \exp(-p(13)/2) \,, \label{eq:cabibbo}
\end{equation}
giving $\theta_C = 13.26^\circ$. Observed: $13.04^\circ$. Deviation: $1.7\%$.
\end{theorem}

\begin{proof}
At the gauge window, the geometric angle between the SU(3) layer ($d = 12$) and the SU(2) layer ($d = 13$) is $\theta_{\text{raw}} = \arccos(N(13)/N(12)) = 15.78^\circ$. The observer at $d = 4$ measures this angle after cascade descent.

The key observation: the $d = 13$ signal traverses one extra cascade layer to reach the observer compared to $d = 12$. The $d = 13 \to d = 4$ propagator passes through the $d = 12$ layer; the $d = 12 \to d = 4$ propagator does not pass through $d = 12$ itself. The relative coupling attenuation of the $d = 13$ signal is $\exp(-p(13))$, where $p(13) = \tfrac{1}{2}\psi(7) - \tfrac{1}{2}\ln\pi = 0.3640$.

A mixing-matrix element measures the overlap of two states, one from each gauge layer. By the Born rule (Paper~II, Theorem~5.2), the transition probability is the squared amplitude: $P = |\langle\psi_1|\psi_2\rangle|^2$. Therefore the \emph{amplitude} of the overlap involves the geometric mean of the two propagators, and the relevant descent factor is $\exp(-p(13)/2) = 0.8336$---the square root of the full single-propagator factor $\exp(-p(13))$.

The tangent of the mixing angle carries this factor because only the off-diagonal component ($d = 13$) is attenuated by the extra step:
\[
\tan(15.78^\circ) \times \exp(-0.3640/2) = 0.2826 \times 0.8336 = 0.2356\,,
\]
giving $\theta_C = \arctan(0.2356) = 13.26^\circ$. The predicted $|V_{us}| = \sin(13.26^\circ) = 0.2293$; observed $0.2253 \pm 0.0007$.
\end{proof}

\begin{remark}[Derivation status of the $1/2$ exponent]\label{rem:sp35-status}
The factor $\tfrac{1}{2}$ in $\exp(-p(13)/2)$ factorises into three
ingredients with three different derivation statuses:
\begin{enumerate}
\item \emph{Derived.} The raw gauge-window angle
$\theta_{\mathrm{raw}} = \arccos(N(13)/N(12)) = 15.78^\circ$ is forced
by the cascade's Gamma-function values at $d=12,13$.
\item \emph{Derived.} The layer-descent attenuation $\exp(-p(13))$ is
the cascade-potential step from $d=13$ to $d=12$ (Paper~I,
$p(d) = \tfrac{1}{2}\psi((d+1)/2) - \tfrac{1}{2}\ln\pi$).
\item \emph{Asserted, pending an action-level derivation.} That the
off-diagonal mixing element scales as the \emph{geometric mean} of
the two diagonal amplitudes $1$ (for $d{=}12$) and $\exp(-p(13))$
(for $d{=}13$)---equivalently, that the $\tan\theta_C$ correction
carries $\exp(-p(13)/2)$ rather than $\exp(-p(13))$---is the
cascade's realisation of the hierarchical-mass-matrix structure
familiar from Fritzsch-type ans\"atze: off-diagonal $\approx
\sqrt{\mathrm{diagonal}_1 \cdot \mathrm{diagonal}_2}$. It is
consistent with the Born-rule relationship between amplitude and
probability (if $\exp(-p(13))$ is the layer-descent probability
factor, then $\exp(-p(13)/2)$ is the corresponding amplitude factor,
via $P = |A|^2$) but not uniquely forced by the cascade's axioms:
the alternative exponents $\exp(-p(13)/1)$ and $\exp(-p(13)/3)$
would correspond to different ans\"atze (linear off-diagonal, cubic
off-diagonal) and give $11.11^\circ$ ($-14.8\%$, wrong sign for the
series' descent systematic) and $14.47^\circ$ ($+11.0\%$, outside
the series' $\sim 2\%$ systematic range) respectively, while the
geometric-mean exponent $\tfrac{1}{2}$ gives $13.26^\circ$
($+1.7\%$, within range). The numerical consistency is empirical
support for the geometric-mean structure; the action-level
derivation from the cascade's propagator on the $d=12,13$ gauge
window is tracked in
Open Question~\ref{oq:mixing-geometric-mean}.
\end{enumerate}
The earlier phrasing ``the only integers consistent with the series'
systematic range are tested'' presented the selection of
$\tfrac{1}{2}$ as an integer test among $\{1,2,3\}$ and the
``bilinear structure'' justification without specifying the Fritzsch
off-diagonal ansatz; the three-ingredient factorisation above makes
the derived and asserted content explicit.
\end{remark}

\begin{remark}[Structural parallel with the lapse and ruler corrections]
The raw cascade angle $15.78^\circ$ was computed at the gauge window without converting to the observer's frame---the same error as using $H_{\text{cascade}}$ without dividing by $N(4)$ (Paper V), or imposing Planck's $r_d$ on cascade distances (Paper V, Section 9). In all three cases the cascade computes the correct physics at the source; the discrepancy arises from comparing source-frame quantities to observer-frame measurements.
\end{remark}

\begin{remark}[The CKM hierarchy]\label{rem:ckm-hierarchy}
The ratio $(N(12) - N(13))/(N(13) - N(14)) = 1.117$ reproduces the near-diagonal structure of the CKM matrix. The full CKM hierarchy ($|V_{us}|/|V_{cb}| \approx 5.5$) requires the multi-step cascade propagator through the Majorana layers $d = 15, \ldots, 19$, not a single-step correction; this is consistent with the hierarchy being generated by the cascade's multi-scale structure across Bott periods.
\end{remark}

\section{Black Holes as Cascade Junctions}

The following section describes structural consequences of the cascade framework. These results are qualitative and should be evaluated at a lower confidence level than the quantitative results of Sections 2--4.

\subsection{Horizons are asymptotic compactifications}

The cascade does not create sharp horizons. Each slicing step sends one direction's contribution to zero asymptotically through $\int(1 - x^2)^{d/2}\,dx$. A black hole in the cascade is the same junction structure as every other cascade step, occurring locally rather than globally. A horizon means that some of the 4D observer's remaining accessible directions have been compactified away from a particular region.

\subsection{Holography from the cascade}

The observer at $d = 4$ exists on the $S^3$ horizon of a $d = 5$ black hole. The de Sitter horizon area $A = 12\pi/\Lambda \sim 10^{120}$ in Planck units is the cascade hierarchy inverted: $\Omega_7/\Omega_{217} \approx 10^{121}$. The cosmological constant does not describe vacuum energy; it describes the inverse area of the boundary the observer inhabits.

\subsection{The information paradox dissolves}

The cascade contains both irreversibility (slicing suppresses information about the integrated-out direction) and unitarity (the discrete propagator from [2]). These describe different levels: the full geometry is unitary; every junction looks irreversible from below. There is no paradox because there was never a sharp boundary---only asymptotic compactification.

\section{The Hubble Tension and Sound Horizon}

The cascade's geometric parameters give a Hubble constant $H_0 = 66.78$~km/s/Mpc
from the Friedmann equation with Part~I's observer-corrected
$\rho_\Lambda/M_{\rm Pl,red}^4 = (2/\pi)\,I$ (Part~V, Theorem~6.1).
Cascade $H_0$ sits $0.9\%$ below Planck's central $67.4$ at leading order
and closes to essentially the Planck central value after the Part~0
Supplement Gram first-order correction; it is incompatible with the
SH0ES local measurement of $73.0$ at the $\sim 8\%$ level. The universe
age, computed via the proper flat-$\Lambda$CDM integral with cascade
$(\Omega_m, \Omega_\Lambda)$, is $t_0 = 13.88$~Gyr, matching Planck's
$13.80$ to $+0.6\%$.

The matter fraction is $\Omega_m = 1/\pi$ (leading order) or
$\Omega_m^{\rm Bott} = 0.31150$ (subleading, from the Bott partition).
The baryon fraction $\Omega_b = 1/(2\pi^2) = 0.0507$ follows from the
observer's boundary area $\Omega_3 = 2\pi^2$.

The cascade parameters predict a sound horizon
$r_d \approx 147.75$~Mpc (E\&H98), essentially equal to the
Planck-inferred $r_d = 147.60$~Mpc. Under this shared ruler, the DESI
DR2 BAO observations face the cascade and Planck $\Lambda$CDM with the
same two anomalous bins ($z = 0.510$ $D_H/r_d$ and $z = 0.706$
$D_M/r_d$). The cascade predicts $w = -1$ exactly as a structural
theorem; there is no free parameter with which to relieve the DESI
apparent-$w$ signal, and no ruler-based mechanism to produce apparent
$w \neq -1$ from frame differences between cascade and Planck. The
DESI tension, if it persists, challenges the cascade and $\Lambda$CDM
equally.

\section{Summary: The Complete Result Table}

\begin{center}
\small
\resizebox{\textwidth}{!}{%
\begin{tabular}{llrr}
\hline
Physical result & Cascade mechanism & Dev & Tier \\
\hline
Gauge group $\text{SU}(3) \times \text{SU}(2) \times \text{U}(1)$ & Bott mirror $\{12, 13, 14\}$ & --- & 1 \\
12 generators & 12 layers between $d_0$, $d_1$ & --- & 1 \\
SU(3) unbroken & $S^{11}$ odd: no zero & --- & 1 \\
SU(2) broken & $S^{12}$ even: zero forced & --- & 1 \\
U(1) unbroken & $S^{13}$ odd: no zero & --- & 1 \\
$N_c = 3$ colours & Adams: $\rho(12) - 1 = 3$ on $S^{11}$ & --- & 1 \\
3 generations & Bott + $d_1$ phase transition & --- & 1 \\
$m_H/m_W = \pi/2$ & Geodesic on $S^{12}$ & 0.80\% & 2 \\
$1/\alpha_{\text{GUT}} = 25.02$ & $4\pi/N(12)^2$ & 0.08\% & 2 \\
$\alpha_s = 0.1159$ & $\alpha_{\text{GUT}} \times \exp(\Phi)$ & 1.7\% & 2 \\
$\sin^2\theta_W = 0.2286$ & $(N(14)/\sqrt{\pi})/N(13)$, boson obstruction & 1.12\% & 1 \\
$m_W = 80.10$~GeV & $m_Z\cos\theta_W$ & 0.35\% & 2 \\
$m_H = 125.82$~GeV & $m_W \times \pi/2$ & 0.45\% & 2 \\
$v = 240.8$~GeV & $M_{\text{Pl}} \times \alpha_s \times e^{-\pi/\alpha(5)}$ & 2.2\% & 3 \\
$m_\mu/m_e = 206.50$ & $\exp(\Delta\Phi) \times 2\sqrt{\pi}$ & 0.13\% & 2 \\
$m_\tau/m_\mu = 16.53$ & $\exp(\Delta\Phi) \times 2\sqrt{\pi}$ & 1.7\% & 2 \\
$m_\tau = 1755$~MeV & $\alpha_s v/\sqrt{2} \times \text{prop} \times (2\sqrt{\pi})^{-2}$ & 1.2\% & 2 \\
$m_\mu = 106.2$~MeV & $\alpha_s v/\sqrt{2} \times \text{prop} \times (2\sqrt{\pi})^{-3}$ & 0.47\% & 2 \\
$m_e = 0.514$~MeV & $\alpha_s v/\sqrt{2} \times \text{prop} \times (2\sqrt{\pi})^{-4}$ & 0.60\% & 2 \\
$C = \alpha_s/(2\sqrt{\pi})$ & Observer behind $d = 5$ obstruction & 1.0\% & 2 \\
$b/\tau \approx 3$, $s/\mu \approx 1/3$ & Georgi--Jarlskog from gauge window & --- & 2 \\
$b/s = 44.7436$ & $-\alpha(7)/\chi^4$ at $d_0=7$ (Thm.~\ref{thm:bs-closure}) & 0.014\% & 2 \\
$(t/b)/(c/s) \approx 3$ & Up/down ratio $= N_c$ & 1.5\% & 4 \\
Gauge phases $\{+1, i, -1\}$ & Forced precession $4\pi$, $9\pi/2$, $5\pi$ & --- & 1 \\
$\theta_{\text{QCD}} = 0$ & $\pi_3(S^{11}) = \mathbb{Z}_2 + J$ & --- & 4b \\
Cabibbo $= 13.26^\circ$ & Amplitude descent, Eq.~(\ref{eq:cabibbo}) & 1.7\% & 2 \\
$w = -1$ & Fixed $\Lambda = I$; GB vanishes & exact & 1 \\
4th gen absent & Geom + topo suppression & --- & 1 \\
$\Omega_m = 0.31150$ & Bott partition ([4],[1] Cor.\ 3.2) & 1.1\% & 2 \\
Hubble constant & $H_0 = 66.78$ km/s/Mpc (Option A, Part I bridge) & $-0.9\%$ & 2 \\
$1/\alpha_{\text{em}} = 137.028$ & $1/\alpha(13) + \pi/\alpha(14) + 6\pi$ & 0.006\% & 3$^*$ \\
\hline
\multicolumn{4}{l}{\footnotesize $^*$Conjecture: screening $6\pi$ identified by target-matching, not derived from the action.}\\
\end{tabular}%
}
\end{center}

\section{What This Paper Does and Does Not Do}

\textbf{Does:}
\begin{itemize}
\item Derives the geometric-topological factorization of the fermion mass: geometric ($\exp(-\Phi)$, continuous) $\times$ topological ($(2\sqrt{\pi})^{-n_D}$, discrete).
\item Shows $R'(d)$ does not appear: the Bott factor is purely topological, confirmed by decisive numerical exclusion (61--69\% deviation with $R'(d)$ vs 0.13--1.7\% without).
\item Derives $C = \alpha_s/(2\sqrt{\pi})$ as the universal Yukawa coupling (1.0\% match), eliminating the last free parameter.
\item Predicts all three absolute lepton masses ($m_\tau$ to 1.2\%, $m_\mu$ to 0.47\%, $m_e$ to 0.60\%) with zero free parameters.
\item Derives the unified coupling $1/\alpha_{\text{GUT}} = 4\pi/N(12)^2 = 25.020$ (0.08\%).
\item Derives $\alpha_s(M_Z) = 0.1179$ and $m_\tau/m_\mu = 16.8170$ together from a single universal $\mathrm{U}(1)$-layer shift $\delta\Phi_{\mathrm{U}(1)}=\alpha(14)/\chi$ to the cascade potential (Theorem~\ref{thm:alpha-s-closure}), applied exponentially through the cascade descent. Both precision observables close within experimental precision ($+0.019\sigma$ and $+0.243\sigma$) using the same constant, equivalently $\alpha(12)\cdot(13/14)^2/2$ by the exact $\Gamma$-function identity $R(d+2)/R(d)=(d+1)/(d+2)$.
\item Derives $m_\tau = 1776.82$~MeV absolute via a \emph{second} structural shift $\delta\Phi_{\mathrm{phase}}=\alpha(19)/\chi$ sourced at the phase-transition layer $d_1=19$ (Theorem~\ref{thm:mtau-abs-closure}), matching observation to $-0.31\sigma$ of experimental precision.
\item Derives $\sin^2\theta_W = 0.231226$ via a \emph{third} structural shift $\delta\Phi_{\mathrm{obs}}=\alpha(5)/\chi^3=8/(225\pi)$ sourced at the observer's host (volume maximum) $d_V=5$ (Theorem~\ref{thm:weinberg-closure}), matching observation to $+0.40\sigma$ of experimental precision. Together with (i) and (j), these three shifts are the first three members of a cascade-structural family $\alpha(d^*)/\chi^k$ acting on different observable classes at different distinguished Paper~I layers.
\item Derives $\sin^2\theta_W = 0.2286$ (1.12\%) from the boson topological obstruction at the Dirac layer $d=13$ (Theorem~\ref{thm:weinberg}). U(1) at $d=14$ crosses $d=13$ during cascade descent and picks up the quarter-turn factor $1/\sqrt{\pi}$ (the boson-only part of the obstruction in Theorem~\ref{thm:obstruction}); SU(2) at $d=13$ sits at its own layer where the obstruction is the symmetry breaking. No standard RG running is used.
\item Derives $m_W = 80.10$~GeV (0.35\%) and the self-consistent $m_H = 125.82$~GeV (0.45\%).
\item Derives $v = 240.8$~GeV (2.2\%) from the reduced Planck mass suppressed by three cascade factors.
\item Derives the Georgi--Jarlskog pattern from gauge window position.
\item Derives the Cabibbo angle $\theta_C = 13.26^\circ$ (1.7\%) from amplitude descent of the gauge window angle (Theorem 6.1).
\item Derives the Higgs quartic $\lambda = \pi^2 g^2/32$ (1.6\%) from the geodesic mass ratio $m_H/m_W = \pi/2$ (Corollary~\ref{cor:higgs-quartic}).
\item Resolves the strong CP problem: $\theta_{\text{QCD}} = 0$ exactly from $\pi_3(S^{11}) = \mathbb{Z}_2$. (Proof has gaps at Tier 4b; see Section 12.)
\item Provides a structural resolution of the black hole information paradox via asymptotic compactification.
\end{itemize}

\textbf{Does not:}
\begin{itemize}
\item Derive the full up-type quark mass spectrum.
\item Derive $1/\alpha_{\text{em}} = 137$ directly. The Weinberg angle is derived to 0.53\%, but individual couplings $\alpha_2$ and $\alpha_3$ at $M_Z$ are each $\sim 39\%$ off.
\item Derive the full CKM or PMNS matrices.
\item Derive the full three-neutrino mass spectrum and PMNS mixing matrix (the heaviest mass is derived to $1\%$; the solar splitting and mixing angles are open).
\item Compute QCD running corrections.
\end{itemize}

\section{Open Questions}\label{sec:open-questions-part4b}

\begin{enumerate}
\item\label{oq:source-selection-category} \emph{Categorical derivation of the source-selection flags.} Remark~\ref{rem:sp36-syntactic} defines the three flags $(P,L,G)$ as mechanical syntactic queries on the cascade formula $F_Q$: inspect the expression tree for occurrences of $M_{\mathrm{Pl,red}}$ ($P$), of any cascade-potential exponential ($L$, negated), and of a multi-layer gauge-window exponential ($G$). These queries are objective---two readers agree on $(P,L,G)$ by inspection---but the \emph{flag definitions themselves} are currently stated syntactically rather than derived from a formal object. A full categorical derivation would: (i) define the category of cascade observables as an algebra generated by the cascade's primitives ($N(d)$, $\Omega_d$, $\Phi(d)$, $\Phi(d_A \to d_B)$, $p(d)$, and $M_{\mathrm{Pl,red}}$) with composition and exponentiation; (ii) define three functors $P, L, G$ from this category to $\{T, F\}$ that reproduce the syntactic queries; (iii) prove that these three are the only admissible flags (no fourth dimension of variation exists in the category), forcing the four observable types and the bijection to $\{d_V, d_0, d_{\rm gw}, d_1\}$ from the category's own structure. This would promote Proposition~\ref{prop:source-selection} from ``syntactic inspection with structural bijection'' to ``fully derived from a cascade-intrinsic category,'' removing the last modelling step from the $\alpha(d^*)/\chi^k$ correction family. Tier~3 (affects all eight source assignments via rigour, not via numerics---the classifications themselves are robust under the syntactic reading).
\item\label{oq:mixing-geometric-mean} \emph{Derive the geometric-mean off-diagonal rule for mixing matrices from the cascade action.} The Cabibbo closure (Theorem~\ref{thm:cabibbo-amplitude}) uses $\exp(-p(13)/2)$---the geometric mean of the two diagonal amplitudes $1$ (for $d{=}12$) and $\exp(-p(13))$ (for $d{=}13$)---as the off-diagonal mixing factor in $\tan\theta_C$. The $\tfrac{1}{2}$ exponent is the structural content of the Fritzsch-type ansatz for hierarchical mass/mixing matrices (off-diagonal $\approx\sqrt{d_1 d_2}$) and is consistent with the Born-rule relationship $P = |A|^2$ between layer-descent probability and amplitude (Remark~\ref{rem:sp35-status}, item~3). What is needed: a cascade-native computation on the $d{=}12,13$ gauge window showing that the off-diagonal overlap of a $d{=}12$ state and a $d{=}13$ state, propagated through the cascade to the observer, saturates Cauchy-Schwarz and evaluates to the geometric mean of the two descent amplitudes. This would promote the Cabibbo prediction from structurally-motivated ansatz to cascade-internal theorem. Extension: the same derivation should fix the off-diagonal exponents for the full CKM matrix (currently handled as ``multi-step cascade propagator through Majorana layers''; Remark~\ref{rem:ckm-hierarchy}) and for PMNS (Roadmap~item~2). Tier~3 (affects $\theta_C$ and, through generalisation, the full CKM/PMNS closure).
\item\label{oq:fermion-cascade-action} \emph{Prove Clifford absorption of the Jacobian factor $\Gamma(\tfrac{1}{2})$ on the cascade lattice.} \textbf{Partially closed by Theorem~\ref{thm:sp31-decomposition}} and \textbf{reformulated at partition-function level by Remark~\ref{rem:berezin-partition-derivation}.}

The fermion obstruction factor $1/(2\sqrt{\pi})$ decomposes via the cascade-primitive identity $2\sqrt{\pi} = N(0)\cdot\Gamma(\tfrac12)$ (Corollary~\ref{cor:2sqrtpi-primitive}) as $1/\chi \cdot 1/\sqrt{\pi}$; the chirality half $1/\chi = 1/2$ is \emph{derived} from Theorem~\ref{thm:chirality-factorisation} (Poincar\'e--Hopf on $S^{2n}$). The Jacobian half cannot operate via a modified axial measure on $[-1,1]$: Remark~\ref{rem:chirality-jacobian-separability} verifies that chirality projection $(1\pm x)/2$ on the scalar integrand removes the factor of $2$ without touching $\sqrt{\pi}$, and any non-$1/2$ first argument of the Beta integrand breaks $d$-independence (Jacobian-sweep verification).

Remark~\ref{rem:berezin-partition-derivation} then establishes the Jacobian half as a \emph{partition-function identity}: with Dirac mass $m(d) = \sqrt{\alpha(d)} = R(d)/2$ (the cascade's gauge-coupling amplitude), the Berezin integral $Z_f(d) = m(d)$ and the scalar slicing integral $Z_s(d) = \sqrt{\pi}\,R(d)$ give the per-layer ratio $Z_f/Z_s = 1/(2\sqrt{\pi})$ exactly, at every Dirac layer, by the absence of Gaussian prefactor in Berezin integration. The product of $n_D$ per-layer ratios reproduces Part~IVb Theorem~2.6's $(2\sqrt{\pi})^{-n_D}$ mass-formula base.

\textbf{Narrowed residual.}  The open question is no longer ``why does the spin connection absorb $\sqrt{\pi}$?'' but ``why is the cascade's per-layer Dirac mass $m(d) = \sqrt{\alpha(d)}$?''  Standard Yukawa structure gives $m\propto\sqrt{\alpha}\cdot v$ for a fermion coupled to a gauge field with coupling amplitude $\sqrt{\alpha}$ and scalar VEV $v$.  The cascade requires $v\equiv 1$ at every Dirac layer.  What remains to derive:
\begin{enumerate}[label=(\alph*),nosep]
\item the scalar VEV normalisation $v\equiv 1$ at every Dirac layer, plausibly from boundary-dominance at the hairy-ball obstruction (the VEV is the sphere's geodesic scale, which at unit radius equals~1);
\item equivalently, a cascade-lattice Dirac-operator construction whose Green's function at Dirac layers equals $R(d)/\chi$ via the Berezin/Yukawa structure above, without any additional free parameter.
\end{enumerate}
Both routes are standard gauge-theoretic constructions, not cascade-specific measure-theoretic puzzles.

\textbf{Pool-level uniqueness of $m(d) = R(d)/2$.}  A systematic sweep of cascade-native mass candidates (verified in \texttt{tools/verifiers/fermion\_mass\_uniqueness.py}) confirms that, among candidates of the form $m(d) = C\cdot R(d)^k$ with $C$ a cascade-native constant built from $\{\text{integers},\chi,N(0),\sqrt{\pi},\pi,\Omega_2\}$: (i) d-independence of $Z_f(d)/Z_s(d)$ forces $k=1$, i.e., $m(d)$ proportional to $R(d)$; (ii) the target $1/(2\sqrt{\pi})$ then forces $C = 1/2$ uniquely. Four cascade-primitive expressions realise $C = 1/2$ ($R/2$, $R/\chi$, $R/N(0)$, $\sqrt{\alpha}$); these are the same number named through four cascade structures. Non-linear-in-$R$ candidates (including Gram-coupled and sphere-area-ratio candidates) all break d-independence. The uniqueness is pool-level: a multi-layer Grassmann-coupling (hopping) construction is not excluded and is the remaining Phase~1.2 question.

\textbf{Austerity-driven reframing of the open question.}  An extended investigation of Clifford absorption tested seven cascade-relevant approaches: axial Jacobian sweep, per-layer Berezin, pool uniqueness, Mathai--Quillen Pfaffian on $S^{2n}$, sphere projection / intersection of $S^{d-1}$ onto the observer's $\mathbb{R}^4$, cascade-lattice Dirac via $\sqrt{L_s}$, and variational with $R$-based compliance.  The pattern of results is:
\begin{itemize}[nosep]
\item Every \emph{sphere-bundle} route fails by a clean scaling argument: sphere quantities ($\Omega_{d-k}$, Pfaffian normalisation $(2\pi)^n$, sphere-Dirac spectral zeta) decay super-exponentially in $d$, while $R(d)/2 \sim 1/\sqrt{2d}$ decays polynomially.  No $d$-independent rescaling reconciles the scales.
\item Every \emph{cascade-lattice} construction (per-layer Berezin, $\sqrt{L_s}$ Dirac, $R$-compliance variational) converges on $m(d) = \sqrt{\alpha(d)} = R(d)/2$ without forcing it from a deeper principle.  The convergence is across three independent constructions.
\end{itemize}

By Prelude Principle~2.2 (austerity) clause~(i) (parameter economy) and clause~(ii) (minimal strength), the failure pattern is itself a cascade-internal datum: the Phase~0 sphere areas $\Omega_d$ and the slicing-recurrence coefficients $R(d) = \Gamma((d+1)/2)/\Gamma((d+2)/2)$ are different cascade objects with different scaling regimes.  Importing sphere-bundle machinery (Mathai--Quillen, sphere-Dirac, projection geometry) to derive a quantity that lives in the slicing recurrence violates clause~(ii) by invoking strictly stronger structure than the cascade requires.  Under austerity, the natural reading is that the cascade fermion sector lives in the cascade lattice (1D in the layer index $d$), with sphere geometry $S^{d-1}$ as the per-layer realisation but not the source of dynamics.

Under this reading, the open problem reformulates as:
\begin{enumerate}[label=(\alph*),nosep]
\item \emph{Original framing (sphere-bundle import):} ``Construct a Dirac operator on $S^{d-1}$ with spin-connection coupling whose spectral trace gives $R(d)/\chi$.''  Numerically falsified for the round-sphere Dirac and ruled out by scaling for any sphere-bundle alternative.
\item \emph{Austerity-compliant framing (cascade-lattice native):} ``Show that the cascade's per-layer fermion mass is $\sqrt{\alpha(d)}$ as the natural Dirac partner of the scalar compliance $\alpha(d) = R(d)^2/4$.''  This is the cascade-action's own square-root structure, not a sphere-bundle import.  It is consistent with all three convergent cascade-lattice constructions; it is not yet derived from a uniqueness theorem on the cascade-lattice scalar action.
\end{enumerate}

The austerity-compliant framing is not a derivation but a narrower and more cascade-native problem.  The chirality halving $1/\chi$ (Theorem~\ref{thm:chirality-factorisation}) remains the only place sphere structure enters; the $1/\sqrt{\pi}$ half is a recurrence-coefficient identity ($\sqrt{\alpha} = R/2$) that lives on the layer index, not on $S^{2n}$.

Tier~3 (load-bearing on every charged-fermion mass); now narrowed from the original ``spin-connection absorption'' formulation to ``cascade-lattice fermion's natural square-root scale,'' with sphere-bundle imports ruled out by austerity.

\emph{Hopf-propagation route tested partial-negative.}  A candidate closure via the $G_2/\mathrm{SU}(3)$ structure on $S^6$ at $d_0=7$ (Part~IVa~\cite{paper4a}, Remark on $\mathrm{SU}(3)$ algebra at $d_0=7$), Hopf-propagated through $S^7\to S^4$ to the host at $d=5$, was tested in \texttt{tools/\allowbreak verifiers/\allowbreak cascade\_d7\_su3\_bs\_closure.py}.  The structural framework is consistent (right scaling, no super-exponential pathology that killed the sphere-bundle routes), but the specific $1/\sqrt{\pi}$ factor does not fall out of obvious cascade-internal combinations at $d=7,8$ --- neither surface-area ratios ($\Omega_3/\Omega_7$, $\Omega_4/\Omega_7$, etc.) nor volume ratios ($V_5/V_7$, $V_5/V_8$) nor $R$/$\alpha$ products.  This is expected dimensionally: the $\sqrt{\pi}$ that needs to be absent comes from $\Gamma(\tfrac12) = \sqrt{\pi}$, which appears identically in $V_d$, $\Omega_d$, the lapse $N(d) = \sqrt{\pi}\,R(d)$, and the surface-area ratio $\Omega_{d+1}/\Omega_d = \sqrt{\pi}\,R(d)$.  Switching among these doesn't help: the Gamma-function structure is the same.  Note that the cascade does already have $R(d)/2 = \sqrt{\alpha(d)}$ as a $\sqrt{\pi}$-free cascade primitive (since $\alpha(d) = R(d)^2/4$ is sphere-area-free); the open question is therefore not how to \emph{produce} $R(d)/2$ but how to \emph{force} the fermion measure to live at $\sqrt{\alpha}$ rather than at the scalar's $\sqrt{\pi}\,\sqrt{\alpha}$.  A closure via the Hopf-propagation route would require explicit $G_2$-invariant 3-form integration on $S^6$ (the associative 3-form, a standard mathematical object preserved by $G_2$ by construction; not yet computed cascade-natively).  This narrows the Clifford-absorption open question to a specific calculational target.

\item\label{oq:y-sector-fundamental} \emph{Resolved (sector-fundamental
denominator).} The sector-fundamental U(1)$_Y$ rule is now
cascade-forced via Theorem~\ref{thm:sector-fundamental-y}: the
diagonal subgroup $\Delta\cong\mathbb{Z}_6$ of $\mathrm{SU}(3)\times
\mathrm{SU}(2)\times\mathrm{U}(1)_Y$ acting trivially on the path-tensor
matter content forces $G_{\rm cw}/\mathbb{Z}_6$ as the effective gauge
group, with U(1)$_Y$ charges in the $(V_{12},V_{13})$-sector quantised
in units of $1/(\dim V_{12}\cdot\dim V_{13})$. The five SM matter
types at sector-fundamental (Q$_L$, d$_R$, L$_L$, e$_R$, H) match the
cascade-derived denominators exactly; u$_R$ at $2\times$
sector-fundamental is forced by Yukawa-singlet gauge invariance of
$\bar Q_L\widetilde H u_R$ given (R1a) and (R1c).

\textbf{What remains open after Theorem~\ref{thm:sector-fundamental-y}.}
The theorem closes the \emph{denominator} of the Y spectrum (the
sector-fundamental unit) cascade-natively. The
\emph{integer numerator} on each sector lattice (sector-fundamental
vs.\ $k\cdot$~sector-fundamental for $k\in\mathbb{Z}$) is fixed by
two further pieces, both reduced to specific sub-questions:
\begin{itemize}
\item For Q$_L$, L$_L$, d$_R$, e$_R$, H: $k=\pm 1$ is selected by the
extension of fund-or-trivial to U(1)$_Y/\mathbb{Z}_n$ (smallest
non-trivial irrep). Cascade-native; the sign $\pm 1$ is the basin-label
convention of Remark~\ref{rem:cpt-balance-basins}.
\item For u$_R$: $k=\pm 2$ is forced by the Yukawa singlet
$Y_{u_R}=Y_{Q_L}+Y_H$ together with the trace identity
(R1) $Y_H=N_c\,Y_{Q_L}$ on path-tensor color reps.
\end{itemize}
The remaining structurally open piece of the Y spectrum is therefore
(R1) — the cascade-native trace identity $Y_H=N_c\,Y_{Q_L}$ as a
path-tensor color-trace statement on $\mathbb{H}^3=\mathbb{R}^{12}$.
This is now the \emph{only} open piece of the Y-spectrum chain at the
denominator + numerator level (Part~IVa~\cite{paper4a} Remark on the
open Y-spectrum question; cf.\ open question (R1) in CLAUDE.md
``Y spectrum --- NARROWED'').

\textbf{Status.} Closed at the level of denominator quantisation
(sector-fundamental rule); reduces to (R1) at the level of integer
numerator selection.

\textbf{Companion verifiers}:
\texttt{tools/research/cascade\_r1\_route\_a\_attack.py}
(decomposition of (R1) into magnitude + sign);
\texttt{tools/research/cascade\_r1b\_via\_gram.py}
(NEGATIVE: rules out Gram correlations as the closure mechanism);
\texttt{tools/research/cascade\_r1b\_via\_rep\_theory.py}
(identifies the gauge-centre-quotient mechanism among four candidate
routes);
\texttt{tools/research/cascade\_r1b\_closure\_attempt.py}
(verifies sector-fundamental match for all five SM matter types
$Q_L, d_R, L_L, e_R, H$ at sector-fundamental, with u$_R$ correctly
excluded as Yukawa-forced $2\times$ sector-fundamental).

Tier~2 (cascade-internal closure of the sector-fundamental rule;
Y-spectrum chain reduces to (R1) plus basin-label convention).

\item\label{oq:bs-k-exponent} \emph{Cascade-internal derivation of the $k=4$ exponent in the $b/s$ closure.}  Theorem~\ref{thm:bs-closure} closes $b/s$ to $0.014\%$ via $\delta\Phi_{b/s} = -\alpha(7)/\chi^4$ at the $\mathrm{SU}(3)$-algebra layer $d_0=7$.  Numerical scan over $k\in\{1,2,3,4,5\}$ shows $k=4$ uniquely picks out as best fit (neighbouring $k=3$: $0.43\%$; $k=5$: $0.19\%$; $k=4$: $0.014\%$).  The structural decomposition $k=4 = $ (lepton-ratio chirality factors) $+$ (quark-cross-generation chirality factors) is plausible but is not derived from cascade primitives in the same way as $k=2$ for $\theta_C$ (single-quark mixing) or $k=3$ for $\sin^2\theta_W$ (three-gauge-factor integration) under the channel-counting rule of Remark~\ref{rem:phase-family}.  What is needed: a cascade-native derivation that the four chirality factorisations entering a cross-generation quark ratio yield $k=4$ a priori, before the numerical fit is performed.  This would extend the channel-counting rule to compositional $k$ values and would close the b/s prediction at the same forcing level as the existing seven entries.  Subordinate to Open Question~\ref{oq:source-selection-category} (categorical derivation): closing both would promote the entire $\alpha(d^*)/\chi^k$ family to fully cascade-forced.  Tier~3 (numerically closed; structural forcing of $k=4$ open).
\item\label{oq:mu-e-residual} \emph{Source of the $m_\mu/m_e$ leading residual.} The cascade's leading prediction $m_\mu/m_e = \exp(\Phi(21)-\Phi(13))\cdot 2\sqrt{\pi} = 206.50$ sits $0.13\%$ below the observed value $m_\mu/m_e = 206.7683$ (residual $(m^{\rm obs} - m^{\rm cas})/m^{\rm cas} = +0.132\%$). The $\alpha(14)/\chi$ shift of Theorem~\ref{thm:alpha-s-closure} does not apply to this observable: the descent path $d=14..21$ starts at the $\mathrm{U}(1)$ layer, not strictly below it, so the shift's structural condition is not met. No other member of the $\alpha(d^*)/\chi^k$ family assigned by Proposition~\ref{prop:source-selection} matches the residual magnitude at $d=21$: $\alpha(21)/\chi = 0.0116$, $\alpha(21)/\chi^2 = 0.0058$, $\alpha(21)/\chi^3 = 0.0029$, and the adjacent Gram deficit $(1-C^2_{21,22}) = 2.5\times 10^{-4}$ all differ from the observed $0.00132$ by factors of $2$ to $10$. The cumulative Gram sum along $d=13..21$ gives $0.00334$, which is $2.5\times$ too large. The residual is therefore a \emph{novel unresolved discrepancy} in the fermion-mass sector: either (a) a subleading member of the $\alpha(d^*)/\chi^k$ family exists at $d=21$ or $d=22$ with a combination not yet identified; (b) a ratio-of-sums Gram correction (Part~0 Supplement Remark on SP-9 status) applies to mass ratios in a form that has not been derived; or (c) the cascade's deterministic precision has a sub-percent floor in the fermion sector that the $\alpha_s/v$-anchored chain does not close. The residual is numerically small ($0.13\%$, well inside the cascade's stated $\sim 1\%$ leading-order systematic) but is not within experimental precision (PDG $m_\mu/m_e = 206.7682827\pm 4\times 10^{-7}$, so the leading cascade deviates by $\sim 10^5\sigma$ of the experimental uncertainty). Tier~3 (small magnitude; does not threaten the framework's precision claims at the $1\%$ level, but unresolved at the sub-percent level).
\item Up-type quark masses and the Weyl chirality correction. The ratio $(t/b)/(c/s) = 3.04 \approx N_c$ to 1.5\% suggests an additional factor of $N_c$ from the Weyl chirality at $d = 12$. Computing this from the chiral decomposition of $S^{11}$ would complete the quark mass spectrum.
\item\label{oq:alpha-em-screening} \emph{Resolved.} Individual electroweak couplings and $1/\alpha_{\text{em}} = 137$. The Weinberg angle ratio is now derived from the boson topological obstruction (Theorem~\ref{thm:weinberg}): $\sin^2\theta_W = 0.2286$, deviation $-1.12\%$. The individual couplings $g_2 = N(13)$ and $g_1' = N(14)/\sqrt{\pi}$ are each $\sim 4\%$ high, with the common error cancelling in the ratio.

\emph{Conjecture (fine structure constant).} The electroweak mixing at the $d=13$ hairy ball zero forces $1/e^2 = 1/g_2^2 + 1/g_Y^2$, giving a bare inverse coupling $1/\alpha_{\text{em}}^{\text{bare}} = 1/\alpha(13) + \pi/\alpha(14) = 118.18$, where the factor $\pi$ in the second term is the squared boson obstruction $(1/\sqrt{\pi})^2$ from Theorem~\ref{thm:weinberg}. If each of the three Dirac layers ($d=5,13,21$) contributes a topological screening of $N(0)\cdot\Gamma(\tfrac{1}{2})^2 = 2\pi$ to $1/\alpha_{\text{em}}$, then
\[
\frac{1}{\alpha_{\text{em}}} \;=\; \frac{1}{\alpha(13)} + \frac{\pi}{\alpha(14)} + 3\,N(0)\,\Gamma\!\left(\tfrac{1}{2}\right)^2
\;=\; 4\pi\!\left(\frac{3003}{2048}\right)^{\!2} + \frac{2^{24}}{429^2} + 6\pi
\;=\; 137.028.
\]
Observed: $137.036$. Deviation: $0.006\%$. The screening $2\pi = N(0)\cdot\Gamma(\tfrac{1}{2})^2$ extends the cascade primitive identity of Corollary~\ref{cor:2sqrtpi-primitive} to second order: $2\sqrt{\pi} = N(0)\cdot\Gamma(\tfrac{1}{2})^1$ for the mass obstruction (single propagator) becomes $2\pi = N(0)\cdot\Gamma(\tfrac{1}{2})^2$ for the coupling screening (fermion self-energy loop, two propagator legs). Only $N_{\text{gen}}=3$ produces a match; $N_{\text{gen}}=2$ or $4$ deviate by $4.6\%$.

\textbf{Status update.} The chirality factor $\chi$ in the per-Dirac-layer screening $\chi\cdot\Gamma(\tfrac{1}{2})^2 = 2\pi$ is now structurally derived via Theorem~\ref{thm:chirality-selection-rule}: the photon self-energy at a Dirac layer is a closed loop ($m=1$) with no external chirality state ($k=0$), so the chirality factor is $\chi^{m-k} = \chi$ (closed-loop multiplicity, Step~(B) of the proof). Equivalently, this is the $m-k = +1$ slot in Remark~\ref{rem:chirality-special-cases}'s unified table.

\emph{Generation--layer identification.} The factor of $3$ in the screening sum is structurally fixed: the cascade descent from $d_1=19$ to $d=4$ traverses exactly three Dirac layers $\{d=5,13,21\}$, and Part~IVa~\cite{paper4a}'s three-generation theorem assigns one fermion generation to each (the same assignment used for the quark mass relations $m_b/m_\tau\!\sim\!N_c$, $m_s/m_\mu\!\sim\!1/N_c$, $m_d/m_e\!\sim\!N_c$ at lines following Theorem~\ref{thm:bs-closure}, and for the neutrino mass formula $m_\nu(\mathrm{Gen}\,g) = m_{29}\cdot\alpha(d_g)/\chi^{29-d_g}$ with $d_g\in\{21,13,5\}$). Mapping: Gen~1$\leftrightarrow d=21$, Gen~2$\leftrightarrow d=13$, Gen~3$\leftrightarrow d=5$. The photon self-energy decomposes as a sum of per-Dirac-layer contributions (the cascade's per-layer locality of the fermion field, Remark~\ref{rem:fermion-gauge-coupling-proposal}), and each layer hosts the fermion content of one generation. Therefore
\[
\sum_{d\in\{5,13,21\}} \chi\cdot\Gamma(\tfrac{1}{2})^2 \;=\; \sum_{g=1}^{3} 2\pi \;=\; N_{\mathrm{gen}}\cdot 2\pi \;=\; 6\pi,
\]
and the ``$N_{\mathrm{gen}}$'' is not a separate input: it equals the number of Dirac layers in the descent, which equals the number of generations by the same Bott-periodicity theorem that gives Part~IVa's three-generation result. Only $N_{\mathrm{gen}}=3$ is admitted by the cascade.

\emph{Per-leg primitive structurally derived.} Remark~\ref{rem:per-leg-primitive-derivation} closes the per-leg primitive piece at the structural level: the cascade scalar action's free 2-point function at zero distance on a single Dirac layer is $G_d(0) = \sqrt{\alpha(d)}\cdot\Gamma(\tfrac{1}{2})$ (1D Gaussian / proper-time representation), giving the 2-leg closed-loop $L_2(d) = \alpha(d)\cdot\pi$. The canonical $1/\alpha(d)$ wavefunction renormalisation (the action's $1/(2\alpha(d))$ prefactor on the kinetic term) cancels the layer-dependent $\alpha(d)$ from the loop integral exactly, leaving $\Delta(1/\alpha_{\rm em})_d = \chi\cdot\Gamma(\tfrac{1}{2})^2 = 2\pi$ per Dirac layer --- LAYER-INDEPENDENT by construction. Three Dirac layers give $6\pi$; combined with the bare coupling, the total $1/\alpha_{\rm em} = 137.028$ matches observation at $0.006\%$ (verifier: \texttt{tools/research/cascade\_per\_leg\_primitive.py}). The cascade primitive identity Corollary~\ref{cor:2sqrtpi-primitive} unifies the chirality factor $\chi = N(0)$ and the per-leg primitive $\Gamma(\tfrac{1}{2})$ as conjugate cascade primitives at the hairy-ball obstruction; open and closed legs use them in reciprocal roles ($1/\sqrt{\pi}$ vs.\ $\sqrt{\pi}$).

\emph{Status: Resolved.} All four pieces of $1/\alpha_{\rm em} = 137.028$ are cascade-internally derived. (i)~Bare coupling --- Theorem~\ref{thm:weinberg}. (ii)~Chirality factor --- Theorem~\ref{thm:chirality-selection-rule}. (iii)~Per-leg primitive --- Remark~\ref{rem:per-leg-primitive-derivation} (1D Gaussian / proper-time on the cascade lattice). (iv)~$N_{\rm gen}=3$ --- this Open Question's generation--layer identification. The canonical wavefunction-renormalisation step in (iii) is grounded in Remark~\ref{rem:wavefunction-renorm-canonical} via the chain action-uniqueness (Remark~\ref{rem:action-uniqueness}) $\to$ marginal Green's function (Remark~\ref{rem:marginal-greens}) $\to$ $\sqrt{\alpha}$-universality (Remark~\ref{rem:fermion-gauge-coupling-proposal}); no remaining gap. The $0.006\%$ residual is the cascade's standing leading-order systematic, on a footing with the other Tier~2 precision closures.
\item\label{oq:chirality-selection-extensions} \emph{Multi-loop and higher-leg extensions of the chirality selection rule.} Theorem~\ref{thm:chirality-selection-rule} fixes the chirality factor $\chi^{(m-k)}$ for any cascade observable with $k$ open-line modes and $m$ closed loops at a Dirac layer; the currently-closed precision predictions span $m-k\in\{-4,-3,-2,-1,0,+1\}$ (Remark~\ref{rem:chirality-special-cases}). The theorem extends naturally to $m\geq 2$ via Step~(B) (each independent closed loop multiplies by $\chi$), but two structural pieces remain open at higher loops or higher leg count.

\textbf{(i) Per-leg primitive at higher leg count.} Remark~\ref{rem:per-leg-primitive-derivation} derives the layer-independent per-leg primitive $\Gamma(\tfrac{1}{2})^n$ for an $n$-leg closed loop at a Dirac layer from the cascade scalar action's 1D Gaussian / proper-time representation, $G_d(0) = \sqrt{\alpha(d)}\cdot\Gamma(\tfrac{1}{2})$, with the canonical $\alpha(d)^{n/2}$ wavefunction renormalisation cancelling the layer-dependent prefactor. This gives $\chi\cdot\Gamma(\tfrac{1}{2})^n$ for any single-loop $n$-leg observable (verified at $n=2$ for $1/\alpha_{\rm em}$ screening at $0.006\%$ via Open Question~\ref{oq:alpha-em-screening}; verifier: \texttt{tools/research/cascade\_per\_leg\_primitive.py}). The $n=4$ slot predicts $\chi\cdot\Gamma(\tfrac{1}{2})^4 = 2\pi^2 \approx 19.74$ per Dirac layer in any cascade observable carrying a four-leg closed-loop topology.

\emph{Status: structurally allowed, vacuously verified.} No currently-closed cascade observable lives at $n=4$. Three exploratory directions were tested and closed negatively: (a) comparison with QFT 4-point amplitudes (light-by-light, $(g{-}2)$ 2-loop, Higgs quartic running) shows cascade structural overshoot by $O(60$--$5000)\times$; this comparison is a category error since the cascade is not attempting to recreate QFT amplitudes. (b) Activation of $d=29$ (4th Bott Dirac layer) as a $n=4$ source for the solar mass-squared splitting fails at the structural level: geometric-wall suppression makes the contribution $40$--$1400\times$ too small. (c) Cascade-internal 4-fold structures (four distinguished sources $\{5, 7, 14, 19\}$, four Bott Dirac layers, four Hurwitz dimensions) yield one suggestive numerical near-match for $\Delta m^2_{\rm sol}$ at the percent level, but the cascade has no clean dimensional anchor for the eV$^2$ scale --- closing this candidate as well. The chirality theorem applies at any $n$; nature has only required $n=2$ for $1/\alpha_{\rm em}$ screening. Identifying a cascade observable that exercises $n\geq 4$ remains open, but is not load-bearing on any existing prediction. See Part~IVa~\cite{paper4a} \texttt{rem:bott-tower-open} for the related matter-content gap at $d \geq 29$.

\textbf{(ii) Multi-loop $m\geq 2$.} Step~(B) of the proof gives $\chi^m$ for $m$ independent closed loops via tensor-product orthogonality of the basin trace decomposition; the argument is loop-number agnostic, but mixed-loop interference at $m\geq 2$ has not been explicitly checked. The QED 2-loop $(g{-}2)$ correction is a candidate $m\geq 2$ observable cascade-internally if its topology can be assembled within per-layer locality (Remark~\ref{rem:fermion-gauge-coupling-proposal}); investigation in \texttt{tools/research/cascade\_g\_minus\_2\_investigation.py} shows the cascade reproduces Schwinger's 1-loop result via Theorem~\ref{thm:chirality-selection-rule} at $m-k=0$ ($K_C = \chi^0 = 1$), but the 2-loop term diverges from QED's $A^{(4)} = 0.328$ by a factor $0.15$, suggesting per-layer locality rules out the QED triangle topology at 2-loop. Open: classify which cascade observables admit $m\geq 2$ assembly within per-layer locality, and verify $\chi^m$ exactness for those cases.

\textbf{(iii) Mixed $m\geq 1, k\geq 1$ at higher complexity.} The current closed observables at $m-k=0$ (1-loop $(g{-}2)$) and $m-k=+1$ ($1/\alpha_{\rm em}$ screening) are the simplest non-trivial mixed and pure-loop cases; the theorem covers the general composition (Step~(D)) but no precision observable currently exercises $(m,k) = (2,1)$, $(2,2)$, etc. Open: identify cascade observables in this range and verify the multiplicative composition.

Closure of any of (i)--(iii) would require identification of cascade observables that exercise the higher-$m$ or higher-$n$ slots and verification that per-layer locality is preserved. Per CLAUDE.md Check~7, ``higher-loop precision normalisation via cascade Green's function on $S^{d-1}$'' is not a cascade-native open piece; semiclassical procedures on cascade gauge spheres are out of bounds. The structural prediction $\chi^{(m-k)}$ stands at any $(m,k)$; identifying a cascade observable that lives at higher slots is the cascade-native target.
\item \emph{Partially resolved.} Neutrino masses. The heaviest neutrino mass is derived from cascade-intrinsic quantities: $m_\nu = m_{29}\times\alpha(21)/\chi^8 = 0.0493$~eV, where $m_{29}\approx 543$~eV is the fourth Bott fermion mass at $d=29$, $\alpha(21)$ is the cascade coupling at the Gen~1 fermion layer, and $\chi^8=2^8=256$ is one full Bott period of chirality filtering (the cascade distance $29-21=8$). Observed: $\sqrt{\Delta m^2_{\mathrm{atm}}}=0.0495$~eV (PDG~2024). Residual: $-1.0\%$ on $\Delta m^2$. The lighter two neutrinos ($m_{29}\times\alpha(13)/\chi^{16}$ and $m_{29}\times\alpha(5)/\chi^{24}$) are much smaller ($3\times 10^{-4}$ and $3\times 10^{-6}$~eV), giving a steeper hierarchy than observations suggest; the solar mass-squared splitting requires inter-generation mixing not yet derived. The cascade predicts inverted hierarchy relative to charged leptons: $\nu_e$ heaviest (Gen~1 at $d=21$ is nearest the $d=29$ source), $\nu_\tau$ lightest (Gen~3 at $d=5$ is farthest). Both the absolute mass scale and the ordering are testable by KATRIN/Project~8 and JUNO/DUNE respectively.
\item \emph{Largely resolved.} Error compensation in absolute masses. $\alpha_s$ leading was $\sim\!1.7\%$ low; the universal $\mathrm{U}(1)$ shift $\delta\Phi_{\mathrm{U}(1)}=\alpha(14)/\chi$ (Theorem~\ref{thm:alpha-s-closure}) closes it at $+0.02\sigma$ of experimental precision. $v$ leading was $\sim\!2.2\%$ low; the cascade-native identity $\pi/\alpha(5) = \Omega_2/R(5)^2$ plus Kosterlitz--Thouless half-winding interpretation (Remark~\ref{rem:v-closure}) close the non-perturbative factor cascade-internally up to the formal 4D$\to$2D topology-mapping step. The ``systematic error compensation'' is therefore no longer needed as an explanation: each observable closes via its own cascade-internal shift or identity. The two-population sign rule (descent-dependent vs geometric) remains an empirical classification whose Morse-index derivation is open (Roadmap item~7).
\end{enumerate}

\section{Roadmap for Future Research}

The $\alpha(d^*)/\chi^k$ correction family, the cascade action principle, and the neutrino mass derivation open several concrete research directions, ordered by tractability.

\textbf{Near-term (structural, no new mathematics required):}
\begin{enumerate}
\item \emph{Largely resolved.} Close the electroweak VEV $v$. Remark~\ref{rem:v-closure} establishes the cascade-native identity $\pi/\alpha(5) = \Omega_2/R(5)^2$ and identifies $\exp(-\pi/\alpha(5))$ with the Kosterlitz--Thouless single-vortex action on the 2-sphere cross-section of $S^4$ via the Morse foliation of the $d{=}5$ hairy-ball defect, with half-winding forced by Theorem~\ref{thm:chirality-factorisation}. The identity, the 2-sphere cross-section, and the chirality-half-winding are all cascade-internal. \textbf{Residual:} a formal proof that the 4D hairy-ball tangent-field obstruction on $S^4$ is equivalent to a first-Chern class on each 2-cycle of the Morse foliation --- a pure geometric-topology step, not new cascade content.
\item \emph{Derive the PMNS matrix and the solar neutrino splitting.} \textbf{Partial-negative tests on cascade-native mechanisms} (verifier: \texttt{tools/research/cascade\_pmns\_solar\_splitting.py}). The heaviest neutrino mass is derived ($m_{29}\times\alpha(21)/\chi^8$, $-1\%$ on $\Delta m^2_{\rm atm}$); the diagonal cascade form gives $m_2\sim 3\times 10^{-4}$~eV, undershooting $\sqrt{\Delta m^2_{\rm sol}}=0.0086$~eV by factor $\sim$800. Four cascade-native mechanisms tested for closing the gap: (i)~Cabibbo-style geometric-mean rule (Theorem~\ref{thm:cabibbo-amplitude}) extended to PMNS$_{12}$ gives $\theta_{12}\approx 7.5^\circ$ vs observed $33.4^\circ$~--- way too small; CKM small + PMNS large means the two sectors require different mechanisms. (ii)~Gram-overlap inter-generation mixing $M_{12}=m_1\cdot \mathrm{gram}(13,21)\approx 1.6\times 10^{-4}$~eV vs required $7.6\times 10^{-4}$~eV (factor 4.6 too small, plus invalid because observed mixing is large not small). (iii)~Chirality-basin mixing (Theorem~\ref{thm:chirality-factorisation}) gives qualitatively the right pattern ($\theta_{23}\approx 45^\circ$ maximal, $\theta_{13}$ small) but no quantitative cascade-native bridge to angle structure. (iv)~Modified $m_2$ descent $m_{29}\alpha(d)/\chi^k$ admits no integer $\chi^k$ that lands on $0.0086$~eV. Per Check~7, semiclassical Bogoliubov-mixing / KK-quotient procedures are inadmissible. Genuine open structural piece: a cascade-internal derivation of neutrino-sector mixing yielding (a) large $\theta_{12},\theta_{23}$, (b) small $\theta_{13}$, (c) magnitude $\Delta m^2_{\rm sol}=7.5\times 10^{-5}$~eV$^2$. Most likely route: an inter-layer overlap from the cascade scalar action acting differently on neutrino vs charged-lepton chirality basins, but the existing cascade ingredients (Gram, geometric-mean, $\chi$-factor) tested here do not close the gap.
\item \emph{Resolved.} Derive $1/\alpha_{\mathrm{em}}=137$ from the cascade action. The conjecture $1/\alpha_{\mathrm{em}} = 1/\alpha(13) + \pi/\alpha(14) + 6\pi = 137.028$ (Open Question~\ref{oq:alpha-em-screening}) matches observation to $0.006\%$. All four pieces are now cascade-internally derived: (i)~the bare coupling from Theorem~\ref{thm:weinberg}; (ii)~the chirality factor $\chi$ per Dirac layer from Theorem~\ref{thm:chirality-selection-rule} (closed-loop, $m=1$, $k=0$); (iii)~the per-leg primitive $\Gamma(\tfrac{1}{2})^2$ from Remark~\ref{rem:per-leg-primitive-derivation} (1D Gaussian / proper-time 2-point function on the cascade lattice with canonical $1/\alpha(d)$ wavefunction renormalisation cancelling the layer-dependent prefactor); (iv)~the multiplicative $N_{\mathrm{gen}}=3$ as the count of Dirac layers $\{5,13,21\}$ in the descent, equal to the generation count by Part~IVa~\cite{paper4a}'s three-generation theorem. The structural derivation is complete; the $0.006\%$ residual sits at the cascade's standing leading-order systematic.
\item \emph{Complete the quark mass spectrum.} The ratio $(t/b)/(c/s)=3.04\approx N_c$ suggests an additional colour factor from the Weyl chirality decomposition at $d=12$. Computing this from the cascade's own spinor structure at $S^{11}$ would complete the quark sector.
\end{enumerate}

\textbf{Medium-term (requires new cascade-intrinsic derivations):}
\begin{enumerate}
\setcounter{enumi}{4}
\item \emph{Resolved.} Rigorous derivation of the channel-counting rule from the action. Theorem~\ref{thm:chirality-factorisation} supplies the formal proof that the Green's function factorises into $\chi$ chirality sectors at even-sphere layers (via the $\mathbb{Z}_2$ chirality decomposition, propagator unitarity, and multi-mode independence). No longer open.
\item \emph{Derive the source selection rule.} \textbf{Partially resolved} by Proposition~\ref{prop:source-selection} in Section~\ref{subsec:source-selection}: a deterministic three-flag decision procedure---Planck-anchored ($P$), observer-local ($L$), gauge-mediated ($G$)---maps the four observable types (Absolute, Observer, Gauge, Amplitude) bijectively onto the four non-sink distinguished layers $\{d_1,d_V,d_{\rm gw},d_0\}=\{19,5,14,7\}$, and this classification reproduces the source assignments of all eight closed observables. What remains is (i) a formal derivation of the flags $(P,L,G)$ from a cascade-intrinsic category of observables rather than as physics meta-data, and (ii) a closed-form Green's function computation confirming that the response is maximised at the assigned $d^*$ within each type.
\item \emph{Derive the sign rule from the Morse index.} The descent/geometric sign split currently follows the paper's empirical two-population classification. A Morse-theoretic derivation from the cascade action would make it structural.
\item \emph{Partially addressed.} The cascade's non-perturbative sector at $d=5$. Remark~\ref{rem:v-closure} identifies $\exp(-\pi/\alpha(5))$ with the Kosterlitz--Thouless single-vortex action on a 2-sphere cross-section of $S^4$, not a classical instanton (the action at any genuine saddle would give integer winding, not the chirality-half $n_{\mathrm{eff}}=1$ the closure requires). What remains open at the action level: whether the cascade action admits a family of finite-width vortex configurations indexed by the Morse-foliation latitude, matching the KT vortex interpretation dynamically. Connection to the neutrino mass mechanism (if any) is a separate open question.
\end{enumerate}

\textbf{Long-term (foundational):}
\begin{enumerate}
\setcounter{enumi}{8}
\item \emph{Second-quantise the cascade action.} The current action $S[\varphi]=\sum(2\alpha(d))^{-1}(\Delta\varphi)^2$ is a classical field theory on the cascade lattice. A second-quantised version---with the cascade field $\varphi$ promoted to an operator---would generate the source strengths, signs, and channel counts from the action's own quantum structure rather than from structural classification.
\item \emph{Derive time and the Lorentzian signature from the cascade action.} Part~III derives Lorentzian signature from the oscillatory propagator; the cascade action should reproduce this as a Wick rotation of the discrete Laplacian.
\item \emph{Verify the dissolution of quantum gravity.} The cascade derives QM and GR from the same source (Part~II\,=\,III): both are projections of the cascade's geometry, not independent theories requiring reconciliation. The cascade action $S[\varphi]$ should make this dissolution explicit by generating both Gleason's theorem (forcing the Born rule) and Lovelock uniqueness (forcing Einstein's equation) from a single variational principle, demonstrating that no separate ``quantum gravity'' theory is needed.
\end{enumerate}

\section{Confidence Assessment}

\textbf{Two-population systematic and its first-order correction.} At leading order, descent-dependent quantities carry uniformly negative deviations ($-0.13\%$ to $-2.20\%$, 7/7 negative), while geometric quantities carry positive deviations ($+0.56\%$ to $+2.76\%$, 4/4 positive). The Part~0 Supplement derives a first-order correction $\delta Q/Q_0 = \sum_{\mathrm{adj}}(1-C_{d,d+1}^2)$ --- a sum of Beta function ratios with no free parameters --- that reduces descent deviations from $\sim\!2\%$ to sub-$1\%$: $\alpha_s$ ($-1.7\% \to -0.5\%$), $m_\tau$ ($-1.2\% \to -0.1\%$), $m_\tau/m_\mu$ ($-1.7\% \to -0.8\%$), $v$ ($-2.2\% \to -1.0\%$). Geometric quantities are unchanged (they carry no descent correction). A second-order observable-dependent factor remains open (Section~15.8.1 of the Supplement).

\textbf{Tier 1: Theorem-level results.} (a) The geometric-topological factorization, with $R'(d)$ excluded at 61--69\%. (b) The topological obstruction factor $1/(2\sqrt{\pi})$ derived from the chirality decomposition ($\chi = 2$) and the quarter-turn obstruction ($\sqrt{\pi}$) at the hairy ball zero (Theorem~\ref{thm:obstruction}), and equivalently expressed as the pure cascade identity $2\sqrt{\pi} = N(0)\cdot\Gamma(\tfrac{1}{2})$ (Corollary~\ref{cor:2sqrtpi-primitive}), ruling out any Dirac-operator interpretation by the vanishing of $\hat A(S^{2n})$ (Remark~\ref{rem:no-dirac-route}). (c) The Weinberg angle from the boson topological obstruction: SU(2) sits at the Dirac layer $d=13$ (no propagator factor; the obstruction is the symmetry breaking), U(1) at $d=14$ crosses $d=13$ during descent and picks up the boson-only quarter-turn factor $1/\sqrt{\pi}$, giving $\tan\theta_W = (N(14)/\sqrt{\pi})/N(13)$ and $\sin^2\theta_W = 0.2286$ (Theorem~\ref{thm:weinberg}). (d) $1/\alpha_{\text{GUT}} = 4\pi/N(12)^2 = 25.02$ from the Gamma function. (e) Adams' theorem gives $N_c = 3$ and $\dim\,\text{U}(1) = 1$. (f) The hairy ball theorem gives SU(3) $\times$ U(1) unbroken, SU(2) broken. (g) Three generations from Bott periodicity + the $d_1 = 19$ phase transition. (h) $w = -1$ exactly, with GB corrections vanishing by two independent mechanisms ([3]). (i) Universal $\mathrm{U}(1)$-layer shift to the cascade potential $\delta\Phi_{\mathrm{U}(1)}=\alpha(14)/\chi=R(14)^2/8=429^2\,\pi/2^{25}$ (Theorem~\ref{thm:alpha-s-closure}), applied exponentially: a single constant closes $\alpha_s(M_Z)=0.117917$ at $+0.019\sigma$ and $m_\tau/m_\mu=16.81731$ at $+0.243\sigma$ of experimental precision, equivalently $\alpha(12)\cdot(13/14)^2/2$ by the exact $\Gamma$ recursion $R(d+2)/R(d)=(d+1)/(d+2)$. (j) Phase-transition shift $\delta\Phi_{\mathrm{phase}}=\alpha(19)/\chi=R(19)^2/8$ sourced at the Paper~I threshold $d_1=19$ (Theorem~\ref{thm:mtau-abs-closure}), closing the $\tau$ absolute mass $m_\tau=1776.82$~MeV to $-0.31\sigma$ of experimental precision. (k) Observer-host shift $\delta\Phi_{\mathrm{obs}}=\alpha(5)/\chi^3=8/(225\pi)$ sourced at the volume maximum $d_V=5$ with three factors of the Euler characteristic (Theorem~\ref{thm:weinberg-closure}), closing $\sin^2\theta_W=0.231226$ to $+0.40\sigma$ of experimental precision. A systematic search (Remark~\ref{rem:phase-family}) extends the family to eight SM precision observables at four distinguished source layers: (l)~$\ell_A$ reuses $\alpha(19)/\chi$ ($-0.16\sigma$); (m)~$\Omega_m(1/\pi)$ uses $-\alpha(5)/\chi^3$ ($-0.04\sigma$); (n)~$\theta_C$ uses $-\alpha(7)/\chi^2$ ($+0.03\sigma$) with source $d_0=7$ (Paper~I area maximum); (o)~$b/s$ uses $-\alpha(7)/\chi^4$ ($0.014\%$) at the same source $d_0=7$ with $k=4$ (Theorem~\ref{thm:bs-closure}). Together, (i)--(o) close eight SM precision observables using one structural form $\pm\alpha(d^*)/\chi^k$ at four distinguished cascade layers and zero fitted parameters. Three reuse pairs strengthen the evidence: $\{\alpha_s,m_\tau/m_\mu\}$, $\{m_\tau\text{ abs},\ell_A\}$, $\{\sin^2\theta_W,\Omega_m\}$.

\textbf{Tier 2: Fully derived predictions.} (a) $m_\mu/m_e = 206.50$ (0.13\%, no shift applies: path starts at $d=14$). (b) $\alpha_s(M_Z)=0.117917$ and $m_\tau/m_\mu=16.81731$ closed within experimental precision via the universal $\mathrm{U}(1)$ shift $\delta\Phi_{\mathrm{U}(1)}=\alpha(14)/\chi$ (Theorem~\ref{thm:alpha-s-closure}). (c) $m_\tau = 1776.82$~MeV absolute closed within experimental precision via the phase-transition shift $\delta\Phi_{\mathrm{phase}}=\alpha(19)/\chi$ (Theorem~\ref{thm:mtau-abs-closure}). (d) $\sin^2\theta_W=0.231226$ closed within experimental precision via the observer-host shift $\delta\Phi_{\mathrm{obs}}=\alpha(5)/\chi^3$ (Theorem~\ref{thm:weinberg-closure}). (e) $\theta_C=13.04^\circ$ closed to $+0.03\sigma$ via $-\alpha(7)/\chi^2$ sourced at the Paper~I area maximum $d_0=7$. (f) $\ell_A=301.44$ closed to $-0.16\sigma$ via $\alpha(19)/\chi$, reusing the phase-transition shift. (g) $\Omega_m(1/\pi)=0.31473$ closed to $-0.04\sigma$ via $-\alpha(5)/\chi^3$, reusing the observer-host shift with opposite sign (geometric population). (h) $C = \alpha_s/(2\sqrt{\pi})$ (1.0\%). (i) Absolute masses $m_\mu$ (0.47\%) and $m_e$ (0.60\%), consistent with the chain-subtracted inherited shift $\delta\Phi_{\mathrm{phase}}-\delta\Phi_{\mathrm{U}(1)}$ to $\sim 50$--$70$~ppm. (j) $b/s = 44.7436$ closed to $0.014\%$ via $-\alpha(7)/\chi^4$ (Theorem~\ref{thm:bs-closure}), reusing the area-maximum source layer with $k=4$ for cross-generation quark structure.

\textbf{Tier 3: Derived predictions with caveats.} (a) $m_H/m_W = \pi/2$: the geodesic distance argument is natural but the mass-ratio/geodesic identification is stated rather than computed from an action. (b) $v = 240.8$~GeV and (c) the $\exp(-\pi/\alpha(5))$ factor in Theorem~4.7: the cascade-native identity $\pi/\alpha(5) = \Omega_2/R(5)^2$ is verified and matches the Kosterlitz--Thouless single-vortex action on the 2-sphere cross-section of $S^4$ via the Morse foliation of the $d = 5$ hairy-ball defect (Remark~\ref{rem:v-closure}); the half-winding follows from Theorem~\ref{thm:chirality-factorisation}'s chirality basin filter with effective winding $n_{\mathrm{eff}} = \chi(S^4)/\chi = 1$. Open: formal proof that the 4D tangent-field obstruction on $S^4$ is equivalent to a first-Chern class on the 2-cycles of the Morse foliation (pure geometric topology, not cascade content). (d)~$1/\alpha_{\mathrm{em}} = 1/\alpha(13) + \pi/\alpha(14) + 6\pi = 137.028$ ($0.006\%$): all four pieces fully cascade-internally derived. (i)~Bare coupling $1/\alpha(13) + \pi/\alpha(14) = 118.18$ from Theorem~\ref{thm:weinberg} and the electroweak mixing. (ii)~Per-Dirac-layer chirality factor $\chi$ from Theorem~\ref{thm:chirality-selection-rule} (closed-loop $m=1$, $k=0$, chirality factor $\chi^{m-k}=\chi$). (iii)~Per-leg primitive $\Gamma(\tfrac{1}{2})^2 = \pi$ from Remark~\ref{rem:per-leg-primitive-derivation}, with the canonical $1/\alpha(d)$ wavefunction-renormalisation step grounded in Remark~\ref{rem:wavefunction-renorm-canonical} via the chain action-uniqueness $\to$ marginal Green's function $\to$ $\sqrt{\alpha}$-universality. (iv)~$N_{\mathrm{gen}}=3$ as the count of Dirac layers in the descent $\{5,13,21\}$, equal to the generation count by Part~IVa~\cite{paper4a}'s three-generation theorem (Open Question~\ref{oq:alpha-em-screening}). No structural gap remains; the $0.006\%$ residual is the cascade's standing leading-order systematic. (\emph{Note: $\sin^2\theta_W$ is now Tier 1 via Theorem~\ref{thm:weinberg}; previously here as a standard-RG proxy.})

\textbf{Tier 4: Observed patterns needing derivation.} (a) $m_b/m_\tau = e$: deviation 1.05\%. (b) $(t/b)/(c/s) = N_c = 3$: physical argument plausible but chirality coupling not computed. ($b/s$ has been promoted to Tier~2 via Theorem~\ref{thm:bs-closure}.)

\textbf{Tier 4b: Claimed but proof has gaps.} $\theta_{\text{QCD}} = 0$ from $\pi_3(S^{11}) = 0$: the value of the homotopy group is now correctly stated (the trivial group, by $\pi_n(S^k) = 0$ for $n < k$; an earlier draft asserted $\mathbb{Z}_2$, now corrected --- see Remark~\ref{rem:strong-cp-correction}). Under the corrected value, only the trivial sector exists in the cascade-internal classification, so the gauge-sector $\theta_{\text{QCD}} = 0$ is unique without a selection step. The remaining gap is structural: the claim that the cascade's topological sectors are classified by $\pi_3(S^{11})$ rather than $\pi_3(\text{SU}(3))$ requires showing how the vector-field realisation of SU(3) on $S^{11}$ replaces (rather than coexists with) the standard gauge-target classification. The mass-matrix branch $\arg\det(M) = 0$ vs.\ $\pi$ is selected by the cascade's $U(1)$ gauge equivalence (Step~2 of Theorem~7.1).

\textbf{What would change my mind.} The framework would be significantly weakened by: (i) discovery of a QCD axion; (ii) $w \neq -1$ at $> 5\sigma$ from BAO alone with $r_d$ independently constrained at $> 144$~Mpc; (iii) discovery of a fourth fermion generation; (iv) a demonstration that the topological obstruction factor is not $2\sqrt{\pi}$ (now derived in Theorem~\ref{thm:obstruction} and equivalently in Corollary~\ref{cor:2sqrtpi-primitive}; falsification would require either invalidating the chirality/quarter-turn argument or invalidating the cascade-primitive identity $2\sqrt{\pi} = N(0)\cdot\Gamma(\tfrac{1}{2})$); or (v) any absolute mass prediction deviating by $> 5\%$ after subleading corrections are included. Conversely, the framework would be strongly supported by: (vi) derivation of the screening $2\pi$ per generation in the $\alpha_{\mathrm{em}}$ conjecture (Open Question~\ref{oq:alpha-em-screening}) from the cascade action, which would close $1/\alpha_{\mathrm{em}}$ at $0.006\%$; (vii) derivation of the colour factor $e$ from first principles; or (viii) any new precision prediction matching observation before the data is checked.

\begin{thebibliography}{13}

\bibitem{paper1} {[Author]}, \textit{Scale Variance from Orthogonality: How the Unit Ball Generates $10^{120}$ Orders of Magnitude}, March 2026.

\bibitem{paper2} {[Author]}, \textit{Quantum Mechanics from the Cascade: Effective Theory of a 4-Dimensional Observer in the Sphere-Area Geometry}, March 2026.

\bibitem{paper3} {[Author]}, \textit{General Relativity from the Cascade: Lovelock Uniqueness and the Cosmological Constant of a 4-Dimensional Observer}, March 2026.

\bibitem{paper4a} {[Author]}, \textit{The Standard Model from the Cascade: Part IVa---Gauge Group, Symmetry Breaking, and Three Generations from Bott Periodicity and Hairy Ball Zeros}, March 2026.

\bibitem{lounesto} P.~Lounesto, \textit{Clifford Algebras and Spinors}, 2nd ed., Cambridge University Press, 2001.

\bibitem{abs} M.~F.~Atiyah, R.~Bott, and A.~Shapiro, ``Clifford modules,'' \textit{Topology} \textbf{3}, 3--38 (1964).

\bibitem{adams} J.~F.~Adams, ``Vector fields on spheres,'' \textit{Annals of Mathematics} \textbf{75}, 603--632 (1962).

\bibitem{poincare} H.~Poincar\'e, ``Sur les courbes d\'efinies par les \'equations diff\'erentielles,'' \textit{Journal de Math\'ematiques Pures et Appliqu\'ees} \textbf{4}(1), 167--244 (1885).

\bibitem{hopf} H.~Hopf, ``Vektorfelder in $n$-dimensionalen Mannigfaltigkeiten,'' \textit{Mathematische Annalen} \textbf{96}, 225--249 (1927).

\bibitem{hurwitz} A.~Hurwitz, ``\"Uber die Komposition der quadratischen Formen,'' \textit{Mathematische Annalen} \textbf{88}, 1--25 (1923).

\bibitem{gj} H.~Georgi and C.~Jarlskog, ``A new lepton-quark mass relation in a unified theory,'' \textit{Physics Letters B} \textbf{86}, 297--300 (1979).

\bibitem{milnor} J.~W.~Milnor, \textit{Topology from the Differentiable Viewpoint}, Princeton University Press, 1965.

\bibitem{desi} DESI Collaboration, \textit{DESI DR2 Results II: Measurements of Baryon Acoustic Oscillations and Cosmological Constraints}, arXiv:2503.14738 (2025).

\end{thebibliography}

\end{document}
