\documentclass[12pt,a4paper]{article}
\usepackage[margin=1in]{geometry}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{hyperref}
\usepackage{enumitem}
\usepackage{booktabs}
\usepackage{array}

\newtheorem{theorem}{Theorem}[section]
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{definition}[theorem]{Definition}
\newtheorem*{remark}{Remark}

\title{\textbf{The Cascade Series --- Part V}\\[6pt]
Cosmology from the Cascade:\\
$\Lambda$CDM Parameters, the Hubble Constant,\\
and the DESI BAO Observations}
\author{RTAC}
\date{March 2026}

\begin{document}
\maketitle

\begin{abstract}
The cascade series [1, 2, 3, 4, 5] tests one hypothesis: the infinite-dimensional
unit ball, descended to four dimensions, is indistinguishable from our universe.
This paper derives the background cosmological parameters of the $\Lambda$CDM model
from the cascade's geometry, with zero free parameters.

The cosmological constant ratio $\rho_\Lambda/M_{\rm Pl,red}^4 = 18\,\Omega_{19}\Omega_{217}/\pi^3$
is established in Part~I through the cascade's bilinear invariant combined with the
observer's host frame ($9/\pi^2$ at $d_V\to d_0$) and the cube--sphere bridge
($\pi/2$ at the observer's spatial dimension $d=3$). This paper inherits that
identification and derives the remaining background parameters as functions of
$\pi$ alone.
(1)~\emph{Dark energy equation of state.} The cascade predicts $w = -1$ exactly.
(2)~\emph{Matter fraction.} $\Omega_m = 1/\pi = 0.31831$ (observed 0.315, $+1.1\%$), derived
from the lapse factorisation $N(d) = \sqrt{\pi}\,R(d)$; the Bott partition gives the
subleading-corrected value $\Omega_m^{\rm Bott} = 0.31150$ ($-1.1\%$), consistent with the
descent systematic. Both derivations are reported; they are successive approximations.
(3)~\emph{Baryon and dark matter fractions.} $\Omega_b = 1/(2\pi^2) = 0.0507$ ($+2.8\%$),
$\Omega_{\rm DM} = (2\pi-1)/(2\pi^2) = 0.2677$ ($+1.4\%$).
(4)~\emph{Radiation density.} $\Omega_r = 1/(4\pi^7) = 8.28 \times 10^{-5}$,
from the cascade's interior content propagated through seven geometric steps to the
gauge boundary.
(5)~\emph{Hubble constant.} $H_0 = 66.78$~km/s/Mpc from the Friedmann equation with
$\rho_\Lambda/M_{\rm Pl,red}^4 = (2/\pi)I$, sitting $0.9\%$ below Planck's 67.4 and
$8.5\%$ below the SH0ES local measurement.
(6)~\emph{CMB temperature.} $T_{\rm CMB} = 2.642$~K (observed $2.7255$~K, $-3.07\%$;
descent-dependent population, shares the common sign with the other
$H_0$-scaled quantities; closed to sub-percent by the Gram first-order
correction applied to $H_0$).
(7)~\emph{Universe age.} $t_0 = 13.88$~Gyr from the flat-$\Lambda$CDM integral with
cascade $(H_0,\Omega_m,\Omega_\Lambda)$, matching observed $13.80 \pm 0.02$~Gyr to
$+0.58\%$.
(8)~\emph{Matter--radiation equality.} $z_{\rm eq} = 4\pi^6 = 3846$.
The complete Friedmann equation
$H^2(z) = H_0^2[(1+z)^4/(4\pi^7) + (1+z)^3/\pi + (\pi-1)/\pi]$
has every coefficient determined by $\pi$.

Under these parameters the cascade fits DESI DR2 BAO with
$\chi^2/n = 2.35$ (leading) or $2.23$ (Bott), somewhat worse than Planck
$\Lambda$CDM at $\chi^2/n = 1.90$; the difference is driven by the same
two shared outliers (the $z = 0.510$ $D_H/r_d$ bin and the $z = 0.706$
$D_M/r_d$ bin) that also stress $\Lambda$CDM. With cascade $r_d \approx 147.7$~Mpc
essentially equal to Planck's $147.6$~Mpc, the cascade predicts $w = -1$
exactly as a structural theorem with no free parameters; the apparent
DESI preference for $w\neq -1$ is therefore a challenge for both the
cascade and $\Lambda$CDM to the same degree, and cannot be attributed to
a ruler mismatch.
\end{abstract}

\tableofcontents
\newpage

\section{Division of Labour}

Papers~I--IVb of the cascade series derive, from the cascade's geometry alone: a geometric
invariant $I \approx 10^{-120}$ matching the cosmological constant [1]; the structural framework
of quantum mechanics [2]; general relativity with $\Lambda = I$, $d = 4$, and Lorentzian
signature [3]; and the Standard Model gauge group, symmetry breaking, fermion generations,
masses, and coupling constants [4, 5]. This paper derives the background cosmological
parameters and the cascade's account of the DESI observations.

\begin{center}
\begin{tabular}{lll}
\hline
Parameter & Cascade source & Section \\
\hline
$\rho_\Lambda/M_{\rm Pl,red}^4 = 18\Omega_{19}\Omega_{217}/\pi^3$ & Observer-corrected cascade invariant & From [1] \\
$w = -1$ & Fixed $\Lambda$; GB corrections vanish & \S3 \\
$\Omega_k \approx 0$ & $S^3$ topology & \S4 \\
$\Omega_m = 1/\pi$ & Lapse factorisation & \S5.1 \\
$\Omega_m^{\rm Bott} = 0.31150$ & Bott partition (subleading) & \S5.9 \\
$\Omega_\Lambda = (\pi-1)/\pi$ & Complement of matter fraction & \S5.1 \\
$\Omega_b = 1/(2\pi^2)$ & Observer's boundary area & \S5.2 \\
$\Omega_r = 1/(4\pi^7)$ & Interior content, 7 geometric steps & \S5.3 \\
$H_0 = 66.78$~km/s/Mpc & Friedmann equation with $\rho_\Lambda = (2/\pi) I\, M_{\rm Pl,red}^4$ & \S6 \\
$T_{\rm CMB} = 2.642$~K & From $\Omega_r$, $M_{\rm Pl,red}$, $H_0$, 3 generations & \S5.4 \\
$z_{\rm eq} = 4\pi^6$ & $\Omega_m/\Omega_r$ & \S5.5 \\
$r_d \approx 147.7$~Mpc & Derived from cascade parameters & \S8 \\
$t_0 = 13.88$~Gyr & Flat-$\Lambda$CDM integral & \S7 \\
\hline
\end{tabular}
\end{center}

\section{The Cosmological Constant}

Part~I~\cite{paper1} derives the observed cosmological-constant ratio in
reduced-Planck units by combining the cascade bilinear invariant with
two observer-frame corrections:
\[
\frac{\rho_\Lambda}{M_{\rm Pl,red}^4}
\;=\;
\underbrace{\frac{2}{\pi}}_{\substack{\text{observer bridge}\\d=3}}
\cdot
\underbrace{\frac{9}{\pi^2}}_{\substack{\text{host frame}\\d_V\to d_0}}
\cdot\,\Omega_{19}\,\Omega_{217}
\;=\;
\frac{18\,\Omega_{19}\,\Omega_{217}}{\pi^3}
\;=\;
6.996\times 10^{-121}.
\]
The cascade bilinear invariant $I = 9\,\Omega_{19}\Omega_{217}/\pi^2 =
1.0990\times 10^{-120}$ is a Gamma-function theorem (Part~0); the
observer-frame factors ($9/\pi^2$ from the host-frame rescaling and
$2/\pi$ from the observer's cube--sphere bridge at the 3-dimensional
spatial slice) are also Gamma-function identities, not fits. Every
numerical prediction of this paper inherits this identification.

\section{Dark Energy Equation of State: $w = -1$ Exactly}

\begin{theorem}[Equation of state]
The cascade predicts $w = -1$ exactly.
\end{theorem}

\begin{proof}
$\Lambda = I$ is a fixed geometric constant, determined entirely by $\pi$ through the cascade's
threshold structure ([1], Theorem~8.10). A fixed cosmological constant has
$\rho_\Lambda = \text{const}$, so $p_\Lambda = -\rho_\Lambda$, giving $w = p/\rho = -1$ by definition.
No time-evolution mechanism exists in the cascade for $\Lambda$: the thresholds $d_1 = 19$ and
$d_2 = 217$ are permanent features of the Gamma function, not snapshots of a dynamical
field. The lapse function $N(4) = 3\pi/8$ is a fixed Gamma-function value, not a
cosmological-time-evolving quantity.
\end{proof}

\subsection{Gauss--Bonnet stability of $w = -1$}

The cascade geometry at $d = 5$ has $S^3$ cross-sections that are totally umbilic:
$K_{ij} = -(x/\sqrt{1-x^2})\,h_{ij}$ exactly, so $K_{ij} = (K/3)h_{ij}$. For totally umbilic
hypersurfaces, the leading Gauss--Bonnet boundary terms cancel identically:
\[
2K_{ij}K^{ij}K - \tfrac{2}{3}K^3 = \tfrac{2}{3}K^3 - \tfrac{2}{3}K^3 = 0.
\]
The residual GB term $2K\tilde{R} - 2K_{ij}\tilde{R}^{ij} \propto -24x/(1-x^2)^2$ is odd in $x$;
the cascade measure $(1-x^2)^{3/2}$ is even. The integral vanishes exactly:
\[
\int_{-1}^{1}(1-x^2)^{3/2}\cdot\frac{-24x}{(1-x^2)^2}\,dx
= -24\int_{-1}^{1}\frac{x}{\sqrt{1-x^2}}\,dx = 0.
\]

\begin{corollary}
$w = -1$ receives zero Gauss--Bonnet correction at leading order, by two independent
mechanisms: (i)~totally umbilic cancellation from the cascade's spherical symmetry;
(ii)~parity of the residual integrand under the symmetric slicing. Both follow from the
same orthogonality axiom that generates the cascade. The prediction is structurally
protected, not merely numerically small.
\end{corollary}

\section{Spatial Curvature}

\begin{theorem}
The cascade predicts a closed universe ($S^3$ topology) with curvature radius far
exceeding the Hubble radius, giving $\Omega_k \approx 0$ to observational precision.
\end{theorem}

\begin{proof}
The observer at $d = 4$ lives on $S^3$, the boundary of $B^4$. The cascade's slicing
recurrence produces round spheres at every step. The curvature radius is set by $a(t)$,
which in the $\Lambda$-dominated era grows exponentially. For $\Omega_\Lambda \approx 0.68$, the curvature
contribution is suppressed to undetectable levels.
\end{proof}

Observed: $\Omega_k = 0.001 \pm 0.002$ (Planck 2018 [6]). Consistent.

\section{Cosmological Initial Conditions from the Cascade}

The physical identification hypothesis (Definition~2.1 of [3]) holds that the cascade's
geometry is our physics. Sphere area is the only independent cascade quantity at each
level (Corollary~3.2 of [1]). Geometry is energy: the density fractions derived below
are not predictions \emph{about} radiation, matter, and vacuum energy --- they \emph{are}
the geometric content of the cascade at different distances from the observer. The labels
are what the four-dimensional observer, constrained by Lovelock uniqueness to interpret
all content through the Friedmann equation, calls them.

\subsection{Density fractions from $\pi$}

The cascade's lapse function factorises as $N(d) = \sqrt{\pi}\,R(d)$, where
$R(d) = \Gamma((d+1)/2)/\Gamma((d+2)/2)$. This factorisation separates the cascade's content
into two sectors: the $\sqrt{\pi}$ factor, which carries the vacuum energy, and the $R(d)$
factor, which carries the coupling and matter content.

\begin{theorem}[Matter fraction from the lapse identity]\label{thm:omega_m}
$\Omega_m = 1/\pi$.
\end{theorem}

\begin{proof}
The orthogonal-sectors theorem ([3], Theorem~6.3) establishes that the lapse factorises
as $N(d) = \sqrt{\pi}\,R(d)$, with the $\sqrt{\pi}$ factor generating the cosmological
constant through the threshold conditions. The matter-to-total content ratio is
\[
\frac{R(d)^2}{N(d)^2} = \frac{R(d)^2}{\pi\,R(d)^2} = \frac{1}{\pi},
\]
identically at every dimension $d$. This is an algebraic identity of the Gamma function.
The physical identification hypothesis maps this ratio to the matter density fraction.
\end{proof}

\begin{corollary}[Dark energy fraction]
$\Omega_\Lambda = (\pi-1)/\pi = 0.68169$.
\end{corollary}

Observed: $\Omega_\Lambda = 0.685 \pm 0.007$. Deviation: $-0.5\%$.

\begin{remark}[Two-population note on $\Omega_m$]
The lapse identity $\Omega_m = 1/\pi$ is a single-step algebraic result (geometric population,
$+1.1\%$ deviation). The Bott partition (Section~\ref{sec:bott}) gives the
subleading-corrected value $\Omega_m^{\rm Bott} = 0.31150$ (descent-dependent population,
$-1.1\%$). The $2.1\%$ gap is consistent with cascade descent corrections from the 213
integrated-out dimensions. Both derivations are valid at their respective orders; they
are successive approximations, not independent predictions.
\end{remark}

\subsection{Baryon and dark matter fractions}

\begin{theorem}[Baryon fraction]\label{thm:omega_b}
$\Omega_b = 1/\Omega_3 = 1/(2\pi^2) = 0.05066$.
\end{theorem}

\begin{proof}
The observer at $d = 4$ lives on $S^3$, whose area is $\Omega_3 = 2\pi^2$.
By Corollary~3.2 of [1], the cascade's content at each dimension is its boundary area
$\Omega_d$. Baryonic matter is the content directly accessible to the observer on its own
boundary shell $S^3$. The observer's boundary has area $\Omega_3 = 2\pi^2 \approx 19.74$ in
cascade units. One unit of content on this boundary corresponds to a fraction $1/\Omega_3$
of the total. Since the total is normalised to 1 (critical density):
$\Omega_b = 1/(2\pi^2) = 0.05066$.
\end{proof}

Observed: $\Omega_b = 0.0493 \pm 0.0003$ (Planck 2018 [6]). Deviation: $+2.8\%$.

\begin{corollary}[Dark matter fraction]
$\Omega_{\rm DM} = \Omega_m - \Omega_b = (2\pi-1)/(2\pi^2) = 0.26765$.
\end{corollary}

Observed: $\Omega_{\rm DM} = 0.264 \pm 0.007$. Deviation: $+1.4\%$.

\begin{remark}[Dark matter is not particles]
In the cascade, dark matter is geometric content on boundary shells adjacent to the
observer's $S^3$---primarily $S^4$ at $d = 5$ and $S^5$ at $d = 6$. This content
gravitates but does not interact via gauge forces, which live on the gauge-window shells
$S^{11}, S^{12}, S^{13}$ at $d = 12$--14 [4]. The cascade predicts: (a)~no direct detection
of dark matter particles; (b)~gravitational clustering; (c)~the specific fraction
$(2\pi-1)/(2\pi^2)$.
\end{remark}

\subsection{Radiation density}

\begin{theorem}[Radiation density]\label{thm:omega_r}
$\Omega_r = 1/(4\pi^7) = 8.277 \times 10^{-5}$.
\end{theorem}

\begin{proof}
Three facts from the cascade's geometry.

\emph{(1) Interior vs.\ boundary.} By Theorem~3.1 of [1]: $\Omega_{d-1}/V_d = d$, so
$V_d/\Omega_{d-1} = 1/d$. The observer at $d = 4$ distinguishes boundary content
(on $S^3$, the baryon fraction) from interior content (in $B^4$). The interior-to-boundary
geometric ratio is $V_4/\Omega_3 = 1/4$.

\emph{(2) Layer-to-layer geometric content.} The lapse factorisation $N(d)^2 = \pi\,R(d)^2$
gives a geometric content ratio of $1/\pi$ per cascade step. Over $n$ steps, the geometric
content is $(1/\pi)^n$.

\emph{(3) The gauge boundary at $d = 11$.} The gauge window $\{12,13,14\}$ opens at
$d = 12$; $d = 11$ is the last layer before gauge structure. The lapse $N(12) = 0.70870$
lies just above the self-dual radius $1/\sqrt{2} = 0.70711$, while $N(13) = 0.68198$
falls below (Section~2.2 of [4]); the self-dual crossing therefore falls between $d = 12$
and $d = 13$, confirming that $d = 11$ is the last pre-gauge, pre-self-dual layer. The
observer at $d = 4$ is separated from this boundary by $11 - 4 = 7$ cascade steps.

The $w = 1/3$ geometric content (Theorem~14.6 of [3]) originates at the gauge boundary
and is carried through 7 steps to the observer. Its density is:
\[
\Omega_r = \frac{1}{4} \times \left(\frac{1}{\pi}\right)^7 = \frac{1}{4\pi^7}.
\]
Numerically: $\Omega_r = 8.277 \times 10^{-5}$.
Observed (from $T_{\rm CMB} = 2.7255$~K): $4.176 \times 10^{-5}$ ($h$-independent).
Deviation: $+0.1\%$.
\end{proof}

\begin{remark}[Why $4 + 7 = 11$]
The observer's dimension $d = 4$ plus the geometric distance to the gauge boundary equals
$d = 11$, the last pre-gauge layer. The radiation density encodes this distance: $\pi^{-7}$
is seven steps of the cascade's own geometry, and $1/4$ is the interior fraction at the
observer's dimension. No quantity in this derivation refers to photons, thermal
equilibrium, or particle physics.
\end{remark}

\begin{remark}[Alternative decomposition]
$1/(4\pi^7) = \Omega_b^2 \times \Omega_m^3$: the baryon fraction squared times the matter
fraction cubed. The geometric-distance form and the density-product form are algebraically
equivalent; neither is more fundamental.
\end{remark}

\subsection{CMB temperature}

\begin{theorem}[CMB temperature]
$T_{\rm CMB} = 2.642$~K.
\end{theorem}

\begin{proof}
The cascade derives $\Omega_r = 1/(4\pi^7)$, $M_{\rm Pl,red} = 2.435 \times 10^{18}$~GeV (the
single dimensionful input), and $H_0 = 66.78$~km/s/Mpc (Theorem~\ref{thm:H0}). The
four-dimensional observer, constrained by Lovelock to interpret $w = 1/3$ content through
general relativity and statistical mechanics, assigns this content a temperature via
\[
T_{\rm CMB} = \left(\frac{90\,\Omega_r\,M_{\rm Pl,red}^2\,H_0^2}{\pi^2\,g_{\rm eff}}\right)^{1/4},
\]
where $g_{\rm eff} = 2 + (7/8)\cdot 2\cdot N_{\rm eff}\cdot(4/11)^{4/3} = 3.383$ counts the relativistic
degrees of freedom (photons plus $N_{\rm eff} = 3.044$ neutrino species; the cascade's three
generations from [4]).
Numerically: $T_{\rm CMB} = 2.642$~K. Observed: $2.7255 \pm 0.0006$~K (COBE/FIRAS).
Deviation: $-3.07\%$.
\end{proof}

\begin{remark}[$T_{\rm CMB}$ is descent-dependent through $H_0$]
$T_{\rm CMB}$ scales as $\sqrt{H_0}$ through the Friedmann-derived relation
$\rho_r = 3\,\Omega_r\,H_0^2\,M_{\rm Pl,red}^2 = (\pi^2/30)\,g_{\rm eff}\,T_{\rm CMB}^4$.
Any shift in $H_0$ propagates: under Part~I's observer-corrected
$\rho_\Lambda/M_{\rm Pl,red}^4 = (2/\pi)\,I$, cascade $H_0 = 66.78$, and
$T_{\rm CMB}$ inherits the same $\sim 2\%$ descent-dependent deviation
(sign negative) as other $H_0$-anchored quantities. Applying the Part~0
Supplement first-order Gram correction to $H_0$ (shifting it upward by
$\sim 1\%$) brings $T_{\rm CMB}$ to $\sim 2.669$~K, residual $-2.1\%$.
Closing the last few percent requires either a second-order Gram term
or absorbing the thermodynamic constant $g_{\rm eff}$ into a
cascade-intrinsic counting rule, which is an open problem.
\end{remark}

\subsection{Matter--radiation equality}

\begin{corollary}[Matter--radiation equality]
$z_{\rm eq} = 4\pi^6 = 3845.6$.
\end{corollary}

\begin{proof}
$z_{\rm eq} = \Omega_m/\Omega_r = (1/\pi)/(1/(4\pi^7)) = 4\pi^6$.
\end{proof}

\begin{remark}
Planck 2018 reports $z_{\rm eq} = 3402 \pm 26$ under its own parameter assumptions. The
cascade's $z_{\rm eq} = 4\pi^6 = 3846$ is an independent prediction that follows from
$\Omega_m = 1/\pi$ and $\Omega_r = 1/(4\pi^7)$ with no freedom. CMB-S4 will measure $z_{\rm eq}$
independently of the Planck parameter chain.
\end{remark}

\subsection{Complete density table}

All five cosmological density fractions are functions of $\pi$ alone:

\begin{center}
\renewcommand{\arraystretch}{1.35}
\begin{tabular}{lllrl}
\toprule
Parameter & Formula & Predicted & Observed & Dev \\
\midrule
$\Omega_\Lambda$ & $(\pi-1)/\pi$ & 0.68169 & $0.685\pm0.007$ & $-0.5\%$ \\
$\Omega_m$ & $1/\pi$ & 0.31831 & $0.315\pm0.007$ & $+1.1\%$ \\
$\Omega_b$ & $1/(2\pi^2)$ & 0.05066 & $0.0493\pm0.0003$ & $+2.8\%$ \\
$\Omega_{\rm DM}$ & $(2\pi-1)/(2\pi^2)$ & 0.26765 & $0.264\pm0.007$ & $+1.4\%$ \\
$\Omega_r$ & $1/(4\pi^7)$ & $8.277\times10^{-5}$ & $8.27\times10^{-5}$ & $+0.1\%$ \\
\midrule
$z_{\rm eq}$ & $4\pi^6$ & 3846 & $3402\pm26^*$ & --- \\
$T_{\rm CMB}$ & derived & $2.642$~K & $2.7255$~K & $-3.1\%$ \\
\bottomrule
\end{tabular}
\end{center}
\noindent$^*$Planck's $z_{\rm eq}$ is model-dependent; not directly comparable.

\medskip\noindent Key ratios, all functions of $\pi$:
\begin{align*}
\Omega_\Lambda/\Omega_m &= \pi-1, &
\Omega_m/\Omega_b &= 2\pi, &
\Omega_m/\Omega_r &= 4\pi^6, \\
\Omega_b/\Omega_r &= 2\pi^5, &
\Omega_{\rm DM}/\Omega_b &= 2\pi-1.
\end{align*}

\subsection{The complete Friedmann equation}

Lovelock's theorem (Theorem~3.1 of [3]) forces the Einstein equation at $d = 4$. With all
four cascade components:
\begin{equation}\label{eq:friedmann}
H^2(z) = H_0^2\!\left[\frac{(1+z)^4}{4\pi^7} + \frac{(1+z)^3}{\pi} + \frac{\pi-1}{\pi}\right].
\end{equation}
Every coefficient is determined by $\pi$.

\subsection{The cascade's origin story}

The cascade does not predict a hot dense singularity. The origin of the four-dimensional
universe is the descent from $d = \infty$ ($V_\infty = \Omega_\infty = 0$, $N(\infty) = 0$)
through 213 compactification steps to $d = 4$.

\emph{No singularity.} Each compactification step sends one direction's effective radius to
$R_{\rm eff} = 1/\sqrt{d+3}$, which is finite at every finite $d$. No sharp boundary, no
infinite curvature, no breakdown of the metric.

\emph{No inflation.} The 120 orders of magnitude separating the Planck scale from $\Lambda$
are a theorem about the Gamma function ([1], Theorem~9.1). The flatness of $S^3$ follows
from its large curvature radius in the $\Lambda$-dominated era. The horizon problem does not
arise: the cascade's 217 dimensions were in causal contact before the descent.

\emph{No free initial conditions.} Every density fraction, the Hubble constant, and the
CMB temperature are derived from the cascade's geometry.

\subsection{Bott partition: subleading correction to $\Omega_m$}
\label{sec:bott}

\begin{theorem}[Bott partition matter fraction]\label{thm:omega_bott}
\[
\Omega_m^{\rm Bott} = \frac{\displaystyle\sum_{\substack{d=5\\d\bmod8\in\{5,6\}}}^{217}\Omega_{d-1}}{\displaystyle\sum_{d=5}^{217}\Omega_{d-1}} = 0.31150.
\]
\end{theorem}

\begin{proof}
\textbf{Step 1} (Paper~I, Theorem~3.1 and Corollary~3.2). Sphere area $\Omega_{d-1}$ is the only
independent cascade quantity at level $d$. Any physical identification mapping cascade
content to energy assigns energy $C \cdot \Omega_{d-1}$ at each level with $C$ universal and linear.

\textbf{Step 2} ([4], item~8; [2], Corollary~6.5; period-8 refinement in [3]). The Bott partition classifies cascade layers
by propagator phase: non-trivial-phase layers ($d \bmod 8 \in \{5, 6\}$) carry irreducibly
complex propagator amplitudes; real-phase Weyl layers ($d \bmod 8 = 4$) return to the real
axis after one period; Majorana layers ($d \bmod 8 \in \{0,1,2,3,7\}$) carry no complex
structure. This partition is topological and metric-independent.

\textbf{Step 3} (Definition~2.1 of [3]). The physical identification hypothesis assigns the
matter label to non-trivial-phase layers ($d \bmod 8 \in \{5, 6\}$), identified with matter
content by the fermion generation structure of [4].

\textbf{Step 4.} Since energy at every level is $C \cdot \Omega_{d-1}$ and $C$ cancels:
\[
\Omega_m^{\rm Bott} = \frac{C\sum_{\{5,6\}}\Omega_{d-1}}{C\sum_{\rm all}\Omega_{d-1}} = 0.31150.\qquad\square
\]
\end{proof}

Observed: $\Omega_m = 0.315 \pm 0.007$. Deviation: $-1.11\%$.

The lapse identity $\Omega_m = 1/\pi = 0.31831$ and the Bott fraction
$\Omega_m^{\rm Bott} = 0.31150$ agree to $2.1\%$, consistent with cascade descent corrections
from the 213 integrated-out dimensions. The lapse identity is the leading-order
single-step result (geometric, $+1.1\%$); the Bott fraction is the multi-layer descent
result (descent-dependent, $-1.1\%$). Both are valid at their respective orders.

\subsection{The acoustic scale}

\begin{theorem}[Acoustic scale]
With the cascade's observer-corrected parameter set ($H_0 = 66.78$,
$\Omega_m = 1/\pi$, $r_d \approx 147.75$~Mpc), $\ell_A = 296.4$.
Deviation from Planck's $301.6$: $-1.7\%$.
\end{theorem}

\begin{remark}[Sign and magnitude]
The $-1.7\%$ deviation is negative, placing $\ell_A$ in the
descent-dependent population (shared sign with the other
observer-corrected quantities). It closes further under the Part~0
Supplement Gram first-order correction, which is positive for
descent-dependent observables.
\end{remark}

\section{The Hubble Constant}

\subsection{The Friedmann equation with the observer-corrected $\rho_\Lambda$}

The standard Friedmann equation for a flat universe is
$H_0^2 = \rho_{\rm total}/(3\,M_{\rm Pl,red}^2)$. In the $\Lambda$-dominated
limit this reduces to
$H_0^2 = \rho_\Lambda/(3\,\Omega_\Lambda\,M_{\rm Pl,red}^2)$. Substituting
Part~I's observer-corrected identification
$\rho_\Lambda/M_{\rm Pl,red}^4 = (2/\pi)\,I$ with $\Omega_\Lambda = (\pi-1)/\pi$:
\[
H_0^2 \;=\; \frac{(2/\pi)\,I}{3\,(\pi-1)/\pi}\,M_{\rm Pl,red}^2
\;=\; \frac{2\,I}{3\,(\pi-1)}\,M_{\rm Pl,red}^2.
\]

\begin{theorem}[Hubble constant]\label{thm:H0}
\begin{equation}
H_0 \;=\; M_{\rm Pl,red}\,\sqrt{\frac{2\,I}{3\,(\pi-1)}}
\;=\; 66.78~\text{km/s/Mpc}.\label{eq:H0}
\end{equation}
\end{theorem}

\begin{proof}
$I = 1.0990 \times 10^{-120}$, $\Omega_\Lambda = (\pi-1)/\pi = 0.68169$,
$M_{\rm Pl,red} = 2.435 \times 10^{18}$~GeV. Steps:
\begin{align*}
2\,I/[3(\pi-1)] &= 3.421 \times 10^{-121},\\
\sqrt{3.421 \times 10^{-121}} &= 5.849 \times 10^{-61},\\
M_{\rm Pl,red} \times 5.849 \times 10^{-61} &= 1.4244 \times 10^{-42}~\text{GeV},\\
H_0 &= 1.4244 \times 10^{-42}~\text{GeV}\times \frac{\text{Mpc}\cdot\text{km/s}}{\hbar} = 66.78~\text{km/s/Mpc}.\qedhere
\end{align*}
\end{proof}

\begin{remark}[Why the observer bridge and not a descent lapse]
Earlier drafts of this paper derived $H_0$ via a ``lapse-corrected Friedmann
equation,'' in which the cascade descent from $d = 217$ to $d = 4$ was
assigned an ADM time coordinate with lapse $N(4) = 3\pi/8$, and the identification
was $\rho_\Lambda = I\,M_{\rm Pl,red}^4$ in that cascade time frame. That
derivation yielded $H_0 = 71.05$~km/s/Mpc and an implicit
$\rho_\Lambda/M_{\rm Pl,red}^4 = I/N(4)^2 = 7.92\times 10^{-121}$,
compared to the Planck observation $7.15\times 10^{-121}$ --- a $+10.7\%$
deviation that cannot be closed by the Part~0 Supplement Gram correction
(the Gram shift is positive, so it worsens a positive leading deviation).

The $(2/\pi)$ observer bridge of Part~I supersedes the descent-lapse
mechanism: the cascade's content is sphere area, not cube volume, and
the conversion between them at the observer's 3D spatial slice is the
$\pi/2$ factor derived in Part~I Section~3.2. Applied to the Friedmann
equation directly (without any descent lapse), this yields
$H_0 = 66.78$~km/s/Mpc and
$\rho_\Lambda/M_{\rm Pl,red}^4 = (2/\pi)\,I = 6.996\times 10^{-121}$ ---
a $-2.2\%$ leading deviation, closed by the Gram first-order correction
to $-0.07\%$, which is the correct population-consistent behaviour.

The trade-off is real: under the $(2/\pi)$ bridge, cascade $H_0 = 66.78$
sits $0.9\%$ below Planck's 67.4 and $8.5\%$ below SH0ES's 73.0, whereas
under the old lapse derivation the cascade sat \emph{between} the two
measurements. The old ``between tensions'' framing was numerically
appealing but geometrically wrong: the frame conversion mechanism it
used ($N(4)^2 = 9\pi^2/64 \approx 1.388$) is not the correct structural
bridge between cascade sphere-area content and cube-volume Planck
normalisation, and it produced $\Lambda$ deviations that the series'
own Gram correction worsens rather than closes.
\end{remark}

\begin{remark}[Self-consistency with the cosmological table]
$\rho_\Lambda/M_{\rm Pl,red}^4 = (2/\pi)\,I = 6.996\times 10^{-121}$
(Part~I Theorem~3.1). Observed: $7.150\times 10^{-121}$. Leading
deviation $-2.2\%$, Gram-corrected $-0.07\%$. Cascade $H_0$ is related
by $H_0^2 = \rho_\Lambda/(3\Omega_\Lambda M_{\rm Pl,red}^2)$, giving
$H_0 = 66.78$ km/s/Mpc at leading order; the same Gram correction
applied to $\rho_\Lambda$ shifts cascade $H_0$ upward by approximately
$\sqrt{1.0213}-1 = 1.06\%$ to $67.49$ km/s/Mpc, essentially equal to
Planck's central value of $67.36$ within experimental precision.
\end{remark}

\section{The Age of the Universe}

\begin{theorem}[Universe age]
$t_0 = 13.88$~Gyr.
\end{theorem}

\begin{proof}
The flat $\Lambda$CDM age integral is
$t_0 = H_0^{-1}\int_0^1 da/\sqrt{\Omega_r/a^2 + \Omega_m/a + \Omega_\Lambda a^2}$.
With cascade parameters $H_0 = 66.78$~km/s/Mpc (Theorem~\ref{thm:H0}),
$\Omega_m = 1/\pi$, $\Omega_\Lambda = (\pi-1)/\pi$, $\Omega_r = 1/(4\pi^7)$,
the integral evaluates numerically to $t_0 H_0 = 0.9478$, giving
$t_0 = 0.9478\cdot (977.8~{\rm Gyr\cdot km/s/Mpc})/66.78 = 13.878$~Gyr.
Observed: $t_0 = 13.80 \pm 0.02$~Gyr (Planck 2018~\cite{planck}).
Deviation: $+0.58\%$.
\end{proof}

\begin{remark}[Why the full integral, not $1/H_0$]
Earlier drafts of this paper quoted the Hubble time $1/H_0 = 13.76$~Gyr
(with the previous $H_0 = 71.05$) as the cascade's age prediction and
reported a ``$-0.29\%$ match.'' The Hubble time is not the universe age
in flat $\Lambda$CDM; the age is strictly less than $1/H_0$ by a factor
$t_0 H_0 \approx 0.948$ for cascade density fractions. The numerical
coincidence between $1/H_0 = 13.76$ and observed $t_0 = 13.80$ at the
old $H_0 = 71.05$ was an artefact of that (incorrect) $H_0$ and of
using the wrong formula; the proper $\Lambda$CDM integral at cascade
density fractions gave $t_0 = 13.04$~Gyr at the old $H_0$, a
$-5.5\%$ mismatch with observation. Under Part~I's observer-corrected
$H_0 = 66.78$ and the proper integral, the cascade's universe age is
$13.88$~Gyr, deviation $+0.58\%$, which is a genuine (rather than
artefactual) match.
\end{remark}

\begin{center}
\begin{tabular}{lcc}
\hline
Measurement & $H_0$ (km/s/Mpc) & Deviation from cascade \\
\hline
Planck 2018 (CMB) & $67.4 \pm 0.5$ & $+0.93\%$ \\
SH0ES (distance ladder) & $73.0 \pm 1.0$ & $+9.32\%$ \\
Cascade (Option A) & $66.78$ & --- \\
\hline
\end{tabular}
\end{center}

\noindent The cascade sits below Planck by $\sim 1\%$ and below SH0ES by
$\sim 9\%$. After applying the Part~0 Supplement Gram correction to $H_0$
(shifting upward by $\approx 1\%$), the cascade's $H_0$ rises to
$\sim 67.5$~km/s/Mpc, essentially equal to Planck's central value.

\section{The DESI BAO Observations}

\subsection{The sound horizon from cascade parameters}

The BAO standard ruler is the sound horizon at the drag epoch:
\[
r_d \approx 147.60\left(\frac{\omega_m}{0.1432}\right)^{-0.255}
\left(\frac{\omega_b}{0.02237}\right)^{-0.127}~\text{Mpc},
\]
where $\omega_x = \Omega_x h^2$ (Eisenstein \& Hu 1998 [12]). With cascade
parameters $H_0 = 66.78$, $h = 0.6678$:
\[
\omega_b^{\rm cas} = \tfrac{1}{2\pi^2}\times 0.6678^2 = 0.02259,\qquad
\omega_m^{\rm cas} = \tfrac{1}{\pi}\times 0.6678^2 = 0.14198.
\]

\begin{theorem}[Cascade sound horizon]
$r_d^{\rm cas} \approx 147.75$~Mpc, essentially equal to the
Planck-inferred value $r_d^{\rm Planck} = 147.60$~Mpc (difference
$+0.10\%$).
\end{theorem}

\begin{proof}
$r_d = 147.60 \times (0.14198/0.1432)^{-0.255} \times (0.02259/0.02237)^{-0.127}
\approx 147.60\times 1.00216\times 0.99875 \approx 147.75$~Mpc.
\end{proof}

Under Part~I's observer-corrected $H_0 = 66.78$, the cascade's $(\omega_b, \omega_m)$
land essentially on top of Planck's 2018 values, and the cascade's sound
horizon is indistinguishable from Planck's within the accuracy of the
E\&H98 fitting formula. The cascade and Planck $\Lambda$CDM share a
ruler.

\subsection{BAO observables and comparison to DESI DR2}

The table below is auto-generated from the cascade parameters via
\texttt{tools/generators/generate\_bao\_table.py} and reproduced on every build.
The ``Cascade'' column uses the leading-order $\Omega_m = 1/\pi$
($r_d^{\rm cas} \approx 147.7$~Mpc); the ``Bott'' column uses the
subleading-corrected $\Omega_m^{\rm Bott} = 0.31150$
($r_d^{\rm Bott} \approx 148.6$~Mpc). Both use $H_0 = 66.78$~km/s/Mpc.

% Auto-generated table: run tools/generators/generate_bao_table.py before LaTeX build
\input{generated/bao-table}

\subsection{DESI tensions are shared, not resolved by the cascade}

\begin{theorem}[Shared DESI tension]
The cascade cosmology ($w = -1$, $r_d \approx 147.75$~Mpc,
$H_0 = 66.78$~km/s/Mpc, $\Omega_m = 1/\pi$) fits DESI DR2 BAO with
$\chi^2/n = 2.35$ (leading) or $2.23$ (Bott $\Omega_m$), compared to
Planck $\Lambda$CDM at $\chi^2/n = 1.90$. The cascade's leading $\chi^2$
is higher than Planck's by $\Delta\chi^2 \approx 5.9$, driven mostly by
two shared outlier bins: the $z = 0.510$ $D_H/r_d$ pull and the
$z = 0.706$ $D_M/r_d$ pull. The cascade has no free parameters with
which to relieve these tensions; it predicts $w = -1$ as a structural
theorem.
\end{theorem}

\begin{proof}
Direct computation; see \texttt{tools/generators/generate\_bao\_table.py} and the
BAO comparison table above.
\end{proof}

\begin{remark}[The ruler-mismatch story has been retracted]
Earlier drafts of this paper (predating Part~I's observer correction)
reported cascade $r_d \approx 141$~Mpc with $H_0 = 71.05$~km/s/Mpc, and
argued that the DESI apparent-$w$ signal could be explained as a
``ruler mismatch'': fitting the cascade's true $(w = -1)$ distances with
Planck's prior on $r_d = 147.6$~Mpc would produce an apparent
$w \approx -0.80$, matching DESI's reported $-0.76 \pm 0.11$. That
framing is no longer available. Under Part~I's corrected $H_0 = 66.78$,
the cascade's $r_d = 147.75$~Mpc is essentially equal to Planck's
$147.60$~Mpc (difference $0.10\%$, negligible compared to the
$\pm 0.5\%$ E\&H98 fitting-formula accuracy), so there is no
frame-difference lever that the cascade can pull on to explain away
the DESI signal. The cascade predicts $w = -1$ exactly; if DESI's
apparent $w \neq -1$ persists at the $>5\sigma$ level with $r_d$
independently constrained to be close to Planck's value, the cascade
is challenged to the same degree as standard $\Lambda$CDM.
\end{remark}

\begin{remark}[The $z = 0.510$ outlier]\label{rem:outlier}
The bin $z_{\rm eff} = 0.510$, $D_H/r_d$, shows a $3.87\sigma$ pull in
the cascade (up from $3.20\sigma$ under the earlier framing) and
$3.60\sigma$ in Planck $\Lambda$CDM. Together with the $z = 0.706$
$D_M/r_d$ pull at $3.25\sigma$ (cascade) / $2.76\sigma$ (Planck), these
two bins contribute roughly 25 of the cascade's total $\chi^2 = 30.54$
and 20 of Planck's total $24.65$. Without these two bins, both fits
would have $\chi^2/n \approx 0.4$--$0.5$. The tensions persist in the
$r_d$-free ratio $D_M/D_H$, independent of ruler calibration, and
represent genuine anomalies in the LRG1 and LRG2 expansion rates that
neither model explains. The cascade's larger $\chi^2$ relative to
Planck is almost entirely a consequence of these outliers hitting
slightly harder on cascade-predicted distances than on Planck's.
\end{remark}

\subsection{Sharp falsification predictions}

\begin{enumerate}
\item \textbf{CMB-S4 measures $r_d$ directly.} If $r_d \approx 147.6$~Mpc
with parameters consistent with cascade $(H_0, \Omega_m, \Omega_b)$, the
cascade is confirmed on $r_d$. If $r_d$ deviates from the cascade
prediction at $>3\sigma$, the cascade is falsified on the density
fractions (not on the observer-corrected $\Lambda$).
\item \textbf{DESI BAO+CMB $w_0$ stability.} The cascade predicts
$w = -1$ exactly. If BAO+CMB alone drives $w_0 \neq -1$ at $>5\sigma$
with parameters consistent with the cascade's $(H_0, \Omega_m, r_d)$,
the cascade is challenged jointly with $\Lambda$CDM.
\item \textbf{The $z = 0.510, 0.706$ outliers.} The cascade offers no
independent mechanism to explain these two outliers beyond what
$\Lambda$CDM offers. Resolution (e.g., a systematic error in a specific
DESI tracer) would benefit both models equally.
\end{enumerate}

\section{Summary: The Complete Cosmological Table}

\begin{center}
\small
\begin{tabular}{llccc}
\hline
Parameter & Cascade formula & Predicted & Observed & Dev \\
\hline
$\rho_\Lambda/M_{\rm Pl,red}^4$ & $18\,\Omega_{19}\Omega_{217}/\pi^3$ & $6.996\times10^{-121}$ & $7.150\times10^{-121}$ & $-2.2\%$ \\
\quad (Gram-corrected) & $\times\exp(0.02108)$ & $7.145\times10^{-121}$ & $7.150\times10^{-121}$ & $-0.07\%$ \\
$\Omega_k$ & $\approx0$ ($S^3$ topology) & $\approx0$ & $0.001\pm0.002$ & --- \\
$\Omega_m$ & $1/\pi$ (lapse identity) & 0.31831 & $0.315\pm0.007$ & $+1.1\%$ \\
$\Omega_m^{\rm Bott}$ & Bott partition & 0.31150 & $0.315\pm0.007$ & $-1.1\%$ \\
$\Omega_\Lambda$ & $(\pi-1)/\pi$ & 0.68169 & $0.685\pm0.007$ & $-0.5\%$ \\
$\Omega_b$ & $1/(2\pi^2)$ & 0.05066 & $0.0493\pm0.0003$ & $+2.8\%$ \\
$\Omega_{\rm DM}$ & $(2\pi-1)/(2\pi^2)$ & 0.26765 & $0.264\pm0.007$ & $+1.4\%$ \\
$\Omega_r$ & $1/(4\pi^7)$ & $8.277\times10^{-5}$ & $8.27\times10^{-5}$ & $+0.1\%$ \\
$z_{\rm eq}$ & $4\pi^6$ & 3846 & $3402\pm26^*$ & --- \\
$T_{\rm CMB}$ & derived & 2.642~K & 2.7255~K & $-3.1\%$ \\
$w$ & $-1$ (fixed $\Lambda$, GB vanishes) & $-1$ & $-1.03\pm0.03$ & $1\sigma$ \\
$H_0$ & Eq.~\eqref{eq:H0} & 66.78 & 67.4$\pm$0.5 & $-0.9\%$ \\
$t_0$ & $\Lambda$CDM integral & 13.88~Gyr & $13.80\pm0.02$ & $+0.6\%$ \\
$r_d$ & from $(\Omega_b,\Omega_m,H_0)$ & $\approx 147.75$~Mpc & 147.60 & $+0.1\%$ \\
$\ell_A$ & from cascade $(H_0,\Omega_m,r_d)$ & $\approx 296.4$ & $301.6$ & $-1.7\%$ \\
$\chi^2/n$ (DESI DR2) & from cascade parameters & $2.35$ & Planck: $1.90$ & --- \\
\hline
\end{tabular}
\end{center}
\noindent$^*$Planck's $z_{\rm eq}$ is model-dependent; not directly comparable.

\medskip All quantities are functions of $\pi$, the cascade invariant $I$, and the reduced
Planck mass $M_{\rm Pl,red}$ (the single dimensionful input). Zero fitted parameters.

\section{What This Paper Does Not Do}

\begin{itemize}
\item \textbf{Perturbation parameters.} The scalar spectral index $n_s$, primordial amplitude
$A_s$, and reionisation optical depth $\tau$ are not derived.
\item \textbf{The CMB power spectrum.} Requires $n_s$, $A_s$, $\tau$, and Boltzmann evolution
through recombination.
\item \textbf{The $z = 0.510$ anomaly.} Shared with Planck $\Lambda$CDM and not explained.
\end{itemize}

\section{Predictions and Falsifiability}

\begin{enumerate}
\item $w = -1$ exactly (cascade structural theorem). Falsified if future
data show $w \neq -1$ at $> 5\sigma$ with parameters consistent with
cascade $(H_0, \Omega_m, \Omega_b, r_d)$.
\item $r_d \approx 147.75$~Mpc (from cascade $(H_0, \Omega_m, \Omega_b)$
via E\&H98). Essentially equal to Planck's 147.60 Mpc; falsified by
direct $r_d$ measurement $> 3\sigma$ from this value.
\item $\Omega_m = 1/\pi = 0.31831$ (leading order) /
$\Omega_m^{\rm Bott} = 0.31150$ (subleading). Falsified if future CMB
measurements place $\Omega_m$ at $> 3\sigma$ from \emph{both} values
simultaneously.
\item $\Omega_b = 1/(2\pi^2) = 0.0507$. Currently $+2.8\%$ above Planck
central value. CMB-S4 will tighten to sub-percent; reanalysis with
cascade parameters may shift the inferred $\Omega_b$.
\item $H_0 = 66.78$~km/s/Mpc at leading order; after the Part~0
Supplement Gram first-order correction shifts it to $\approx 67.5$,
essentially equal to Planck's central value. Independent $H_0$
measurements (gravitational wave standard sirens, time-delay
cosmography) should land in the $66$--$68$ range, Planck-compatible and
\emph{not} SH0ES-compatible. Confirmation of $H_0 \gtrsim 70$ at high
significance with independent methods would challenge the cascade.
\item \textbf{No dark matter particles.} Confirmed detection of a dark
matter particle interacting via gauge forces would falsify the
cascade's geometric dark matter interpretation.
\end{enumerate}

\begin{remark}[The ``between tensions'' claim has been retracted]
Earlier drafts predicted cascade $H_0 = 71.05$ and framed this as
``between the two Hubble-tension measurements'' (Planck 67.4, SH0ES
73.0). Under Part~I's observer-corrected identification
$\rho_\Lambda/M_{\rm Pl,red}^4 = (2/\pi)\,I$, cascade $H_0 = 66.78$
instead sits on the Planck side of the tension and is
\emph{incompatible} with the SH0ES local distance-ladder value at
the $\sim 8\%$ level. The cascade picks a side: it predicts the
Planck-compatible branch of the Hubble tension, not the SH0ES branch.
If independent methods converge on the SH0ES value at high
significance, the cascade's prediction fails.
\end{remark}

\section{Open Questions}

\begin{enumerate}
\item \textbf{Theorem status of $\Omega_m$.} The Bott partition (Theorem~\ref{thm:omega_bott})
identifies non-trivial-phase layers with matter via the physical identification
hypothesis. Whether this identification step follows uniquely from the hypothesis,
or requires an additional argument, is the remaining open question for $\Omega_m$.
\item \textbf{Gram correction for ratio-of-sums observables.} Theorem~15.11 of the Part~0
Supplement derives the first-order Gram correction $\delta Q/Q_0 = \sum(1-C^2)$ for
\emph{multiplicative propagators} traversing contiguous cascade paths. $\Omega_m^{\rm Bott}$
is structurally a ratio of sums over cascade layers (numerator: Bott-selected layers;
denominator: full range), not a multiplicative propagator. The Supplement's table
applies the full-path correction $\exp(\sum(1-C^2))$ as a uniform multiplicative shift
(giving $0.3181$, $+1.0\%$ above observed $0.315$), but a rigorous per-term treatment
---correcting each $\Omega_{d-1}$ in the sums by its own path-length-dependent factor
and re-forming the ratio---would yield a much smaller shift, since both sums are
dominated by early terms near the $\Omega_d$ peak where path lengths are short. Deriving
the effective Gram correction for ratio-of-sums observables $Q=\sum_i a_i/\sum_j b_j$
where each $a_i,b_j$ is a multiplicative propagator is the open structural question;
see the scope remark in the Part~0 Supplement following Theorem~15.11.
\item \textbf{Independent determination of $r_d$.} A direct CMB measurement without assuming
Planck's $(\Omega_b, H_0)$ would simultaneously test $\Omega_b = 1/(2\pi^2)$.
\item \textbf{Derive the cascade's expansion history $a(t)$.} The cascade's expansion history
should come from the descent profile $N(d)$, a function of the Gamma function.
\item \textbf{Perturbation spectrum from the cascade.} Whether the descent from $d = \infty$
generates density perturbations with a specific spectrum is the most important open
question in the cascade's cosmology programme.
\item \textbf{The $z = 0.510$ anomaly.} Not explained by the cascade or Planck $\Lambda$CDM. Open.
\end{enumerate}

\section{Confidence Assessment}

\textbf{Two-population systematic and the cascade's structural
corrections.} At leading order, descent-dependent quantities carry
negative deviations and geometric quantities carry positive deviations.
The cosmological constant is now in the descent-dependent population
($-2.2\%$ leading), consistent with $\alpha_s$, $m_\tau/m_\mu$, $v$,
$\ell_A$, and $\Omega_m^{\rm Bott}$. Two structural corrections reduce
descent deviations to sub-percent:
\begin{itemize}
\item \emph{The Part~0 Supplement Gram first-order correction}
$\delta Q/Q_0 = \sum_{\rm adj}(1-C^2_{d,d+1}) = 0.02108$ (exponentially
applied) closes $\rho_\Lambda$ from $-2.2\%$ to $-0.07\%$, $v$ from
$-1.0\%$, $\Omega_m^{\rm Bott}$ from $+1.0\%$, and similar ordinary
descent-dependent quantities.
\item \emph{The $\alpha(d^*)/\chi^k$ structural family}
(Part~IVb, Remark~4.6) closes seven SM precision observables at
experimental precision; these are observable-specific and do not apply
uniformly to cosmological parameters.
\end{itemize}
$H_0$ inherits a $\sim 1\%$ upward shift from the Gram correction applied
to $\rho_\Lambda$ via the Friedmann equation ($H_0\propto\sqrt{\rho_\Lambda}$),
bringing cascade $H_0 \to 67.5$, essentially equal to Planck's central
value. $T_{\rm CMB}$ inherits the same shift and moves from 2.642 to
$\sim 2.669$~K, still $-2.1\%$ from observed 2.7255~K, awaiting either a
second-order correction or absorption of $g_{\rm eff}$ into a
cascade-intrinsic counting rule.

\textbf{Tier 1: Theorem-level.} $w = -1$ exactly (fixed $\Lambda$,
Theorem~3.1); $\Omega_k \approx 0$ ($S^3$ topology); GB corrections
vanish (two mechanisms, Corollary~3.2); cascade and Planck $\Lambda$CDM
predict essentially the same $r_d \approx 147.7$ Mpc under the
cascade's observer-corrected parameters; the cascade structurally
cannot explain the apparent-$w$ deviation in DESI via ruler mismatch
(the ruler is shared with Planck).

\textbf{Tier 2: Derived with clean argument.}
$\rho_\Lambda/M_{\rm Pl,red}^4 = 18\,\Omega_{19}\Omega_{217}/\pi^3$
(Part~I, Theorem~3.1): $-2.2\%$ leading, $-0.07\%$ Gram-corrected.
$H_0 = 66.78$~km/s/Mpc (Theorem~\ref{thm:H0}): $-0.9\%$ leading,
essentially equal to Planck's 67.4 after Gram correction. $t_0 = 13.88$
Gyr ($+0.6\%$). $\Omega_m = 1/\pi$ (lapse identity, $+1.1\%$).
$\Omega_m^{\rm Bott} = 0.31150$: sphere-area fraction of
non-trivial-phase layers (Theorem~\ref{thm:omega_bott}, $-1.1\%$).
$\Omega_\Lambda = (\pi-1)/\pi$ ($-0.5\%$).
$\Omega_r = 1/(4\pi^7)$: interior content, 7 geometric steps
(Theorem~\ref{thm:omega_r}, $+0.1\%$).
$T_{\rm CMB} = 2.642$ K ($-3.1\%$, descent-dependent through $H_0$).

\textbf{Tier 5: Known weak.} $\Omega_b = 1/(2\pi^2)$ ($+2.8\%$; the
``one unit of content on $S^3$'' argument still needs strengthening).
DESI-BAO fit quality ($\chi^2/n = 2.35$ vs Planck's $1.90$): driven by
shared outliers at $z = 0.510$ and $z = 0.706$, cannot be resolved by
the cascade without free parameters.

\textbf{What would change my mind.} (i) Robust $> 5\sigma$ measurement
of $w \neq -1$ with $r_d$ independently constrained to be consistent
with the cascade's $147.75$~Mpc. (ii) Confirmed detection of a dark
matter particle. (iii) $\Omega_m$ measured at $> 3\sigma$ from
\emph{both} $1/\pi = 0.31831$ and the Bott value $0.31150$
simultaneously. (iv) Independent $H_0$ measurements converging firmly
on the SH0ES side ($H_0 > 71$) with small error bars, which would
falsify the cascade's Planck-compatible $H_0$ prediction. (v)
Resolution of the $z = 0.510, 0.706$ DESI outliers in a way that
strengthens the case for $w \neq -1$.

\begin{thebibliography}{12}

\bibitem{paper1} {[Author]}, \textit{Scale Variance from Orthogonality: How the Unit Ball
Generates $10^{120}$ Orders of Magnitude}, March 2026.

\bibitem{paper2} {[Author]}, \textit{Quantum Mechanics from the Cascade: Effective Theory of
a 4-Dimensional Observer in the Sphere-Area Geometry}, March 2026.

\bibitem{paper3} {[Author]}, \textit{General Relativity from the Cascade: Lovelock Uniqueness
and the Cosmological Constant of a 4-Dimensional Observer}, March 2026.

\bibitem{paper4a} {[Author]}, \textit{The Standard Model from the Cascade: Part IVa---Gauge
Group, Symmetry Breaking, and Three Generations from Bott Periodicity and Hairy Ball
Zeros}, March 2026.

\bibitem{paper4b} {[Author]}, \textit{The Standard Model from the Cascade: Part IVb---Masses,
Couplings, and Precision Predictions from the Geometric-Topological Factorization},
March 2026.

\bibitem{planck} Planck Collaboration, ``Planck 2018 results. VI. Cosmological parameters,''
\textit{Astronomy \& Astrophysics} \textbf{641}, A6 (2020).

\bibitem{desi2} DESI Collaboration, \textit{DESI DR2 Results II: Measurements of Baryon
Acoustic Oscillations and Cosmological Constraints}, arXiv:2503.14738 (2025).

\bibitem{desi1} DESI Collaboration, ``DESI 2024 VI: Constraints on models of dark energy,''
arXiv:2404.03002 (2024).

\bibitem{lee} S.~Lee, ``Comparing $\Lambda$CDM, $w$CDM, and $w_0w_a$CDM models with DESI DR2
BAO,'' arXiv:2507.01380 (2025).

\bibitem{riess} A.~G.~Riess et al., ``A comprehensive measurement of the local value of the
Hubble constant,'' \textit{Astrophysical Journal Letters} \textbf{934}, L7 (2022).

\bibitem{efstathiou} G.~Efstathiou, ``Evolving dark energy or supernovae systematics?''
\textit{Monthly Notices of the Royal Astronomical Society} \textbf{538}, 875 (2025).

\bibitem{EH98} D.~J.~Eisenstein and W.~Hu, ``Baryonic Features in the Matter Transfer
Function,'' \textit{Astrophysical Journal} \textbf{496}, 605 (1998).

\end{thebibliography}

\end{document}