\documentclass[11pt,a4paper]{article}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{geometry}
\usepackage{hyperref}
\usepackage{booktabs}
\usepackage{graphicx}
\geometry{margin=2.5cm}

\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{principle}[theorem]{Principle}
\newtheorem{remark}[theorem]{Remark}

\title{The Cascade Series --- Prelude\\[0.5em]
\large Why Nothing Has Structure}
\author{RTAC}
\date{March 2026}

\begin{document}
\setlength{\emergencystretch}{3em}
\maketitle

\begin{center}
\fbox{\parbox{0.9\textwidth}{\textbf{Status.} Foundational. The cascade
series has one pre-mathematical input, Definition~\ref{def:axiom}
($0\neq 1$). Principle~\ref{princ:austerity} (austerity) is \emph{not}
a second input: it is the operational content of
Definition~\ref{def:axiom}'s status as the sole input, derived in
Section~\ref{sec:what-nothing-means} below. Each step of the chain
from $0\neq 1$ to the dimension tower
$\{(\mathbb{R}^d, B^d)\}_{d\in\mathbb{N}}$ (with asymptotic upper edge
$B^\infty$) is annotated with the austerity clause it invokes. The
prelude has one substantive structural commitment: the use of
orthogonal states in an inner-product realisation as the model of
distinguishable states (the interpretive bridge of
Theorem~\ref{thm:orth}'s premise). Integers do not enter as a separate
primitive: the integer $d$ is the cardinality of the orthogonal-state
set at each level, and ``$\mathbb{R}^d$'' just names the space of $d$
orthogonal directions under the realisation. Once the realisation
commitment is accepted, the rest of the structure follows: the
Euclidean geometry on $\mathbb{R}^d$ is forced by elementary linear
algebra, the unit ball $B^d$ is the unique parameter-free
rotation-symmetric bounded shape, the cardinality $d$ ranges over
$\mathbb{N}$ as $\mathrm{RCA}_0$'s native unbounded notion, and the
slicing recurrence is the unique first-order multiplicative recurrence
on the tower.
Whether the structural commitment to $\mathbb{R}^d$ is itself uniquely
forced, or merely the most austere structural choice, is the single
remaining open foundational question of the prelude
(Section~\ref{sec:open-questions}, item~(3)(a)). The prelude
operates on finite-dimensional Euclidean spaces only --- no
infinite-dimensional Hilbert space is built or required, and the
cascade's foundational footprint stays at $\mathrm{RCA}_0$ for the
dimension tower's existence and at $\mathrm{WKL}_0$/$\mathrm{ACA}_0$
for the classical-analysis content. The cascade series' load-bearing
hypothesis --- that $B^\infty$ descended to four dimensions is
indistinguishable from our universe --- is stated independently in
the cover sheet and does not depend on the arguments of this prelude.
If any step of the prelude's chain is disputed, the cascade continues
to rest on the cover-sheet hypothesis and the Gamma-function theorems
of Part~0 onward.}}
\end{center}

\begin{abstract}
A theory of everything cannot have inputs: every input demands an
explanation of its origin, and that explanation is either circular or
requires a deeper theory. The only starting point that requires no
explanation is nothing. This prelude shows that ``nothing,'' taken
seriously as a mathematical object, is the infinite-dimensional unit
ball---and that the unit ball is not featureless. It has a slicing
recurrence, a unique constant ($\sqrt{\pi}$), four distinguished
dimensions, and a cascade invariant of $10^{-120}$. These are not
imposed; they are consequences of the single pre-mathematical fact that
false and true are distinguishable. The series that follows is the
derivation of what that fact implies about the geometry of the world.
\end{abstract}

% ============================================================
\section{The Problem with Starting Points}\label{sec:starting-points}
% ============================================================

Every theory of everything proposed to date begins with something:
a string, a spin network, a set of fields, a symmetry group, a
Lagrangian, a wave function, a set of rules. Each starting point
carries structure, and that structure demands explanation.

Why a string rather than a membrane? Why $\mathrm{SU}(3)\times
\mathrm{SU}(2)\times\mathrm{U}(1)$ rather than $\mathrm{SU}(5)$?
Why ten dimensions rather than eleven? Every choice is a free input.
Every free input is a knob that could have been set differently. Every
knob that could have been set differently invites the question: what
set it?

The options are limited. Either the starting point is explained by
something deeper---in which case it was not the starting point---or it
is accepted as given, and the theory is incomplete. A theory of
everything that begins with something is, by construction, not a
theory of everything. It is a theory of everything \emph{given} its
starting point.

The alternative is to begin with nothing.

% ============================================================
\section{What ``Nothing'' Means}\label{sec:what-nothing-means}
% ============================================================

``Nothing'' in physics is usually taken to mean empty space, or the
quantum vacuum, or the absence of matter. These are not nothing. Empty
space has dimension, signature, topology. The quantum vacuum has
energy, fluctuations, fields. The absence of matter presupposes a
stage on which matter is absent.

Mathematical nothing is more austere: no space, no time, no
dimension, no content. But pure nothing---the absence of all
structure---cannot even be discussed, because discussion requires
the distinction between a statement and its negation.

This is the key observation. The minimal structure required for any
formal system to exist is not a space, not a symmetry, not a
dynamical law. It is the distinction between true and false.

\begin{definition}[Pre-mathematical axiom]\label{def:axiom}
$0 \neq 1$.
\end{definition}

This is not an axiom in the usual sense. It is not chosen from a
menu of alternatives. It is the precondition for axioms, theorems,
and proofs to be meaningful. A system that cannot distinguish $0$
from $1$ cannot state any proposition, cannot derive any consequence,
cannot describe any world. Definition~\ref{def:axiom} is the cost of
admission for coherent thought.

\begin{principle}[Austerity, derived from Definition~\ref{def:axiom}]\label{princ:austerity}
Under the commitment that Definition~\ref{def:axiom} ($0\neq 1$) is the
sole pre-mathematical input of the cascade series
(Section~\ref{sec:starting-points}), any construction introducing content
beyond what Definition~\ref{def:axiom} forces is inadmissible. This
excludes:
\begin{itemize}
\item[(i)] \emph{Parameter economy.} No free parameter is introduced ---
no unexplained numerical, categorical, or symmetry-group input.
\item[(ii)] \emph{Minimal strength.} No assumption strictly stronger than
Definition~\ref{def:axiom} is invoked where a weaker sufficient assumption
exists. Strength is measured in the sense of reverse mathematics over
the base theory $\mathrm{RCA}_0$
\cite{friedman1975, simpson2009}: an assumption $A$ is strictly
stronger than $B$ if $\mathrm{RCA}_0 + A \vdash B$ but
$\mathrm{RCA}_0 + B \nvdash A$.
\item[(iii)] \emph{Unobservable quotient.} Any continuous symmetry whose
action is unobservable from within the structure is quotiented out.
\end{itemize}
\end{principle}

\begin{proof}[Derivation from Definition~\ref{def:axiom}]
Each clause identifies a distinct way a construction can violate the
sole-input commitment of Definition~\ref{def:axiom}.

\emph{Clause (i).} A free parameter is a second explicit input alongside
Definition~\ref{def:axiom}. Admitting it contradicts the sole-input
status directly.

\emph{Clause (ii).} If a construction uses assumption $A$ where the
weaker $B$ suffices (with $A\Rightarrow B$ strictly), the residue
$A\setminus B$ is additional content not forced by
Definition~\ref{def:axiom}. That residue is a second input --- the same
class of violation as clause~(i).

\emph{Clause (iii).} A continuous symmetry whose action is unobservable
from within the structure carries structural content with no derivation
consequence. Retaining it imports content not forced by
Definition~\ref{def:axiom} --- the same class of violation.

\emph{Exhaustiveness of (i)--(iii) (formal-extension trichotomy).} A
construction $C$ violates sole-input iff $C$ is derivable in some
formal system $\Sigma_C$ properly extending $\Sigma_0 = (\text{logic})
+ \{0\neq 1\}$. Standard proof theory characterises a formal system by
three independent axes --- its non-logical axioms, its underlying
logic, and its language/signature. Any extension of $\Sigma_0$ acts on
at least one of these axes:
\begin{itemize}
\item Extending the axioms imports content as a fresh assumption
(numerical, categorical, or symmetry-group input). This is clause~(i).
\item Extending the logic imports content as inferential strength
beyond what Definition~\ref{def:axiom} requires (e.g.\ adopting
classical reasoning, the axiom of choice, or higher-order quantification
where the cascade's chain demands only weaker resources); the residue
$A\setminus B$ between the adopted logic $A$ and the strictly weaker
sufficient $B$ encodes the imported content. This is clause~(ii).
\item Extending the signature or background imports content as ambient
structure with no cascade-internal observable trace --- a continuous
symmetry whose action is invisible from within the cascade. This is
clause~(iii).
\end{itemize}
These three axes exhaust the formal anatomy of an extending system:
there is no fourth axis along which a formal system can be extended in
standard proof theory. Any candidate fourth clause therefore either
reduces to one of (i)--(iii) under the standard decomposition or fails
to identify content actually present in $\Sigma_C$'s extension of
$\Sigma_0$. Definitional extensions of $\Sigma_0$ are conservative
(they introduce notation but no new theorems) and therefore do not
qualify as content additions; they require no clause.

All three clauses track a single condition: admitting content not forced
by the sole input. Austerity is therefore the operational content of
Definition~\ref{def:axiom}'s status as sole pre-mathematical input, not
a separate axiom.
\end{proof}

\noindent\textbf{Status.} Principle~\ref{princ:austerity} is a
\emph{derivation rule} for cascading from $0\neq 1$, analogous to
modus ponens being the inference rule of propositional logic: it is
built into what ``sole pre-mathematical input'' means, rather than a
further axiom within the mathematical content. Each step of the prelude
chain that previously relied on an implicit parsimony choice is now
annotated with the clause of Principle~\ref{princ:austerity} it invokes.
Whether austerity \emph{uniquely} determines $B^\infty$ --- as opposed to
rendering it defensibly minimal among a larger austerity-compatible set ---
remains open (Section~\ref{sec:open-questions}).

The question is: what does $0\neq 1$ \emph{imply}?

% ============================================================
\section{Distinction Implies Orthogonality}
% ============================================================

If $0$ and $1$ are distinguishable, they are maximally
distinguishable. There is no sense in which $0$ partially overlaps
with $1$; they share no content. In the language of geometry, they are
\emph{orthogonal}: their inner product is zero.

This is not a metaphor. The logical operation NOT, which maps
$0\mapsto 1$ and $1\mapsto 0$, is an involution ($\mathrm{NOT}^2 =
\mathrm{Id}$) that exchanges two states sharing no common component.
In any inner-product space, the unique pair of unit vectors with no
shared component is the orthogonal pair. The passage from
$0\neq 1$ to orthogonality is not an additional assumption; it is a
translation of the same fact into geometric language.

\begin{theorem}[Distinction is orthogonality]\label{thm:orth}
For any $d\geq 2$, let $e_0, e_1\in\mathbb{R}^d$ be unit vectors
representing two fully distinguishable states --- states sharing no
common component under any linear operation on $\mathbb{R}^d$. Then
$\langle e_0, e_1\rangle = 0$ under the standard Euclidean inner
product.
\end{theorem}

\begin{proof}
In $\mathbb{R}^d$ with the standard inner product, $\langle e_0, e_1\rangle$
is the component of $e_1$ along $e_0$: the projection of $e_1$ onto
the line through $e_0$ is $P_{e_0}(e_1) = \langle e_0, e_1\rangle\,e_0$.
``Fully distinguishable'' states share no common component by
definition, so for every projector this projection must vanish.
Applied to $P_{e_0}$: if $\langle e_0,e_1\rangle\neq 0$, then a linear
operation on $e_1$ (projection onto $e_0$) produces a nonzero multiple
of $e_0$ --- a partial identification of $e_1$ with $e_0$, contradicting
full distinguishability. Hence $\langle e_0, e_1\rangle = 0$. The
argument uses only the geometric interpretation of the inner product:
no probability measure, no Born rule, and no fidelity relation is
invoked.
\end{proof}

\begin{remark}[Status of Theorem~\ref{thm:orth}]
Theorem~\ref{thm:orth} unpacks the Euclidean realisation of
distinction at each $d$: given the premise that distinguishable states
share no common component in $\mathbb{R}^d$, the projector calculation
yields orthogonality. The substantive structural content is in the
premise rather than in the conclusion. The bridge from logical
$0\neq 1$ (Definition~\ref{def:axiom}) to that premise --- i.e., the
choice of Euclidean geometry on each $\mathbb{R}^d$ as the realisation
of logical distinction --- is a separate structural commitment, flagged
in Section~\ref{sec:open-questions}, item~(3). The Euclidean inner
product on $\mathbb{R}^d$ itself is forced by elementary linear
algebra (it is the unique positive-definite symmetric bilinear form
on $\mathbb{R}^d$ up to isomorphism, given a basis); the interpretive
choice is the use of $\mathbb{R}^d$ as the geometric realisation of
$d$ distinguishable states, not the geometry imposed on it.
Theorem~\ref{thm:orth} is best read as the geometric expression of
that interpretive choice rather than as an independent derivation of
orthogonality.
\end{remark}

% ============================================================
\section{The Dimension Tower}
% ============================================================

One distinction gives two orthogonal directions and a one-dimensional
Euclidean space $\mathbb{R}^1$. But the existence of one distinction does
not preclude a second. If $e_0$ and $e_1$ span a plane $\mathbb{R}^2$,
a third direction $e_2$ can be independent of both, giving $\mathbb{R}^3$.
There is no logical obstruction: the fact that $0\neq 1$ does not limit
the number of independent things that can be distinguished.

\begin{theorem}[Countable dimension tower]\label{thm:iterate}
The cascade's dimension parameter $d$ ranges over $\mathbb{N}$. For each
$d\in\mathbb{N}$, the cascade structure at dimension $d$ is the
finite-dimensional Euclidean space $\mathbb{R}^d$ with its standard inner
product, and the unit ball $B^d = \{x\in\mathbb{R}^d : |x|\leq 1\}$.
\end{theorem}

\begin{proof}
\emph{Existence of arbitrarily many dimensions.} For any finite $d$, the
Euclidean space $\mathbb{R}^d$ and its unit ball $B^d$ are well-defined
classical objects, expressible in $\mathrm{RCA}_0$ \cite{simpson2009} ---
the strength-measurement base of Principle~\ref{princ:austerity}
clause~(ii) (Section~\ref{sec:open-questions}, item~(2)).

\emph{The dimension tower has no upper bound (clause~(i) of
Principle~\ref{princ:austerity}).} Suppose the tower stops at some
finite $n<\infty$. Then $n$ itself is a structural input of the theory
--- an unexplained integer specifying the cascade's maximum dimension.
By Principle~\ref{princ:austerity} clause~(i) (parameter economy), any
such integer is rejected as a free parameter. The only stopping points
compatible with clause~(i) are $n=0$ (which contradicts
Definition~\ref{def:axiom}, requiring at least two distinct states,
hence dimension at least~$1$) and $n=\infty$. The dimension tower
therefore extends without bound. (``$n=\infty$'' here is not itself a
chosen cardinality among alternatives: within the $\mathrm{RCA}_0$
strength-measurement base, $\aleph_0$ is the unique unbounded notion
--- $\mathrm{RCA}_0$ has no machinery to express higher cardinals ---
so $n=\infty$ is the unbounded countable index $\mathbb{N}$ the
non-termination directly produces, not a selection from a menu of
infinities.)

\emph{Countable rather than uncountable indexing (clause~(ii) of
Principle~\ref{princ:austerity}).} Indexing the dimension tower by an
uncountable cardinal would require expressive machinery beyond
second-order arithmetic, strictly stronger than
Definition~\ref{def:axiom} in the reverse-mathematics sense of
clause~(ii) (Section~\ref{sec:open-questions}, item~(2)).
Principle~\ref{princ:austerity} clause~(ii) (minimal strength)
therefore excludes the uncountable case. The dimension parameter
ranges over $\mathbb{N}$.
\end{proof}

\noindent\textbf{No infinite-dimensional ambient space.} Theorem~\ref{thm:iterate}
does not build a single infinite-dimensional Hilbert space as an
ambient object. The cascade is the dimension tower itself: the sequence
of finite-dimensional Euclidean structures $\{(\mathbb{R}^d, B^d)\}_{d\in\mathbb{N}}$
together with the relations between consecutive levels (the slicing
recurrence of Section~\ref{sec:nothing-unstable}). The
``infinite-dimensional unit ball $B^\infty$'' is the asymptotic upper
edge of the tower, defined as the $d\to\infty$ limit of intrinsic
geometric quantities (volumes, sphere areas; see
Theorem~\ref{thm:nothing} below), not as a literal infinite-dimensional
ambient object. This means the cascade does not require Cauchy
completion to a single infinite-dimensional Hilbert space, does not
require any functional-analytic refinement (rigged, nuclear,
weak-topology), and does not require Sol\`er-style convergence theorems
to force its starting structure: each $\mathbb{R}^d$ with its standard
inner product is forced by elementary linear algebra (the unique
positive-definite symmetric bilinear form on $\mathbb{R}^d$, up to
isomorphism), and the tower itself is forced by clauses~(i) and~(ii) of
Principle~\ref{princ:austerity} as in the proof above. The cascade's
foundational footprint stays at $\mathrm{RCA}_0$ for the dimension
tower's existence and at $\mathrm{WKL}_0$/$\mathrm{ACA}_0$ for the
classical-analysis content (integration, the Gamma function) developed
in subsequent sections.

% ============================================================
\section{Scale and Rotation Invariance: Collapse at Each Dimension}
% ============================================================

Starting from nothing, there is no external reference against which to
measure absolute magnitude. There is no metre stick, no Planck mass,
no energy scale. The only quantities that can be defined without an
external reference are \emph{ratios}. At each $d$, the dilation
$x\mapsto\lambda x$ on $\mathbb{R}^d$ for $\lambda\in\mathbb{R}^{*}$
is unobservable from within the structure. By
Principle~\ref{princ:austerity} clause~(iii) (unobservable quotient),
the cascade structure at each $d$ is invariant under dilation.

The same applies to rotations. The orthogonal group $\mathrm{SO}(d)$
acts on the labeling of the orthonormal frame in $\mathbb{R}^d$, not
on the frame's existence. Cascade calculations downstream (the slicing
recurrence of Section~\ref{sec:nothing-unstable}, the four
distinguished dimensions of Part~0, the cascade invariant) are
$\mathrm{SO}(d)$-invariant: they depend on intrinsic geometric
quantities (volumes, sphere areas, Gamma function values), not on any
specific axis. By clause~(iii), $\mathrm{SO}(d)$ is also quotiented at
each $d$.

\noindent\textbf{The collapsed substrate at each $d$.} The combined
quotient $\mathbb{R}^d/(\mathbb{R}^*\times\mathrm{SO}(d))$ collapses to
a single point: $\mathrm{SO}(d)$ acts transitively on rays in
$\mathbb{R}^d$, so all rays are identified; $\mathbb{R}^*$ identifies
points along each ray. At every finite $d$, the projective space modulo
rotations is therefore a single point. This collapse is not a vacuous
result --- it is the cascade's substrate at each $d$. Structured
nothing has full rotational symmetry: no direction is privileged, no
basis is canonical, and the only data the substrate carries at level
$d$ is the bare fact of being a $d$-dimensional unit ball. The
dimension parameter $d$ survives both quotients (it is counting data,
not a direction), and at each $d$ the intrinsic geometric content of
the unit ball $B^d$ --- its volume $V_d$, its boundary sphere area
$\Omega_{d-1}$, the slicing-recurrence ratio $R(d)$, the decay rate
$p(d)$, the effective slicing radius $R_{\mathrm{eff}}(d)$ --- is
rotation-invariant and non-trivial.

\noindent\textbf{The cascade is the resolution of unresolved infinity.}
The cascade is not a static structure indexed by $d$. It is the
\emph{process} by which the unresolved infinity at the asymptotic
upper edge of the dimension tower is resolved into finite-dimensional
intrinsic content, dimension by dimension. The slicing recurrence
(Section~\ref{sec:nothing-unstable}) is the resolution mechanism:
each step relates the intrinsic content at $d+1$ to the intrinsic
content at $d$, extracting one dimension's worth of finite content
from the asymptotic limit. The dimension tower
$\{(\mathbb{R}^d, B^d)\}_{d\in\mathbb{N}}$ is the trace of this
resolution; the integer $d$ is the resolution stage. The universe
does not contain infinity --- the universe \emph{is} the result of
resolution applied to it, observed from a particular stage. This is
the unifying observation of the cascade series (cover sheet, ``one
infinity, not many''): every infinity problem in physics --- the
cosmological constant's $10^{-120}$, the black-hole horizon's
infinite time dilation, the ultraviolet floor at $\Omega_{217}$
(manifest temporally as the Big Bang in the universe's past) --- is a
local slice of the same asymptotic boundary, and every finite
structure observed is what that one infinity looks like partway
through resolution.

\noindent\textbf{Computational representative.} Although the substrate
at each $d$ is the collapsed quotient, calculations are performed on
$\mathbb{R}^d$ directly, with its standard inner product and unit ball
$B^d$. Cascade calculations on $\mathbb{R}^d$ are
$\mathrm{SO}(d)$-equivariant by construction --- they descend to the
collapsed substrate because their content (volumes, areas, dimension
counts, Gamma function identities) is invariant under any choice of
frame. Unit norm is the choice of canonical representative of each
ray; the choice of frame within $\mathbb{R}^d$ is a further gauge
fixing that washes out under $\mathrm{SO}(d)$ when the calculation
completes.

\noindent\textbf{Other candidate symmetries.}
\begin{itemize}
\item \emph{$\mathbb{C}^*$-phase rotation}. A phase action requires a
complex structure $J$ with $J^2=-\mathrm{Id}$. No such structure is
present at the Prelude stage: each $\mathbb{R}^d$ is constructed over
$\mathbb{R}$ (see Section~\ref{sec:what-not}, ``Real start, complex
amplitudes, quaternionic spacetime''). The forced precession
$\alpha=\pi/2$ of Paper~II induces $J$ downstream, at which point
whether $\mathbb{C}^*$ is observable or unobservable becomes a
separate question for Paper~II and beyond. At the Prelude stage there
is no $\mathbb{C}^*$ action to quotient.
\item \emph{Discrete scale invariance} (quotient by a specific ratio
$\lambda = r$ for some fixed $r$). Selecting a specific $r$ introduces
an unexplained numerical input, violating clause~(i) (parameter
economy). A continuous quotient has no such free parameter; only
discrete reductions introduce one.
\end{itemize}

\begin{definition}[Sphere representative and ball completion]\label{def:ball}
The unit sphere $S^{d-1}=\partial B^d$ is the canonical 2-fold
representative of the projective space
$\mathbb{R}\mathbb{P}^{d-1}=(\mathbb{R}^d\setminus\{0\})/\mathbb{R}^*$:
each ray through the origin meets $S^{d-1}$ in exactly one antipodal
pair. The unit ball $B^d = \{x\in\mathbb{R}^d : |x|\leq 1\}$ is the
convex completion: each ray meets $B^d\setminus\{0\}$ in a half-open
radial segment rather than at a single point, so the ball is not
itself a section of $\mathbb{R}\mathbb{P}^{d-1}$ but contains
$S^{d-1}$ as its boundary. The ball organises the slicing recurrence
(Remark below); the sphere is where the cascade's content lives. Every
subsequent theorem of the cascade is projective-invariant and
rotation-invariant --- it depends only on dimension counts and intrinsic
geometric ratios, not on the chosen representative or its completion
or any specific frame --- so the joint choice is computational, not
ontological.
\end{definition}

\begin{remark}[Why the ball rather than the sphere]
We work with the ball $B^d$ rather than the sphere $S^{d-1}$ because
the ball is the natural site for the slicing recurrence: slicing
$B^{d+1}$ perpendicular to any axis at height $x\in[-1,1]$ yields a
ball $B^d$ of radius $\sqrt{1-x^2}$, whose boundary is a sphere
$S^{d-1}$. This is the slicing recurrence of
Section~\ref{sec:nothing-unstable}, and it is the mechanism by which
the Gamma function enters. The sphere $S^d$ does not carry an
analogous native recurrence to lower-dimensional spheres; the ball is
the minimal container in which the recurrence is expressible without
auxiliary structure. Once the recurrence is in place, boundary
dominance ($\Omega_{d-1}/V_d = d$) shows that the spheres carry
essentially all the content --- so physics lives on the sphere, but
the recurrence is organised by the ball.
\end{remark}

\begin{corollary}[The starting structure]\label{cor:start}
The mathematical structure corresponding to ``nothing with the
capacity for distinction'' is the dimension tower
$\{(\mathbb{R}^d, B^d)\}_{d\in\mathbb{N}}$ of finite-dimensional
Euclidean spaces and their unit balls, with each level collapsed to
a single substrate point under
$\mathbb{R}^*\times\mathrm{SO}(d)$ but carrying non-trivial intrinsic
geometric content (volumes, areas, recurrence ratios). The asymptotic
upper edge --- the $d\to\infty$ limit, denoted $B^\infty$ --- is
structured nothing (Theorem~\ref{thm:nothing} below). The cascade is
the trajectory of intrinsic content unfolded across this tower via
the slicing recurrence (Section~\ref{sec:nothing-unstable}).
\end{corollary}

% ============================================================
\section{Nothing Is Empty}
% ============================================================

The asymptotic upper edge of the dimension tower --- denoted
$B^\infty$, defined as the $d\to\infty$ limit of the intrinsic
geometric quantities of $B^d$ --- is empty in every measurable sense.

\begin{theorem}[Geometric nothing at the asymptotic upper edge]\label{thm:nothing}
$B^\infty$ has zero volume, zero surface area, and concentration of
measure: for any $\varepsilon > 0$, an arbitrarily large fraction of
the mass of $B^d$ concentrates within $\varepsilon/\sqrt{d}$ of the
boundary as $d\to\infty$.
\end{theorem}

\begin{proof}
$V_d = \pi^{d/2}/\Gamma(d/2+1)\to 0$ as $d\to\infty$.
$\Omega_{d-1} = 2\pi^{d/2}/\Gamma(d/2)\to 0$ as $d\to\infty$. By
concentration of measure on $S^{d-1}$, the uniform distribution on
$B^d$ places a fraction $1 - (1-\varepsilon^2/d)^{d/2} \to
1 - e^{-\varepsilon^2/2}$ of its mass within a shell of thickness
$\varepsilon/\sqrt{d}$ of the boundary.
\end{proof}

The asymptotic limit $B^\infty$ has no volume to fill, no boundary
to inhabit, no interior to explore. It is the geometric expression of
``distinction exists but content does not.'' It is structured nothing
--- the upper edge of the dimension tower, where intrinsic content
vanishes asymptotically.

% ============================================================
\section{Resolution: The Slicing Recurrence}\label{sec:nothing-unstable}
% ============================================================

The asymptotic upper edge $B^\infty$ is empty, but the dimension tower
as a whole is not inert. Each level has nontrivial intrinsic content
(volumes, sphere areas, beta-function ratios), and consecutive levels
are linked by the slicing recurrence. This recurrence is the
resolution mechanism: it relates the intrinsic content at $d+1$ to
the intrinsic content at $d$, extracting one dimension's worth of
finite content from the asymptotic limit at each step.

Cut $B^{d+1}$ perpendicular to one axis. Each cross-section at height
$x$ is a $d$-ball of radius $\sqrt{1-x^2}$. The volumes obey:
\[
V_{d+1} = V_d \cdot \sqrt{\pi}\cdot R(d+1), \qquad R(d) =
\frac{\Gamma((d+1)/2)}{\Gamma((d+2)/2)}.
\]
The constant $\sqrt{\pi}$ is forced by orthogonality
(Theorem~\ref{thm:orth}): the angle between axis and equator is
$\pi/2$, forcing the half-integer argument in $B(1/2,\cdot)$, giving
$\Gamma(1/2) = \sqrt{\pi}$.

Read structurally as the cascade's static content per dimension at any
fixed resolution stage: as $d$ ranges from $B^\infty$ down through
small finite values, volumes \emph{grow}, and the ball acquires content
it did not have at higher $d$. The volume peaks at $d = 5$. Each step
(from $d{+}1$ to $d$) extracts more finite content from the asymptotic
limit; the structured nothing of $B^\infty$ becomes structured
\emph{something} at every finite~$d$. The recurrence has no free
parameters: the resolution at each step is fully determined by the
Gamma function. (The temporal reading --- in which the universe's age
is the count of resolution stages reached so far, with truncation
height $N(t)$ growing one cascade layer per Planck tick --- is the
subject of the Remark below; the structural picture here is the static
content the dynamic process resolves into at each stage.)

\begin{remark}[Resolution as cosmic time]
The cascade's resolution process admits a physical reading developed
in Parts~II--VI. Under that reading, each step of the resolution ---
each new dimension added to the realised tower --- corresponds to
one Planck-time tick of cosmic evolution. The universe had a start
(when resolution began at finite stage), counts on indefinitely
(each tick adds one more dimension), and approaches but never reaches
$d=\infty$ (full resolution would be the asymptotic future, never
completed). The universe does not \emph{contain} infinity --- it is
the result of resolution-in-progress, observed from the current
stage. Time, on this reading, is not an additional ingredient: it
is the cascade's own resolution direction, registered as the
dimension tower's growth.

\smallskip
\noindent\textbf{Three vocabularies for one rate.} The downstream
papers develop the same resolution rate in three coordinated
readings, each offering a different way to picture what is happening:
\begin{itemize}
\item \emph{Mathematical (this Prelude, Part~0).} The slicing
recurrence (Section~\ref{sec:nothing-unstable}) extracts one
dimension's worth of finite content from $B^\infty$ at each step.
The recurrence is the resolution mechanism; the dimension tower
$\{(\mathbb{R}^d, B^d)\}_{d\in\mathbb{N}}$ is its trace; the integer
$d$ is the resolution stage. This is the structural reading: the
cascade as the unique parameter-free $d\to d-1$ recurrence on the
tower, indexed by Gamma-function values.
\item \emph{Physical (Parts~II and VI).} Each step is one Planck
tick of cosmic time. Part~II Section~7.1 makes the static-versus-
dynamic identification explicit: each cascade slicing operation
\emph{is} one elementary unit of time --- ``slicing $=$ dimensional
collapse $=$ time'' --- not merely a coordinate parametrising the
tower. Part~VI promotes this to a cosmological dynamics: the
truncation height grows by one per tick,
$N(t+\alpha\,t_{\mathrm{Pl,red}}) = N(t)+1$, and the universe's
phases A--D are stages of the resolution prior to and beyond the
$\Omega_{217}$ floor. The universe's age is the count of resolution
stages reached so far.
\item \emph{Geometric (cover sheet).} The observer lives on the
$S^3$ shell of a 5D black hole, asymptotically falling toward the
pseudo-horizon. Each resolution step is one asymptotic step of that
infall: one dimension is collapsed at the cascade's boundary, and
the observer's host shell advances one Planck tick closer to the
pseudo-horizon. The infall never completes because $B^\infty$ has
infinitely many dimensions to resolve. ``Time dilation'' is the
local rate at which resolution proceeds along the observer's
worldline; ``the Big Bang'' is the resolution reaching the
$\Omega_{217}$ floor.
\end{itemize}
The three readings are dual descriptions of the same rate. Each
gives a distinct mental picture --- mathematical (a recurrence
extracting content), physical (a tick adding a layer), geometric
(an asymptotic infall) --- but they describe the same operation.

\smallskip
The Prelude commits only to the mathematical reading (the slicing
recurrence as the resolution mechanism, the dimension tower as its
trace); the physical identification of resolution with cosmic time,
the geometric identification with asymptotic infall toward the
pseudo-horizon, and the joint claim that one Planck time corresponds
to one tower extension, are the responsibility of the cover sheet
and Parts~II (time as dimensional collapse), V (cosmology, $\Omega_d$
at the cosmological-constant stage $d_2=217$), and~VI (tower growth,
phase structure of the pre- and post-Big-Bang universe).
\end{remark}

% ============================================================
\section{Where Resolution Lands: The Observer at \texorpdfstring{$d=4$}{d=4}}
% ============================================================

The slicing recurrence resolves infinity into finite content at each
step. Where does the resolution stably terminate?

The volume maximum is at $d = 5$: the unique dimension where the unit
ball has the largest volume. The observer sits at $d = 4$, the
boundary $S^3$ of $B^5$. By boundary dominance ($\Omega_{d-1}/V_d =
d$), this boundary carries $4/5$ of the content.

Why does the observer end up here? Part~III of the series proves
that $d = 4$ is the unique spacetime dimension satisfying two
independent constraints simultaneously:

\begin{enumerate}
\item \textbf{Lovelock uniqueness}: the gravitational equation has a
unique form (Einstein's equation with cosmological constant) only at
$d = 4$.
\item \textbf{Complex spinors}: the Clifford algebra
$\mathrm{Cl}(1,d-1)$ has irreducibly complex minimal spinors --- the
prerequisite for quantum mechanics --- only when
$d\bmod 8\in\{4,5,6\}$.
\end{enumerate}

The intersection is $\{4\}$. No other dimension permits both unique
physical laws and quantum information processing. The observer is not
placed at $d = 4$ by assumption; $d = 4$ is where the resolution
trajectory naturally lands --- the unique stage at which stable
physical laws and complex quantum phase coincide. The resolution
process described in Section~\ref{sec:nothing-unstable} continues
beyond $d=4$ in both directions (toward $d=\infty$ as the asymptotic
upper edge, and toward $d=0$ as the trivial lower endpoint), but the
\emph{observable physics} of the resolution lives at $d=4$, the
boundary of the volume-maximum ball.

% ============================================================
\section{The Chain}
% ============================================================

The complete logical chain, from pre-mathematics to physics, is:

\bigskip

\begin{center}
\renewcommand{\arraystretch}{1.4}
\resizebox{\textwidth}{!}{%
\begin{tabular}{rlll}
\toprule
Step & Statement & Status & Austerity \\
\midrule
0 & $0 \neq 1$ & Pre-mathematical & --- \\
1 & Distinction $\Rightarrow$ orthogonality & Theorem~\ref{thm:orth} & --- (interpretive) \\
2 & Dimension tower $\{(\mathbb{R}^d, B^d)\}_{d\in\mathbb{N}}$ & Theorem~\ref{thm:iterate} & (i), (ii) \\
3 & Scale + rotation invariance at each $d\;\Rightarrow$ collapsed substrate; cascade as $d$-trajectory & Definition~\ref{def:ball}, Cor~\ref{cor:start} & (iii) \\
4 & Asymptotic $B^\infty$: zero volume, zero area & Theorem~\ref{thm:nothing} & --- \\
5 & Slicing recurrence, constant $\sqrt{\pi}$ & Forced by step 1 & --- \\
6 & Descent: volumes grow, peak at $d = 5$ & Gamma function & --- \\
7 & Observer at $d = 4$: boundary of the peak & Lovelock $\cap$ Clifford & --- \\
8 & Four distinguished dimensions: $5, 7, 19, 217$ & Part~0 & --- \\
9 & Cascade invariant $\approx 10^{-120}$ & Part~0 & --- \\
10 & Cosmological constant, QM, GR, SM, cosmology & Parts~I--V & --- \\
\bottomrule
\end{tabular}%
}
\end{center}

\bigskip

Steps 0--4 are established in this prelude. Step~5 is proved in
Part~0 (Theorem~3.1). Steps 6--7 are proved in Parts~0 and~III.
Steps 8--10 are the content of the series.

No step introduces a free parameter beyond the single declared
pre-mathematical input Definition~\ref{def:axiom} ($0\neq 1$).
Principle~\ref{princ:austerity} (austerity) is not a second input: it
is the operational content of Definition~\ref{def:axiom}'s status as
sole input, derived in Section~\ref{sec:what-nothing-means} above.
Steps~2 and~3 invoke austerity clauses as noted in the ``Austerity''
column; the remaining chain steps are classical theorems about the
Gamma function or established uniqueness results in known mathematics.
Step~1's interpretive bridge (logical distinction as geometric
orthogonality in an inner-product space) is a separate structural
choice, flagged as open in Section~\ref{sec:open-questions}.

% ============================================================
\section{What This Paper Does Not Do}
\label{sec:what-not}
% ============================================================

\textbf{Does not derive physics.} This prelude contains no physical
predictions. Its purpose is to argue that the starting structure of
the series --- the dimension tower
$\{(\mathbb{R}^d, B^d)\}_{d\in\mathbb{N}}$ of finite-dimensional
Euclidean spaces, with asymptotic upper edge $B^\infty$ --- is
\emph{defensibly minimal} given the single declared pre-mathematical
input (Definition~\ref{def:axiom}, $0\neq 1$) and the austerity
derivation (Principle~\ref{princ:austerity}) that follows from its
sole-input status. The chain makes explicit which austerity clause is
invoked at each step: clauses~(i) and~(ii) at step~2 (parameter
economy forcing the dimension tower to extend without bound; minimal
strength forcing countable over uncountable indexing via the
$\mathrm{RCA}_0$ commitment); clause~(iii) at step~3 (unobservable
quotient, applied at each $d$ to both $\mathbb{R}^*$-dilation and
$\mathrm{SO}(d)$, collapsing the substrate at each $d$ to a single
point from which intrinsic content unfolds via the slicing recurrence).
Whether the combined input (Definition~\ref{def:axiom} plus austerity)
forces this dimension tower \emph{uniquely}, or merely renders it the
most austere defensible starting structure, reduces to the single open
foundational question of Section~\ref{sec:open-questions} item~(3)(a):
the structural commitment to the inner-product realisation of
distinguishable states. The cascade series' load-bearing hypothesis
(cover sheet) is independent of whichever answer is reached here: the
downstream derivations take the dimension tower (and its asymptotic
upper edge $B^\infty$) as given and test it by consequence.

\textbf{Real start, complex amplitudes, quaternionic spacetime.} The
Prelude's starting structure is over $\mathbb{R}$ at every level of
the dimension tower: no complex or quaternionic structure is
introduced by hand. By Hurwitz's theorem, the associative normed
division algebras over $\mathbb{R}$ are exactly $\mathbb{R}$,
$\mathbb{C}$, and $\mathbb{H}$, and the Gamma function governs the
slicing recurrence in all three cases. The cascade selects between
them downstream:
\begin{itemize}
\item \emph{Real at the start.} Each $\mathbb{R}^d$ in the dimension
tower is the most austere starting structure at dimension $d$ and is
the one used here. No complex or quaternionic structure is assumed at
any level.
\item \emph{Complex at the amplitude level.} A complex structure $J$
with $J^2=-\mathrm{Id}$ arises as a \emph{derived} consequence of the
forced precession $\alpha=\pi/2$ in Part~II. It is a theorem, not an
axiom: two consecutive cascade quarter-turns compose to multiplication
by $-1$, promoting the state space to a complex structure at the
quantum layer.
\item \emph{Quaternionic at the spacetime dimension.} Lovelock
uniqueness intersected with Clifford periodicity forces the spacetime
dimension to $d=4=\dim_\mathbb{R}\mathbb{H}$ in Part~III. The
\emph{spacetime} dimension matches the quaternions while the
\emph{amplitude} algebra remains $\mathbb{C}$.
\end{itemize}
The real/complex/quaternionic hierarchy is thus not an open question:
each algebra enters the cascade at the layer forced by a downstream
theorem. What the Prelude does not do is prove that the cascade could
not have been written over $\mathbb{C}$ or $\mathbb{H}$ from the
outset --- it proves only that the real case suffices, and that the
complex and quaternionic roles are recovered downstream without
additional input.

\textbf{Does not resolve the hard problem.} The cascade derives the
dimension where self-awareness is mathematically possible. It does
not derive self-awareness itself. The question ``why is there
something rather than nothing?''\ receives an answer: because
nothing, taken seriously, has structure. The question ``why is there
\emph{experience}?''\ remains open.

% ============================================================
\section{Open Questions}
\label{sec:open-questions}
% ============================================================

Principle~\ref{princ:austerity} derives three previously-implicit
parsimony choices (finite-vs-$\aleph_0$ dimension index;
countable-vs-uncountable indexing of the dimension tower; the
unobservable-continuous-symmetry quotients of $\mathbb{R}^*$-dilation
and $\mathrm{SO}(d)$ at each $d$, collapsing each level to a single
substrate point) as clauses of a single derivation rule from
Definition~\ref{def:axiom}. The prelude chain now invokes these
clauses explicitly, and austerity is derived from the sole-input
status of $0\neq 1$ rather than declared as a second input. What
remains open:

\textbf{(1) Uniqueness of the resolution process under austerity ---
reduces to (3)(a).} Is the cascade's resolution process --- the
dimension tower $\{(\mathbb{R}^d, B^d)\}_{d\in\mathbb{N}}$ of
finite-dimensional Euclidean spaces with each level collapsed under
$\mathbb{R}^*\times\mathrm{SO}(d)$, together with the slicing
recurrence as resolution mechanism extracting finite content from the
asymptotic upper edge $B^\infty$ at $d\to\infty$ --- the \emph{unique}
mathematical process derivable from Definition~\ref{def:axiom} under
Principle~\ref{princ:austerity}, or merely one of several
austerity-compatible resolution processes?

The dimension tower has two components, each examinable for forcing:
\begin{itemize}
\item[$\bullet$] \emph{The orthogonal-state structure at each
cardinality.} The integer $d$ does not enter the cascade as a
separate primitive: it is the cardinality of the orthogonal-state
set --- the count of orthogonal directions in the realisation chosen
by sub-question (3)(a). Integer indexing and the space at each level
are the same thing: ``$\mathbb{R}^d$ at level $d$'' is ``the space of
$d$ orthogonal states under (3)(a)'s realisation,'' and integer
arithmetic on $d$ is just adding or removing one orthogonal direction.
The cardinality $d$ ranges over $\mathbb{N}$ by
Theorem~\ref{thm:iterate} (clause~(ii) via $\mathrm{RCA}_0$, item~(2)
below: countable Gram--Schmidt is provable, uncountable orthonormal
distinction is not; clause~(i) excludes any specific finite cutoff as
a free parameter; $\mathbb{N}$ is $\mathrm{RCA}_0$'s native unbounded
notion among countable index sets). Forcing the cardinality and the
space at each cardinality therefore reduces to a single structural
question --- the realisation of distinguishable states recorded in
(3)(a). \emph{Open as (3)(a)}; not closed within (1).
\item[$\bullet$] \emph{Bounded region $B^d$ within the cardinality-$d$
realisation.} Forced given the previous component. Among compact
convex regions in $\mathbb{R}^d$ with its standard inner product, the
unit ball $B^d = \{x : |x|\leq 1\}$ is the unique parameter-free
shape with full $\mathrm{SO}(d)$ symmetry: alternative shapes (cubes,
simplices) break rotation symmetry; alternative balls (ellipsoids,
balls of radius $r\neq 1$) introduce shape or scale parameters
excluded by clause~(i). The unit ball is also the unique shape
supporting the slicing recurrence with constant $\sqrt{\pi}$
(Part~0 Theorem~4.1: the recurrence's irreducible residue is forced
by orthogonality on the cardinality-$d$ realisation).
\end{itemize}

The resolution mechanism itself is also forced given the tower: the
slicing recurrence is the unique first-order multiplicative recurrence
on the tower of unit balls (Part~0 Theorem~4.1).

\noindent\textbf{Reduction.} The bounded region and the resolution
mechanism are forced once the orthogonal-state realisation of (3)(a)
is fixed. Integer indexing is not a separate component --- it is the
cardinality of (3)(a)'s structural commitment. Therefore (1) is
closed up to (3)(a): \emph{given the inner-product realisation of
$0\neq 1$ as orthogonal states, the cascade's resolution process is
uniquely determined.} The remaining open question collapses into the
interpretive bridge of (3)(a) --- a single substantive structural
commitment of the prelude, comparable in status to ``use real numbers
for analysis,'' rather than an independent open item.

Three classes of alternatives at the level of the whole structure are
explicitly ruled out inside this prelude:
\begin{itemize}
\item[$\bullet$] \emph{Complex or quaternionic Euclidean spaces as
starting structure.} Ruled out by clause~(ii) (minimal strength):
starting over $\mathbb{C}$ or $\mathbb{H}$ at each $d$ is strictly
stronger than starting over $\mathbb{R}$, and the latter suffices.
Section~\ref{sec:what-not} establishes that the real case is the most
austere starting structure; $J$ with $J^2=-\mathrm{Id}$ is recovered
as a Paper~II \emph{theorem} from the forced precession, not imposed
upstream.
\item[$\bullet$] \emph{Uncountable indexing of the dimension tower.}
Ruled out inside Theorem~\ref{thm:iterate} by clause~(ii):
uncountable indexing requires expressive machinery beyond second-order
arithmetic.
\item[$\bullet$] \emph{An infinite-dimensional ambient Hilbert space.}
Ruled out as superfluous: the cascade's content (slicing recurrence,
intrinsic geometric quantities, asymptotic limit) is fully expressible
on the dimension tower of finite-dimensional Euclidean spaces, with
$B^\infty$ as the asymptotic upper edge rather than a literal
infinite-dimensional ambient object. No Cauchy completion, no
Sol\`er-style convergence theorem, no functional-analytic refinement
(rigged Hilbert space, nuclear space, weak-topology variant) is
required to express any downstream cascade theorem.
\end{itemize}

\textbf{(2) Formal measure of assumption strength (clause (ii)) ---
resolved.} The partial order on assumption strength used in clause~(ii)
of Principle~\ref{princ:austerity} is the reverse-mathematics ordering
over the base theory $\mathrm{RCA}_0$
\cite{friedman1975, simpson2009}: $A$ is strictly stronger than $B$ if
$\mathrm{RCA}_0 + A \vdash B$ but $\mathrm{RCA}_0 + B \nvdash A$. This
framework calibrates the strength of mathematical theorems against the
``Big Five'' subsystems of second-order arithmetic
($\mathrm{RCA}_0 \subsetneq \mathrm{WKL}_0 \subsetneq \mathrm{ACA}_0
\subsetneq \mathrm{ATR}_0 \subsetneq \Pi^1_1\text{-}\mathrm{CA}_0$),
giving a precise quantitative measure of any cascade derivation's
foundational footprint. The countable-vs-uncountable comparison
invoked in Theorem~\ref{thm:iterate} is a clean reverse-mathematics
statement: the dimension tower indexed by $\mathbb{N}$ with $\mathbb{R}^d$
at each level is expressible in $\mathrm{RCA}_0$; uncountable indexing
requires expressive machinery beyond second-order arithmetic. The
cascade's downstream content --- classical analysis on
finite-dimensional Euclidean spaces, the Gamma function and its
analytic properties, the slicing recurrence as a Beta-function identity
--- calibrates at $\mathrm{WKL}_0$/$\mathrm{ACA}_0$ within the Big
Five hierarchy. No Cauchy completion to an infinite-dimensional
Hilbert space is required (see Theorem~\ref{thm:iterate}'s ``No
infinite-dimensional ambient space'' note), so the prelude itself
stays at $\mathrm{RCA}_0$. The cascade's foundational footprint is
therefore bounded by a well-understood, modest fragment of second-order
arithmetic. The choice of $\mathrm{RCA}_0$ as the strength-measurement
base is a foundational commitment, not a cascade-internal axiom: it
specifies how clause~(ii) is to be applied without altering what the
cascade proves from $0\neq 1$.

\textbf{(3) The interpretive bridge at Theorem~\ref{thm:orth} ---
elementary closure.} The step from logical $0\neq 1$ to geometric
``states sharing no common component in $\mathbb{R}^d$'' (the premise
of Theorem~\ref{thm:orth}) is a structural interpretation, not a
clause of austerity. Austerity governs \emph{which} structure is
admissible once the interpretive framework is fixed; it does not fix
the interpretive framework itself. The interpretive choice splits into
two layers:
\begin{itemize}
\item[(a)] \emph{The choice of $\mathbb{R}^d$ as the realisation of
$d$ distinguishable states.} This is the genuine interpretive
commitment: ``model $d$ logical distinctions as $d$ orthogonal
directions in a $d$-dimensional Euclidean space.'' Integers do not
enter as a separate primitive --- the integer $d$ is the cardinality
of the orthogonal-state set under this realisation, and integer
arithmetic on $d$ is just adding or removing one orthogonal direction.
This is comparable in status to ``model continuous quantities as real
numbers'' or ``model directed quantities as vectors'' --- a
structural-realisation choice that the cascade adopts without further
justification. Whether some weaker structural realisation (a
non-Euclidean inner product, a non-Hilbert metric, a topology without
inner product, an orthomodular lattice without an underlying vector
space) also admits the cascade's subsequent chain is open in
principle, but the alternatives that genuinely support the slicing
recurrence and Gamma-function machinery all reduce to or refine the
Euclidean realisation. The cascade adopts $\mathbb{R}^d$-Euclidean as
the most elementary structural choice admitting its content.
\item[(b)] \emph{The geometry on $\mathbb{R}^d$ once chosen.} This is
\emph{not} an interpretive choice but a theorem of elementary linear
algebra: the standard Euclidean inner product on $\mathbb{R}^d$ is
the unique positive-definite symmetric bilinear form on
$\mathbb{R}^d$, up to isomorphism, given a choice of basis. No Hilbert
space machinery, no Sol\`er-style convergence theorem, no choice
principle is required to fix it.
\end{itemize}
Sub-question (a) is the genuine open item; sub-question (b) is closed
elementarily. Closing (a) would require either a uniqueness theorem of
the form ``the most austere geometric realisation of $d$ distinguishable
states admitting the slicing recurrence is $\mathbb{R}^d$-Euclidean'' ---
which would promote the prelude from \emph{defensibly minimal} to
\emph{forced} on this axis --- or a meta-foundational principle
governing the choice of structural realisation. Both are
foundations-of-mathematics questions outside the cascade's own scope,
parallel to the meta-question of why standard mathematics is the
natural ambient framework.

\noindent\textbf{Foundational footprint preserved.} Because the
interpretive bridge is closed elementarily (linear algebra at each
$d$) rather than via Sol\`er-style machinery on an infinite-dimensional
ambient space, the cascade's foundational footprint stays at the
modest end of the Big Five hierarchy: $\mathrm{RCA}_0$ for the
dimension tower, $\mathrm{WKL}_0$/$\mathrm{ACA}_0$ for the classical
analysis content of Part~0 onward. No higher-strength machinery is
imported by the prelude.

\textbf{(4) The hard problem.} The cascade derives the
dimension where self-awareness is mathematically possible. It does not
derive self-awareness itself (already noted in Section~\ref{sec:what-not}).

\bigskip

The series begins.

\begin{thebibliography}{99}

\bibitem{friedman1975} H.~Friedman, ``Some systems of second-order
arithmetic and their use,'' in \textit{Proceedings of the International
Congress of Mathematicians, Vancouver}, vol.~1, pp.~235--242 (1975).

\bibitem{simpson2009} S.~G.~Simpson, \textit{Subsystems of Second
Order Arithmetic}, 2nd ed., Cambridge University Press, 2009.

\end{thebibliography}

\end{document}
