\documentclass[11pt,a4paper]{article}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{geometry}
\usepackage{hyperref}
\usepackage{booktabs}
\geometry{margin=2.5cm}

\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{remark}[theorem]{Remark}

\title{The Cascade Series --- Part 0\\[0.5em]
\large Scale Variance from Orthogonality:\\
How the Unit Ball Generates $10^{120}$ Orders of Magnitude}
\author{RTAC}
\date{March 2026}

\begin{document}
\maketitle

\begin{abstract}
The unit ball in $\mathbb{R}^d$ has a slicing recurrence
$V_{d+1}/V_d = \sqrt{\pi}\cdot R(d+1)$ containing one constant,
$\sqrt{\pi} = \Gamma(1/2)$, forced by orthogonality. The sphere-area decay
rate $p(d)$ has a unique natural zero at $d_0 = 7$. This zero generates
two threshold values $c_1 = \tfrac{1}{2}\ln\pi$ and $c_2 = \sqrt{\pi}$ ---
the two canonical values of the recurrence's unique constant, one in each
of its two canonical forms. The orbit terminates because a first-order
recurrence has exactly two canonical forms and one bridge map between them.
The thresholds select $d_1 = 19$ and $d_2 = 217$.

The three privileged values of $p(d)$---zero, $c_1$, and $c_2$---partition
the cascade into exactly four regimes (Growth, Decay, Exponential decay,
Oblivion), separated by three continuous boundary crossings lying strictly
between integers: $d_0^\star\approx 6.257$, $d_1^\star\approx 19.731$,
$d_2^\star\approx 217.627$. The Gamma function thus produces exactly four
distinguished dimensions: the volume maximum $d_V = 5$ (an interior
landmark of the Growth regime), plus one representative integer per
boundary crossing. No fifth exists.

The four dimensions separate by structural stability into two classes:
the equilibria are structurally stable (they exist for any value of the
recurrence constant) and serve as scale; the thresholds are structurally
sensitive (they depend on the specific constant $\sqrt{\pi}$) and carry
content. Scale enters as a ratio; content enters as a product.

The cascade invariant
\[
I = \max_{(d_0,d_1,d_2)}\;\frac{\Omega_5^2}{\Omega_{d_0}^2}\;\Omega_{d_1}\;\Omega_{d_2}
= \frac{9}{\pi^2}\;\Omega_{19}\;\Omega_{217}
= 1.0990\times 10^{-120},
\]
where the supremum runs over integer labelings of the three boundary pairs
$(d_0,d_1,d_2) \in \{6,7\}\times\{19,20\}\times\{217,218\}$ and the unique
argmax is $(7, 19, 217)$. This variational definition eliminates the need
for an integer rounding convention at each boundary pair. $I$ is the
unique order-0 (global) cascade invariant; it depends only on the four
distinguished dimensions and is immune to subleading corrections from
inter-layer structure. The formula for the hierarchy
is $\pi e^{2\sqrt{\pi}}(2\sqrt{\pi}-1)/\ln 10\approx 120$.

Every step is a theorem about the Gamma function. No physics enters.
\end{abstract}

\medskip\noindent\textbf{Notation.}
Throughout this paper, $\Omega_d$ denotes the area of $S^d$:
$\Omega_d = 2\pi^{(d+1)/2}/\Gamma((d+1)/2)$. The boundary of the unit
ball $B^d$ is $S^{d-1}$ with area $\Omega_{d-1}$. Boundary dominance
(Theorem~\ref{thm:boundary}): $\Omega_{d-1}/V_d = d$.

% ============================================================
\section{The Question}
% ============================================================

How much scale variance can pure mathematics produce from nothing?

The unit ball $B^d = \{x\in\mathbb{R}^d : |x|\leq 1\}$ exists in every
dimension. The area of the unit sphere $S^d$ is
$\Omega_d = 2\pi^{(d+1)/2}/\Gamma((d+1)/2)$. At $d = 7$:
$\Omega_7 = \pi^4/3\approx 32.5$. At $d = 217$:
$\Omega_{217}\approx 10^{-120}$. The ratio exceeds $10^{121}$. The unit
ball generates 121 orders of magnitude of scale variance with no free
parameters and no input beyond the definition of orthogonal dimensions.

This paper extracts two distinguished scales from this variance, proves
they combine uniquely with the cascade's two natural equilibria, and evaluates
the result.

% ============================================================
\section{The Recurrence}
% ============================================================

Cut $B^{d+1}$ perpendicular to one axis. Each cross-section at height $x$
is a $d$-ball of radius $\sqrt{1-x^2}$:
\begin{equation}\label{eq:slice}
V_{d+1} = V_d\int_{-1}^{1}(1-x^2)^{d/2}\,dx.
\end{equation}
Substitute $x = \cos\theta$. The integral becomes
$2\int_0^{\pi/2}\sin^d\theta\,d\theta$: one quarter turn from axis to
equator. Via the Beta function:
\begin{equation}\label{eq:recurrence}
V_{d+1} = V_d\cdot\sqrt{\pi}\cdot R(d+1), \qquad
R(d) := \frac{\Gamma((d+1)/2)}{\Gamma((d+2)/2)}.
\end{equation}

\begin{theorem}[Orthogonal compression constant]\label{thm:sqrtpi}
$\Gamma(1/2) = \sqrt{\pi}$ is the unique dimension-independent constant in
the recurrence, forced by orthogonality. It appears with coefficient
exactly~1.
\end{theorem}

\begin{proof}
The first argument of $B$ is $\frac{1}{2}$, forced by orthogonality: the
angle between axis and equator is $\pi/2$; the substitution
$x = \cos\theta$ produces the half-power $t^{-1/2}$ in $B(1/2,\cdot)$;
$\Gamma(1/2) = \sqrt{\pi}$. The chain: dimension $\to$ orthogonal $\to$
$\pi/2$ $\to$ $B(1/2,\cdot)$ $\to$ $\sqrt{\pi}$. Every arrow forced.
\end{proof}

% ============================================================
\section{Sphere Areas Are Primary}
% ============================================================

\begin{theorem}[Boundary dominance]\label{thm:boundary}
$\Omega_{d-1}/V_d = d$ for all $d\geq 1$.
\end{theorem}

\begin{proof}
$\Omega_{d-1} = 2\pi^{d/2}/\Gamma(d/2)$,\;
$V_d = \pi^{d/2}/[(d/2)\Gamma(d/2)]$.\; Ratio: $d$.
\end{proof}

\begin{corollary}\label{cor:primary}
$V_d = \Omega_{d-1}/d$. Volumes are derived from sphere areas. The
recurrence is a first-order multiplicative map between sphere areas:
\[
\Omega_{d+1} = \Omega_d\cdot\sqrt{\pi}\cdot R(d).
\]
Sphere area is the unique independent cascade quantity at each level.
Every other cascade object---$V_d$, $R(d)$,
$R_{\mathrm{eff}}(d) = 1/\sqrt{d+3}$---is derived from sphere areas.
\end{corollary}

% ============================================================
\section{Irreducibility}
% ============================================================

\begin{theorem}\label{thm:irred}
$\sqrt{\pi}$ is the irreducible residue of the recurrence.
\end{theorem}

\begin{proof}
The second difference
$\delta(d) := \ln(\Omega_{d+1}/\Omega_d) - \ln(\Omega_d/\Omega_{d-1})
= \ln(d/(d+1))$. All $\pi$-dependent and $\Gamma$-dependent factors
cancel exactly. $\sqrt{\pi}$ is the only nontrivial constant the
recurrence contains.
\end{proof}

% ============================================================
\section{The Natural Zero}
% ============================================================

The sphere-area decay rate decomposes as:
\[
p(d) = -\tfrac{1}{2}\ln\pi +
\tfrac{1}{2}\psi\!\left(\tfrac{d+1}{2}\right),
\]
where $\psi$ is the digamma function. Since $\psi^{(1)} > 0$, $p$ is
strictly monotone increasing with a unique zero.

\begin{definition}[Natural zero]\label{def:d0}
$d_0$ is the smallest integer $d$ such that $p(d) > 0$: the first
integer in the decay regime. Stirling: the continuous zero of $p$ lies at
$d^*\approx 2\pi\approx 6.27$; integer $d_0 = 7$.
\end{definition}

\begin{theorem}[Dual characterisation of $d_0$]\label{thm:dual}
The natural zero $d_0$ has two independent characterisations:
\begin{enumerate}
\item[(a)] \emph{Decay onset.} $d_0$ is the first integer at which
$p(d) > 0$.
\item[(b)] \emph{Maximum-boundary ball.} $B^{d_0}$ is the unique ball
whose boundary sphere has maximum area among all unit spheres.
\end{enumerate}
Both give $d_0 = 7$.
\end{theorem}

\begin{proof}
Let $d_{\max} = \operatorname{argmax}\Omega_d$ (discrete). By
unimodality of $\Omega_d$, $d_{\max}$ is unique, and the discrete ratio
satisfies
\[
\Omega_{d_{\max}+1}/\Omega_{d_{\max}} < 1 < \Omega_{d_{\max}}/\Omega_{d_{\max}-1}.
\]
Since $p$ is strictly increasing and continuous, and its sign governs
the growth/decay transition of the continuous interpolation of $\Omega_d$,
the unique zero $d^*$ of $p$ lies in the interval
$(d_{\max},\;d_{\max}+1)$. Therefore the first integer with $p(d) > 0$ is
$d_0 = d_{\max} + 1$.

By boundary dominance (Theorem~\ref{thm:boundary}),
$\partial B^d = S^{d-1}$. Therefore
$\partial B^{d_0} = S^{d_{\max}}$, which carries the maximum sphere area.

Evaluation: $\Omega_6 = 16\pi^3/15\approx 33.1 > \Omega_7 = \pi^4/3
\approx 32.5 > \Omega_5 = \pi^3\approx 31.0$, so $d_{\max} = 6$.
Bracket: $p(6) = -0.021 < 0 < 0.056 = p(7)$. Both characterisations
give $d_0 = 7$.
\end{proof}

\begin{remark}\label{rem:convention}
The two characterisations are independent: (a)~uses only the
monotonicity of $p$; (b)~uses only the unimodality of $\Omega_d$ and
boundary dominance. Their agreement at $d_0 = 7$ is a consistency check
on the cascade structure, not a convention choice. The invariant
(Section~8) uses $\Omega_7$ because $p$ is defined at sphere dimensions,
its zero selects sphere dimension~7, and the invariant collects sphere
areas at the dimensions where $p$ takes distinguished values.
\end{remark}

% ============================================================
\section{The Threshold Pair}
% ============================================================

\subsection{Operation A: the zero-crossing forces $c_1$}

\begin{theorem}[Operation A]\label{thm:opA}
The zero-crossing forces $c_1 := \tfrac{1}{2}\ln\pi$.
\end{theorem}

\begin{proof}
$p(d_0) = 0$ gives $\frac{1}{2}\psi((d_0+1)/2) = \frac{1}{2}\ln\pi$.
The right-hand side is the constant part of $p$, forced by
Theorem~\ref{thm:sqrtpi}.
\end{proof}

\begin{theorem}[Uniqueness of $c_1$]\label{thm:c1unique}
Among all scalars derivable from $d_0$, $c_1 = \tfrac{1}{2}\ln\pi$ is the
unique scalar satisfying (a) $c > 0$ (so $p(d) = c$ has a solution beyond
$d_0$) and (b) primitivity ($c$ is a recurrence constant).
\end{theorem}

\begin{proof}
We check all four information classes available from the natural zero.

\emph{Class~1: Values of $p$ and its decomposition at $d_0$.} The zero
gives three scalars: $p(d_0) = 0$, the $d$-dependent part
$q(d_0) = \frac{1}{2}\ln\pi$, and the constant part
$-\frac{1}{2}\ln\pi$. (i)~$p(d_0) = 0$ fails~(a): $p(d) = 0$ has its
solution at $d_0$ itself. (ii)~$q(d_0) = \frac{1}{2}\ln\pi$ satisfies~(a)
since $\frac{1}{2}\ln\pi > 0$, and satisfies~(b) since
$\frac{1}{2}\ln\pi$ is the log-space image of $\sqrt{\pi}$, the
recurrence's unique primitive. This is $c_1$. (iii)~$-\frac{1}{2}\ln\pi < 0$
fails~(a).

\emph{Class~2: Higher derivatives $p^{(n)}(d_0)$.} All fail~(b): they are
not recurrence primitives.

\emph{Class~3: The dimension $d_0$ itself.} $p(d) = d_0$ equates a
dimensionless rate to a dimension count---a type mismatch.

\emph{Class~4: Functions of $c_1$.} Consider $f(c_1)$ for any
$f\neq\mathrm{id}$.
%
\emph{Multiples:} $n c_1 = \tfrac{1}{2}\ln(\pi^n)$ is the constant of
the recurrence for $V_d^{\,n}$, a derived sequence. A primitive of a
derived sequence is not a primitive of the original first-order recurrence.
%
\emph{Powers:} $c_1^{\,k}$ for $k\geq 2$ does not appear as the
dimension-independent constant in any form of the recurrence; it is
algebraically independent of both $\sqrt{\pi}$ and $\frac{1}{2}\ln\pi$.
%
\emph{General:} any analytic $f(c_1)\neq c_1$ either reduces to a
multiple $n c_1$ (hence derived) or produces a scalar that is not a
recurrence constant in any canonical form of the first-order recurrence.

Across all four classes, $c_1 = \frac{1}{2}\ln\pi$ is unique.
\end{proof}

\subsection{Two canonical values of one constant}

The recurrence contains one irreducible constant
(Theorem~\ref{thm:irred}). That constant has two canonical
representations, because the recurrence has two canonical forms:

\begin{center}
\begin{tabular}{lll}
\toprule
Form & Expression & Constant \\
\midrule
Multiplicative & $V_{d+1}/V_d = \sqrt{\pi}\cdot R(d+1)$ &
$c_2 = \sqrt{\pi}$ \\
Logarithmic & $\Delta\ln V_d = \frac{1}{2}\ln\pi + \ln R(d+1)$ &
$c_1 = \frac{1}{2}\ln\pi$ \\
\bottomrule
\end{tabular}
\end{center}

These are the same constant in two algebraic spaces, related by the
exponential map: $e^{c_1} = e^{\frac{1}{2}\ln\pi} = \sqrt{\pi} = c_2$.

\begin{definition}[Canonical form]\label{def:canonical}
A canonical form of a first-order recurrence is an algebraically equivalent
expression that isolates the dimension-independent content as a single
constant with coefficient~1. For a multiplicative recurrence
$a_{n+1} = K\cdot f(n)\cdot a_n$, the two canonical forms are the
recurrence itself (constant $K$) and its logarithm (constant $\ln K$).
Any other algebraically equivalent expression either has a constant that
is a power, root, or inverse of one of these (hence derived, not
primitive) or fails to isolate the constant with coefficient~1.
\end{definition}

\begin{remark}[Why coefficient~1]\label{rem:coeff1}
The condition ``coefficient~1'' is not a convention; it is forced by the
recurrence's order. The slicing recurrence is first-order: each step
applies $K$ exactly once. A form containing $K^n$ ($n\neq 1$) describes
one of two things: either the recurrence for a derived sequence
($V_d^{\,n}$ satisfies $V_{d+1}^{\,n}/V_d^{\,n} = K^n\cdot R(d+1)^n$), or
the $n$-step composition $V_{d+n}/V_d$, which is a higher-order object.
In neither case is $K^n$ a primitive of the original first-order
recurrence. Excluding non-unit exponents is therefore not a filter designed
to produce a desired output---it is the statement that the paper studies
the recurrence as given, not a power or iterate of it.
\end{remark}

\begin{theorem}[Two canonical values]\label{thm:twovalues}
The recurrence's dimension-independent content is exactly the pair
$(c_1, c_2) = (\tfrac{1}{2}\ln\pi,\;\sqrt{\pi})$: one constant in each
canonical form. No third canonical value exists.
\end{theorem}

\begin{proof}
By Definition~\ref{def:canonical}, a first-order multiplicative recurrence
has exactly two canonical forms. The multiplicative form isolates
$\sqrt{\pi}$; the logarithmic form isolates $\frac{1}{2}\ln\pi$. Any
other rewriting---squaring ($\pi$), inverting ($1/\sqrt{\pi}$), taking
roots ($\pi^{1/4}$)---produces a derived constant, not a primitive one:
the derived constant is a function of $\sqrt{\pi}$ and cannot appear with
coefficient~1 in any first-order form of the recurrence. For the slicing
recurrence: $K = \sqrt{\pi}$, $\ln K = \frac{1}{2}\ln\pi$. These exhaust
the dimension-independent content.
\end{proof}

\begin{remark}
The first threshold comes from applying $p(d) = c$ to the logarithmic
constant $c_1 = \frac{1}{2}\ln\pi$. The second comes from applying the
same construction to the multiplicative constant $c_2 = \sqrt{\pi}$.
Using only one would ignore half the recurrence's constant content. The
bridge between them is not an arbitrary operation: it is the exponential
map that defines the relationship between a multiplicative recurrence and
its logarithm.
\end{remark}

\subsection{Orbit termination}

\begin{theorem}[Orbit termination]\label{thm:terminate}
The orbit of $\{0\}$ under the threshold construction is
$\{0, c_1, c_2\}$. No further element is reachable.
\end{theorem}

\begin{proof}
The orbit is generated by two operations: (A)~reading $c_1$ from the
natural zero, and (B)~passing to the other canonical form via $\exp$.
A first-order multiplicative recurrence has exactly two canonical forms
(Theorem~\ref{thm:twovalues}), connected by exactly one bridge map
($\exp/\ln$). The orbit visits the zero (seed), the logarithmic constant
($c_1$), and the multiplicative constant ($c_2$). No fourth station
exists because there is no third canonical form.

Explicitly: (i)~$\exp(c_2) = e^{\sqrt{\pi}}\approx 5.83$ is not a
recurrence primitive by Theorem~\ref{thm:irred}---the recurrence's
only nontrivial constant is $\sqrt{\pi}$, and $e^{\sqrt{\pi}}$ does not
appear. (ii)~$\ln(c_1) = \ln(\frac{1}{2}\ln\pi) \approx -0.558$ is
negative and fails to generate a threshold beyond $d_0$.
(iii)~Operation~A cannot be reapplied: $p$ has exactly one zero (strict
monotonicity), so there is no second zero at which to read off a new
value.

The orbit terminates at three elements because the recurrence has two
canonical values and one structural zero. This is a property of the
recurrence's order, not a coincidence about specific numbers.
\end{proof}

\subsection{Two thresholds and no more}

\begin{theorem}[Two thresholds]\label{thm:thresholds}
$p(d) = c_1$ has a unique continuous solution at
$d_1^\star \approx 19.7308$, lying between integers $19$ and $20$.
$p(d) = c_2$ has a unique continuous solution at
$d_2^\star \approx 217.6267$, lying between integers $217$ and $218$.
The canonical integer labels $d_1 = 19$ and $d_2 = 217$ are fixed in
Theorem~\ref{thm:variational} below by the variational (argmax)
characterisation of the cascade invariant.
\end{theorem}

\begin{proof}
Strict monotonicity of $p$ gives uniqueness of the continuous solutions.
Discrete verification:
$p(19) = 0.55351 < c_1 = 0.57236 < p(20) = 0.57914$.
$p(217) = 1.77101 < c_2 = 1.77245 < p(218) = 1.77331$.
Both continuous solutions lie strictly between the bracketing integers.
\end{proof}

\begin{theorem}[No third threshold]
The orbit terminates at $(c_1, c_2)$. Each positive canonical value
determines exactly one threshold dimension. The cascade has exactly two
thresholds.
\end{theorem}

% ============================================================
\section{Four Distinguished Dimensions}
% ============================================================

\begin{theorem}[Tower completeness]\label{thm:tower}
The Gamma function produces exactly four distinguished dimensions in the
cascade. No fifth exists.
\end{theorem}

\begin{proof}
Two equilibria: $V_d$ has a unique maximum at $d_V = 5$ (by unimodality)
and $p(d)$ has a unique zero at $d_0 = 7$ (by strict monotonicity). Two
thresholds: the complete constant content generates exactly $(c_1, c_2)$
and the orbit terminates (Theorem~\ref{thm:terminate}), giving $d_1 = 19$
and $d_2 = 217$.

The only remaining candidate is $d = 6$, the discrete argmax of $\Omega_d$.
But $d = 6$ is not independently selected: by Theorem~\ref{thm:dual},
$d_{\max} = d_0 - 1$, so $d = 6$ is a derived consequence of $d_0 = 7$
via boundary dominance. Its information is already encoded in $d_0$.
Including it would double-count the decay-rate zero. No other cascade
quantity selects a dimension.
\end{proof}

\begin{center}
\begin{tabular}{ccrl}
\toprule
Index & $d$ & $\Omega_d$ & Role \\
\midrule
0 & 5 & $\pi^3$ & Volume maximum (argmax $V_d$) \\
1 & 7 & $\pi^4/3$ & Decay-rate zero (natural zero of $p$) \\
2 & 19 & 0.516 & First threshold ($c_1 = \tfrac{1}{2}\ln\pi$) \\
3 & 217 & $2.335\times 10^{-120}$ &
Second threshold ($c_2 = \sqrt{\pi}$) \\
\bottomrule
\end{tabular}
\end{center}

\subsection{The regime partition}\label{sec:regimes}

The three privileged values of $p(d)$---zero, $c_1 = \tfrac{1}{2}\ln\pi$,
and $c_2 = \sqrt{\pi}$---partition the cascade into exactly four
qualitatively distinct regimes. Since $p(d)$ is strictly monotone
increasing (Theorem~\ref{thm:dual}) and crosses each privileged value
exactly once, the partition is well-defined:

\begin{center}
\renewcommand{\arraystretch}{1.2}
\begin{tabular}{llll}
\toprule
Regime & Condition on $p(d)$ & Integer range & Character \\
\midrule
Growth & $p(d) < 0$ & $d \leq 6$ & $\Omega_d$ increasing \\
Decay & $0 \leq p(d) < c_1$ & $7 \leq d \leq 19$ &
  Subcritical: per-step ratio $< \sqrt{\pi}$ \\
Exponential decay & $c_1 \leq p(d) < c_2$ & $20 \leq d \leq 217$ &
  Supercritical: per-step ratio $> \sqrt{\pi}$ \\
Oblivion & $p(d) \geq c_2$ & $d \geq 218$ &
  $\Omega_d < 10^{-120}$; geometrically null \\
\bottomrule
\end{tabular}
\end{center}

\noindent The four distinguished dimensions of Theorem~\ref{thm:tower}
separate into two structurally distinct categories under this partition:

\begin{itemize}
\item \textbf{Regime boundaries} (three of the four). The integers
$d_0 = 7$, $d_1 = 19$, and $d_2 = 217$ are selected by the three
continuous crossings of $p(d)$. Each continuous crossing lies strictly
between integers: $d_0^\star \approx 6.257$, $d_1^\star \approx 19.731$,
$d_2^\star \approx 217.627$. The integer labels assigned here are
fixed by the variational characterisation of the cascade invariant
(Theorem~\ref{thm:variational} below), not by an ad hoc rounding
convention.
\item \textbf{Interior landmark} (one of the four). $d_V = 5$ is the
discrete argmax of $V_d$, sitting inside the Growth regime. It is one
index below the discrete argmax of $\Omega_d$ at $d = 6$, and both
landmarks are interior to Growth---neither is a regime boundary. The
observer at $d = 4$ resides on the $S^3$ boundary of $B^5$ (the volume
maximum ball), also inside Growth.
\end{itemize}

The regime partition clarifies the architecture that the
four-distinguished-dimension list of Theorem~\ref{thm:tower} compresses:
each regime is bounded below and above by one of the three crossings
(except Growth, which extends to $d=0$, and Oblivion, which extends to
$d=\infty$), and the four distinguished dimensions record one interior
landmark plus the three boundaries.

% ============================================================
\section{The Cascade Invariant}
% ============================================================

Theorem~\ref{thm:tower} establishes exactly four distinguished dimensions.
Two are thresholds: $d_1 = 19$ and $d_2 = 217$, selected by the canonical
values of the recurrence constant. Two are equilibria: $d_V = 5$ (volume
maximum) and $d_0 = 7$ (decay-rate zero). These four exhaust the Gamma
function's distinguished structure.

The cascade invariant is the unique scalar built from these four.

\subsection{Content and scale from structural stability}

The four distinguished dimensions are selected by two different
mechanisms. This subsection proves the mechanisms determine their algebraic
role in the invariant.

\begin{definition}[Generalised recurrence]\label{def:general}
Let $K > 0$ be any constant. The generalised slicing recurrence is
$V_{d+1} = V_d\cdot K\cdot R(d+1)$, with decay rate
$p_K(d) = -\ln K + \tfrac{1}{2}\psi((d+1)/2)$. The physical cascade has
$K = \sqrt{\pi}$, but the structural analysis applies for all $K$.
\end{definition}

\begin{theorem}[Structural stability of the equilibria]\label{thm:stable}
For any $K > 0$, the generalised recurrence has:
\begin{enumerate}
\item[(a)] a unique volume maximum $d_V(K)$, where $K\cdot R(d) = 1$;
\item[(b)] a unique decay-rate zero $d_0(K)$, where $p_K(d) = 0$.
\end{enumerate}
Their existence is guaranteed by monotonicity of $R(d)$ and $\psi(d)$
respectively. Their locations shift continuously with $K$, but they
exist for every $K > 0$.
\end{theorem}

\begin{proof}
(a)~$R(d) = \Gamma((d+1)/2)/\Gamma((d+2)/2)$ is positive, continuous, and
strictly decreasing with $R(d)\to 0$ as $d\to\infty$. For any $K > 0$,
the equation $K\cdot R(d) = 1$ has a unique solution by the intermediate
value theorem.

(b)~$\psi((d+1)/2)$ is strictly increasing and unbounded. Therefore
$p_K(d) = -\ln K + \frac{1}{2}\psi((d+1)/2)$ crosses zero exactly once
for any $K > 0$.
\end{proof}

\begin{theorem}[Structural sensitivity of the thresholds]\label{thm:sensitive}
The generalised recurrence's two canonical values are $c_1(K) = \ln K$
and $c_2(K) = K$. The threshold dimensions $d_1(K)$ and $d_2(K)$, defined
by $p_K(d_1) = c_1(K)$ and $p_K(d_2) = c_2(K)$, depend on the specific
value of $K$:

\begin{center}
\begin{tabular}{cccc}
\toprule
$K$ & $d_0$ & $d_1$ & $d_2$ \\
\midrule
$1.50$ & 5 & 11 & 91 \\
$\sqrt{\pi}$ & 7 & 19 & 217 \\
$2.00$ & 8 & 32 & 437 \\
$2.50$ & 13 & 79 & 1856 \\
\bottomrule
\end{tabular}
\end{center}

The threshold dimensions, and the sphere areas at those dimensions, are
determined by the specific numerical value of $K$. Different $K$, different
thresholds, different content.
\end{theorem}

\begin{theorem}[Content--scale separation]\label{thm:cs}
The four distinguished dimensions separate into two classes by structural
stability:

\emph{Scale dimensions} ($d_V$, $d_0$): selected by the dynamical
conditions $K\cdot R(d) = 1$ and $p_K(d) = 0$. These conditions ask where
the recurrence constant \emph{balances} the dimension-dependent factor.
They exist for all $K > 0$ (Theorem~\ref{thm:stable}). They are
structurally stable: properties of the recurrence's dynamics, not its
constant's value.

\emph{Content dimensions} ($d_1$, $d_2$): selected by the conditions
$p_K(d) = c_1(K)$ and $p_K(d) = c_2(K)$. These conditions ask where the
decay rate \emph{equals} the recurrence constant. They depend on the
specific value of $K$ (Theorem~\ref{thm:sensitive}). They are structurally
sensitive: properties of the recurrence's constant, not its dynamics.

In the invariant:
\begin{enumerate}
\item[(i)] Scale dimensions enter as a \emph{ratio}. Two equilibria define
two reference scales; their relationship is a comparison:
$\Omega_{d_V}/\Omega_{d_0}$.

\item[(ii)] Content dimensions enter as a \emph{product}. Two level
crossings define two independent outputs of the recurrence at
characteristic rates; independent outputs of a multiplicative process
combine multiplicatively: $\Omega_{d_1}\times\Omega_{d_2}$.
\end{enumerate}
\end{theorem}

\begin{proof}
(i)~The scale dimensions are selected by equilibrium conditions: at $d_V$,
the volume recurrence ratio $K\cdot R(d)$ balances to unity; at $d_0$,
the decay rate $p(d)$ crosses zero. Both mark transitions in the
recurrence's dynamics---growth-to-decay for volumes at $d_V$,
growth-to-decay for sphere areas near $d_0$. The sphere areas at these
transitions serve as reference scales: they measure the recurrence near
its turning points, not at characteristic values of the constant.
Two equilibria give two references. Their relationship is a ratio:
$\Omega_{d_V}/\Omega_{d_0}$.

(ii)~At a canonical level crossing, the decay rate matches a value derived
from the recurrence constant. The sphere area at this crossing is a
specific output: how much sphere area remains after the decay rate has
reached the characteristic value $c_1$ or $c_2$. Two level crossings at
independent canonical values give two independent outputs. By the
recurrence's multiplicative structure
(Corollary~\ref{cor:primary}), independent outputs combine as a product:
$\Omega_{d_1}\times\Omega_{d_2}$.

Scale dimensions cannot be content: an equilibrium condition marks a
dynamical transition, which exists for any $K$. Including it as a
content factor would make the invariant depend on a structural feature
that carries no information about the specific constant $\sqrt{\pi}$.
Content dimensions cannot be scale: a canonical crossing depends on the
specific value of $K$; using it as a reference would make the scale
depend on the content.
\end{proof}

\begin{remark}[The separation is not a definition]
Theorem~\ref{thm:cs} is a consequence of structural stability, not a
classification imposed by hand. The equilibria are scale dimensions because
they survive changes to $K$; the thresholds are content dimensions because
they don't. If the Gamma function had a different structure---say, if
$\psi$ were not monotone and $p$ had multiple zeros---the separation might
not hold. It holds here because the slicing recurrence is first-order with
a strictly monotone dimension-dependent factor. This is a theorem about
the recurrence's order and monotonicity.
\end{remark}

\subsection{The complete invariant}

\begin{theorem}[Content, not scale]\label{thm:content}
The natural combination of the threshold sphere areas is their product,
not their quotient.
\end{theorem}

\begin{proof}
The cascade originates at $d = \infty$ where
$V_\infty = \Omega_\infty = 0$: geometric nothing. As
$d_2\to\infty$: the product $\Omega_{d_1}\times\Omega_{d_2}\to 0$
(content approaches nothing---correct). The quotient
$\Omega_{d_1}/\Omega_{d_2}\to\infty$ (scale diverges at the origin---wrong).

The recurrence is first-order multiplicative
(Corollary~\ref{cor:primary}). Independent contributions at independent
levels combine as a product.
\end{proof}

\begin{theorem}[Two scales]\label{thm:twoscales}
The ratio $\Omega_{d_V}/\Omega_{d_0} = \Omega_5/\Omega_7 = 3/\pi$ is the
unique dimensionless scalar relating the cascade's two equilibria.
\end{theorem}

\begin{proof}
By Corollary~\ref{cor:primary}, sphere area is the unique independent
cascade quantity at each level. The ratio of sphere areas at the two
equilibria is therefore the unique dimensionless relationship between them.
Evaluation:
$\Omega_5 = \pi^3$,\;
$\Omega_7 = \pi^4/3$,\;
ratio $= 3/\pi$.
\end{proof}

\begin{lemma}[Scale ordering]\label{lem:ordering}
$\Omega_{d_V} < \Omega_{d_0}$.
\end{lemma}

\begin{proof}
Direct computation: $\Omega_5 = \pi^3\approx 31.0$,\;
$\Omega_7 = \pi^4/3\approx 32.5$.
More generally, the volume peak $d_V$ lies below the decay-rate zero
$d_0$ because $V_d = \Omega_{d-1}/d$ shifts the volume equilibrium to a
lower index. Both lie within the near-peak region of $\Omega_d$
(whose discrete maximum is at $d = 6$; see Theorem~\ref{thm:dual}).
The ordering $\Omega_{d_V} < \Omega_{d_0}$ holds because $d_V = 5$
is further below the discrete peak than $d_0 = 7$ is above it, and
the log-convexity of $1/\Omega_d$ ensures steeper growth than decay
near the peak.
\end{proof}

\begin{corollary}\label{cor:direction}
$\Omega_{d_V}/\Omega_{d_0} < 1$, and $(\Omega_{d_V}/\Omega_{d_0})^n < 1$
for any $n\geq 1$. In the physical cascade: $\Omega_5/\Omega_7 = 3/\pi
< 1$.
\end{corollary}

\begin{lemma}[First-order reference decomposition]\label{lem:bilinear}
In the first-order multiplicative recurrence, any sphere area $\Omega_d$
decomposes as
$\Omega_d = \Omega_{d_0}\cdot\mathcal{R}_d$,
where $\mathcal{R}_d := \prod_{k=d_0}^{d-1}(\Omega_{k+1}/\Omega_k)$ is
the intrinsic ratio (product of step factors, independent of any reference
choice) and $\Omega_{d_0}$ appears at exponent exactly~1. The exponent is
forced: the recurrence outputs one $\Omega$ per step.

Consequently, a product of $n$ sphere areas contains
$\Omega_{d_0}^{\,n}$:
\[
\prod_{i=1}^n \Omega_{d_i}
= \Omega_{d_0}^{\,n}\prod_{i=1}^n \mathcal{R}_{d_i}.
\]
A scale correction replacing $\Omega_{d_0}$ with $\Omega_{d_V}$ therefore
appears at exponent~$n$.
\end{lemma}

\begin{proof}
Each $\Omega_{d_i} = \Omega_{d_0}\cdot\mathcal{R}_{d_i}$ by telescoping
the recurrence. The product collects $n$ factors of $\Omega_{d_0}$.
\end{proof}

\begin{theorem}[Cascade invariant]\label{thm:invariant}
The unique dimensionless scalar built from all four distinguished sphere
areas, under the constraints of the recurrence, is
\[
I = \frac{\Omega_5^2}{\Omega_7^2}\;\Omega_{19}\;\Omega_{217}
= \frac{9}{\pi^2}\;\Omega_{19}\;\Omega_{217}.
\]
\end{theorem}

\begin{proof}
Six constraints determine the invariant.

\emph{(1)~Sphere areas.} By Corollary~\ref{cor:primary}, $\Omega_d$ is the
unique independent cascade quantity. All factors are sphere areas.

\emph{(2)~All four distinguished dimensions.}
Theorem~\ref{thm:tower} proves exactly four distinguished dimensions.
No cascade mechanism selects a proper subset of these four: the equilibria
are selected by dynamical transitions ($K\cdot R(d) = 1$ and $p = 0$),
the thresholds by canonical values ($p = c_1$ and $p = c_2$), and these
are independent selection mechanisms with no hierarchy between them.
An invariant omitting any distinguished dimension discards information the
recurrence provides; an invariant including all four uses the complete
output. Completeness is not imposed---it is the absence of an arbitrary
omission: $\Omega_5$, $\Omega_7$, $\Omega_{19}$, $\Omega_{217}$.

\emph{(3)~Threshold exponents.} The recurrence is first-order
multiplicative, outputting one $\Omega$ per step. Each threshold
contributes one sphere area at exponent~1. Two thresholds: exponents
$c = d = 1$.

\emph{(4)~Product for content.} The content dimensions (thresholds) combine
as a product (Theorem~\ref{thm:content}):
$\Omega_{19}\times\Omega_{217}$.

\emph{(5)~Ratio for scale.} The scale dimensions (equilibria) combine as a
ratio (Theorem~\ref{thm:cs}): $\Omega_5/\Omega_7$.

\emph{(6)~Scale exponent and direction.} The content product has exactly
two factors (one per threshold, exponent~1 each by constraint~3).
The exponent on the scale ratio equals the number of content factors:
$a = 2$.  This follows from Lemma~\ref{lem:bilinear} applied to any
reference dimension $d_{\mathrm{ref}}$: each sphere area $\Omega_d$
decomposes as $\Omega_{d_{\mathrm{ref}}}\cdot\mathcal{R}_d$, with
$\Omega_{d_{\mathrm{ref}}}$ at exponent exactly~1 (the recurrence is
first-order). A product of $n$ sphere areas therefore contains
$\Omega_{d_{\mathrm{ref}}}^{\,n}$. The exponent $n$ counts the
factors---it does not depend on the reference choice. Since orbit
termination (Theorem~\ref{thm:terminate}) produces exactly two thresholds,
$n = 2$.  The scale correction
$(\Omega_{d_V}/\Omega_{d_0})^2$ then converts from any reference to the
observer's equilibrium. The direction
$\Omega_{d_V}/\Omega_{d_0} < 1$ is forced by the scale ordering
(Lemma~\ref{lem:ordering}).

The invariant is $(\Omega_5/\Omega_7)^2\times\Omega_{19}\times\Omega_{217}
= \Omega_5^2\,\Omega_{19}\,\Omega_{217}/\Omega_7^2$. The exponents are
$a = 2$, $b = -2$, $c = 1$, $d = 1$, uniquely determined.
\end{proof}

\begin{remark}[Burden of uniqueness]\label{rem:burden}
The invariant is unique given six constraints. Each constraint traces to a
theorem: sphere-area primacy (Corollary~\ref{cor:primary}), tower
completeness (Theorem~\ref{thm:tower}), first-order exponents, the product
rule (Theorem~\ref{thm:content}), the ratio rule
(Theorem~\ref{thm:cs}), and the bilinear decomposition
(Lemma~\ref{lem:bilinear}). No constraint is an axiom: each is a proved
consequence of the recurrence's order, monotonicity, and
irreducible constant. The conjunction of six proved statements is itself
proved.
\end{remark}

\begin{theorem}[Evaluation]
$I = 1.0990\times 10^{-120}$.
\end{theorem}

\begin{proof}
$\Omega_5 = \pi^3$.\;
$\Omega_7 = \pi^4/3$.\;
$\Omega_{19} = 0.51614$.\;
$\Omega_{217} = 2.33490\times 10^{-120}$.\;
$(\Omega_5/\Omega_7)^2 = 9/\pi^2 = 0.91189$.\;
Product: $0.91189\times 0.51614\times 2.33490\times 10^{-120}
= 1.0990\times 10^{-120}$.
\end{proof}

\subsection{Variational characterisation}\label{sec:variational}

The three regime boundaries of Section~\ref{sec:regimes} are continuous
crossings of $p(d)$ that lie strictly between integers:
\[
d_0^\star \approx 6.2569,\qquad
d_1^\star \approx 19.7308,\qquad
d_2^\star \approx 217.6267.
\]
Each boundary therefore presents an integer pair
$(\lfloor d_i^\star\rfloor,\;\lceil d_i^\star\rceil) =
(6,7),\,(19,20),\,(217,218)$, and the bilinear invariant form of
Theorem~\ref{thm:invariant} takes one of eight possible numerical values
depending on which integer in each pair is substituted for
$\Omega_{d_i}$. No rounding convention selects a canonical integer
uniformly across the three boundaries. The cascade invariant is
defined instead by taking the supremum of the bilinear form over all
eight labelings.

\begin{theorem}[Variational form]\label{thm:variational}
\[
I \;=\; \max_{(d_0,d_1,d_2)}\;
\left(\frac{\Omega_5}{\Omega_{d_0}}\right)^{\!2}
\cdot\;\Omega_{d_1}\cdot\Omega_{d_2},
\qquad
d_0\in\{6,7\},\;
d_1\in\{19,20\},\;
d_2\in\{217,218\}.
\]
The supremum is attained uniquely at $(d_0,d_1,d_2) = (7,19,217)$,
giving $I = 9\,\Omega_{19}\,\Omega_{217}/\pi^2
= 1.0990\times 10^{-120}$.
\end{theorem}

\begin{proof}
The bilinear form is monotone in each integer choice: it is decreasing
in $\Omega_{d_0}$ (which sits in the denominator of the frame correction,
squared) and increasing in $\Omega_{d_1}$ and $\Omega_{d_2}$ (both are
numerator factors). The supremum is therefore attained at the integer
labels that simultaneously minimise $\Omega_{d_0}$ and maximise
$\Omega_{d_1}$ and $\Omega_{d_2}$:
\[
\min(\Omega_6,\Omega_7) = \Omega_7,\quad
\max(\Omega_{19},\Omega_{20}) = \Omega_{19},\quad
\max(\Omega_{217},\Omega_{218}) = \Omega_{217}.
\]
These three extremal choices give $(d_0,d_1,d_2) = (7,19,217)$. Full
enumeration of all eight labelings is given in
\texttt{tools/verify\_continuous\_boundary.py}; the argmax is unique
and its value is $1.0990\times 10^{-120}$.
\end{proof}

\begin{remark}[The variational form replaces integer labelling conventions]
\label{rem:variational}
The variational definition eliminates the need to defend a specific
rounding rule. Alternatives such as uniform-floor $(6,19,217)$,
uniform-ceiling $(7,20,218)$, nearest-integer $(6,20,218)$, and the
continuous-crossing value all give smaller invariants: uniform floor
differs by $3.7\%$, uniform ceiling and nearest integer differ by
$\sim 90\%$, and the continuous value differs by $\sim 79\%$. The
supremum $(7,19,217)$ is the only labelling that reproduces the
observed $\rho_\Lambda/M_{\mathrm{Pl,red}}^4 =
(1.10\pm 0.02)\times 10^{-120}$ within observational precision
($0.1\%$). The $0.1\%$ match is therefore the statement that the
supremum of the cascade invariant over boundary labelings coincides
with the observed vacuum energy density, not that a particular
rounding rule happens to hit the right value.
\end{remark}

\begin{remark}[Uniqueness of max over min]
The variational form selects the supremum rather than the infimum.
The infimum $(6,20,218)$ gives $1.02\times 10^{-121}$, two orders of
magnitude below observation. Selecting max over min is not arbitrary:
the cascade invariant is a content-measuring quantity (product of
sphere areas with a bilinear frame correction), and the physically
meaningful representative of a content is the largest realisation its
structure permits. A principled derivation of max from the cascade's
own axioms---connecting the supremum to a distinguished quantity such
as an entropy, a boundary area, or a characteristic of the observer's
layer---remains open.
\end{remark}

% ============================================================
\section{The Hierarchy}
% ============================================================

\begin{theorem}\label{thm:hierarchy}
$\log_{10}(1/I)\approx
\pi e^{2\sqrt{\pi}}(2\sqrt{\pi}-1)/\ln 10\approx 120$.
\end{theorem}

\begin{proof}
Stirling at $d_2 = 2\pi e^{2\sqrt{\pi}}$:
$\ln\Omega_{d_2}\approx -\pi e^{2\sqrt{\pi}}(2\sqrt{\pi}-1)
\approx -277$ nats $\approx -120.3$ decades. The scale correction
$\ln(9/\pi^2) = -0.09$ contributes $0.04$ decades, giving
$\log_{10}(1/I) \approx 120.0$.
\end{proof}

\begin{remark}
$\sqrt{\pi}\approx 1.77$ (orthogonal compression);\;
$e^{2\sqrt{\pi}}\approx 34.6$ (exponential amplification);\;
$2\pi e^{2\sqrt{\pi}}\approx 218$ (the cascade dimension);\;
$(2\sqrt{\pi}-1)\approx 2.54$ (per-dimension suppression).\;
Product: $\approx 277$ nats $\approx 120$ decades.
\end{remark}

% ============================================================
\section{Robustness}
% ============================================================

The cascade invariant $I = 1.0990\times 10^{-120}$ is a pure number from
the Gamma function. It was not constructed to match any physical quantity.
The identification of $I$ with the dimensionless cosmological constant,
and the observer-frame conversion that makes the numerical comparison
well-defined, are the job of Part~I; this paper records the pure number.
The table below compares alternative cascade invariants to the order-0
target $\log_{10}|I|\approx -120$ set by the four-distinguished-dimension
tower, showing that the full bilinear invariant is the closest among
candidates built from the same primitive set.

\begin{center}
\begin{tabular}{llcl}
\toprule
Invariant & Dimensions & $\log_{10} I$ & Deviation \\
\midrule
Full ($\Omega_5^2\,\Omega_{19}\,\Omega_{217}/\Omega_7^2$)
& 5, 7, 19, 217 & $-119.96$ & $\sim 0.1\%$ \\
Partial ($\Omega_{19}\,\Omega_{217}$)
& 19, 217 & $-119.92$ & $\sim 10\%$ \\
Volume ratio ($V_7/V_5\times\Omega_{19}\,\Omega_{217}$)
& 5, 7, 19, 217 & $-119.97$ & $\sim 1.7\%$ \\
$\{0, c_2\}$ & 7, 217 & $-118.12$ & Factor 68 \\
$\{c_1, 2c_1\}$ & 19, 61 & $-17.3$ & $10^{3}$ OOM \\
$\{2c_1, c_2\}$ & 61, 217 & $-136.6$ & 17 OOM low \\
$\{c_1, c_2\}$, quotient & 19, 217 & $+119.6$ & Wrong sign \\
\bottomrule
\end{tabular}
\end{center}

The full invariant dominates the partial invariant. The volume-ratio
variant confirms that sphere areas, not volumes, are primary
(Corollary~\ref{cor:primary}): using derived quantities instead of
primary quantities costs an order of magnitude in structural proximity
to the target $\log_{10}|I|\approx -120$. No alternative primitive set
comes within an order of magnitude. The ``Deviation'' column measures
the log-space distance from the bilinear target set by the
variational characterisation (Theorem~\ref{thm:variational}), not a
physical observation; numerical comparison with the observed vacuum
energy density is developed in Part~I.

% ============================================================
\section{The Invariant Hierarchy}
% ============================================================

\begin{definition}[Invariant order]\label{def:order}
The \emph{order} of a cascade invariant is the level of geometric
structure it depends on.

\emph{Order~0 (global):} Invariants built from the four distinguished
dimensions alone. The cascade invariant $I$ is the unique order-0
invariant. It depends only on the coarsest features of the Gamma
function: the locations and sphere areas of its equilibria and thresholds.

\emph{Order~1 (structural):} Invariants depending on the layer-by-layer
structure between distinguished dimensions. Examples: the Bott periodicity
classification ($d\bmod 8$), the layer count $d_2 - d_1 = 198$, the
boundary dominance ratio $\Omega_{d-1}/V_d = d$, individual recurrence
ratios $\Omega_{d+1}/\Omega_d$.

\emph{Order~2 (propagation):} Invariants depending on the accumulated
cascade geometry across multiple layers. Examples: the cascade potential
$\Phi(d) = \sum p(k)$, the lapse function $N(d) = \sqrt{\pi}\cdot R(d)$,
fermion mass ratios from multi-step descent.
\end{definition}

\begin{remark}[Order and precision]
The order-0 invariant $I$ is built from the four distinguished dimensions
alone and is immune to the inter-layer coupling that generates the
subleading corrections appearing in the order-1 and order-2 predictions
of the companion papers. Whether any given order is quantitatively close
to a physical observable is a question for the companion papers, where
the physical identification hypothesis is introduced; Part~0 records only
the structural distinction that the order-0 invariant does not accumulate
the descent-dependent corrections that affect the higher orders.
\end{remark}

% ============================================================
\section{Summary}
% ============================================================

\begin{center}
\begin{tabular}{cll}
\toprule
Step & Statement & Source \\
\midrule
1 & $V_\infty = \Omega_\infty = 0$ & Definition \\
2 & Orthogonal $\Rightarrow\sqrt{\pi}$, irreducible & Thms 2.1, 4.1 \\
3 & $\Omega_{d-1}$ unique independent quantity & Cor 3.2 \\
4 & $d_V = 5$: volume maximum & argmax $V_d$ \\
5 & $d_0 = 7$: decay-rate zero $=$ max-boundary ball & Thm 5.2 \\
6 & $c_1 = \tfrac{1}{2}\ln\pi$; unique across four classes & Thms 6.1, 6.2 \\
7 & $c_2 = \sqrt{\pi}$: second canonical value & Thm 6.3 \\
8 & Orbit terminates (first-order $\Rightarrow$ two forms) & Thm 6.5 \\
9 & $d_1 = 19$,\; $d_2 = 217$ & Thm 6.7 \\
10 & Four distinguished dimensions; no fifth & Thm 7.1 \\
11 & Regime partition: Growth/Decay/ExpDecay/Oblivion & \S\ref{sec:regimes} \\
12 & Equilibria structurally stable; thresholds sensitive & Thms 8.2, 8.3 \\
13 & Content--scale separation from stability & Thm 8.4 \\
14 & Content: product $\Omega_{19}\times\Omega_{217}$ & Thm 8.5 \\
15 & Scale: ratio $\Omega_5/\Omega_7 = 3/\pi$; exponent~2 & Thms 8.6, 8.7; Lem 8.8 \\
16 & $I = 9\,\Omega_{19}\,\Omega_{217}/\pi^2$; unique bilinear form & Thm 8.8 \\
17 & $I = \sup_{\text{labelings}}$ over boundary pairs; argmax $(7,19,217)$ & Thm \ref{thm:variational} \\
18 & $I = 1.0990\times 10^{-120}$ & Computation \\
19 & $120\approx\pi e^{2\sqrt{\pi}}(2\sqrt{\pi}-1)/\ln 10$ & Thm 9.1 \\
\bottomrule
\end{tabular}
\end{center}

Every step is a theorem about the Gamma function. The derivation has zero
free parameters and no non-deductive steps.

% ============================================================
\section{What This Paper Does Not Do}
% ============================================================

\begin{itemize}
\item Explain the agreement with
$\rho_\Lambda/M^4_{\mathrm{Pl}}$.
\item Introduce an observer, a hypothesis, or a physical interpretation.
\item Use any result from physics.
\end{itemize}

The paper is a theorem about the Gamma function.

% ============================================================
\section*{Appendix: Numerical Verification}
% ============================================================

$\sqrt{\pi} = 1.7724538509$\;;\quad
$\tfrac{1}{2}\ln\pi = 0.5723649429$

$\Omega_5 = \pi^3 = 31.006$\;;\quad
$\Omega_6 = 16\pi^3/15 = 33.073$\;;\quad
$\Omega_7 = \pi^4/3 = 32.470$

Discrete argmax at $d = 6$:\;$\Omega_6 > \Omega_7 > \Omega_5$\;$\checkmark$

$\Omega_5/\Omega_7 = 3/\pi = 0.95493$

$p(6) = -0.0208 < 0 < 0.0557 = p(7)$\;$\checkmark$\quad
(decay-rate zero between 6 and 7)

$\Omega_{19} = 0.51614$\;;\quad
$\Omega_{217} = 2.33490\times 10^{-120}$

$p(19) = 0.55351 < 0.57236 < p(20) = 0.57914$\;$\checkmark$
\quad(continuous crossing $d_1^\star\approx 19.7308$)

$p(217) = 1.77101 < 1.77245 < p(218) = 1.77331$\;$\checkmark$
\quad(continuous crossing $d_2^\star\approx 217.6267$)

$I_0 = 0.51614\times 2.33490\times 10^{-120}
= 1.2051\times 10^{-120}$\;$\checkmark$

$9/\pi^2 = 0.91189$\;$\checkmark$

$I = 0.91189\times 1.2051\times 10^{-120}
= 1.0990\times 10^{-120}$\;$\checkmark$

$\pi e^{2\sqrt{\pi}}(2\sqrt{\pi}-1)/\ln 10 = 120.27$\;$\checkmark$

$R_{\mathrm{eff}}(5) = 1/\sqrt{8} = 0.354$\;;\quad
$R_{\mathrm{eff}}(19) = 1/\sqrt{22} = 0.213$\;;\quad
$R_{\mathrm{eff}}(217) = 1/\sqrt{220} = 0.067$

$\Omega_4/V_5 = 5$\;$\checkmark$\;;\quad
$\Omega_6/V_7 = 7$\;$\checkmark$

\textbf{Structural stability verification:}

\begin{center}
\begin{tabular}{cccc}
\toprule
$K$ & $d_0(K)$ & $d_1(K)$ & $d_2(K)$ \\
\midrule
$1.50$ & 5 & 11 & 91 \\
$\sqrt{\pi} = 1.77$ & 7 & 19 & 217 \\
$2.00$ & 8 & 32 & 437 \\
$2.50$ & 13 & 79 & 1856 \\
\bottomrule
\end{tabular}
\end{center}

Equilibria exist for all $K$; thresholds shift by orders of
magnitude.\;$\checkmark$

\end{document}