\documentclass[12pt,a4paper]{article}
\usepackage[margin=1in]{geometry}
\usepackage{amsmath,amssymb}
\usepackage{hyperref}
\usepackage{booktabs}
\usepackage{array}
\usepackage{graphicx}

\title{\textbf{The Cascade Series}}
\author{RTAC}
\date{March 2026}

\begin{document}
\setlength{\emergencystretch}{3em}
\maketitle
\thispagestyle{empty}

\section*{The Thought Experiment}

A black hole has a horizon. An infalling observer sees infinite time dilation there.
But the black hole evaporates in finite time. These two facts are in tension.

The resolution is asymptotic: there is no sharp horizon. Infalling matter never
crosses --- from outside, it asymptotically approaches the horizon; from the infaller's
perspective, the black hole evaporates and the horizon recedes. Both observers agree:
the infaller asymptotically tracks a shrinking shell where the rate of infall balances
the rate of evaporation.

This shell is the physical content of the ``horizon.'' On it sits a two-dimensional
spatial surface ($S^2$ for a Schwarzschild hole), and the time direction is the asymptotic
balance itself --- the direction along which infall and shrinkage compete. What if
this shell is a $2 + 1$ dimensional universe?

If a 4D black hole produces a $2 + 1$ shell universe, does a 5D black hole produce
a $3 + 1$ shell universe? Is that us? Are we living on the $S^3$ horizon of a 5D black
hole? If so, time is the forced direction --- the compactification that the cascade
must undergo, not a coordinate we choose. Moving orthogonally, along the shell,
slows progress in this forced direction but can never stop it completely. This is time
dilation, from the geometry of a shell.

If yes, then the 5D black hole sits inside a 6D space --- another shell of a 7D object.
Follow the tower upward: each shell is the horizon of the next. The tower terminates
at $d = \infty$: the infinite-dimensional unit ball, which has zero volume, zero surface
area, no interior. Geometric nothing. A natural starting point.

That was the end of the intuition. The question was: what happens if you work
back down from infinity?

\section*{One Infinity, Not Many}

The thought experiment begins with an infinity---the infinite time dilation at a
black hole horizon---and resolves it asymptotically: the infaller never crosses,
because the horizon recedes as the hole evaporates. What looked like a horizon is
really a shrinking shell the infaller asymptotically tracks; it is the
\emph{infall} that takes forever to complete, not the evaporation. The evaporation
is finite. Following the tower upward, the thought experiment then arrives at
another infinity: $B^\infty$, the starting point. These are not two infinities.
They are the same infinity, viewed from two different positions along its descent.

This is the unifying observation of the series. Every ``infinity problem'' in
physics is a local slice of the same asymptotic boundary. The cosmological
constant is $\sim 10^{-120}$ because it is the cascade's natural floor after
descending from $d=\infty$. The Standard Model mass hierarchy is the descent rate
$\exp(-\Phi)$ measured at different cascade depths. Black hole horizons are where
the cascade's descent locally approaches its asymptotic limit---the same limit
the rest of the universe is descending toward, just at a steeper local rate.
Ultraviolet divergences hit a natural floor at $\Omega_{217}$. The Big Bang
is the cascade's resolution crossing that floor in the universe's past;
the asymptotic upper edge $B^\infty$ remains the unreached future of
resolution, approached but never completed (as black hole horizons are
unreachably approached going forward).

Physics has treated these as separate problems for a century and made local
fixes to each. The cascade says they are one problem with one resolution: there
was only ever one infinity, and it is the infinite-dimensional starting object
of the series. Every finite structure we observe---spacetime, matter, forces,
couplings, masses---is what that one infinity looks like while it is partway
through resolving itself. The universe is not a place where infinity was avoided.
It is the unique asymptotic completion of the infinity that had to exist,
observed from a particular finite depth.

\textbf{Time is the rate at which the resolution proceeds.} Each Planck tick
corresponds to one more dimension of $B^\infty$ resolved into a definite
cascade layer; the truncation height $N(t)$ above the observer grows by one
per tick (Part~VI). Time is not a coordinate the cascade lives in but a
counter of resolved dimensions: the universe is in-progress arithmetic
against an unbounded denominator. Three vocabularies describe the same rate,
each offering a different way to picture what is happening:
\begin{itemize}
\item \textbf{Mathematical} (Prelude, Part~0): the slicing recurrence
extracts one dimension's worth of finite content from $B^\infty$ at each
step; the integer $d$ is the resolution stage.
\item \textbf{Physical} (Parts~II and~VI): each step is one Planck tick of
cosmic time --- ``slicing $=$ dimensional collapse $=$ time'' --- and the
truncation height grows one layer per tick.
\item \textbf{Geometric} (this thought experiment): the observer's $S^3$
shell takes one asymptotic step toward the 5D black hole's pseudo-horizon
per tick. ``Time dilation'' is the local rate at which resolution proceeds
along the observer's worldline; ``the Big Bang'' is the resolution
reaching the $\Omega_{217}$ floor.
\end{itemize}
The three readings are dual descriptions of the same operation.

\section*{The Starting Point}

The thought experiment arrives at the infinite-dimensional unit ball from above,
by following a tower of black hole shells upward. But there is a deeper reason
to start there, established in the Prelude: \emph{there is nowhere else to start.}

A theory of everything cannot have inputs. Every input demands an explanation of
its origin, and that explanation either requires a deeper theory or is circular.
The only starting point that requires no explanation is nothing.

Nothing --- taken seriously as a mathematical object --- is not featureless.
The precondition for any formal system is the distinction between true and false:
$0\neq 1$. Distinction implies orthogonality (zero mutual information between
distinguishable states). Orthogonality iterates without bound (no logical
obstruction to countable independence). The absence of external scale forces unit
norm. The result is the infinite-dimensional unit ball: zero volume, zero
surface area, no interior. Structured nothing.

The unit ball is not chosen from a menu of starting points. It is the unique
mathematical structure that corresponds to ``nothing with the capacity for
distinction,'' derived from the precondition for coherent thought. The series
is the derivation of what that structure implies.

\section*{The Hypothesis}

\begin{center}
\textbf{The infinite-dimensional unit ball, descended to four dimensions, is indistinguishable from our universe.}
\end{center}

It is tested by deriving, from the cascade's geometry alone, the cosmological
constant, the dimension and signature of spacetime, quantum mechanics, and the
Standard Model gauge group and its symmetry breaking, three fermion generations,
fermion masses, and gauge coupling constants. It further derives the background
cosmological parameters --- including a Hubble constant between the two competing
measurements and an account of the DESI baryon acoustic oscillation observations
without dynamical dark energy. Every prediction is a test of the hypothesis.

\section*{The Physical Content}

If the hypothesis is correct, then every particle in the Standard Model is a
\emph{stable projection of higher-dimensional geometry into 4D spacetime}.

Specifically: the cascade's Bott periodicity anchors fermion generations at
fixed dimensions ($d = 5, 13, 21$); Adams' theorem anchors gauge bosons at
$d = 12, 13, 14$; the Gamma function's critical points anchor the cascade's
structural landmarks ($d_V = 5$, $d_0 = 7$, $d_1 = 19$, $d_2 = 217$). These
positions are topologically rigid---they cannot be continuously deformed.
What the observer at $d = 4$ perceives as ``a particle with mass $m$'' is the
projection amplitude of one of these features through the cascade's geometric
and topological attenuation channels: $\exp(-\Phi)$ for the decay of sphere
areas along the descent, and $(2\sqrt{\pi})^{-n_D}$ for each hairy ball
obstruction crossed. Heavier particles project from nearer layers; lighter
particles from deeper ones. The mass hierarchy \emph{is} the projection hierarchy.

The projection determines \emph{all} particle properties, not only mass.
\textbf{Spin} is the Clifford type at the particle's cascade layer: complex
spinor ($d\bmod 8=5$, from $O(d)$ Bott periodicity) for fermions, Adams vector fields for gauge
bosons, the hairy ball zero for the Higgs. \textbf{Charge} is determined by
which gauge layers the projection passes through: colour from $d=12$, weak
isospin from $d=13$, hypercharge from $d=14$. \textbf{Chirality} follows from
the $\mathbb{Z}_2$ decomposition at even-sphere layers ($\chi(S^{2n})=2$).
\textbf{Mixing angles} arise when the projection crosses multiple generation
or gauge layers.

The higher dimensions are not hidden containers of extra content. $B^\infty$
has zero volume, zero area, no interior. What the observer sees as particles,
forces, and coupling constants is the \emph{shape of the emptiness}---the
structured geometry of nothing, projected onto the observer's shell. Stability
is topological: the features cannot be removed because the theorems that
position them (Bott, Adams, Poincar\'{e}--Hopf, the Gamma function) are rigid.

\textbf{Particle content: what exists and what does not.} The cascade produces
exactly the Standard Model---three fermion generations, the gauge group
$\mathrm{SU}(3)\times\mathrm{SU}(2)\times\mathrm{U}(1)$, its breaking
pattern, and one Higgs doublet. The fourth Bott fermion layer at $d=29$,
suppressed to $\sim 0.5$~eV by the $d_1=19$ phase transition, sits at the
neutrino mass scale; whether this provides the neutrino mass mechanism is an
open question (Part~IVb, Open Question~5). The cascade predicts that
no supersymmetric partners exist (the topology has no pairing mechanism), no
extra gauge bosons (Adams' theorem is unique), no extra Higgs bosons (one
hairy ball zero), no axion ($\theta_{\mathrm{QCD}}=0$ topologically), and no
dark matter particles (the cascade's own geometry provides the missing
gravitational content), and no gravitons (the metric is a state property,
never promoted to an operator; Part~II\,=\,III, Section~6). Discovery of
any of these would falsify the framework.

The series that follows is the quantitative derivation of this picture.

\input{generated/predictions-table}

\bigskip

\section*{The Series}

\textbf{Prelude: Why Nothing Has Structure.} From $0\neq 1$ to the
dimension tower $\{(\mathbb{R}^d, B^d)\}_{d\in\mathbb{N}}$ of
finite-dimensional Euclidean spaces, with $B^\infty$ as the asymptotic
upper edge: distinction $\to$ orthogonality $\to$ countable dimension
tower $\to$ collapsed substrate at each $d$ under
$\mathbb{R}^*\times\mathrm{SO}(d)$. The cascade is the resolution of
the unresolved infinity at $B^\infty$ into finite intrinsic content,
dimension by dimension, via the slicing recurrence. No step introduces
a free parameter. No step selects from alternatives. One substantive
structural commitment --- the inner-product realisation of
distinguishable states, comparable to ``use real numbers for
analysis'' --- and the rest follows. Foundational footprint at
$\mathrm{RCA}_0$ for the dimension tower; no infinite-dimensional
ambient Hilbert space required.

\textbf{Part 0: Scale Variance from Orthogonality.} Pure mathematics. The slicing
recurrence of the unit ball contains one constant ($\sqrt{\pi}$), whose natural zero
generates two thresholds at $d = 19$ and $d = 217$. The Gamma function produces exactly
four distinguished dimensions: the volume maximum ($d = 5$), the area maximum ($d = 7$),
and the two thresholds. No fifth exists. The cascade invariant
$\Omega_{19}\times\Omega_{217} = 1.2051\times 10^{-120}$ is forced. Every step is a theorem
about the Gamma function. No physics enters.

\textbf{Part I: The Cosmological Constant from the Observer's Frame.} The hypothesis
enters. We observe four dimensions. The observer at $d = 4$ lives on the $S^3$ boundary
of $d = 5$, the volume maximum. Two observer-frame corrections connect the cascade's
pure-number invariant to the physical ratio measured in reduced-Planck units: the host
frame correction $(\Omega_5/\Omega_7)^2 = 9/\pi^2$ (from cascade reference $d_0 = 7$ to
observer's host $d_V = 5$), and the cube--sphere bridge
$\Omega_2/V_3^{\mathrm{cube}} = 4\pi/8 = \pi/2$ at the observer's spatial dimension
$d = 3$ (converting cascade sphere-area content to the cube-volume normalisation of
the reduced Planck mass):
\[
\frac{\rho_\Lambda}{M^4_{\mathrm{Pl,red}}}
= \frac{2}{\pi}\cdot\frac{9\;\Omega_{19}\;\Omega_{217}}{\pi^2}
= \frac{18\;\Omega_{19}\;\Omega_{217}}{\pi^3}
= 6.996\times 10^{-121}.
\]
Observed (Planck 2018): $(7.150\pm 0.13)\times 10^{-121}$. Leading deviation: $-2.2\%$,
placing $\Lambda$ in the descent-dependent population. After the Part~0 Supplement
first-order Gram correction, residual $-0.07\%$, inside the Planck $1\sigma$
uncertainty. The vacuum is 198 discrete layers of geometry with a natural floor at
$\Omega_{217}\approx 10^{-120}$.

\textbf{Part II: Quantum Mechanics from the Cascade.} Consecutive slicing axes are forced
orthogonal, producing complex structure, Hilbert space, the Born rule, and unitary
evolution --- without quantum postulates.

\textbf{Part III: General Relativity from the Cascade.} The cascade's complex spinor
structure intersected with Lovelock uniqueness forces $d = 4$, Lorentzian signature, and
Einstein's equation. The dark energy equation of state is $w = -1$ exactly.

\textbf{Part II\,=\,III: Quantum Gravity without Quantising Gravity.} The quantum and
gravitational projections share the same source, propagator, state space, and hypothesis.
Gleason forces the Born rule; Lovelock forces Einstein's equation; both are unique at
$d = 4$. The standard QM/GR conflicts dissolve. Black hole entropy $S = A/4$ is hidden
geometry from boundary dominance: the factor $1/4$ is $1/d$ at the observer's dimension.

\textbf{Part IVa: The Standard Model from the Cascade (Gauge Group).} Bott periodicity,
Adams' theorem, and the hairy ball theorem jointly force
$\mathrm{SU}(3)\times\mathrm{SU}(2)\times\mathrm{U}(1)$, its symmetry breaking pattern,
and three fermion generations.

\textbf{Part IVb: The Standard Model from the Cascade (Masses and Couplings).} The
geometric-topological factorisation of the fermion mass gives fourteen precision
predictions --- all sub-3\%, zero free parameters.

\textbf{Part V: Cosmology from the Cascade.} The cascade derives
$\Omega_m = 1/\pi$, $\Omega_b = 1/(2\pi^2)$, $\Omega_r = 1/(4\pi^7)$,
and $H_0 = 66.78$~km/s/Mpc from the Friedmann equation with Part~I's
observer-corrected $\rho_\Lambda/M_{\rm Pl,red}^4 = (2/\pi)\,I$. Cascade
$H_0$ sits $0.9\%$ below Planck's central $67.4$ at leading order and
closes to essentially the Planck value under the Part~0 Supplement Gram
first-order correction; it is not compatible with the SH0ES $73.0$. The
cascade's $r_d \approx 147.75$~Mpc is essentially equal to Planck's
$147.60$~Mpc, so the cascade and Planck share a ruler; the cascade
predicts $w = -1$ as a structural theorem and offers no ruler-based
alternative account of the DESI apparent-$w$ signal, which challenges
both cascade and $\Lambda$CDM in the same way. The Friedmann equation
has every coefficient determined by $\pi$. Universe age
$t_0 = 13.88$~Gyr ($+0.6\%$).

\bigskip

The hypothesis is currently not falsified. Papers~II--III and II\,=\,III
are exact: every result matches observation with no approximation and no
discrepancy. Papers~I, IV, and V use leading-order approximations; at
leading order their predictions separate into two populations --- descent-dependent
quantities with uniformly negative deviations, geometric quantities with
positive deviations. The cosmological constant now lives in the
descent-dependent population ($-2.2\%$ leading, $-0.07\%$ after the
Part~0 Supplement Gram first-order correction), alongside $\alpha_s$,
$m_\tau/m_\mu$, $v$, $\ell_A$, and $\Omega_m^{\rm Bott}$. Part~IVb then derives a structural
family of cascade potential shifts $\delta\Phi = \alpha(d^*)/\chi$, sourced at
distinguished cascade layers from Part~0's four-dimension tower and weighted by the
same Euler characteristic $\chi(S^{2n})=2$ that appears in the $2\sqrt{\pi}$ fermion
obstruction. Eight Standard Model precision observables close within experimental
precision with zero fitted parameters, using shifts sourced at four
of Paper~0's distinguished dimensions:
$\alpha(14)/\chi$ closes $\alpha_s$ and $m_\tau/m_\mu$ ($d=14$,
$\mathrm{U}(1)$ gauge layer);
$\alpha(19)/\chi$ closes $m_\tau$ absolute and $\ell_A$ ($d_1=19$,
phase transition);
$\alpha(5)/\chi^3$ closes $\sin^2\theta_W$ and $\Omega_m$ ($d_V=5$,
volume maximum, opposite signs for the two populations);
$\alpha(7)/\chi^2$ closes $\theta_C$ and $\alpha(7)/\chi^4$ closes $b/s$ ($d_0=7$,
area maximum, two chirality powers).
All are closed forms in $\Gamma$ function values; three reuse pairs
share the same shift across independent observables.
A discrete elastic action on the cascade lattice
($S=\sum(2\alpha(d))^{-1}(\Delta\varphi)^2$ with $\varphi=\ln\Omega_d$)
generates this family: distinguished layers act as sources, the Green's
function decays as $\alpha(d^*)$, and each independent cascade channel
filters the signal by $1/\chi$ (one of two chirality basins selected). Falsification requires finding a precision
observable whose correction falls outside $\pm\alpha(d^*)/\chi^k$ at a
distinguished layer, or demonstrating that the channel-counting rule
($k=$ number of cascade sectors in the observable) fails for a new quantity.

\bigskip

The thought experiment that opens this page is the physical content of the series'
central result. The observer is on the $S^3$ shell of a 5D black hole, partway through
an asymptotic compactification that never completes. The cosmological constant is the
cascade's geometry measured from that shell. Time dilation is the local rate of the
compactification. Black hole horizons are where the transition locally completes. The
cosmological constant and time dilation are proportional, in the ratio $d = 5$. The
Gamma function knows how big the universe is because the universe is the unit ball,
descended.

\end{document}